### 2018-11-23 Paper_reorganized.Rmd
Reorganized paper

---
title             : "Quantitative involvement moderates the interplay between mothers' versus fathers' qualitative parenting behaviors and children's emotional adjustment development"
shorttitle        : "QUALITATIVE AND QUANTITATIVE INVOLVEMENT"

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Renske Keizer"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Utrecht University, Methodology & Statistics"
  - id            : "2"
    institution   : "Erasmus University Rotterdam, Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["DSEM_2017.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : yes

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
mask <- TRUE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

Emotional adjustment, defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001], is an important developmental outcome, integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013; @aldaoEmotionregulationStrategiesPsychopathology2010]. Fathers have been hypothesized to play a unique role in children's emotional adjustment development, because they encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000]. The classical notion of the “attachment relationship” is rooted in security and exploration (REF). Fatherhood researchers have posited that, whereas the mother-child attachment relationship primarily provides warmth and security, exploration might be somewhat more central to the father-child relationship, which can has thus been construed as an “activation relationship” [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a broad spectrum of emotions, thereby structuring opportunities to expand children's emotional repertoire.

## Fatherhood theory

The past decades have seen major strides in the consolidation of theories to guide fatherhood research. Initially, such research was inspired by two societal shifts: Increasing divorce rates, and mothers' greater labor force participation. Both shifts impacted the extent of fathers' involvement in childrearing. Early research focused primarily on the *quantity* of paternal involvement: Whether fathers were present or absent in divorced families, which constitutes a dichotomous operationalization of quantitative involvement, and how much they contributed to the work of childcare in intact families, i.e., a continuous operationalization [@lambBiosocialPerspectivePaternal1987]. Over time, the notion of the father as a coparent gained ground [@pleckFatherhoodIdealsUnited1997], according to which fathers share the responsibilities of childrearing with mothers, and might, to some extent, adopt roles that were historically fulfilled by mothers. To keep pace with changing conceptualizations of fatherhood, research on father involvement evolved to encompass *qualitative* dimensions as well, including warmth and control [or: demandingness; @maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].

Although it is now widely acknowledged that both the quality and quantity of parental involvement are relevant factors in child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated. This might be, in part, because research interests have historically varied across discipline lines: Developmental psychology has focused primarily on how qualitative parenting behaviors influence child development, whereas sociology and gender studies have focused more on involvement quantity, and its implications for maternal wellbeing [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement in relation to children's emotional adjustment.

## Qualitative parenting and emotional adjustment

### Parenting effects

The standard family environment model posits that the quality of parents' involvement is a proximal mechanism, shaping children's development [@amatoParentalDivorceMarital2008]. Qualitative parenting is typically examined in terms of the constituent dimensions of the parenting styles taxonomy [@baumrindChildCarePractices1967]: Responsiveness and demandingness [@maccobySocializationContextFamily1983]. Responsiveness refers to warm and supportive parenting, which substantial evidence indicates is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Two aspects of demandingness investigated in the present study are consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. The literature on links between demandingness and emotional adjustment is more limited, and mixed [@morrisRoleFamilyContext2007]; in part because some forms of demandingness help children meet societal expectations, whereas other forms are overly restrictive, and undermine children's autonomy needs [@barberCentralityControlParenting2013]. A fourth dimension, which has consistently been linked with children's emotional maladjustment, but is distinct from the parenting styles typology, is negative parenting [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. We operationalized negative parenting as one of its specific subdimensions: parental anger. The anger parents express in interactions with children might reflect faillure to regulate their own negative emotions, and models the use of anger in communication, and undermines children's emotion regulation [@scaramellaParentingChildrenDistress2008].

<!-- 
Inductive reasoning is thought to be integral to children's development of moral emotions [@HoffmanEmpathymoraldevelopment2000], and might thus play a positive role general emotional development as well.
-->

Although some research suggests that both parents' qualitative parenting behaviors influence children in similar ways [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique roles of mothers versus fathers. Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010]. Furthermore, whether for evolutionary or cultural reasons [@bogelsDoesFatherKnow2011], mothers and fathers may tend to adopt somewhat specialized roles in terms of qualitative parenting: The aforementioned notions of "attachment" and "activation" relationships posit that mothers empathize more with their children, and  provide more warmth and support, whereas fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior [@majdandzicFathersChallengingParenting2014; @paquetteTheorizingFatherchildRelationship2004]. Empirical research supports these unique roles for mothers and fathers in children's emotional adjustment development: Mothers tend to encourage the expression of emotions, and respond warmly and supportively, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for emotional adjustment in adolescence, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. This differentiated pattern was also found with regard to younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Furthermore, there is some evidence that mothers' negative parenting has a stronger negative impact upon children's emotional adjustment than fathers' negative parenting [@changHarshParentingRelation2003]; perhaps because this behavior is incongruous with mothers' prototypical role as an attachment figure. Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.

### Child effects

According to the transactional view of family relationships, parents and children reciprocally influence one another [@kerrPerceivedParentingStyle2012; @amatoParentalDivorceMarital2008]. This view of the family as a dynamic system whose members reciprocally influence one another has become widely accepted [@kerrPerceivedParentingStyle2012]. Nevertheless, examining bi-directional effects is not yet standard practice in the childhood literature [@flouriFathersBehaviorsChildren2010]. Available empirical evidence suggests that children's adjustment problems elicit less warmth, more demandingness, and more negative parenting from parents [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @LenguaBidirectionalassociationstemperament2005]. In the adolescent literature, bi-directional effects are more commonly examined. This might be because the notion of children influencing parents is more salient in adolescence, when parent-child relationships become increasingly egalitarian. Findings from the adolescent literature are broadly in line with findings in childhood; suggesting that emotional adjustment problems predicts decreased parental support, and greater control [@OtterpohlCrossLaggedRelationsParenting2015; @VanderGiessenPerceivedAutonomySupport2014; @Wijsbroekdirectioneffectsperceived2011]. Thus, parents' responses to maladjustment appear to be a combination of punitive responses, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness. Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively [@Dixaffectiveorganizationparenting1991]. We might thus expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.

### Assessing the direction of effects

Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if there are stable between-family differences in the variables studied [@Hamakercritiquecrosslaggedpanel2015]. Recent methodological innovations provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@Hamakercritiquecrosslaggedpanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@AsparouhovDynamicStructuralEquation2018].

Several studies have re-examined the direction of effects between parents and children using these innovative methods, but mostly in adolescent samples. One such study found that predictive effects between parenting and delinquency adjustment that were apparent using classical methods, were found to be explained by correlations between stable between-family differences in parenting and adjustment, once these components were included in the model [@KeijsersParentalmonitoringadolescent2016]. Another study examined the interplay between maternal and paternal qualitative parenting dimensions and adolescents' emotional adjustment, and similarly found that most associations manifested as between-family level correlations. Furthermore, child effects were found to substantially outnumber parenting effects [@vanlissaRoleFathersMothers2018a]. The authors suggested that this might be due to the relative equality of adolescent-parent relationships, as compared to parent-child relationships. If this interpretation is correct, we might expect to find more prominent parenting effects in the present study, with a younger sample. These preliminary findings highlight the fact that the direction of effects between parents and children, while controlling for stable between-family differences, remains an important issue to be examined. At this stage, however, it is difficult to formulate specific hypotheses about directionality, and the level at which associations between parenting and child development should occur.

<!-- Kroll: No evidence for effect of quantitative involvement This might help explain why some studies, focusing on absolute involvement, have concluded that fathers' quantitative involvement does not predict children's adjustment when controlling for mothers' involvement>> Kroll actually used absolute involvement and "who does most"-questions, which measure relative involvement -->

## Quantitative involvement

### Operationalizing quantitative involvement

Quantitative involvement has been researched extensively in terms of the amount of time spent caring for the child [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the the way childcare tasks are divided is more relevant for children's adjustment than the absolute amount of time invested by each parent [@KrollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003]. Indeed, research has shown that the equity of the childcare distribution affects child development directly [@deutschPaternalParticipationChild2001], and impacts the emotional climate of the interparental relationship [@CraigDoesFatherCare2006]. When childcare is unequally distributed, women often perform a disproportionate share of the work. This strains the marital relationship [@shockleyItNotWhat2018], and the resulting stress erodes parenting quality, particularly for parents of difficult children [@SheltonInterparentalconflictnegative2008].

There are also methodological advantages to focusing on the distribution of childcare. Operationalizing quantitative involvement in terms of both parents' absolute involvement, e.g., as hours spent caring for the child, incurs a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care [see @deutschPaternalParticipationChild2001]. This confound is mitigated by operationalizing quantitative involvement in terms of the childcare distribution, and controlling for total joint care. Some prior research has attempted to capture the childcare distribution directly, by asking which has most responsibility for certain tasks, on a continuum from the mother doing everything to the father doing everything [@KrollEarlyFatherInvolvement2016]. When quantitative involvement is measured in terms of absolute involvement, however, an elegant statistical solution can disentangle the childcare distribution from total joint care. The so-called "latent mean and difference" approach, which is typically used to capture the difference between consecutive longitudinal measurements [@McArdleLatentVariableModeling2009], can be utilized to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other. This approach taps directly into a theoretically relevant construct, and prevents confounding. 
## Main effects of quantitative involvement

### Effects on qualitative involvement

In many countries, mothers continue to spend more time with their children than fathers [@hofferth28669resident], particularly when it comes to routine care tasks [@craigHowMothersFathers2011]. Fathers' involvement is on the increase, however [@YeungChildrenTimeFathers2001]. The amount of time parents spend with children is likely to predict the quality of parenting behaviors [@sabattiniRelationMothersFathers2004], through practice and experience. Indeed, more involved parents tend to experience greater parenting self-efficacy [@GialloMothersfathersinvolvement2013]. We might thus hypothesize that greater total joint care should positively predict both parents' parenting quality.

By the same logic, we might expect that, when one parent is more involved than the other, the parenting quality of the more involved parent would improve. However, an argument to the contrary can be made as well. Retrospective reports by 19-year old participants indicate that an unequal childcare distribution tends to go hand in hand with suboptimal qualitative parenting behaviors - specifically, increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. Greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship. Childcare responsibilities are a heavy burden if they come down on a single parent's shoulders [@deutschPaternalParticipationChild2001]. Mothers thus benefit directly when fathers relieve some of this burden [@PleckPaternalinvolvementRevised2004]. Consequently, mothers tend to be more satisfied when fathers are more involved [@PharesFathersMothersInvolvement2009]. An unfair distribution of childcare, on the other hand, is associated with greater maternal anger and resentment [@RossGenderParenthoodAnger1996]. This strains the interparental relationship, interferes with sensitive parenting, and undermines child wellbeing [@davies2006interparental]. In other words, when care is distributed relatively equally, the entire family system will likely be more harmonious. We might thus expect greater differential involvement to aversely affect both parents' parenting quality, as well as children's emotional adjustment.

### Effects on emotional adjustment

Although the quantity of paternal involvement has received substantial interest in the fields of sociology and gender studies, this literature has focused primarily on its implications for mothers' time use and wellbeing. Consequences for children's emotional outcomes have received less attention to date. Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @BronfenbrennerEcologyHumanDevelopment1979]. This suggests that the total amount of care provided should be positively associated with children's emotional adjustment. Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their wellbeing [@deutschPaternalParticipationChild2001].

<!--Fathers are as capable as mothers of caring for their children **(Lamb 1997; Yeung et al. 2001)**, and on average, -->

<!-- dit mogelijk vóór de effecten van quant plaatsen? -->

## Interactions between involvement quality and quantity

Interactions between involvement quality and quantity have largely been overlooked in the aforementioned theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@JaffeeLifeFatherBenefits2003]. However, It is conceivable that children's emotional adjustment might be differently impacted by paternal control if that father is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent. Conversely, children's emotional adjustment might elicit different qualitative parenting behaviors from hands-on fathers, who are aware of the reasons for their children's emotional troubles, than from out-of-touch fathers, who might perceive maladjustment as a problem behavior that should be sanctioned. To capture patterns like these, a model should account for interactions between involvement quality and quantity.

In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@FeinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the role of parent. Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families. One dimension of coparenting is involvement quantity, which Feinberg defines as the "division of labor" in childcare. Note that this definition parallels our operationalization of quantitative involvement. The division of childcare labor is proposed to affect children's adjustment and parents' qualitative childrearing behaviors directly, and to interact with the link between these two constructs.

There is some preliminary empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@JiaTransactionalRelationsFather2012]. Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@WangParentalInvolvementAfrican2014]. Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.

## Sex differences

<!-- pubertal timing and associated changes in hormone levels during adolescence are related to sex-typed attributes, including emotional distress (Brooks-Gunn, Graber, & Paikoff, 1994; Susman, Dorn, & Chrousos, 1991).-->

There is substantial evidence that parents tend to raise boys and girls differently. Prior research has indicated that parents tend to be more controlling with girls than with boys **(e.g., Pomerantz & Ruble, 1998; Smetana & Daddis, 2002)**, and express more negative parenting with boys than with girls **(Straus and Stewart 1999)**. Another point to consider is the interplay between parent and child sex. According to social learning theory, the same-sex parent plays a stronger role in a child's development **(Bussey & Bandura, 1999)**, particularly in the socialization of sex-typed behavior, such as display rules for emotions [@Brodysocializationgenderdifferences2000]. Thus, we might expect child sex to predict the same-sex parents' qualitative involvement, and associations between parenting and emotional adjustment might be stronger for same-sex parent-child dyads. There is also evidence that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships **(Lovas, 2005; Schoppe-Sullivan et al., 2006)**. These findings suggest that boys tend to receive a "tougher" upbringing than girls. In sum, it is clear that sex differences are a relevant factor in the interplay between parenting and emotional adjustment. Specifically, we might expect to find parents, display more responsiveness, structure, and inductive reasoning with girls than with boys. Boys, on the other hand, might receive more negative parenting. Furthermore, we might expect associations between qualitative parenting behaviors and emotional adjustment to be stronger between children and the same-sex parent.

There is little evidence for sex differences in emotional adjustment in young children, as indicated by a meta-analysis which found no sex differences in the temperament dimension of negative affectivity, based primarily on infant and toddler samples [@Else-questGenderdifferencestemperament2006]. Evidence for such sex differences is sometimes found in older children [@SilkAdolescentsemotionregulation2003], but these might appear and intensify in puberty [@HillIntensificationGenderRelatedRole1983]. We consequently did not expect to find sex differences in children's average levels of emotional adjustment.

## The present study

The present study aimed to make several contributions to the literature on fathers' role in their children's emotional adjustment development. First, we included both mothers and fathers, in order to tease apart their common and unique roles. Second, our operationalization of involvement incorporated both qualitative involvement, in terms of specific parenting behaviors, and quantitative involvement, operationalized as total joint childcare and differential involvement. Third, we employed recent methodological advances to disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative parenting behaviors and emotional adjustment. Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.

Based on our review of the literature, we expected specific dimensions of qualitative involvement to be associated with emotional adjustment at both the within- and between-family level. The literature suggests that warmth, consistency, and inductive reasoning should be positively associated, whereas negative parenting should be negatively associated, with emotional adjustment. With regard to child effects, we might expect some of these paths to be reversed: Positive adjustment might lead to diminished inductive reasoning and parental anger. We further hypothesized that quantitative involvement would predict children's adjustment directly: Greater joint total care should predict better emotional adjustment, whereas a more skewed distribution of care should predict poorer emotional adjustment. Greater joint care should also predict both parents' qualitative involvement, in the sense of lower parental anger, and greater consistency, inductive reasoning, and warmth. A more skewed distribution of care, on the other hand, should show these associations only for the more involved parent. Importantly, we also hypothesized that involvement quality and quantity would interact, such that a more skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children, and parents' adaptive responses to children's maladjustment. In other words, we expected that an unfair childcare distribution would predict suboptimal parenting, and parental responses to maladjustment, due to the strain such a distribution places upon the entire family system.  Finally, we accounted for potential effects of child sex.

# Methods

## Participants

We employed data from the`r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. A detailed description of the study design, sample, and procedure is available at `r masked(" http://data.growingupinaustralia.gov.au/")`.

Our target population consisted of native `r masked("Australian")`, non-minority, intact, biological families. This yielded a subsample of 1884 families. At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$). Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$; 669 father-only, 41 mother-only). The modal annual income for fathers was between \$26,000-\$51,999 (N = 663), and for mothers, \$25,999 or less per year (N = 1152).

## Measures

### Quantitative involvement.

Measures of quantitative involvement varied from wave to wave, due to children's changing care needs. We used reports at T1, when both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" (e.g.: "In the past month how often did you help this child brush his/her teeth?").

### Qualitative involvement.

Parents completed measures of several qualitative parenting dimensions. Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @PatersonAssociationBehaviouralAdjustment2001]. Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth **(NLSCY; Statistics Canada, 2000)**. For these instruments, parents reported on their own behavior using five-point scales, ranging from "never or almost never" to "always or almost always".

### Emotional adjustment.

Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the Strengths and Difficulties Questionnaire, reported by teachers on a three-point scale from "not true" to "certainly true" **(e.g., "Often unhappy, downhearted or tearful"; Goodman, 1999)**. Scale descriptives are displayed in Table X, and scale score intercorrelations are reported in the online supplement SX.

## Scale construction

We calculated mean scores for all of the study variables. Scale descriptive statistics are provided in Supplementary Table 1, and scale intercorrelations are provided in Supplementary Table 2. To establish that metric measurement invariance could be imposed across measurement waves, and across mothers and fathers, we estimated structurally invariant CFA's for all of the study variables, and then imposed constraints on the factor loadings across time and reporter. Although chi-square difference tests indicated that the fit was significantly worse for metric invariance models than for structural invariance models for all scales (all $ps \leq$ .03$), chi-square is overpowered in large samples. The difference in fit between the structural invariance models and metric invariance models was smaller than .01 according to conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). This indicates that it is defensible to assume that the questionnaires measure the same constructs across time and reporter.

## Strategy of analyses

We used dynamic structural equation modeling (DSEM) to address our research questions. This approach analyzes longitudinal data as a multilevel, autoregressive model. First, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences, and within-individual fluctuations. In the multilevel model, stable individual differences are captured at the between level, by random intercepts which are based on all scores originating from the same individual. At the within level, fluctuations from these random intercepts are analyzed at the within level. Moreover, time-lagged copies of all within-level variables are created, in order to estimate a cross-lagged autoregressive model. The coefficients of this cross-lagged autoregressive model are free to vary between individuals, and can thus be predicted by between level moderators, such as quantitative involvement and child sex.

This approach is particularly suited to address our research questions, for several reasons. First, controlling for stable between-individual and between-family differences allows us to obtain more accurate estimates of mothers' and fathers' unique roles in child development: Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach estimates the interplay between qualitative parenting behaviors and emotional adjustment as a random effect, varying between families. We can directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment, by including total joint childcare, and the childcare distribution, as between-family level predictors of the within-family cross-lagged effects between these variables. Child sex was included as a moderator in the same way.

DSEM models can only be estimated using Bayesian methods. We ran two chains of Mplus 8.0's Gibbs sampler for 50000 iterations each. We used mildly informative priors for the variance/covariance matrices to aid convergence, and  regularizing priors centered on zero for all regression coefficients. We assessed convergence through visual inspections of the chains and posterior distributions. We report the posterior median, and a 95% highest posterior density interval (HPDI). This interval is the "narrowest" possible interval of the posterior distribution that contains 95% of samples from the posterior distribution. It represents the most likely population values of the parameter in question, given the model, priors, and data. Results are best interpreted by examining the effect size (standardized regression coefficients and correlations), and the width of the HPDI. However, for convenience's sake, we focus on interpreting results whose HPDI did not include zero. Significance is not a Bayesian concept, but we will use the shorthand "significant" to refer to parameters whose 95\% HPDI did not include zero.

# Results

Standardized and unstandardized model parameters are reported in Table 1. A graphical summary of the parameters discussed in the Results section is displayed in Figure 1.

<!-- structure notes
1) mothers and fathers, in order to tease apart their common and unique roles. 
2) incorporated quantitative involvement, in terms of total joint childcare and the childcare distribution, and qualitative involvement
3) interactions between qualitative and quantitative involvement
4) assess the direction of effects between parenting behaviors and emotional adjustment.

HYP:
3) Qualitative parenting associated with emotional adjustment at both the within- and between-family level
   * responsiveness, structure, and reasoning positively associated, negative parenting negatively associated
   * Correlations, and regressions 
     * Parent effects
   * Child effects, some paths reversed: Maladjustment should elicit decreased responsiveness, but increased structure, reasoning, and negative parenting. 
1) Interactions between involvement quality and quantity
   * skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children (parenting effects), and parents' adaptive responses to children's maladjustment (child effects).
   * unfair childcare distribution would predict suboptimal parenting
2) Direct effects of quantitative involvement on adjustment and parenting
   2.a) Adjustment
   * Greater joint total care should predict better emotional adjustment
   * a more skewed distribution of care should predict poorer emotional adjustment
   2.b) Greater joint care should also predict both parents' qualitative involvement
   * greater responsiveness and structure, and lower negative parenting.
   * A more skewed distribution of care, on the other hand, should show these associations only for the more involved parent.

4) child sex
-->

## Average associations between qualitative parenting dimensions and emotional adjustment

We hypothesized that, at both the within- and between-family levels, anger would be negatively associated, and consistency, inductive reasoning, and warmth would be positively associated with emotional adjustment.

### Between-family correlations

At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment. The random intercept of each family is based on their average level across all waves of the study. Correlations between random intercepts are estimated across all families in the sample.

In support of hypotheses, we found significant negative correlations between maternal and paternal anger, and emotional adjustment, as well as positive correlations between maternal and paternal consistency, and children's emotional adjustment. Contrary to hypotheses, we found significant negative correlations between maternal and paternal inductive reasoning, and children's emotional adjustment. This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment.

<!--(Discussion: Although inductive reasoning is considered to be a disciplinary behavior that stimulates children's positive development (better ref here), the fact that it is a disciplinary behavior might explain the negative association with emotional adjustment.).-->

Finally, contrary to hypotheses, we found no significant correlations between maternal and paternal warmth, and children's emotional adjustment, at the between-family level.

### Within-family correlations

At the within-family level, we examined correlations between the residuals of qualitative parenting behaviors and emotional adjustment. Within-family correlations are based on residuals of the study variables, across all waves and families. Contrary to our expectations, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. In partial support of hypotheses, we did find significant positive associations of adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. In support of hypotheses, we found significant positive within-wave associations between emotional adjustment and fathers' and mothers' warmth. Notably, these associations were nearly twice as strong for fathers' warmth, as compared to mothers' warmth.

### Within-family predictive effects

At the within-family level, we examined over-time cross-lagged effects between qualitative parenting and emotional adjustment. The intercepts of these random effects reflect the average cross-lagged effect across families. The variances indicate whether the effect differs in strength (and sign) between families.

#### Parenting effects

None of the intercepts of within-families parenting effects differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment. Thus, on average across families, mothers' warmth had a positive significant predictive effect on children's emotional adjustment. The variances of all parenting effects were significant, however, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were notably larger for fathers than for mothers, indicating that fathers have a much more variable effect on their children through specific parenting behaviors than mothers do.

#### Child effects

Contrary to hypotheses, the intercepts of the effects of emotional adjustment on maternal qualitative parenting behaviors were significant and positive, not negative, for  anger, non-significant for consistency, and positive for inductive reasoning. In line with hypotheses, the intercept for maternal warmth was positive and significant. Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year. 
In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts for child effects on paternal parenting were significant.
The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' childrearing behaviors differs between families.

## Associations between quantitative involvement and adjustment and parenting

We examined regression coefficients from total join involvement and relative involvement to the random intercepts of emotional adjustment and qualitative parenting behaviors. We found that greater joint care was negatively associated with emotional adjustment, counter to our hypothesis. Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment. With regard to the distribution of childcare, however, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.

With regard to parenting, we did not find the hypothesized negative effects of joint total involvement on parental anger, but, in partial support of our hypotheses, we did find that, for both parents, joint involvement positively predicted greater consistency, inductive reasoning, and warmth. Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers.

With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict reduced anger, and greater consistency, inductive reasoning, and warmth, only for the more involved parent. For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers.
For fathers' parenting behavior, we found that, in line with this hypothesis, fathers displayed less consistency, inductive reasoning, and warmth when mothers were relatively more involved than fathers.

## Interactions between involvement quality and quantity

To examine interactions between involvement quality and quantity, we tested whether between-family differences in relative and joint total quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment.

### Moderation of parenting effects

With regard to the moderation of parenting effects by quantitative involvement, we found no effects for anger, consistency, inductive reasoning, and paternal warmth. The only significant effects we found were for maternal warmth. Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively when both parents' total joint involvement was higher. This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement.

### Moderation of child effects

With regard to the moderation of child effects by quantitative involvement, we found no effects for anger and consistency. For inductive reasoning, we found that adjustment was more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning.

We further found that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when both parents' joint involvement was higher. In other words, the more both parents are more involved, the more negative the cross-lagged effect of children's adjustment on maternal discipline will be.

For warmth, we found that the effect of adjustment on maternal warmth was negatively attenuated, both when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers.

<!--	(Discussion: i.e., if the child is well-adjusted, that leads to less maternal maternal warmth, but only when both parents are more involved. Paradoxical, but keep in mind that maternal and paternal warmth are higher in families with greater joint involvement. So perhaps this means that mothers of maladjusted children, in families where joint involvement is low, coddle their children.)-->

## Effects of child sex

To explore sex differences, we included child sex as a predictor of the within-family cross-lagged effects, and of the random intercepts for emotional adjustment and qualitative parenting behaviors.

The only cross-lagged effects significantly moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning. Specifically, emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls. In other words, if children display better adjustment than usual in one year, their parents decrease inductive reasoning more strongly in the next year when children are male.

With regard to sex differences in the random intercepts, we found no sex differences in emotional adjustment. For qualitative parenting behaviors, we found that both mothers and fathers displayed more anger and inductive reasoning, and fathers displayed less warmth, towards sons as compared to daughters. No significant differences were found for both parents' consistency, and for maternal warmth.

# Discussion

We set out to disentangle mothers' versus fathers' unique roles in children's emotional adjustment development. In doing so, we incorporated both qualitative and quantitative measures of parental involvement, and examined interactions between these two aspects of involvement, which have historically been examined mostly in isolation. To this end, we used innovative methods which enabled us to include between-family differences in quantitative involvement, operationalized as total joint care and the relative childcare distribution, as predictors of within-family cross-lagged effects between qualitative involvement behaviors and emotional adjustment.

## Associations between qualitative parenting and emotional adjustment

We hypothesized that, at both the within- and between-family levels, anger would be negatively associated, and consistency, inductive reasoning, and warmth would be positively associated with emotional adjustment. Breaking down the results by parenting behavior, we found mixed evidence for these hypothesized patterns.

With regard to parental anger, we found mixed evidence for the predicted negative associations with emotional adjustment. We primarily found support for these associations at the between-family level (i.e., in terms of the random intercept correlations). This means that, in families where mothers or fathers display low anger, children tend to be more emotionally well-adjusted. Contrary to our expectations, however, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger. Similarly, the average within-family predictive effects of parental anger on adjustment were non-significant. This suggests that the brunt of negative associations between anger and adjustment are explained by differences between families. One potential explanation for the absence of within-family associations might be a ceiling effect; it might be the case that, once the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger no longer make much of a difference. Another explanation might be that any further effects of fluctuations in anger play out on a different time scale than the one captured by our study; for example, if such effects primarily play out immediately, in dyadic interactions between parents and children. Finally, we found that, on average, emotional adjustment showed a paradoxical positive predictive effect on maternal anger. This means that, if children demonstrated better emotional adjustment than usual in one year, their mothers tended to display more anger than usual in the next year. The literature on expectation-based loss aversion offers a possible explanation for this puzzling finding [@KoszegiModelReferenceDependentPreferences]. Note that, on average, emotional adjustment showed no-to-negative stability coefficients in all models. That means that, when children demonstrated better emotional adjustment than usual in one year, this was more likely than not to be followed by a year with worse adjustment. Mothers, on the other hand, might expect their children's emotional adjustment to be relatively stable. When a child falls short of the mother's expectations for stable emotional adjustment, the perceived "lost utility" resulting from having to deal with a child that is less well-adjusted than expected might render a mother more prone to experiencing anger towards the child.

We found partial but straightforward support for the hypothesized positive associations of parental consistency with emotional adjustment; again primarily at the between-family level, where both maternal and paternal consistency were positively correlated with children's emotional adjustment. This means that, in families where parents tend to display high consistency, children tend to be more emotionally well-adjusted. In terms of within-family correlations, we found the predicted links for mothers', but not fathers', consistency. On average, we found no support for within-family predictive links between consistency and adjustment. It is interesting that we found only between-family links and no over-time predictive effects. There are several reasons why between-family links might be observed: For instance, these can reflect the accumulated effect of within-family causal effects at an earlier age. However, our study included children from the age of six, and - although it is possible that consistency did, on average, affect children's emotional development prior to the preschool years, it seems unlikely that this effect would have dissipated entirely by the time our study commenced. Alternatively, associations at the between-family level can reflect stable differences between families that impact both constructs; for example, it has been proposed that the economic hardship associated with low SES diminishes parents' capacity for consistent parenting, while at the same time, undermining children's socio-emotional adjustment [@McLoydImpactEconomicHardship1990]. The fact that we observed no within-family predictive effects, despite including children from a very young age, might suggest that the "causal" role of consistency in children's emotional adjustment is more limited than prior research has concluded based on methods that did not control for stable between-family differences [REF].

With regard to inductive reasoning, we found mixed support for the hypothesized positive associations with emotional adjustment: Associations were opposite in sign at the between- versus within-family levels. At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment. This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior, although a constructive one that stimulates children's internalization of rules and expectations **(better ref here)**. Parents' reports of engaging in inductive reasoning might thus be confounded with how often they discipline their children. Disciplining children might result in, be precipitated by, or share a common cause with emotional maladjustment. Paradoxically, at the within-family level, emotional adjustment was positively correlated with paternal inductive reasoning, and emotional adjustment had a positive prospective effect on maternal inductive reasoning. In other words, increased maladjustment goes hand in hand with increased inductive reasoning. The fact that between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment might suggest that between-family differences in inductive reasoning tap more strongly into the amount of discipline parents engage in, whereas the within-family fluctuations might more strongly reflect changes in the way parents discipline their children.

Finally, we found partial, but unequivocal, support for the hypothesized positive associations between warmth and emotional adjustment. These links were absent at the between-family level, but abundant at the within-family level. Specifically, we found within-family correlations for both parents' warmth; an average prospective effect of maternal warmth on children's emotional adjustment, and child effects of adjustment on both parents' warmth. In other words, well-adjusted children evoke more parental warmth. The only socialization effect that was significant on average, which suggests that it is applicable to most families, was this predictive effect from maternal warmth to children's emotional adjustment. However, it is important to note that *all* within-family cross-lagged effects had significant variance components, highlighting the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.

## Quantitative involvement and emotional adjustment

The present study operationalized quantitative involvement in terms of total joint care provided by parents, and the childcare distribution, which captures differences between mothers and fathers. This operationalization addresses a criticism raised in the literature, that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. For total joint care, we found that, counter to our hypothesis, joint care was *negatively* associated with between-family differences in children's emotional adjustment. One potential explanation of this finding is suggested by self-determination: Perhaps excessive levels of involvement undermine children fundamental needs for autonomy and a sense of competence, thereby leading to increased emotional maladjustment [@DeciFacilitatingoptimalmotivation2008]. Research on "helicopter parenting" provides some support for this interpretation, as psychological wellbeing has been found to be lower in children of overly involved parents [@SchiffrinHelpingHoveringEffects2014]. It should be noted, however, that this literature is mostly cross-sectional, and focused on adolescents. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were more involved than fathers - even after controlling for the total amount of childcare provided by parents. Moreover, mothers were more involved than fathers, on average, as indicated by the positive significant mean of the latent variable capturing differential involvement. This suggests that the majority of children experienced the disadvantage in emotional adjustment associated with differential involvement. Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001]. Of course, causality cannot be inferred from correlational findings at the between-family level. Future research should examine potential pathways, such as whether fathers shy away from childcare responsibilities when children display poor emotional adjustment, or whether children's adjustment suffers when they feel like both parents are not equally invested in their wellbeing, or whether differential involvement strains the marital relationship, and this strain in turn spills over and affects children's emotional adjustment.

## Interactions between qualitative and quantitative involvement

We found limited evidence that within-family parenting effects - cross-lagged effects from qualitative involvement behaviors on emotional adjustment - were moderated by quantitative involvement. One such effect, which was in line with predictions, is that the effect of maternal warmth on adjustment was contingent on both total joint care, and the relative childcare distribution. Mothers' warmth predicted children's adjustment more strongly if mothers were relatively more involved than fathers. This suggests that having a warm mother is of greater developmental importance with regard to children's emotional adjustment if fathers are less involved with the day-to-day minutiae of childcare. Conversely, we also found that mothers' warmth was less strong of a predictor of adjustment development when both parents were more involved. This suggests that maternal warmth is less essential when both parents are highly involved. Together, these findings are reminiscent of a "buffering" effect, where mothers' warmth can make up for any disadvantages incurred by having a relatively uninvolved father.

Interactions between qualitative and quantitative involvement were more abundant in the context of child effects on parents; i.e., the way parents respond to children's emotional adjustment or maladjustment. Specifically, we found that when mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Considering that inductive reasoning is a form of discipline, this finding suggests that, when mothers are relatively more involved in care tasks, fathers take more of a role as disciplinarian.

<!-- the domains of discipline and play tend to be the areas in which fathers feel most comfortable (Parke 2000).
 Mothers’ satisfaction with the division of labor within the household was consistently related to greater levels of paternal responsibility in all four domains (school work, discipline, daily care, fun activities), but fathers’ satisfaction was only related to greater paternal involvement in two of the four domains (discipline and fun activities). 
-->
This finding is complemented by the negative effect of total joint involvement on the link from emotional adjustment to maternal inductive reasoning, which suggests that mothers also have more of a role as disciplinarian in families where both parents are highly involved. We also found that emotional adjustment became less predictive of maternal warmth when parents' joint involvement was higher, and when mothers' relative involvement was greater compared to fathers. This finding can be interpreted in the context of the sizable average effect of adjustment on maternal warmth. On average, increases in emotional adjustment predicted substantial increases in maternal warmth. In families where mothers' involvement was higher - either in terms of joint, or relative involvement - maternal warmth was thus less contingent on children's adjustment.

## Differences between fathers and mothers

Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development. These differences all played out at the within-family level, however. Stable between-family differences in either parent's qualitative parenting were all similarly associated with children's adjustment.

At the within-family level we found stronger associations with emotional adjustment for maternal consistency, and for paternal reasoning and warmth. As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample. This might help explain the stronger associations for maternal consistency - a behavior which implies, by definition, repeated interactions with the child. For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can be expressed in one-off interactions, such as when disciplining a child (inductive reasoning), or during play time (warmth). Moreover, the fact that the association between inductive reasoning and emotional adjustment was stronger for fathers than mothers is in line with the classic view of the father as disciplinarian [REF].

With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers. This is in line with prior research, which suggested that warmth is an important socialization mechanism for mothers [REF mezelf].

With regard to child effects on parenting, we found substantial evidence that mothers' parenting was more contingent on children's emotional adjustment than fathers' parenting. Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.

<!-- For anger, the one difference between mothers and fathers was in Within-family predictive effects suggested that, over-time changes in mothers' anger were more contingent on children's emotional adjustment than fathers' anger. We explained this finding in the context of expectation-based loss aversion, resulting from children falling short of mothers' expectations for positive adjustment. The fact that this effect was present for mothers, but not for fathers, might be explained by the fact that mothers typically spend more time with the child [REF] - as evidenced in our study by the significant and positive mean for the latent difference between maternal and paternal involvement. A parent who spends more time with the child might develop stronger expectations for the child's future behavior. -->

## Direction of effects

Historically, the field of developmental psychology focused primarily on how parents "shape", or cultivate, their children's development. Nowadays it is commonly acknowledged that parents and children exert mutual influences. However, commonly used developmental methods have been criticized for failing to properly reveal the direction of effects: When stable between-family differences are not accounted for, estimates of cross-lagged effects can be substantially biased [@hamakerCritiqueCrosslaggedPanel2015]. One recent study examined the interplay between qualitative parenting behaviors and emotion regulation development in adolescence, and found that child effects on parents substantially outweighed parent effects on children [@vanlissaRoleFathersMothers2018a]. The authors speculated that this finding might be due to the relatively egalitarianism of parent-child relationships in adolescence, as compared to earlier childhood. The present study casts doubt on this interpretation, as we found that child effects outweighed parenting effects, even at a young age. Although still preliminary, the prominence of child effects across two large longitudinal studies, conducted on samples from different continents and at different age ranges, invites the question: Are parents shaping children's development, or are they responding to it? An extreme interpretation of the findings from our study and this preceding work might be that parents' influence is limited to "setting the stage" for children's development, through their stable differences from other parents, and from there on, are merely adjusting to children's developmental needs. This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations. For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time. Consequently, stable parent characteristics can cause change in children. Future research might examine alternative theoretical conceptualizations of parental influences in child development. Another point to consider is that parent effects on children might play out on a different time scale than is typically captured in macro-longitudinal studies. If parents shape their children through specific behaviors in dyadic interactions, which are not captured by annual reflective self-reports, then we might fail to capture parenting effects. Future research might test this conjecture by applying DSEM methods to dyadic interaction data.

<!--To what extent these findings should be interpreted as "child-driven" effects, versus "parents adjusting to children's developmental needs".-->

<!--. This suggests that reciprocal dynamic processes might be less prominent than previous research suggested. Instead, third variables, such as socio-economic status, or shared genetic factors, might play a larger role - another outstanding issue on which further research is required (REF Flouri). -->

## Sex differences

In line with prior research, we found that, on average, parents raise boys and girls differently. Specifically, we found that mothers and fathers both displayed more angry parenting and inductive reasoning with sons than with daughters, and that fathers displayed less warmth towards sons. These findings suggest that parents are stricter towards their sons. Parenting effects on children were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors. Child effects on parenting, however, were affected: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were boys. In other words, parents relax their verbal discipline more in response to positive emotional adjustment when children are male, than when they are female. When viewed in the context of the fact that parents are generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.

## Strengths and Limitations

This study had several important strengths, which allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated measures of quantitative involvement, in terms of the amount of time spent in childcare, as well as qualitative involvement, in terms of specific parenting behaviors. This allowed us to provide a more nuanced overview of the way parental involvement is linked with emotional adjustment development, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Additionally, although some prior studies have also incorporated both quality and quantity of parenting, this work has typically included the absolute amounts of maternal and paternal involvement. A limitation of this approach is that it is not possible to distinguish the effect the childcare distribution from the effect of the total amount of care children receive [see @deutschPaternalParticipationChild2001]. We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach provides more accurate assessments of dynamic reciprocal processes within the family. Our findings contributed to a growing literature that suggests that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. Furthermore, teachers' reports should be relatively unbiased, because they are likely to be "blind" to parents' childrearing behavior, and have no stake in the parent-child relationship.

One potential limitation of the present study is that teachers' reports of emotional adjustment might be more "noisy" than self- or parent-reports. A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment. Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are known to be poor judges of others' emotional states [REF Ickes]. Consequently, it is important to consider the fact that our effect sizes might be attenuated by noisy measures of emotional adjustment. Teachers' reports do have clear benefits compared to alternatives: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies. Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983]. We focused on specific parenting behaviors for two main reasons. First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction. This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in childrearing. Moreover, the methods we used to disentangle stable between-family differences from within-family processes are not suitable for categorical variables, such as parenting style. Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation. Finally, one potential limitation is that we only included quantitative involvement in the first year of the study. This was done for pragmatic reasons: Children's developmental needs change rapidly, especially in the first years of life, and consequently, it would not be possible to establish measurement invariance for quantitative involvement from age 6 to 12. For instance, whereas 6-year olds might need help brushing their teeth, 12-year olds typically do not. Our approach assumes that a "snapshot" of quantitative involvement at age 6 is representative for parents' general level of involvement. Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact. Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range, so measurement invariance for quantitative involvement can be assured.

## Conclusions

We set out to assess whether the quantity of parental care interacted with within-family links between qualitative parenting behaviors and children's emotional adjustment. Our findings suggest that, in families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development. In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment. This suggests that, when differential involvement was higher, mothers more strongly adopted the classic role of a warm "attachment figure", and fathers that of an authority figure. When both parents were more involved, mothers' warmth was less strong of a predictor for emotional adjustment, and mothers' inductive reasoning also became more contingent on adjustment, suggesting that mothers in such families played more similar roles to fathers.

We also found substantial evidence that both parents showed more desirable patterns of qualitative parenting behaviors - behaviors that are positively associated with children's adjustment - when total joint involvement was higher. When mothers were more involved than fathers, however, mothers' expressed more warmth, whereas fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings suggest that all family members might benefit from a more equal distribution of childcare tasks.

With regard to mothers' versus fathers' unique roles in children's emotional development, we found that, in line with the notion of the mother as a warm, supportive attachment figure, maternal warmth and consistency were associated with children's emotional adjustment. Evidence for fathers' unique role was relatively limited; significant within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role in the capacity of disciplinarian. The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - fathers' role is more context dependent.

<!--We used `r cite_r("r-references.bib")` for all our analyses.-->

\newpage

# References
```{r create_r-references}
#r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2019-05-01 Draft_1.Rmd
Final draft before submission

---
title             : "Quantitative involvement as a moderator of the interplay between mothers’ and fathers’ qualitative parenting behaviors, and children's emotional adjustment development"
shorttitle        : "Quantitative and qualitative involvement"

author:
  - name          : "Caspar J. van Lissa"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Renske Keizer"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Utrecht University, dept. Youth & Family"
  - id            : "2"
    institution   : "Erasmus University Rotterdam, dept. Family Sociology"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

keywords          : "father, involvement, parenting, emotional adjustment, dynamic structural equation modeling"
wordcount         : "X"

bibliography      : ["DSEM_2017.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
library(foreign)
library(dplyr)
library(psych)
library(MplusAutomation)
library(reshape2)
library(stringr)
library(data.table)

source("C:/Users/lissa102/Dropbox/R/RFunctions.r")
```

```{r read_data, include = FALSE}
ifnofile("lsacfiles.RData", {
  lsacfiles <- readLsacdata()
  saveRDS(lsacfiles, "lsacfiles.RData")
}, {
  lsacfiles<-readRDS("lsacfiles.RData")
})

ifnofile("target.families.HICID.RData", {
  relationship.cp<-getLsacdata(lsacfiles=lsacfiles, 
                             invar=c("zf02m2", "zf02m3",       #Sex of parent 1 and 2
                             "zf09m1", "zf09m2", "zf09m3",     #Country of birth
                             "zf12m2", "zf12m1"),              #Indigenous status
                             variant=c("[a-g]f03m1",
                                       "id12a$",               #Interpreter
                                       "f11m[123]",            #English spoken in the home
                                       "^[a-g]prel$",          #study child relationship to parents, 1=2 bio parents
                                       "^[a-g]paraway$",       #1=both parents currently live in the home
                                       "[a-g]parpart$",        #2=two parents who are partners
                                       "[a-g]parsit$",         #1=both bio parents
                                       "[a-g]p2$",             #study child has two parents in the home 1=yes
                                       "[a-g]f0[67]m[123]",    #Relationships between parents and child
                                       "[a-g]f15m[23]",        #Stopped living with the study child 
                                       "[a-g]af05m1$"))        #Study child living away

  sep<-getLsacdata(lsacfiles=lsacfiles, 
                               invar=c("dsep$"))        #Study child living away

  ScreenFun<-function(x, values){
    ifelse(!all(is.na(x))&&all(x[!is.na(x)] %in% values), 1, 0)
  }

  relationship.cp[, "filter_parentsex" := 3-zf02m2]
  relationship.cp[, "filter_parentsex" := apply(.SD, 1, ScreenFun, values=1), 
                  .SDcols=c("zf02m3", "filter_parentsex")]

  relationship.cp[, "filter_parpresent" := apply(.SD, 1, ScreenFun, values=1), 
                  .SDcols=grep("^[a-g]paraway$", names(relationship.cp), value = TRUE)]

  relationship.cp[, "filter_bothbio" := apply(.SD, 1, ScreenFun, values=1), 
                  .SDcols=grep("^[a-g]prel$", names(relationship.cp), value = TRUE)]

  relationship.cp[, "filter_birthcountry" := apply(.SD, 1, ScreenFun, values=1101), 
                  .SDcols=grep("f09m[123]$", names(relationship.cp), value = TRUE)]

  relationship.cp[, "filter_indigenous" := apply(.SD, 1, ScreenFun, values=1), 
                  .SDcols=grep("zf12m[12]$", names(relationship.cp), value = TRUE)]

  relationship.cp[, "filter_interpreter" := apply(.SD, 1, ScreenFun, values=4), 
                  .SDcols=grep("id12a", names(relationship.cp), value = TRUE)]

  relationship.cp[, "filter_english_language" := apply(.SD, 1, ScreenFun, values=c(0, 1201)), 
                  .SDcols=grep("f11m[123]$", names(relationship.cp), value = TRUE)]

  relationship.cp[, "filter" := apply(.SD, 1, ScreenFun, values=1), 
                  .SDcols=grep("^filter", names(relationship.cp), value = TRUE)]

  table(relationship.cp$filter)

  target.families<-relationship.cp[filter==1,]$hicid

  saveRDS(target.families, "target.families.HICID.RData")
}, {
  target.families <- readRDS("target.families.HICID.RData")
})

ifnofile("filtered_data.RData", {
  study.variables<-getLsacdata(lsacfiles=lsacfiles, variant=c("[a-g]bseoi$",
    "[a-g]wseoi$",
    "^[a-g]se03[mft]\\d\\w$", #Socio-emotional adjustment
    "^[a-g]pa04[mf]\\d$",
    "^[a-g]pa10[mf]\\d", #Involvement. Note: These variables end in "a", not in a number
    "^[a-g]pa11[mf]\\d$",#Consistent parenting
    "^[a-g]pa13[mf]\\d$",#Angry parenting
    "^[a-g]pa15[mf]\\d$",#Overprotection
    "^[a-g]pa18[mf]\\d$",#Monitoring
    "^[a-g]pa09[mf]\\d$",#Inductive reasoning
    "^[a-g]pa03[mf]\\d$"))

    filtered.data<-study.variables[hicid %in% target.families,]

    filtered.data <- filtered.data[cohort=="K",]
    filtered.data[, which( colSums(is.na(filtered.data)) == nrow(filtered.data) ) := NULL]
    saveRDS(filtered.data, "filtered_data.RData")
}, {
  filtered.data <- readRDS("filtered_data.RData")
})

# Waarschijnlijk is dit om descriptives (age) te krijgen. Kan waarschijnlijk vervangen worden met een compleet stukje demographics
agevars <- relationship.cp[hicid %in% target.families & cohort=="K", .SD, .SDcols=grep("[d-g]f03m1", names(relationship.cp))]
agevars[, lapply(.SD, function(x){round(c(mean(x, na.rm = TRUE), sd(x, na.rm = TRUE)), 2)})]
# Eind blokje

```

```{r score_scales, include = FALSE}
ifnofile("scales.RData", {
  # Get list of scales from file
  source("LSAC.scales.list.R")

  # Remove misscoded 0s in scale data
  for (j in unlist(scales.list)){
    set(filtered.data, which(filtered.data[[j]] == 0), j, NA)
  }
  ###Reverse code items
  revthese <-
    c(
      "(^[a-g]se03[mft]3\\w$)",
      "([a-g]se03[mft]5[bc])",
      "([a-g]pa11[mf][345])",
      "([a-g]pa13[mf]2)",
      "([a-g]pa18[mf]4)",
      "([a-g]pa10[mf]\\d)"      #Involvement
    )
  revthese<-unlist(sapply(revthese, grep, names(filtered.data), value=TRUE))

  filtered.data[, (revthese) := lapply(.SD, reverse.code, keys=-1), .SDcols=revthese]

  #Compute mean scores
  scales <- doReliability(subset(filtered.data, select=gsub("^-", "", unlist(scales.list))), keys.list=scales.list, name="reliabilitiestest")
  scales$table_descriptives
  saveRDS(scales, "scales.RData")
}, {
  scales <- readRDS("scales.RData")
})
```

```{r mplus_data, include = FALSE}
scaledata <- cbind(kdata[,.SD, .SDcols=c("hicid", "zf02m1")], data.table(scales$scores))
setnames(scaledata, c("hicid", "zf02m1"), c("family", "sex"))
names(scaledata)
### Select these variables from scaledata
scaledata[is.na(scaledata)]<-NA

scaledata[ , c("Fin2", "Fin3", "Fin4") := Fin1] #Make involvement constant across time, to be used as between-families variable
scaledata[ , c("Min2", "Min3", "Min4") := Min1]

# centereddata<-scaledata
#cols <- names(scaledata)[-c(1, 2)]
# centereddata[, (cols) := lapply(.SD, scale), .SDcols=cols]
# describe(centereddata)
# centereddata <- melt(centereddata, id.vars=c("family", "sex"))
# centereddata$time <- as.numeric(str_sub(centereddata$variable, -1, -1))
# centereddata$variable <- str_sub(centereddata$variable, 1, -2)
# longdat <- dcast(centereddata, family + sex + time ~ variable)
#longdat$time<-as.numeric(longdat$time)
#longdat <- longdat[ , c("family", "sex", "time", "Man", "Fan", "Mco", "Fco", "Min", "Fin", "Mre", "Fre", "Mwa", "Fwa", "con", "emo", "pee", "pro")]
#prepareMplusData(longdat, "22-08-17_longdat_in_constant_centered.dat", inpfile = TRUE)

scaledata <- melt(scaledata, id.vars=c("family", "sex"))
scaledata$time <- as.numeric(str_sub(scaledata$variable, -1, -1))
scaledata$variable <- str_sub(scaledata$variable, 1, -2)

longdat <- dcast(scaledata, family + sex + time ~ variable)
longdat$time<-as.numeric(longdat$time)
longdat <- longdat[ , c("family", "sex", "time", "Man", "Fan", "Mco", "Fco", "Min", "Fin", "Mre", "Fre", "Mwa", "Fwa", "emo")]

prepareMplusData(longdat, "involvement_recode_emo.dat", inpfile = TRUE, hashfilename = TRUE)

round(cor(longdat[, .SD, .SDcols = grep("^M", names(longdat), value = TRUE)], use="pairwise.complete.obs"),2)

createModels("21-08-17_template_dsem.txt")
setwd("C:/data/Wave 5 GR CD/target sample/DSEM/From_template")
runModels()

plot(density(longdat$Fco, na.rm = TRUE))

fullscales<-data.table(scales$scores)
var <- "^Mwar"
tmp<-fullscales[, .SD, .SDcols = grep(var, names(fullscales))]
tmp <- tmp[apply((sapply(tmp, is.na)), 1, function(x){length(which(x))}) < 4, ]
ICC(tmp, missing=FALSE)
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

Fathers have been hypothesized to play a unique role in children's emotional adjustment development, because they encourage children to challenge themselves, in play and in life (Flouri, 2010; Hagan& Kuebli, 2007; Sandseter & Kennair, 2011 ; Clarke-Stewart, 1980; Hossain & Roopnarine, 1994; Paquette, Carbonneau, Dubeau, Bigras, & Tremblay, 2003; Parke, 1996) (Panksepp, Burgdorf, Turner, & Gordon, 2003) (Fletcher, St. George, & Freeman, 2013; Grossmann, Grossmann, Kindler, & Zimmermann, 2008; Paquette & Dumont, 2013). Emotional adjustment, defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@GoodmanPsychometricpropertiesstrengths2001], is an important developmental outcome, because it is integral to social competence, academic achievement, and the absence of psychopathology (Gross, 2013; Lin, Lawrence, & Gorrell, 2003) Aldao, Nolen-Hoeksema, & Schweizer, 2010). By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a wide range of unpredictable emotions, thereby structuring opportunities to expand children's emotional repertoire. Compared to the familiar notion of an "attachment relationship", grounded in comfort and support, the father-child relationship can thus be construed as an "activation relationship" (REF).

## Fatherhood theory

The past decades have seen major strides in the consolidation of theories to guide fatherhood research. Initially, such research was inspired by two major societal shifts: Increasing divorce rates, and mothers’ greater labor force participation, both of which impacted the extent of fathers’ participation in childrearing. Consequently, early research focused primarily on the quantity of paternal involvement: Whether fathers were present or absent in divorced families, and how much they contributed to the work of childcare in intact families [@Lambbiosocialperspectivepaternal1987]. Over time, the notion of the father as a coparent gained ground (Burgess 1997; Cabrera and TamisLeMonda 1999; Coleman and Ganong 2004; Pleck and Pleck 1997), and conceptualizations of father involvement evolved to encompass qualitative dimensions as well, including warmth, responsiveness, and control [@PleckPaternalinvolvementRevised2004]. Nowadays, it is widely acknowledged that both the quantity and quality of parental involvement are relevant factors in child development [@FaganShouldresearchersconceptualize2014; @PleckPaternalinvolvementRevised2004]. Interactions between quantity and quality have rarely been considered, however, perhaps in part because, historically, interest in these dimensions has varied across discipline lines. Sociology and gender studies, for instance, have primarily focused on involvement quantity [@CraigDoesFatherCare2006; @FaganShouldresearchersconceptualize2014], whereas developmental psychology has focused on how fathers’ qualitative parenting behaviors influence child development. The present study attempts to bridge these disciplines, by investigating whether the division of quantitative involvement between mothers and fathers moderates the influence of their qualitative parenting behaviors in children’s emotional adjustment development over time.

## Interactions between involvement quantity and quality

One remaining limitation of the aforementioned theoretical models of fatherhood is that they have not explicitly addressed potential interactions between involvement quantity and quality [@JaffeeLifeFatherBenefits2003]. For example, children's emotional adjustment might be differently impacted by a father's efforts to exert control if he is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent. Conversely, children's emotional adjustment might elicit different qualitative parenting behaviors from hands-on fathers, who which struggles give rise to their children's maladjustment, than from out-of-touch fathers, who might perceive maladjustment as a problem behavior. With regard to such interactions between involvement quantity and quality, existing theories of father involvement are complemented by Feinberg’s coparenting model [@FeinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the role of parent. Although the coparenting literature typically focuses on divorced families it holds insights for intact families, too. One dimension of coparenting is involvement quantity, which Feinberg defines as the ``division of labor’’ in childcare. According to Feinberg’s model, coparenting is directly associated with children’s adjustment and parents’ qualitative childrearing behaviors, as well as the link between these constructs. There is some preliminary empirical support for such interactions. For instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship (Jia et al., 2012). Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@WangParentalInvolvementAfrican2014]. Thus, applying this insight from the coparenting literature, we might expect involvement quantity to be associated with parenting quality and emotional adjustment, and to moderate prospective links between qualitative parenting dimensions and children's emotional adjustment.

## Operationalizing quantitative involvement

A second important insight from coparenting theory is that the *distribution* of childcare labor - that is, parents' relative contributions to childrearing - might be more relevant for predicting children's adjustment than either parent's absolute level of quantitative involvement (REF Kroll). The equity of the childcare labor distribution has been found to affect children's development directly (Deutsch et al., 2001), and to impact the emotional climate of the interparental relationship (Craig, 2006). When childcare is unequally distributed, women often perform a disproportionate share of the work (Craig, 2006). This strains the marital relationship [@ShockleyItnotwhat2018], and the resulting stress erodes parenting quality, particularly for difficult children (Jenkins et al., 2003). A second reason for focusing on the distribution of childcare, is that examining parents' absolute involvement introduces a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care (see also Deutsch, 2001). Operationalizing quantitative involvement in terms of the childcare distribution, while controlling for total joint care provided by both parents, mitigates this confound. One way in which prior research has attempted to capture the childcare distribution is by asking who has most responsibility for certain task, on a continuum from the mother doing everything to the father doing everything [@KrollEarlyFatherInvolvement2016]. An elegant alternative, which statistically disentangles quantity from the distribution of care, is to use a latent mean and difference approach. This technique is typically used to capture the difference between consecutive longitudinal measurements [@McArdleLatentVariableModeling2009], but it can also be implemented to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other. This approach taps directly into a theoretically relevant construct, and prevents confounding.

Kroll: No evidence for effect of quantitative involvement This might help explain why some studies, focusing on absolute involvement, have concluded that fathers' quantitative involvement does not predict children's adjustment when controlling for mothers' involvement>> Kroll actually used absolute involvement and "who does most"-questions, which measure relative involvement

## Main effects of quantitative involvement

According to bioecological theories of development, children require a certain amount of care, and who provides this care is, to some extent, flexible [@BronfenbrennerEcologyHumanDevelopment1979]. This suggests that the amount of total joint care provided should be positively associated with children's emotional adjustment. Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they will likely feel less supported than when both parents are equally invested in their wellbeing [@DEUTSCHPaternalParticipationChild2001].

Parents' quantitative involvement is also likely to be associated with the quality of their involvement. First, it stands to reason that, as parents spend more time with their children, the quality of their parenting will increase through practice and experience. Thus, greater total joint care should positively predict parenting quality, whereas greater differential involvement should positively predict parenting quality of the more involved parent only. Differential involvement might also impact involvement quality through its effect on the marital relationship. Childcare responsibilities are a heavy burden if they come down on a single parent's shoulders [@DEUTSCHPaternalParticipationChild2001]. Consequently, an unfair distribution of childare is associated with greater maternal anger and resentment, and reduced marital quality (Goodnow, 1998; Ross and Van Willigen 1996). The resulting strain on the interparental relationship interferes with warm, sensitive parenting [@FeinbergInternalStructureEcological2003]. Fathers are as capable as mothers of caring for their children (Lamb 1997; Yeung et al. 2001), and on average, their involvement in childcare is growing (Cabrera et al. 1999). When fathers do take on childcare responsibilities, this benefits mothers directly by relieving some of this burden (Pleck, 2010). Thus, when care is distributed relatively equally, the entire family system will likely be more harmoneous.

## The interplay between qualitative parenting behaviors and emotional adjustment

The standard family environment model posits that the quality of parents' involvement is a proximal mechanism, shaping children's development (Amato and Cheadle 2008). Studies following this model have found that fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence (Flouri 2010). Moreover, there is evidence for unique roles of fathers' versus mothers' qualitative parenting behaviors, in line with the aforementioned distinction between the "activation relationship" versus "attachment relationship" [@PaquetteTheorizingfatherchildrelationship2004]. Fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior, whereas mothers empathize more with their children, and provide more warmth and support (Majdandžić, Möller, de Vente, Bögels, & van den Boom, 2014; Paquette, 2004). Empirical research supports these unique roles for mothers and fathers in children's emotional adjustment development: Mothers tend to encourage the expression of emotions, and respond warmly and supportively, whereas fathers structure and sanction emotion expressions to a greater extent (Cassano, Perry-Parrish, & Zeman, 2007 Eisenberg et al., 1999). Furthermore, mothers' supportive behavior was found to be a better predictor of children’s adjustment than their controlling behavior, whereas the inverse held for fathers (McDowell, Parke, & Wang, 2003). Paradoxically, there is also evidence that mothers' negative parenting has a stronger negative impact upon children's emotional adjustment than fathers' negative parenting [@ChangHarshParentingRelation2003]; perhaps because this behavior is incongruous with mothers' prototypical role as an attachment figure. Based on these findings, we might also expect warmth and support to play a stronger role in mother-child relationships, and structuring behaviors to play a stronger role in father-child relationships.

## Child effects

The notion of child effects was initially introduced by Bell (1986), who noted that correlations between parenting and child adjustment might be equally well explained by children affecting parents. There is now substantial evidence to suggest that children's adjustment problems elicit less support, more structure, and more negative parenting from parents [@GadeyneLongitudinalRelationsParenting2004], @NepplIntergenerationalContinuityParenting2009, LenguaBidirectionalassociationstemperament2005]. Child effects have been more consistently investigated in the adolescent literature than in the childhood literature. This might be because the notion of children influencing parents is more salient in adolescence, when parent-child relationships become increasingly egalitarian. Nonetheless, findings from the adolescent literature are broadly in line with the childhood literature, suggesting that emotional adjustment problems predicts decreased parental support, and greater control (Otterpohl & Wild, 2015; Van der Giessen, Branje, & Meeus, 2014; Wijsbroek, Hale, Raaijmakers, & Meeus, 2011). Thus, parents' responses to maladjustment appear to be a combination of punitive responses, including decreased support and increased negative parenting, and corrective responses, such as increased structure and control. Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively (Dix, 1991). **We might thus expect emotional adjustment to predict....**

## Direction of effects

According to the transactional view of family relationships, parents and children reciprocally influence one another (Kerr et al., 2012; Amato and Cheadle 2008). This view of the family as a dynamic system whose members reciprocally influence one another is now widely accepted [@KerrPerceivedparentingstyle2012]. Nevertheless, it is not yet common practice to examine bi-directional effects. Consequently, there has been a call for research to tease apart the direction of effects between parents and children (REF Flouri). When studies do consider such effects, they often employ cross-lagged panel models. These models have recently been criticized for misrepresenting over-time predictive effects, if there are stable between-family differences in the variables studied. Thus, the direction of effects between parents and children remains an issue to be examined. Recent methodological innovations hold the potential to estimate dynamic reciprocal processes within families more accurately, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model, and dynamic structural equation modeling (DSEM); a multilevel generalization of this model. Several studies have re-examined the direction of effects between parents and children using these models, but mostly in adolescent samples. One study is noteworthy, because it used the random-intercept cross-lagged panel model to examine the interplay between maternal and paternal qualitative parenting dimensions, and adolescents' emotional adjustment. Results indicated that mothers' role in adolescents' emotional adjustment development occurred primarily through support, whereas fathers influenced adolescents through behavioral control (Van Lissa et al., under review). Interestingly, over-time predictive effects were sparse, and most associations between study variables appeared to be due to stable between-family differences. This suggests that reciprocal dynamic processes might be less prominent than previous research suggested. Instead, third variables, such as socio-economic status, or shared genetic factors, might play a larger role - another outstanding issue on which further research is required (REF Flouri). Furthermore, child effects were found to substantially outnumber parenting effects; a finding which might be due to the relative equality of adolescent-parent relationships when compared to parent-child relationships. If this is the case, we might expect to find more prominent parenting effects in the present study on younger children.

## Differences between sons and daughters

<!-- pubertal timing and associated changes in hormone levels during adolescence are related to sex-typed attributes, including emotional distress (Brooks-Gunn, Graber, & Paikoff, 1994; Susman, Dorn, & Chrousos, 1991).-->

At a young age, there is little evidence for sex differences in emotional adjustment, as indicated by a meta-analysis which found no sex differences in the temperament dimension of negative affectivity, based primarily on infant and toddler samples [@Else-questGenderdifferencestemperament2006]. In older children, there is substantial evidence for such sex differences [@SilkAdolescentsemotionregulation2003], but these appear to intensify in puberty [@HillIntensificationGenderRelatedRole1983]. There is substantial evidence, however, parents tend to raise boys and girls differently. For example, parents tend control girls to a greater extent than boys (e.g., Pomerantz & Ruble, 1998; Smetana & Daddis, 2002), and exert more negative parenting with boys than with girls (Straus and Stewart 1999). Another point to consider is the interplay between parent and child sex. According to social learning theory, the same-sex parent plays a stronger role in a child’s development (Bussey & Bandura, 1999), particularly in the socialization of sex-typed behavior, such as display rules for emotions [@Brodysocializationgenderdifferences2000]. Thus, we might expect child sex to predict the same-sex parents' qualitative involvement, and associations between parenting and emotional adjustment might be stronger for same-sex parent-child dyads. There is also evidence that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships (Lovas, 2005; Schoppe-Sullivan et al., 2006). These findings suggest that boys tend to receive a "tougher" upbringing than girls. In sum, it is clear that sex differences are a relevant factor in the interplay between parenting and emotional adjustment. Specifically, we might expect to find parents, display more responsiveness, structure, and inductive reasoning with girls than with boys. Boys, on the other hand, might receive more negative parenting. Furthermore, we might expect associations between qualitative parenting behaviors and emotional adjustment to be stronger between children and the same-sex parent.

## The present study

The present study aimed to make several contributions to the literature on fathers' role in their children's emotional adjustment development. First, we included both mothers and fathers, in order to tease apart their common and unique roles. Second, we incorporated both quantitative involvement, in terms of total joint childcare and the childcare distribution, and qualitative involvement, in terms of specific parenting behaviors. Third, we examined interactions between quantitative and qualitative involvement, which have been largely overlooked in prior work. Fourth, we employed recent methodological advances to disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between parenting behaviors and emotional adjustment.

Based on our review of the literature, we expected that involvement quantity and quality would interact, such that a more skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children, and parents' adaptive responses to children's maladjustment. In other words, we expected that an unfair childcare distribution would predict suboptimal parenting, and parental responses to maladjustment, due to the strain such a distribution places upon the entire family system. We further hypothesized that quantitative involvement would predict children's adjustment directly: Greater joint total care should predict better emotional adjustment, whereas a more skewed distribution of care should predict poorer emotional adjustment. Greater joint care should also predict both parents' qualitative involvement, in the sense of greater responsiveness and structure, and lower negative parenting. A more skewed distribution of care, on the other hand, should show these associations only for the more involved parent. Finally, we expected specific qualitative parenting dimensions to be associated with emotional adjustment at both the within- and between-family level. The literature suggests that responsiveness, structure, and reasoning should be positively associated, whereas negative parenting should be negatively associated, with emotional adjustment. With regard to child effects, we might expect some of these paths to be reversed: Maladjustment should elicit decreased responsiveness, but increased structure, reasoning, and negative parenting. We also accounted for potential effects of child sex.

# Methods

## Participants

<!--As a detailed description of the LSAC study has been given on pages 4-9 of this edition of Family Matters, only a brief overview is provided here (see also the LSAC Annual Report, FaCS, 2005). The LSAC study is following two cohorts of children from urban and rural areas of all states of Australia through early and mid childhood. It aims to provide comprehensive data on Australian children’s wellbeing and progress that can be used to underpin policy development and service provision for young children and their families (FaCS 2005). The two LSAC cohorts are broadly representative of the Australian population, although there are trends for the cohorts to be slightly more advantaged than the general population. The study uses weighting to adjust for this. Data from the first LSAC survey wave, collected in 2004, are used here. The younger cohort of 5104 children was aged between 3 and 18 months, with an average age of 8.8 months. The older cohort (N=4976 children) was aged between 51 and 67 months, and averaged 56.9 months. -->

We employed data from the K-cohort of the Longitudinal Study of Australian Children (LSAC), waves 2-5, which were collected during biannual home visits, and self-report questionnaires sent by mail, from 2006-2012. LSAC employed a clustered sampling design to obtain a sample broadly representative of the Australian population (see, Soloff et al., 2005). A detailed description of the study design and procedure is available at http://data.growingupinaustralia.gov.au/.

Our target population consisted of native Australian, non-minority, intact, biological families. This yielded a subsample of 1884 families. Children were approximately 6 years old at T1, $M = 6.28$, $SD = .45$, and 12 at T4 ($M = 12.42$, $SD = .49$).

REPHRASE AND PROVIDE APPROPRIATE DEMOGRAPHICS: **Socio-demographic characteristics of our study sample are described in Table X. There were clear differences between mothers and fathers in terms of socio-demographic and work characteristics. For both mothers and fathers, the most common family employment type was a couple dual-earner family. Of couple fathers, 29 per cent were the sole breadwinner compared to less than three per cent of couple mothers. The median annual income range for families was between $78,000 and $114,399. However, the median individual income for mothers was $32,500 and $62,782 for fathers, indicating a large income gap.**

## Design
Quantitative involvement was measured once, at T1. Emotional adjustment and four qualitative parenting dimensions were measured biannually.

## Measures

### Emotional adjustment

Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the Strengths and Difficulties Questionnaire, reported by teachers on a three-point scale from "not true" to "certainly true" (e.g., "Often unhappy, downhearted or tearful"; Goodman, 1999). Scale descriptives are displayed in Table X, and scale score intercorrelations are reported in the online supplement SX.

### Quantitative involvement

Both parents rated their involvement in four different child care on a five-point scale ranging from "once a day or more" to "not at all", using the self-report questionnaire (e.g.: "In the past month how often did you help get this child ready for school?"). **REPORT CFA, measurement invariance**

### Qualitative involvement

Parents completed measures of several qualitative parenting dimensions.
Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire (CRQ; Paterson & Sanson, 1999). Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth (NLSCY; Statistics Canada, 2000). For these instruments, parents reported on their own behavior using five-point scales, ranging from "never or almost never" to "always or almost always".

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("DSEM_2017.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "DSEM_2017.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2019-05-01 Paper.Rmd
Final draft before submission

--- 2018-11-23 Paper_reorganized.Rmd
@@
 ---
-title             : "Quantitative involvement moderates the interplay between mothers' versus fathers' qualitative parenting behaviors and children's emotional adjustment development"
-shorttitle        : "QUALITATIVE AND QUANTITATIVE INVOLVEMENT"
+title             : "Quantitative involvement moderates mothers' versus fathers' roles in children's emotional adjustment development"
+shorttitle        : "QUANTITATIVE AND QUALITATIVE INVOLVEMENT"
 
@@
 lineno            : yes
-mask              : yes
+mask              : no
 
 class             : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 ---
@@
 
-Emotional adjustment, defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001], is an important developmental outcome, integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013; @aldaoEmotionregulationStrategiesPsychopathology2010].
-Fathers have been hypothesized to play a unique role in children's emotional adjustment development, because they encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+Fathers have been hypothesized to play a unique role in children's emotional adjustment development, because they encourage children to challenge themselves, in play and in life [@FlouriFathersbehaviorschildren2010; @BogelsDoesFatherKnow2011; @ParkeFatherinvolvement2000].
+Emotional adjustment, defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@GoodmanPsychometricpropertiesstrengths2001], is an important developmental outcome, because it is integral to social competence, academic achievement, and mental health [@GrossHandbookemotionregulation2013; @AldaoEmotionregulationstrategiespsychopathology2010].
+By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a broad spectrum of emotions, thereby structuring opportunities to expand children's emotional repertoire.
 The classical notion of the “attachment relationship” is rooted in security and exploration (REF).
-Fatherhood researchers have posited that, whereas the mother-child attachment relationship primarily provides warmth and security, exploration might be somewhat more central to the father-child relationship, which can has thus been construed as an “activation relationship” [@paquetteFatherChildActivationRelationship2013].
-By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a broad spectrum of emotions, thereby structuring opportunities to expand children's emotional repertoire.
+Fatherhood researchers have posited that, whereas the mother-child attachment relationship primarily provides warmth and security, exploration might be somewhat more central to the father-child relationship, which can has thus been construed as an “activation relationship” [@PaquetteFatherChildActivationRelationship2013].
 
@@
 Both shifts impacted the extent of fathers' involvement in childrearing.
-Early research focused primarily on the *quantity* of paternal involvement: Whether fathers were present or absent in divorced families, which constitutes a dichotomous operationalization of quantitative involvement, and how much they contributed to the work of childcare in intact families, i.e., a continuous operationalization [@lambBiosocialPerspectivePaternal1987].
-Over time, the notion of the father as a coparent gained ground [@pleckFatherhoodIdealsUnited1997], according to which fathers share the responsibilities of childrearing with mothers, and might, to some extent, adopt roles that were historically fulfilled by mothers.
-To keep pace with changing conceptualizations of fatherhood, research on father involvement evolved to encompass *qualitative* dimensions as well, including warmth and control [or: demandingness; @maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
+Early research focused primarily on the *quantity* of paternal involvement: Whether fathers were present or absent in divorced families, which constitutes a dichotomous operationalization of quantitative involvement, and how much they contributed to the work of childcare in intact families, i.e., a continuous operationalization [@Lambbiosocialperspectivepaternal1987].
+Over time, the notion of the father as a coparent gained ground [@PleckFatherhoodidealsUnited1997], according to which fathers share the responsibilities of childrearing with mothers, and might, to some extent, adopt roles that were historically fulfilled by mothers.
+To keep pace with changing conceptualizations of fatherhood, research on father involvement evolved to encompass *qualitative* dimensions as well, including warmth and control [or: demandingness; @MaccobySocializationcontextfamily1983; @PleckPaternalinvolvementRevised2004].
+Nowadays, it is widely acknowledged that both the quantity and quality of parental involvement are relevant factors in child development [@FaganShouldresearchersconceptualize2014; @PleckPaternalinvolvementRevised2004].
+Interactions between quantity and quality have rarely been considered, however, perhaps in part because, historically, interest in these dimensions has varied across discipline lines.
+Sociology and gender studies, for instance, have primarily focused on involvement quantity [@CraigDoesFatherCare2006; @FaganShouldresearchersconceptualize2014], whereas developmental psychology has focused on how qualitative parenting behaviors influence child development.
+The present study attempts to bridge these disciplines, by investigating whether the division of quantitative involvement between mothers and fathers moderates the interplay between qualitative parenting behaviors and children's emotional adjustment.
 
-Although it is now widely acknowledged that both the quality and quantity of parental involvement are relevant factors in child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated.
-This might be, in part, because research interests have historically varied across discipline lines: Developmental psychology has focused primarily on how qualitative parenting behaviors influence child development, whereas sociology and gender studies have focused more on involvement quantity, and its implications for maternal wellbeing [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement in relation to children's emotional adjustment.
+## Interactions between involvement quantity and quality
 
-## Qualitative parenting and emotional adjustment
+The aforementioned theoretical models of fatherhood are limited, in the sense that they have underemphasized potential interactions between involvement quantity and quality [@JaffeeLifeFatherBenefits2003].
+For example, children's emotional adjustment might be differently impacted by a father's efforts to exert control if he is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent.
+Conversely, children's emotional adjustment might elicit different qualitative parenting behaviors from hands-on fathers, who are aware of the reasons for their children's emotional troubles, than from out-of-touch fathers, who might perceive maladjustment as a problem behavior that should be sanctioned.
+To capture these patterns, a model should account for interactions between involvement quantity and quality.
 
-### Parenting effects
+With regard to such interactions, existing theories of father involvement are complemented by Feinberg's coparenting model [@FeinbergInternalStructureEcological2003].
+Coparenting refers to the ways parents relate to each other in the role of parent.
+Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families.
+One dimension of coparenting is involvement quantity, which Feinberg defines as the "division of labor" in childcare.
+This dimension is proposed to affect children's adjustment and parents' qualitative childrearing behaviors directly, and to moderate the link between these two constructs.
 
-The standard family environment model posits that the quality of parents' involvement is a proximal mechanism, shaping children's development [@amatoParentalDivorceMarital2008].
-Qualitative parenting is typically examined in terms of the constituent dimensions of the parenting styles taxonomy [@baumrindChildCarePractices1967]: Responsiveness and demandingness [@maccobySocializationContextFamily1983].
-Responsiveness refers to warm and supportive parenting, which substantial evidence indicates is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].
-Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior.
-Two aspects of demandingness investigated in the present study are consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
-The literature on links between demandingness and emotional adjustment is more limited, and mixed [@morrisRoleFamilyContext2007]; in part because some forms of demandingness help children meet societal expectations, whereas other forms are overly restrictive, and undermine children's autonomy needs [@barberCentralityControlParenting2013].
-A fourth dimension, which has consistently been linked with children's emotional maladjustment, but is distinct from the parenting styles typology, is negative parenting [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
-We operationalized negative parenting as one of its specific subdimensions: parental anger.
-The anger parents express in interactions with children might reflect faillure to regulate their own negative emotions, and models the use of anger in communication, and undermines children's emotion regulation [@scaramellaParentingChildrenDistress2008].
+There is some preliminary empirical support for such interactions between involvement quantity and quality.
+For instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@JiaTransactionalRelationsFather2012].
+Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@WangParentalInvolvementAfrican2014].
+Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.
 
-<!-- 
-Inductive reasoning is thought to be integral to children's development of moral emotions [@HoffmanEmpathymoraldevelopment2000], and might thus play a positive role general emotional development as well.
--->
+## Operationalizing quantitative involvement
 
-Although some research suggests that both parents' qualitative parenting behaviors influence children in similar ways [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique roles of mothers versus fathers.
-Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010].
-Furthermore, whether for evolutionary or cultural reasons [@bogelsDoesFatherKnow2011], mothers and fathers may tend to adopt somewhat specialized roles in terms of qualitative parenting: The aforementioned notions of "attachment" and "activation" relationships posit that mothers empathize more with their children, and  provide more warmth and support, whereas fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior [@majdandzicFathersChallengingParenting2014; @paquetteTheorizingFatherchildRelationship2004].
-Empirical research supports these unique roles for mothers and fathers in children's emotional adjustment development: Mothers tend to encourage the expression of emotions, and respond warmly and supportively, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for emotional adjustment in adolescence, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
-This differentiated pattern was also found with regard to younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
-Furthermore, there is some evidence that mothers' negative parenting has a stronger negative impact upon children's emotional adjustment than fathers' negative parenting [@changHarshParentingRelation2003]; perhaps because this behavior is incongruous with mothers' prototypical role as an attachment figure.
-Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.
+A second important insight from coparenting theory is this notion of the "division of labor", pertaining to childcare.
+This notion suggests that parents' relative contributions to childrearing might be more relevant for predicting children's adjustment than parent's absolute quantitative involvement [@KrollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004].
+Indeed, research has shown that the equity of the childcare labor distribution affects child development directly [@DeutschPaternalParticipationChild2001], and impacts the emotional climate of the interparental relationship [@CraigDoesFatherCare2006].
+When childcare is unequally distributed, women often perform a disproportionate share of the work.
+This strains the marital relationship [@ShockleyItnotwhat2018], and the resulting stress erodes parenting quality, particularly for difficult children [@SheltonInterparentalconflictnegative2008].
 
-### Child effects
-
-According to the transactional view of family relationships, parents and children reciprocally influence one another [@kerrPerceivedParentingStyle2012; @amatoParentalDivorceMarital2008].
-This view of the family as a dynamic system whose members reciprocally influence one another has become widely accepted [@kerrPerceivedParentingStyle2012].
-Nevertheless, examining bi-directional effects is not yet standard practice in the childhood literature [@flouriFathersBehaviorsChildren2010].
-Available empirical evidence suggests that children's adjustment problems elicit less warmth, more demandingness, and more negative parenting from parents [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @LenguaBidirectionalassociationstemperament2005].
-In the adolescent literature, bi-directional effects are more commonly examined.
-This might be because the notion of children influencing parents is more salient in adolescence, when parent-child relationships become increasingly egalitarian.
-Findings from the adolescent literature are broadly in line with findings in childhood; suggesting that emotional adjustment problems predicts decreased parental support, and greater control [@OtterpohlCrossLaggedRelationsParenting2015; @VanderGiessenPerceivedAutonomySupport2014; @Wijsbroekdirectioneffectsperceived2011].
-Thus, parents' responses to maladjustment appear to be a combination of punitive responses, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness.
-Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively [@Dixaffectiveorganizationparenting1991].
-We might thus expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.
-
-### Assessing the direction of effects
-
-Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if there are stable between-family differences in the variables studied [@Hamakercritiquecrosslaggedpanel2015].
-Recent methodological innovations provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@Hamakercritiquecrosslaggedpanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@AsparouhovDynamicStructuralEquation2018].
-
-Several studies have re-examined the direction of effects between parents and children using these innovative methods, but mostly in adolescent samples.
-One such study found that predictive effects between parenting and delinquency adjustment that were apparent using classical methods, were found to be explained by correlations between stable between-family differences in parenting and adjustment, once these components were included in the model [@KeijsersParentalmonitoringadolescent2016].
-Another study examined the interplay between maternal and paternal qualitative parenting dimensions and adolescents' emotional adjustment, and similarly found that most associations manifested as between-family level correlations.
-Furthermore, child effects were found to substantially outnumber parenting effects [@vanlissaRoleFathersMothers2018a].
-The authors suggested that this might be due to the relative equality of adolescent-parent relationships, as compared to parent-child relationships.
-If this interpretation is correct, we might expect to find more prominent parenting effects in the present study, with a younger sample.
-These preliminary findings highlight the fact that the direction of effects between parents and children, while controlling for stable between-family differences, remains an important issue to be examined.
-At this stage, however, it is difficult to formulate specific hypotheses about directionality, and the level at which associations between parenting and child development should occur.
+There are also methodological advantages to focusing on the distribution of childcare.
+It has been noted that operationalizing quantitative involvement in terms of both parents' absolute involvement, e.g., as hours spent caring for the child, incurs a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care [see @DeutschPaternalParticipationChild2001].
+This confound is mitigated by operationalizing quantitative involvement in terms of the childcare distribution, while controlling for total joint care.
+Some prior research has attempted to capture the childcare distribution; for example, by asking who has most responsibility for certain task, on a continuum from the mother doing everything to the father doing everything [@KrollEarlyFatherInvolvement2016].
+When quantitative involvement has been measured in terms of absolute involvement, there is an elegant statistical solution to disentangle the childcare distribution from total joint care.
+The so-called "latent mean and difference" approach, which is typically used to capture the difference between consecutive longitudinal measurements [@McArdleLatentVariableModeling2009], can be utilized to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other.
+This approach taps directly into a theoretically relevant construct, and prevents confounding.
 
@@
 
-## Quantitative involvement
-
-### Operationalizing quantitative involvement
-
-Quantitative involvement has been researched extensively in terms of the amount of time spent caring for the child [see @pleckPaternalInvolvementRevised2004].
-However, some have argued that the the way childcare tasks are divided is more relevant for children's adjustment than the absolute amount of time invested by each parent [@KrollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003].
-Indeed, research has shown that the equity of the childcare distribution affects child development directly [@deutschPaternalParticipationChild2001], and impacts the emotional climate of the interparental relationship [@CraigDoesFatherCare2006].
-When childcare is unequally distributed, women often perform a disproportionate share of the work.
-This strains the marital relationship [@shockleyItNotWhat2018], and the resulting stress erodes parenting quality, particularly for parents of difficult children [@SheltonInterparentalconflictnegative2008].
-
-There are also methodological advantages to focusing on the distribution of childcare.
-Operationalizing quantitative involvement in terms of both parents' absolute involvement, e.g., as hours spent caring for the child, incurs a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care [see @deutschPaternalParticipationChild2001].
-This confound is mitigated by operationalizing quantitative involvement in terms of the childcare distribution, and controlling for total joint care.
-Some prior research has attempted to capture the childcare distribution directly, by asking which has most responsibility for certain tasks, on a continuum from the mother doing everything to the father doing everything [@KrollEarlyFatherInvolvement2016].
-When quantitative involvement is measured in terms of absolute involvement, however, an elegant statistical solution can disentangle the childcare distribution from total joint care.
-The so-called "latent mean and difference" approach, which is typically used to capture the difference between consecutive longitudinal measurements [@McArdleLatentVariableModeling2009], can be utilized to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other.
-This approach taps directly into a theoretically relevant construct, and prevents confounding.
 ## Main effects of quantitative involvement
@@
 Greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship.
-Childcare responsibilities are a heavy burden if they come down on a single parent's shoulders [@deutschPaternalParticipationChild2001].
+Childcare responsibilities are a heavy burden if they come down on a single parent's shoulders [@DeutschPaternalParticipationChild2001].
 Mothers thus benefit directly when fathers relieve some of this burden [@PleckPaternalinvolvementRevised2004].
@@
 This suggests that the total amount of care provided should be positively associated with children's emotional adjustment.
-Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their wellbeing [@deutschPaternalParticipationChild2001].
+Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they will likely feel less supported than when both parents are equally invested in their wellbeing [@DeutschPaternalParticipationChild2001].
 
@@
 
-## Interactions between involvement quality and quantity
+## Qualitative parenting and emotional adjustment
 
-Interactions between involvement quality and quantity have largely been overlooked in the aforementioned theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@JaffeeLifeFatherBenefits2003].
-However, It is conceivable that children's emotional adjustment might be differently impacted by paternal control if that father is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent.
-Conversely, children's emotional adjustment might elicit different qualitative parenting behaviors from hands-on fathers, who are aware of the reasons for their children's emotional troubles, than from out-of-touch fathers, who might perceive maladjustment as a problem behavior that should be sanctioned.
-To capture patterns like these, a model should account for interactions between involvement quality and quantity.
+The standard family environment model posits that the quality of parents' involvement is a proximal mechanism, shaping children's development [@AmatoParentalDivorceMarital2008].
+Qualitative parenting has often been examined in terms of the dimensions underlying parenting styles [@BaumrindChildcarepractices1967]: Responsiveness and demandingness [@MaccobySocializationcontextfamily1983].
+Responsiveness refers to warm and supportive parenting, which substantial evidence indicates is positively associated with children's socio-emotional adjustment [@KhalequePerceivedParentalAcceptanceRejection2002; @Morrisrolefamilycontext2007].
+Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior.
+Two aspects of demandingness investigated in the present study are consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
+The literature on links between demandingness and emotional adjustment is more limited, and mixed [@Morrisrolefamilycontext2007]; in part because some forms of demandingness help children meet societal expectations, whereas other forms are overly restrictive, and undermine children's autonomy needs [@Barbercentralitycontrolparenting2013].
+A fourth dimension which has consistently been linked with children's emotional maladjustment - but is distinct from the parenting styles typology - is negative parenting, operationalized here as one of its constitutuent components: parental anger [@ChangHarshParentingRelation2003; @EisenbergParentalReactionsChildren1999].
+The anger parents express in interactions with children might reflect faillure to regulate their own negative emotions, and models the use of anger in communication, and undermines children's emotion regulation [@ScaramellaParentingChildrenDistress2008].
 
-In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@FeinbergInternalStructureEcological2003].
-Coparenting refers to the ways parents relate to each other in the role of parent.
-Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families.
-One dimension of coparenting is involvement quantity, which Feinberg defines as the "division of labor" in childcare.
-Note that this definition parallels our operationalization of quantitative involvement.
-The division of childcare labor is proposed to affect children's adjustment and parents' qualitative childrearing behaviors directly, and to interact with the link between these two constructs.
+<!-- 
+Inductive reasoning is thought to be integral to children's development of moral emotions [@HoffmanEmpathymoraldevelopment2000], and might thus play a positive role general emotional development as well.
 
-There is some preliminary empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@JiaTransactionalRelationsFather2012].
-Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@WangParentalInvolvementAfrican2014].
-Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.
+-->
 
-## Sex differences
+Although some research suggests that both parents' qualitative parenting behaviors influence children in similar ways [@LambDevelopmentSignificanceFatherChild2004; @FaganShouldresearchersconceptualize2014], there is also evidence for unique roles of mothers versus fathers.
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@FlouriFathersbehaviorschildren2010].
+Furthermore, whether for evolutionary or cultural reasons [@BogelsDoesFatherKnow2011], mothers and fathers may tend to adopt somewhat specialized roles in terms of qualitative parenting: The aforementioned notions of "attachment" and "activation" relationships posit that mothers empathize more with their children, and  provide more warmth and support, whereas fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior [@MajdandzicFathersChallengingParenting2014; @PaquetteTheorizingfatherchildrelationship2004].
+Empirical research supports these unique roles for mothers and fathers in children's emotional adjustment development: Mothers tend to encourage the expression of emotions, and respond warmly and supportively, whereas fathers structure and sanction emotion expressions to a greater extent [@CassanoInfluenceGenderParental2007; @EisenbergParentalReactionsChildren1999].
+Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for emotional adjustment in adolescence, whereas this pattern was reversed for fathers [@VanLissarolefathersmothersinpress].
+This differentiated pattern was also found with regard to younger children's psychological adjustment [@McDowellDifferencesMothersFathers2003].
+Furthermore, there is some evidence that mothers' negative parenting has a stronger negative impact upon children's emotional adjustment than fathers' negative parenting [@ChangHarshParentingRelation2003]; perhaps because this behavior is incongruous with mothers' prototypical role as an attachment figure.
+Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.
+
+## Child effects
+
+One major limitation of correlational research is that associations between parenting and children's adjustment might be explained equally well by child effects on parents.
+There is now substantial evidence to suggest that children's adjustment problems elicit less warmth, more demandingness, and more negative parenting from parents [@GadeyneLongitudinalRelationsParenting2004; @NepplIntergenerationalContinuityParenting2009; @LenguaBidirectionalassociationstemperament2005].
+Child effects have been more consistently investigated in the adolescent literature than in the childhood literature.
+This might be because the notion of children influencing parents is more salient in adolescence, when parent-child relationships become increasingly egalitarian.
+Nonetheless, findings from the adolescent literature are broadly in line with the childhood literature, suggesting that emotional adjustment problems predicts decreased parental support, and greater control [@OtterpohlCrossLaggedRelationsParenting2015; @VanderGiessenPerceivedAutonomySupport2014; @Wijsbroekdirectioneffectsperceived2011].
+Thus, parents' responses to maladjustment appear to be a combination of punitive responses, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness.
+Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively [@Dixaffectiveorganizationparenting1991].
+We might thus expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.
+
+## Direction of effects
+
+According to the transactional view of family relationships, parents and children reciprocally influence one another [@KerrPerceivedparentingstyle2012; @AmatoParentalDivorceMarital2008].
+This view of the family as a dynamic system whose members reciprocally influence one another is now widely accepted [@KerrPerceivedparentingstyle2012].
+Nevertheless, it is not yet common practice to examine bi-directional effects [@FlouriFathersbehaviorschildren2010].
+Moreover, methods commonly used to examine direction of effects have recently been criticized for yielding biased results if there are stable between-family differences in the variables studied [@Hamakercritiquecrosslaggedpanel2015].
+Recent methodological innovations provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@Hamakercritiquecrosslaggedpanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@AsparouhovDynamicStructuralEquation2018].
+
+Several studies have re-examined the direction of effects between parents and children using these innovative methods, but mostly in adolescent samples.
+One such study found that predictive effects between parenting and delinquency adjustment that were apparent using classical methods, were found to be explained by correlations between stable between-family differences in parenting and adjustment, once these components were included in the model [@KeijsersParentalmonitoringadolescent2016].
+Another study examined the interplay between maternal and paternal qualitative parenting dimensions and adolescents' emotional adjustment, and similarly found that most associations manifested as between-family level correlations.
+Furthermore, child effects were found to substantially outnumber parenting effects [@VanLissarolefathersmothersinpress].
+The authors suggested that this might be due to the relative equality of adolescent-parent relationships, as compared to parent-child relationships.
+If this interpretation is correct, we might expect to find more prominent parenting effects in the present study, with a younger sample.
+These preliminary findings highlight the fact that the direction of effects between parents and children, while controlling for stable between-family differences, remains an important issue to be examined.
+At this stage, however, it is difficult to formulate specific hypotheses about directionality, and the level at which associations between parenting and child development should occur.
+
+## Differences between sons and daughters
 
@@
 
+1) Interactions
+2) Effect of QUANT
+3) Effect of QUAL
+4) Direction of effects
+5) Sex differences
+
+(1)	Qual effects
+(2)	Quant effects
+(3)	Interactions
+(4)	Child effects
+(5)	Direction of effects
+(6)	Sex differences
+
 The present study aimed to make several contributions to the literature on fathers' role in their children's emotional adjustment development.
 First, we included both mothers and fathers, in order to tease apart their common and unique roles.
-Second, our operationalization of involvement incorporated both qualitative involvement, in terms of specific parenting behaviors, and quantitative involvement, operationalized as total joint childcare and differential involvement.
-Third, we employed recent methodological advances to disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative parenting behaviors and emotional adjustment.
-Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
+Second, our operationalization of involvement incorporated both quantitative dimensions, in terms of total joint childcare and differential involvement, and qualitative dimensions, in terms of specific parenting behaviors.
+Third, we examined interactions between quantitative and qualitative involvement, which have been largely overlooked in prior work.
+Fourth, we employed recent methodological advances to disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative parenting behaviors and emotional adjustment.
 
-Based on our review of the literature, we expected specific dimensions of qualitative involvement to be associated with emotional adjustment at both the within- and between-family level.
-The literature suggests that warmth, consistency, and inductive reasoning should be positively associated, whereas negative parenting should be negatively associated, with emotional adjustment.
-With regard to child effects, we might expect some of these paths to be reversed: Positive adjustment might lead to diminished inductive reasoning and parental anger.
+Based on our review of the literature, we expected that involvement quantity and quality would interact, such that a more skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children, and parents' adaptive responses to children's maladjustment.
+In other words, we expected that an unfair childcare distribution would predict suboptimal parenting, and parental responses to maladjustment, due to the strain such a distribution places upon the entire family system.
 We further hypothesized that quantitative involvement would predict children's adjustment directly: Greater joint total care should predict better emotional adjustment, whereas a more skewed distribution of care should predict poorer emotional adjustment.
@@
 A more skewed distribution of care, on the other hand, should show these associations only for the more involved parent.
-Importantly, we also hypothesized that involvement quality and quantity would interact, such that a more skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children, and parents' adaptive responses to children's maladjustment.
-In other words, we expected that an unfair childcare distribution would predict suboptimal parenting, and parental responses to maladjustment, due to the strain such a distribution places upon the entire family system.
-Finally, we accounted for potential effects of child sex.
+Finally, we expected specific qualitative parenting dimensions to be associated with emotional adjustment at both the within- and between-family level.
+The literature suggests that warmth, consistency, and inductive reasoning should be positively associated, whereas negative parenting should be negatively associated, with emotional adjustment.
+With regard to child effects, we might expect some of these paths to be reversed: Positive adjustment might lead to diminished inductive reasoning and parental anger.
+We also accounted for potential effects of child sex.
 
@@
 
-Measures of quantitative involvement varied from wave to wave, due to children's changing care needs.
-We used reports at T1, when both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" (e.g.: "In the past month how often did you help this child brush his/her teeth?").
+Due to children's changing care needs and increasing independence, the operational definition of quantitative involvement varied from wave to wave.
+We therefore used reports at T1 as an indicator of between-family differences in parents' quantitative involvement.
+Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" (e.g.: "In the past month how often did you help this child brush his/her teeth?").
 
@@
 
-We used dynamic structural equation modeling (DSEM) to address our research questions.
-This approach analyzes longitudinal data as a multilevel, autoregressive model.
-First, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences, and within-individual fluctuations.
-In the multilevel model, stable individual differences are captured at the between level, by random intercepts which are based on all scores originating from the same individual.
-At the within level, fluctuations from these random intercepts are analyzed at the within level.
-Moreover, time-lagged copies of all within-level variables are created, in order to estimate a cross-lagged autoregressive model.
+We used multilevel, autoregressive structural equation modelling to address our research questions [@asparouhovDynamicStructuralEquation2018].
+In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
+In the multilevel model, these stable individual differences are represented by random intercepts at the between-persons level.
+Fluctuations from these random intercepts are analyzed at the within level.
+Time-lagged copies of all within-level variables are created, in order to estimate a cross-lagged autoregressive model.
 The coefficients of this cross-lagged autoregressive model are free to vary between individuals, and can thus be predicted by between level moderators, such as quantitative involvement and child sex.
 
-This approach is particularly suited to address our research questions, for several reasons.
-First, controlling for stable between-individual and between-family differences allows us to obtain more accurate estimates of mothers' and fathers' unique roles in child development: Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts.
-Secondly, this approach estimates the interplay between qualitative parenting behaviors and emotional adjustment as a random effect, varying between families.
-We can directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment, by including total joint childcare, and the childcare distribution, as between-family level predictors of the within-family cross-lagged effects between these variables.
+This approach has several advantages over the classic fixed-time cross-lagged panel model.
+First, controlling for stable between-individual and between-family differences allows us to obtain more accurate estimates of mothers' and fathers' unique roles in child development [@hamakerCritiqueCrosslaggedPanel2015; @berryPracticalInterpretabilityCrosslagged2017]: Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts.
+Secondly, this approach allows us to directly test our hypothesis that between-individual differences in quantitative involvement would interact with the interplay between qualitative involvement and emotional adjustment.
+To this end, we included total joint childcare, and the childcare distribution, as between-family level predictors, and examined their predictive effect on cross-lagged effects between qualitative parenting behaviors and emotional adjustment.
 Child sex was included as a moderator in the same way.
 
-DSEM models can only be estimated using Bayesian methods.
-We ran two chains of Mplus 8.0's Gibbs sampler for 50000 iterations each.
-We used mildly informative priors for the variance/covariance matrices to aid convergence, and  regularizing priors centered on zero for all regression coefficients.
+All analyses were conducted in Mplus 8.0, controlled from R using MplusAutomation.
+In Mplus, this type of multilevel autoregressive model can currently only be estimated using Bayesian methods.
+We used Gibbs sampling to estimate the model, with two chains of 50000 iterations each.
+We used mildly informative priors for the variance/covariance matrices to aid convergence, and regularizing priors, centered on zero, for all regression coefficients.
 We assessed convergence through visual inspections of the chains and posterior distributions.
@@
 2) incorporated quantitative involvement, in terms of total joint childcare and the childcare distribution, and qualitative involvement
-3) interactions between qualitative and quantitative involvement
+3) interactions between quantitative and qualitative involvement
 4) assess the direction of effects between parenting behaviors and emotional adjustment.
@@
    * Child effects, some paths reversed: Maladjustment should elicit decreased responsiveness, but increased structure, reasoning, and negative parenting. 
-1) Interactions between involvement quality and quantity
+1) Interactions between involvement quantity and quality
    * skewed distribution of quantitative involvement would interfere with the effects of qualitative parenting behaviors on children (parenting effects), and parents' adaptive responses to children's maladjustment (child effects).
@@
 This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment.
+Finally, contrary to hypotheses, we found no significant correlations between maternal and paternal warmth, and children's emotional adjustment, at the between-family level.
 
 <!--(Discussion: Although inductive reasoning is considered to be a disciplinary behavior that stimulates children's positive development (better ref here), the fact that it is a disciplinary behavior might explain the negative association with emotional adjustment.).-->
-
-Finally, contrary to hypotheses, we found no significant correlations between maternal and paternal warmth, and children's emotional adjustment, at the between-family level.
 
@@
 
-## Associations between quantitative involvement and adjustment and parenting
+## Interactions between involvement quantity and quality
 
-We examined regression coefficients from total join involvement and relative involvement to the random intercepts of emotional adjustment and qualitative parenting behaviors.
-We found that greater joint care was negatively associated with emotional adjustment, counter to our hypothesis.
-Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
-With regard to the distribution of childcare, however, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.
-
-With regard to parenting, we did not find the hypothesized negative effects of joint total involvement on parental anger, but, in partial support of our hypotheses, we did find that, for both parents, joint involvement positively predicted greater consistency, inductive reasoning, and warmth.
-Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers.
-
-With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict reduced anger, and greater consistency, inductive reasoning, and warmth, only for the more involved parent.
-For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers.
-For fathers' parenting behavior, we found that, in line with this hypothesis, fathers displayed less consistency, inductive reasoning, and warmth when mothers were relatively more involved than fathers.
-
-## Interactions between involvement quality and quantity
-
-To examine interactions between involvement quality and quantity, we tested whether between-family differences in relative and joint total quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment.
+To examine interactions between involvement quantity and quality, we tested whether between-family differences in relative and joint total quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment.
 
@@
 
+## Associations between quantitative involvement and adjustment and parenting
+
+We examined regression coefficients from total join involvement and relative involvement to the random intercepts of emotional adjustment and qualitative parenting behaviors.
+We found that greater joint care was negatively associated with emotional adjustment, counter to our hypothesis.
+Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
+With regard to the distribution of childcare, however, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.
+
+With regard to parenting, we did not find the hypothesized negative effects of joint total involvement on parental anger, but, in partial support of our hypotheses, we did find that, for both parents, joint involvement positively predicted greater consistency, inductive reasoning, and warmth.
+Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers.
+
+With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict reduced anger, and greater consistency, inductive reasoning, and warmth, only for the more involved parent.
+For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers.
+For fathers' parenting behavior, we found that, in line with this hypothesis, fathers displayed less consistency, inductive reasoning, and warmth when mothers were relatively more involved than fathers.
+
 ## Effects of child sex
@@
 We set out to disentangle mothers' versus fathers' unique roles in children's emotional adjustment development.
-In doing so, we incorporated both qualitative and quantitative measures of parental involvement, and examined interactions between these two aspects of involvement, which have historically been examined mostly in isolation.
+In doing so, we incorporated both quantitative and qualitative measures of parental involvement, and examined interactions between these two aspects of involvement, which have historically been examined mostly in isolation.
 To this end, we used innovative methods which enabled us to include between-family differences in quantitative involvement, operationalized as total joint care and the relative childcare distribution, as predictors of within-family cross-lagged effects between qualitative involvement behaviors and emotional adjustment.
+
+## Interactions between quantitative and qualitative involvement
+
+With regard to our primary aim of examining interactions between quantitative and qualitative involvement, we found limited evidence that within-family parenting effects - cross-lagged effects from qualitative involvement behaviors on emotional adjustment - were moderated by quantitative involvement.
+One such effect, which was in line with predictions, is that the effect of maternal warmth on adjustment was contingent on both total joint care, and the relative childcare distribution.
+Mothers' warmth predicted children's adjustment more strongly if mothers were relatively more involved than fathers.
+This suggests that having a warm mother is of greater developmental importance with regard to children's emotional adjustment if fathers are less involved with the day-to-day minutiae of childcare.
+Conversely, we also found that mothers' warmth was less strong of a predictor of adjustment development when both parents were more involved.
+This suggests that maternal warmth is less essential when both parents are highly involved.
+Together, these findings are reminiscent of a "buffering" effect, where mothers' warmth can make up for any disadvantages incurred by having a relatively uninvolved father.
+
+Interactions between quantitative and qualitative involvement were more abundant in the context of child effects on parents; i.e., the way parents respond to children's emotional adjustment or maladjustment.
+Specifically, we found that when mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment.
+Considering that inductive reasoning is a form of discipline, this finding suggests that, when mothers are relatively more involved in care tasks, fathers take more of a role as disciplinarian.
+
+<!-- the domains of discipline and play tend to be the areas in which fathers feel most comfortable (Parke 2000).
+ Mothers’ satisfaction with the division of labor within the household was consistently related to greater levels of paternal responsibility in all four domains (school work, discipline, daily care, fun activities), but fathers’ satisfaction was only related to greater paternal involvement in two of the four domains (discipline and fun activities).
+-->
+This finding is complemented by the negative effect of total joint involvement on the link from emotional adjustment to maternal inductive reasoning, which suggests that mothers also have more of a role as disciplinarian in families where both parents are highly involved.
+We also found that emotional adjustment became less predictive of maternal warmth when parents' joint involvement was higher, and when mothers' relative involvement was greater compared to fathers.
+This finding can be interpreted in the context of the sizable average effect of adjustment on maternal warmth.
+On average, increases in emotional adjustment predicted substantial increases in maternal warmth.
+In families where mothers' involvement was higher - either in terms of joint, or relative involvement - maternal warmth was thus less contingent on children's adjustment.
 
@@
 The present study operationalized quantitative involvement in terms of total joint care provided by parents, and the childcare distribution, which captures differences between mothers and fathers.
-This operationalization addresses a criticism raised in the literature, that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001].
+This operationalization addresses a criticism raised in the literature, that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@DeutschPaternalParticipationChild2001].
 For total joint care, we found that, counter to our hypothesis, joint care was *negatively* associated with between-family differences in children's emotional adjustment.
@@
 This suggests that the majority of children experienced the disadvantage in emotional adjustment associated with differential involvement.
-Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001].
+Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@ShockleyItnotwhat2018], our research is one of few to examine the implications for children's adjustment [see also @DeutschPaternalParticipationChild2001].
 Of course, causality cannot be inferred from correlational findings at the between-family level.
 Future research should examine potential pathways, such as whether fathers shy away from childcare responsibilities when children display poor emotional adjustment, or whether children's adjustment suffers when they feel like both parents are not equally invested in their wellbeing, or whether differential involvement strains the marital relationship, and this strain in turn spills over and affects children's emotional adjustment.
-
-## Interactions between qualitative and quantitative involvement
-
-We found limited evidence that within-family parenting effects - cross-lagged effects from qualitative involvement behaviors on emotional adjustment - were moderated by quantitative involvement.
-One such effect, which was in line with predictions, is that the effect of maternal warmth on adjustment was contingent on both total joint care, and the relative childcare distribution.
-Mothers' warmth predicted children's adjustment more strongly if mothers were relatively more involved than fathers.
-This suggests that having a warm mother is of greater developmental importance with regard to children's emotional adjustment if fathers are less involved with the day-to-day minutiae of childcare.
-Conversely, we also found that mothers' warmth was less strong of a predictor of adjustment development when both parents were more involved.
-This suggests that maternal warmth is less essential when both parents are highly involved.
-Together, these findings are reminiscent of a "buffering" effect, where mothers' warmth can make up for any disadvantages incurred by having a relatively uninvolved father.
-
-Interactions between qualitative and quantitative involvement were more abundant in the context of child effects on parents; i.e., the way parents respond to children's emotional adjustment or maladjustment.
-Specifically, we found that when mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment.
-Considering that inductive reasoning is a form of discipline, this finding suggests that, when mothers are relatively more involved in care tasks, fathers take more of a role as disciplinarian.
-
-<!-- the domains of discipline and play tend to be the areas in which fathers feel most comfortable (Parke 2000).
- Mothers’ satisfaction with the division of labor within the household was consistently related to greater levels of paternal responsibility in all four domains (school work, discipline, daily care, fun activities), but fathers’ satisfaction was only related to greater paternal involvement in two of the four domains (discipline and fun activities).
--->
-This finding is complemented by the negative effect of total joint involvement on the link from emotional adjustment to maternal inductive reasoning, which suggests that mothers also have more of a role as disciplinarian in families where both parents are highly involved.
-We also found that emotional adjustment became less predictive of maternal warmth when parents' joint involvement was higher, and when mothers' relative involvement was greater compared to fathers.
-This finding can be interpreted in the context of the sizable average effect of adjustment on maternal warmth.
-On average, increases in emotional adjustment predicted substantial increases in maternal warmth.
-In families where mothers' involvement was higher - either in terms of joint, or relative involvement - maternal warmth was thus less contingent on children's adjustment.
 
@@
 With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers.
-This is in line with prior research, which suggested that warmth is an important socialization mechanism for mothers [REF mezelf].
+This is in line with prior research, which suggested that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].
 
@@
 Nowadays it is commonly acknowledged that parents and children exert mutual influences.
-However, commonly used developmental methods have been criticized for failing to properly reveal the direction of effects: When stable between-family differences are not accounted for, estimates of cross-lagged effects can be substantially biased [@hamakerCritiqueCrosslaggedPanel2015].
-One recent study examined the interplay between qualitative parenting behaviors and emotion regulation development in adolescence, and found that child effects on parents substantially outweighed parent effects on children [@vanlissaRoleFathersMothers2018a].
+However, commonly used developmental methods have been criticized for failing to properly reveal the direction of effects: When stable between-family differences are not accounted for, estimates of cross-lagged effects can be substantially biased [REF HAMAKER].
+One recent study examined the interplay between qualitative parenting behaviors and emotion regulation development in adolescence, and found that child effects on parents substantially outweighed parent effects on children [REF mezelf].
 The authors speculated that this finding might be due to the relatively egalitarianism of parent-child relationships in adolescence, as compared to earlier childhood.
@@
 This allowed us to provide a more nuanced overview of the way parental involvement is linked with emotional adjustment development, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs.
-Additionally, although some prior studies have also incorporated both quality and quantity of parenting, this work has typically included the absolute amounts of maternal and paternal involvement.
-A limitation of this approach is that it is not possible to distinguish the effect the childcare distribution from the effect of the total amount of care children receive [see @deutschPaternalParticipationChild2001].
+Additionally, although some prior studies have also incorporated both quantity and quality of parenting, this work has typically included the absolute amounts of maternal and paternal involvement.
+A limitation of this approach is that it is not possible to distinguish the effect the childcare distribution from the effect of the total amount of care children receive [see @DeutschPaternalParticipationChild2001].
 We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement.
@@
 Teachers' reports do have clear benefits compared to alternatives: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies.
-Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983].
+Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@MaccobySocializationcontextfamily1983].
 We focused on specific parenting behaviors for two main reasons.
@@
 Our approach assumes that a "snapshot" of quantitative involvement at age 6 is representative for parents' general level of involvement.
-Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact.
+Consequently, we are unable to assess whether within-person fluctuations in quantitative and qualitative involvement interact.
 Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range, so measurement invariance for quantitative involvement can be assured.
@@
 Evidence for fathers' unique role was relatively limited; significant within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role in the capacity of disciplinarian.
-The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - fathers' role is more context dependent.
+The aforementioned interaction effects between quantitative and qualitative involvement, however, seem to suggest that - above all - fathers' role is more context dependent.
 

### 2019-05-01 Paper_reorganized.Rmd
Final draft before submission

@@
 ---
-title             : "Quantitative involvement moderates the interplay between mothers' versus fathers' qualitative parenting behaviors and children's emotional adjustment development"
-shorttitle        : "QUALITATIVE AND QUANTITATIVE INVOLVEMENT"
+title             : "Quantitative involvement moderates associations between qualitative parenting behaviors and children's emotional adjustment development"
+shorttitle        : "INVOLVEMENT AND EMOTIONAL ADJUSTMENT"
 
@@
 library("papaja")
+library(motley)
+source("rfunctions.R")
+load("results.RData")
+load("table_coefficients.RData")
 mask <- TRUE
@@
 
-Emotional adjustment, defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001], is an important developmental outcome, integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013; @aldaoEmotionregulationStrategiesPsychopathology2010].
-Fathers have been hypothesized to play a unique role in children's emotional adjustment development, because they encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
-The classical notion of the “attachment relationship” is rooted in security and exploration (REF).
-Fatherhood researchers have posited that, whereas the mother-child attachment relationship primarily provides warmth and security, exploration might be somewhat more central to the father-child relationship, which can has thus been construed as an “activation relationship” [@paquetteFatherChildActivationRelationship2013].
-By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a broad spectrum of emotions, thereby structuring opportunities to expand children's emotional repertoire.
+Emotional adjustment is defined as the absence of internalizing symptoms and manifest signs of emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
+This important developmental outcome is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013; @aldaoEmotionregulationStrategiesPsychopathology2010].
+According to the fatherhood literature, emotional adjustment development is one domain in which fathers might play a unique role, because fathers encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+The classical notion of the "attachment relationship" is rooted in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Fatherhood researchers have posited that, whereas the mother-child attachment relationship primarily provides warmth and security, exploration might be somewhat more central to the father-child relationship, which can has thus been construed as an "activation relationship" [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers are thought to stimulate a broad spectrum of emotions, which expands children's emotional repertoire, and engages their developing regulatory abilities.
 
@@
 Although it is now widely acknowledged that both the quality and quantity of parental involvement are relevant factors in child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated.
-This might be, in part, because research interests have historically varied across discipline lines: Developmental psychology has focused primarily on how qualitative parenting behaviors influence child development, whereas sociology and gender studies have focused more on involvement quantity, and its implications for maternal wellbeing [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement in relation to children's emotional adjustment.
+This might be, in part, because research interests have historically varied across discipline lines: Developmental psychology has focused primarily on how qualitative parenting behaviors influence child development, whereas sociology and gender studies have focused more on involvement quantity, and its implications for maternal well-being and gender equality [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
@@
 Responsiveness refers to warm and supportive parenting, which substantial evidence indicates is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].
+
+The literature on links between demandingness and emotional adjustment is more limited, and mixed, than the literature on responsiveness [@morrisRoleFamilyContext2007].
+This appears to be, in part, because some forms of demandingness help children meet societal expectations, whereas other forms are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013].
 Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior.
-Two aspects of demandingness investigated in the present study are consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
-The literature on links between demandingness and emotional adjustment is more limited, and mixed [@morrisRoleFamilyContext2007]; in part because some forms of demandingness help children meet societal expectations, whereas other forms are overly restrictive, and undermine children's autonomy needs [@barberCentralityControlParenting2013].
-A fourth dimension, which has consistently been linked with children's emotional maladjustment, but is distinct from the parenting styles typology, is negative parenting [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
-We operationalized negative parenting as one of its specific subdimensions: parental anger.
-The anger parents express in interactions with children might reflect faillure to regulate their own negative emotions, and models the use of anger in communication, and undermines children's emotion regulation [@scaramellaParentingChildrenDistress2008].
+Two specific aspects of demandingness, investigated in the present study, are consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
+Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990].
+Inductive reasoning is thought to promote the internalization of parental expectations, and is integral to children's development of moral emotions [@hoffmanEmpathyMoralDevelopment2000].
+It will thus likely play a positive role in general emotional development as well.
 
-<!-- 
-Inductive reasoning is thought to be integral to children's development of moral emotions [@HoffmanEmpathymoraldevelopment2000], and might thus play a positive role general emotional development as well.
--->
+Negative parenting is an additional qualitative parenting dimension which is regularly included in the parenting styles literature, and appears to be negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+We operationalized negative parenting as one of its specific subcomponents: parental anger.
+It is thought that the anger parents express in interactions with children reflects a faillure to regulate their own negative emotions, models children's use of anger in communication, and undermines children's emotion regulation [@scaramellaParentingChildrenDistress2008].
 
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010].
 Although some research suggests that both parents' qualitative parenting behaviors influence children in similar ways [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique roles of mothers versus fathers.
-Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010].
-Furthermore, whether for evolutionary or cultural reasons [@bogelsDoesFatherKnow2011], mothers and fathers may tend to adopt somewhat specialized roles in terms of qualitative parenting: The aforementioned notions of "attachment" and "activation" relationships posit that mothers empathize more with their children, and  provide more warmth and support, whereas fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior [@majdandzicFathersChallengingParenting2014; @paquetteTheorizingFatherchildRelationship2004].
+Whether for evolutionary or cultural reasons [@bogelsDoesFatherKnow2011], mothers and fathers sometimes appear to fulfill somewhat specialized roles in terms of qualitative parenting: The aforementioned notions of "attachment" and "activation" relationships posit that mothers empathize more with their children, and  provide more warmth and support, whereas fathers activate children by encouraging them to take risks and overcome obstacles, and live up to certain standards for behavior [@majdandzicFathersChallengingParenting2014; @paquetteTheorizingFatherchildRelationship2004].
 Empirical research supports these unique roles for mothers and fathers in children's emotional adjustment development: Mothers tend to encourage the expression of emotions, and respond warmly and supportively, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
@@
 Nevertheless, examining bi-directional effects is not yet standard practice in the childhood literature [@flouriFathersBehaviorsChildren2010].
-Available empirical evidence suggests that children's adjustment problems elicit less warmth, more demandingness, and more negative parenting from parents [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @LenguaBidirectionalassociationstemperament2005].
+Available empirical evidence suggests that children's adjustment problems elicit less warmth, more demandingness, and more negative parenting from parents [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @lenguaBidirectionalAssociationsTemperament2005].
 In the adolescent literature, bi-directional effects are more commonly examined.
 This might be because the notion of children influencing parents is more salient in adolescence, when parent-child relationships become increasingly egalitarian.
-Findings from the adolescent literature are broadly in line with findings in childhood; suggesting that emotional adjustment problems predicts decreased parental support, and greater control [@OtterpohlCrossLaggedRelationsParenting2015; @VanderGiessenPerceivedAutonomySupport2014; @Wijsbroekdirectioneffectsperceived2011].
+Findings from the adolescent literature are broadly in line with findings in childhood; suggesting that emotional adjustment problems predicts decreased parental support, and greater control [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014; @wijsbroekDirectionEffectsPerceived2011].
 Thus, parents' responses to maladjustment appear to be a combination of punitive responses, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness.
-Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively [@Dixaffectiveorganizationparenting1991].
+Moreover, parents' responses to adjustment problems are contingent on children's age: Because younger children are perceived to be less competent than older children, mothers respond to their misbehavior more inductively, and less punitively [@dixAffectiveOrganizationParenting1991].
 We might thus expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.
@@
 
-Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if there are stable between-family differences in the variables studied [@Hamakercritiquecrosslaggedpanel2015].
-Recent methodological innovations provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@Hamakercritiquecrosslaggedpanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@AsparouhovDynamicStructuralEquation2018].
+Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if there are stable between-family differences in the variables studied [@hamakerCritiqueCrosslaggedPanel2015].
+Recent methodological innovations provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].
 
 Several studies have re-examined the direction of effects between parents and children using these innovative methods, but mostly in adolescent samples.
-One such study found that predictive effects between parenting and delinquency adjustment that were apparent using classical methods, were found to be explained by correlations between stable between-family differences in parenting and adjustment, once these components were included in the model [@KeijsersParentalmonitoringadolescent2016].
+One such study found that predictive effects between parenting and delinquency adjustment that were apparent using classical methods, were found to be explained by correlations between stable between-family differences in parenting and adjustment, once these components were included in the model [@keijsersParentalMonitoringAdolescent2016].
 Another study examined the interplay between maternal and paternal qualitative parenting dimensions and adolescents' emotional adjustment, and similarly found that most associations manifested as between-family level correlations.
@@
 
+<!--
+Better quality parental relationships are usually associated with more involved fathering (Belsky 1984;
+Bouchard & Lee 2000; Coiro & Emery 1998; McBride & Mills 1993; Verhoeven et al. 2007). Further, when
+relationships are better, this is also evident in more positive or sensitive parenting styles and more cooperative
+and supportive co‑parenting (Floyd, Gilliom & Costigan 1998; Verhoeven et al. 2007).
+-->
+
 ## Quantitative involvement
@@
 
-Quantitative involvement has been researched extensively in terms of the amount of time spent caring for the child [see @pleckPaternalInvolvementRevised2004].
-However, some have argued that the the way childcare tasks are divided is more relevant for children's adjustment than the absolute amount of time invested by each parent [@KrollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003].
-Indeed, research has shown that the equity of the childcare distribution affects child development directly [@deutschPaternalParticipationChild2001], and impacts the emotional climate of the interparental relationship [@CraigDoesFatherCare2006].
+Quantitative involvement has been researched extensively in terms of the absolute amount of time parents spend, or tasks they perform, in caring for the child [see @pleckPaternalInvolvementRevised2004].
+However, some have argued that the the way childcare tasks are divided is more relevant for children's adjustment than the absolute amount of time invested by each parent [@krollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003].
+Indeed, research has shown that the equity of the childcare distribution affects child development directly [@deutschPaternalParticipationChild2001], and impacts the emotional climate of the interparental relationship [@craigDoesFatherCare2006].
 When childcare is unequally distributed, women often perform a disproportionate share of the work.
-This strains the marital relationship [@shockleyItNotWhat2018], and the resulting stress erodes parenting quality, particularly for parents of difficult children [@SheltonInterparentalconflictnegative2008].
+This strains the marital relationship [@shockleyItNotWhat2018], and the resulting stress erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].
+Thus, operationalizing quantitative involvement in terms of relative involvement might capture patterns of greater relevance.
 
 There are also methodological advantages to focusing on the distribution of childcare.
-Operationalizing quantitative involvement in terms of both parents' absolute involvement, e.g., as hours spent caring for the child, incurs a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care [see @deutschPaternalParticipationChild2001].
-This confound is mitigated by operationalizing quantitative involvement in terms of the childcare distribution, and controlling for total joint care.
-Some prior research has attempted to capture the childcare distribution directly, by asking which has most responsibility for certain tasks, on a continuum from the mother doing everything to the father doing everything [@KrollEarlyFatherInvolvement2016].
-When quantitative involvement is measured in terms of absolute involvement, however, an elegant statistical solution can disentangle the childcare distribution from total joint care.
-The so-called "latent mean and difference" approach, which is typically used to capture the difference between consecutive longitudinal measurements [@McArdleLatentVariableModeling2009], can be utilized to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other.
+Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Any effect that appears to be due to fathers' greater involvement might, in fact, be due to children receiving greater a greater total amount of parental care [see @deutschPaternalParticipationChild2001].
+This risk is exacerbated by the fact that, within families, fathers' absolute involvement is strongly associated with mothers' absolute involvement [@russellCoparentingEarlySchoolage1994].
+Operationalizing quantitative involvement in terms of the childcare distribution, and controlling for total joint care, mitigates this confound.
+Some prior research has attempted to capture the childcare distribution directly, by asking which parent is most responsible for certain tasks, on a continuum from the mother doing everything to the father doing everything [@krollEarlyFatherInvolvement2016].
+When quantitative involvement has been measured in terms of absolute involvement, however, an elegant statistical solution can be applied to disentangle the childcare distribution from total joint care.
+The so-called "latent mean and difference" approach, which is typically used to capture the difference between consecutive longitudinal measurements [@mcardleLatentVariableModeling2009], can be utilized to re-express mothers' and fathers' absolute quantitative involvement into a "latent mean" of involvement, which represents the total joint care provided by both parents, and latent differential involvement, which represents the extent to which one parent is more involved than the other.
 This approach taps directly into a theoretically relevant construct, and prevents confounding.
+
+A final consideration in the operationalization of quantitative involvement is that the type of tasks parents perform in the context of child care change rapidly as children develop [@wangATUSFathersInvolvement2009].
+By the time children are school-aged, their dependence on instrumental care might be relatively minor.
+When the behaviors constituting quantitative involvement change over time, it is not clear how one might establish measurement invariance on a macro-longitudinal time scale.
+Consequently, in the present study, we took quantitative involvement at the first measurement wave as an indicator of between-individual differences in quantitative involvement.
+This means that we cannot investigate within-family dynamics pertaining to change in quantitative involvement.
+
 ## Main effects of quantitative involvement
@@
 In many countries, mothers continue to spend more time with their children than fathers [@hofferth28669resident], particularly when it comes to routine care tasks [@craigHowMothersFathers2011].
-Fathers' involvement is on the increase, however [@YeungChildrenTimeFathers2001].
+Fathers' involvement is on the increase, however [@yeungChildrenTimeFathers2001].
 The amount of time parents spend with children is likely to predict the quality of parenting behaviors [@sabattiniRelationMothersFathers2004], through practice and experience.
-Indeed, more involved parents tend to experience greater parenting self-efficacy [@GialloMothersfathersinvolvement2013].
+Indeed, more involved parents tend to experience greater parenting self-efficacy [@gialloMothersFathersInvolvement2013a].
 We might thus hypothesize that greater total joint care should positively predict both parents' parenting quality.
+This hypothesis is also supported by the notion of parental "buffering", which holds that, when both parents are more involved, they monitor each other's parenting behaviors, thereby ensuring that children are raised well [@mclanahanConsequencesFatherAbsence2014].
 
-By the same logic, we might expect that, when one parent is more involved than the other, the parenting quality of the more involved parent would improve.
+The notion that parenting quality improves through practice [@sabattiniRelationMothersFathers2004] might suggests that, when one parent is more involved than the other, the more involved parent would display more desirable qualitative parenting behaviors.
 However, an argument to the contrary can be made as well.
-Retrospective reports by 19-year old participants indicate that an unequal childcare distribution tends to go hand in hand with suboptimal qualitative parenting behaviors - specifically, increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
+The notion of parental buffering would suggest that both parents' qualitative parenting behaviors might suffer when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014].
+Indeed, retrospective reports by 19-year old participants indicate that an unequal childcare distribution tends to go hand in hand with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
 Greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship.
 Childcare responsibilities are a heavy burden if they come down on a single parent's shoulders [@deutschPaternalParticipationChild2001].
-Mothers thus benefit directly when fathers relieve some of this burden [@PleckPaternalinvolvementRevised2004].
-Consequently, mothers tend to be more satisfied when fathers are more involved [@PharesFathersMothersInvolvement2009].
-An unfair distribution of childcare, on the other hand, is associated with greater maternal anger and resentment [@RossGenderParenthoodAnger1996].
+Mothers thus benefit directly when fathers relieve some of this burden [@pleckPaternalInvolvementRevised2004].
+Consequently, mothers tend to be more satisfied when fathers are more involved [@pharesFathersMothersInvolvement2009].
+Moreover, balanced co-parenting appears to be associated with more desirable qualitative parenting behaviors; at least in terms of parental warmth [@bondsExaminingProcessWhich2007; @barnettInterdependenceParentingMothers2008].
+An unfair distribution of childcare, on the other hand, is associated with greater maternal anger and resentment [@rossGenderParenthoodAnger1996].
 This strains the interparental relationship, interferes with sensitive parenting, and undermines child wellbeing [@davies2006interparental].
@@
 Consequences for children's emotional outcomes have received less attention to date.
-Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @BronfenbrennerEcologyHumanDevelopment1979].
+Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
 This suggests that the total amount of care provided should be positively associated with children's emotional adjustment.
@@
 
-Interactions between involvement quality and quantity have largely been overlooked in the aforementioned theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@JaffeeLifeFatherBenefits2003].
-However, It is conceivable that children's emotional adjustment might be differently impacted by paternal control if that father is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent.
+Interactions between involvement quality and quantity have largely been overlooked in the aforementioned theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@jaffeeLifeFatherBenefits2003].
+Children's emotional adjustment might be differently impacted by paternal control if that father is generally highly involved with the day-to-day minutiae of childrearing, than when he is generally absent.
 Conversely, children's emotional adjustment might elicit different qualitative parenting behaviors from hands-on fathers, who are aware of the reasons for their children's emotional troubles, than from out-of-touch fathers, who might perceive maladjustment as a problem behavior that should be sanctioned.
@@
 
-In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@FeinbergInternalStructureEcological2003].
+In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@feinbergInternalStructureEcological2003].
 Coparenting refers to the ways parents relate to each other in the role of parent.
@@
 
-There is some preliminary empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@JiaTransactionalRelationsFather2012].
-Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@WangParentalInvolvementAfrican2014].
+There is some preliminary empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012].
+Furthermore, in adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014].
 Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.
@@
 There is substantial evidence that parents tend to raise boys and girls differently.
-Prior research has indicated that parents tend to be more controlling with girls than with boys **(e.g., Pomerantz & Ruble, 1998; Smetana & Daddis, 2002)**, and express more negative parenting with boys than with girls **(Straus and Stewart 1999)**.
+Prior research has indicated that parents tend to be more controlling with girls than with boys [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and express more negative parenting with boys than with girls [@strausCorporalPunishmentAmerican1999].
 Another point to consider is the interplay between parent and child sex.
-According to social learning theory, the same-sex parent plays a stronger role in a child's development **(Bussey & Bandura, 1999)**, particularly in the socialization of sex-typed behavior, such as display rules for emotions [@Brodysocializationgenderdifferences2000].
+According to social learning theory, the same-sex parent plays a stronger role in a child's development [@busseySocialCognitiveTheory1999], particularly in the socialization of sex-typed behavior, such as display rules for emotions [@brodySocializationGenderDifferences2000].
 Thus, we might expect child sex to predict the same-sex parents' qualitative involvement, and associations between parenting and emotional adjustment might be stronger for same-sex parent-child dyads.
-There is also evidence that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships **(Lovas, 2005; Schoppe-Sullivan et al., 2006)**.
+There is also evidence that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
 These findings suggest that boys tend to receive a "tougher" upbringing than girls.
@@
 
-There is little evidence for sex differences in emotional adjustment in young children, as indicated by a meta-analysis which found no sex differences in the temperament dimension of negative affectivity, based primarily on infant and toddler samples [@Else-questGenderdifferencestemperament2006].
-Evidence for such sex differences is sometimes found in older children [@SilkAdolescentsemotionregulation2003], but these might appear and intensify in puberty [@HillIntensificationGenderRelatedRole1983].
+There is little evidence for sex differences in emotional adjustment in young children, as indicated by a meta-analysis which found no sex differences in the temperament dimension of negative affectivity, based primarily on infant and toddler samples [@else-questGenderDifferencesTemperament2006a].
+Evidence for such sex differences is sometimes found in older children [@silkAdolescentsEmotionRegulation2003], but these might appear and intensify in puberty [@hillIntensificationGenderRelatedRole1983].
 We consequently did not expect to find sex differences in children's average levels of emotional adjustment.
@@
 `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population.
-A detailed description of the study design, sample, and procedure is available at `r masked(" http://data.growingupinaustralia.gov.au/")`.
+Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.
 
-Our target population consisted of native `r masked("Australian")`, non-minority, intact, biological families.
-This yielded a subsample of 1884 families.
+Our target population consisted of a subsample of 1884 native `r masked("Australian")`, non-minority, intact, biological families.
 At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$).
@@
 
-### Quantitative involvement.
-
-Measures of quantitative involvement varied from wave to wave, due to children's changing care needs.
-We used reports at T1, when both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" (e.g.: "In the past month how often did you help this child brush his/her teeth?").
-
 ### Qualitative involvement.
@@
 Parents completed measures of several qualitative parenting dimensions.
-Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @PatersonAssociationBehaviouralAdjustment2001].
-Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth **(NLSCY; Statistics Canada, 2000)**.
+Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001].
+Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
 For these instruments, parents reported on their own behavior using five-point scales, ranging from "never or almost never" to "always or almost always".
+
+### Quantitative involvement.
+
+Measures of quantitative involvement varied from wave to wave, due to children's changing care needs [@wangATUSFathersInvolvement2009].
+We used reports at T1 as an indicator of between-family differences in quantitative involvement.
+Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].
 
@@
 
-Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the Strengths and Difficulties Questionnaire, reported by teachers on a three-point scale from "not true" to "certainly true" **(e.g., "Often unhappy, downhearted or tearful"; Goodman, 1999)**.
-Scale descriptives are displayed in Table X, and scale score intercorrelations are reported in the online supplement SX.
+Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the Strengths and Difficulties Questionnaire, reported by teachers on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].
 
@@
 To establish that metric measurement invariance could be imposed across measurement waves, and across mothers and fathers, we estimated structurally invariant CFA's for all of the study variables, and then imposed constraints on the factor loadings across time and reporter.
-Although chi-square difference tests indicated that the fit was significantly worse for metric invariance models than for structural invariance models for all scales (all $ps \leq$ .03$), chi-square is overpowered in large samples.
+Chi-square difference tests did indicate significantly worse fit for metric invariance models than for structural invariance models, (all $ps \leq$ .03$).
+However, chi-square difference tests are overpowered in large samples.
 The difference in fit between the structural invariance models and metric invariance models was smaller than .01 according to conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
@@
 
-We used dynamic structural equation modeling (DSEM) to address our research questions.
-This approach analyzes longitudinal data as a multilevel, autoregressive model.
-First, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences, and within-individual fluctuations.
-In the multilevel model, stable individual differences are captured at the between level, by random intercepts which are based on all scores originating from the same individual.
-At the within level, fluctuations from these random intercepts are analyzed at the within level.
-Moreover, time-lagged copies of all within-level variables are created, in order to estimate a cross-lagged autoregressive model.
+We used multilevel, autoregressive structural equation modelling to address our research questions [@asparouhovDynamicStructuralEquation2018].
+In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
+In the multilevel model, these stable individual differences are represented by random intercepts at the between-persons level.
+Fluctuations from these random intercepts are analyzed at the within level.
+Time-lagged copies of all within-level variables are created, in order to estimate a cross-lagged autoregressive model.
 The coefficients of this cross-lagged autoregressive model are free to vary between individuals, and can thus be predicted by between level moderators, such as quantitative involvement and child sex.
 
-This approach is particularly suited to address our research questions, for several reasons.
-First, controlling for stable between-individual and between-family differences allows us to obtain more accurate estimates of mothers' and fathers' unique roles in child development: Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts.
-Secondly, this approach estimates the interplay between qualitative parenting behaviors and emotional adjustment as a random effect, varying between families.
-We can directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment, by including total joint childcare, and the childcare distribution, as between-family level predictors of the within-family cross-lagged effects between these variables.
-Child sex was included as a moderator in the same way.
+This approach has several advantages over the classic fixed-time cross-lagged panel model.
+First, controlling for stable between-individual and between-family differences allows us to obtain more accurate estimates of mothers' and fathers' unique roles in child development [@hamakerCritiqueCrosslaggedPanel2015; @berryPracticalInterpretabilityCrosslagged2017]: Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts.
+Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
+In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.
 
-DSEM models can only be estimated using Bayesian methods.
-We ran two chains of Mplus 8.0's Gibbs sampler for 50000 iterations each.
-We used mildly informative priors for the variance/covariance matrices to aid convergence, and  regularizing priors centered on zero for all regression coefficients.
+All analyses were conducted in Mplus 8.0, controlled from R using MplusAutomation.
+In Mplus, this type of multilevel autoregressive model can currently only be estimated using Bayesian methods.
+We used Gibbs sampling to estimate the model, with two chains of 50000 iterations each.
+We used mildly informative priors for the variance/covariance matrices, and regularizing priors centered on zero for all regression coefficients.
 We assessed convergence through visual inspections of the chains and posterior distributions.
@@
 However, for convenience's sake, we focus on interpreting results whose HPDI did not include zero.
-Significance is not a Bayesian concept, but we will use the shorthand "significant" to refer to parameters whose 95\% HPDI did not include zero.
+Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate; i.e., for a positive estimate, this is the proportion of the posterior below zero.
+A complete overview of the standardized and unstandardized model parameters is provided in Supplementary Table 4.
 
@@
 
-Standardized and unstandardized model parameters are reported in Table 1. A graphical summary of the parameters discussed in the Results section is displayed in Figure 1.
-
-<!-- structure notes
+<!--structure notes
 1) mothers and fathers, in order to tease apart their common and unique roles. 
@@
 
-## Average associations between qualitative parenting dimensions and emotional adjustment
+## Associations between qualitative parenting dimensions and emotional adjustment
 
@@
 
-### Between-family correlations
+### Between-family level correlations
 
-At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment.
+At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabbetweencor)).
 The random intercept of each family is based on their average level across all waves of the study.
 Correlations between random intercepts are estimated across all families in the sample.
-
-In support of hypotheses, we found significant negative correlations between maternal and paternal anger, and emotional adjustment, as well as positive correlations between maternal and paternal consistency, and children's emotional adjustment.
+Results provided mixed support for our hypotheses: Contrary to hypotheses, we found no significant correlations between maternal and paternal warmth and children's emotional adjustment.
+In support of hypotheses, we did find significant positive correlations between maternal and paternal consistency, and children's emotional adjustment.
 Contrary to hypotheses, we found significant negative correlations between maternal and paternal inductive reasoning, and children's emotional adjustment.
 This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment.
+Finally, as hypothesized, we found significant between-family level negative correlations between maternal and paternal anger, and emotional adjustment.
+
+```{r tabbetweencor, results = "asis"}
+cor_tables <- lapply(results, corTable)
+
+table_between_cor <- lapply(cor_tables, `[[`, "Within.Std.Averaged.Over.Clusters")
+table_between_cor <- order_by_name(table_between_cor)
+table_between_cor <- do.call(cbind, table_between_cor)
+table_between_cor <- cbind(paste0(1:3, ". ", rename_pars(rownames(table_between_cor))), table_between_cor)
+table_between_cor <- as.matrix(table_between_cor)
+colnames(table_between_cor) <- c("Parameter", rep(paste0(1:3, "."), 4))
+rownames(table_between_cor) <- NULL
+table_between_cor <- gsub("\\*+", "*", table_between_cor)
+table_between_cor <- table_between_cor[, -seq(2, 11, by = 3)]
+write.csv(table_between_cor, "table_between_cor.csv", row.names = FALSE)
+apa_table(table_between_cor, caption = 
+            "Between-families level correlations", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.", col_spanners = colspan_fun(table_between_cor, skip = 1)
+          )
+
+```
 
@@
 
-Finally, contrary to hypotheses, we found no significant correlations between maternal and paternal warmth, and children's emotional adjustment, at the between-family level.
-
 ### Within-family correlations
 
-At the within-family level, we examined correlations between the residuals of qualitative parenting behaviors and emotional adjustment.
+We also examined correlations between qualitative parenting behaviors and emotional adjustment at the within-family level (Table \@ref(tab:tabwithincor)).
 Within-family correlations are based on residuals of the study variables, across all waves and families.
-Contrary to our expectations, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.
-In partial support of hypotheses, we did find significant positive associations of adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
 In support of hypotheses, we found significant positive within-wave associations between emotional adjustment and fathers' and mothers' warmth.
 Notably, these associations were nearly twice as strong for fathers' warmth, as compared to mothers' warmth.
+In partial support of hypotheses, we did find significant positive associations of adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
+Finally, contrary to our expectations, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.
+
+```{r tabwithincor, results = "asis"}
+table_within_cor <- lapply(cor_tables, function(x){x[["Between"]][3:5, 3:5]})
+table_within_cor <- order_by_name(table_within_cor)
+table_within_cor <- do.call(cbind, table_within_cor)
+table_within_cor <- cbind(paste0(1:3, ". ", rename_pars(rownames(table_within_cor))), table_within_cor)
+table_within_cor <- as.matrix(table_within_cor)
+colnames(table_within_cor) <- c("Parameter", rep(paste0(1:3, "."), 4))
+rownames(table_within_cor) <- NULL
+table_within_cor <- gsub("\\*+", "*", table_within_cor)
+table_within_cor <- table_within_cor[, -seq(2, 11, by = 3)]
+write.csv(table_within_cor, "table_within_cor.csv", row.names = FALSE)
+apa_table(table_within_cor, caption = 
+             "Within-family residual correlations", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.", col_spanners = colspan_fun(table_within_cor, skip = 1)
+          )
+```
 
@@
 
-None of the intercepts of within-families parenting effects differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment.
+None of the intercepts of within-families parenting effects differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment (Table \@ref(tab:tabparenteff)).
 Thus, on average across families, mothers' warmth had a positive significant predictive effect on children's emotional adjustment.
@@
 
+```{r tabparenteff, results = "asis"}
+tabparenteff <- rbind(tab_std_sections$`Intercepts of parenting effects`, 
+                      tab_std_sections$`Between-family level residual variances`[c(1, 5:6),])
+
+tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
+tabparenteff$Parameter[c(1,4)] <- c("Random intercepts of parenting effects", "Variance of parenting effects")
+reg_table(tabparenteff, caption = "Parenting effects on emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+```
+
 #### Child effects
 
-Contrary to hypotheses, the intercepts of the effects of emotional adjustment on maternal qualitative parenting behaviors were significant and positive, not negative, for  anger, non-significant for consistency, and positive for inductive reasoning.
-In line with hypotheses, the intercept for maternal warmth was positive and significant.
+In line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabchildeff)).
 Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year.
+Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant.
+As hypothesized, the intercept for maternal inductive reasoning was positive.
+Finally, counter to hypotheses, the intercept for maternal anger was positive, not negative.
+
 In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant.
 None of the other intercepts for child effects on paternal parenting were significant.
+
 The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' childrearing behaviors differs between families.
 
-## Associations between quantitative involvement and adjustment and parenting
+```{r tabchildeff, results = "asis"}
+tabchildeff <- rbind(tab_std_sections$`Intercepts of child effects`,
+                      tab_std_sections$`Between-family level residual variances`[c(1, 7:8),])
 
-We examined regression coefficients from total join involvement and relative involvement to the random intercepts of emotional adjustment and qualitative parenting behaviors.
-We found that greater joint care was negatively associated with emotional adjustment, counter to our hypothesis.
+tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
+tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects", "Variance of child effects")
+reg_table(tabchildeff, caption = "Child effects on qualitative parenting", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+```
+
+## Associations between quantitative involvement and parenting and adjustment
+
+We examined regression coefficients from total joint involvement and relative involvement to the random intercepts of emotional adjustment and qualitative parenting behaviors (Table \@ref(tab:tabinvolvement)).
+With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning.
+Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers.
+We did not find the hypothesized negative effects of joint total involvement on parental anger.
+
+Counter to our hypothesis, we found that greater joint care was negatively associated with emotional adjustment.
 Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
@@
 
-With regard to parenting, we did not find the hypothesized negative effects of joint total involvement on parental anger, but, in partial support of our hypotheses, we did find that, for both parents, joint involvement positively predicted greater consistency, inductive reasoning, and warmth.
-Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers.
+With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict greater warmth, consistency, and inductive reasoning, and reduced anger, only for the more involved parent.
+For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers.
+For fathers, we found that, in line with this hypothesis, fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.
 
-With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict reduced anger, and greater consistency, inductive reasoning, and warmth, only for the more involved parent.
-For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers.
-For fathers' parenting behavior, we found that, in line with this hypothesis, fathers displayed less consistency, inductive reasoning, and warmth when mothers were relatively more involved than fathers.
+```{r tabinvolvement, results = "asis"}
+tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
+tabinv$Parameter[c(1,4)] <- c("Effects of joint quantitative involvement on:",
+                              "Effects of relative quantitative involvement on:")
+tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
+reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+```
 
@@
 
-To examine interactions between involvement quality and quantity, we tested whether between-family differences in relative and joint total quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment.
+To examine interactions between involvement quality and quantity, we tested whether between-family differences in quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvementmod)).
+
+```{r tabinvolvementmod, results = "asis"}
+tabinvmod <- rbind(
+  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
+  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
+tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
+tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
+                              "Moderating effects of relative involvement on parenting effects of:",
+                              "Moderating effects of joint involvement on child effects on:",
+                              "Moderating effects of relative involvement on child effects on:")
+reg_table(tabinvmod, caption = "Moderating effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+```
 
@@
 
-With regard to the moderation of parenting effects by quantitative involvement, we found no effects for anger, consistency, inductive reasoning, and paternal warmth.
-The only significant effects we found were for maternal warmth.
+With regard to the moderation of parenting effects by quantitative involvement, the only significant effects we found were for maternal warmth.
 Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively when both parents' total joint involvement was higher.
 This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement.
+We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.
 
@@
 
-With regard to the moderation of child effects by quantitative involvement, we found no effects for anger and consistency.
-For inductive reasoning, we found that adjustment was more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
+With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers.
+We found no effects for consistency.
+For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when both parents' joint involvement was higher.
+In other words, when both parents were  more involved, the cross-lagged effect of children's adjustment on maternal discipline was more negative.
+Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
 Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning.
-
-We further found that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when both parents' joint involvement was higher.
-In other words, the more both parents are more involved, the more negative the cross-lagged effect of children's adjustment on maternal discipline will be.
-
-For warmth, we found that the effect of adjustment on maternal warmth was negatively attenuated, both when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers.
+We found no moderating effects for anger.
 
@@
 
-To explore sex differences, we included child sex as a predictor of the within-family cross-lagged effects, and of the random intercepts for emotional adjustment and qualitative parenting behaviors.
+To explore sex differences, we included child sex as a predictor of the random intercepts for emotional adjustment and qualitative parenting behaviors, and of the within-family cross-lagged effects.
+
+With regard to sex differences in the random intercepts, we found no sex differences in emotional adjustment (Table \@ref(tab:tabsex)).
+For qualitative parenting behaviors, we found that both mothers and fathers displayed less warmth, and more inductive reasoning and anger, towards sons as compared to daughters.
+No significant differences were found for maternal warmth, and for both parents' consistency.
 
@@
 
-With regard to sex differences in the random intercepts, we found no sex differences in emotional adjustment.
-For qualitative parenting behaviors, we found that both mothers and fathers displayed more anger and inductive reasoning, and fathers displayed less warmth, towards sons as compared to daughters.
-No significant differences were found for both parents' consistency, and for maternal warmth.
+```{r tabsex, results = "asis"}
+tabsex <- rbind(
+  tab_std_sections$`Sex differences in parenting and adjustment`,
+  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
+tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
+tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
+tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
+                              "Sex differences in parenting effects of:",
+                              "Sex differences in parenting effects on:")
+reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+```
 
@@
 In doing so, we incorporated both qualitative and quantitative measures of parental involvement, and examined interactions between these two aspects of involvement, which have historically been examined mostly in isolation.
-To this end, we used innovative methods which enabled us to include between-family differences in quantitative involvement, operationalized as total joint care and the relative childcare distribution, as predictors of within-family cross-lagged effects between qualitative involvement behaviors and emotional adjustment.
+To this end, we used innovative methods to include between-family differences in quantitative involvement, operationalized as total joint care and the relative childcare distribution, as predictors of within-family cross-lagged effects between qualitative involvement behaviors and emotional adjustment.
 
@@
 
-We hypothesized that, at both the within- and between-family levels, anger would be negatively associated, and consistency, inductive reasoning, and warmth would be positively associated with emotional adjustment.
+We hypothesized that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
 Breaking down the results by parenting behavior, we found mixed evidence for these hypothesized patterns.
+
+We found partial, but unequivocal, support for the hypothesized positive associations between warmth and emotional adjustment.
+These links were absent at the between-family level, but abundant at the within-family level.
+Specifically, we found within-family correlations for both parents' warmth; an average prospective effect of maternal warmth on children's emotional adjustment, and child effects of adjustment on both parents' warmth.
+In other words, well-adjusted children evoke more parental warmth.
+This was the only socialization effect that was significant on average, which suggests that it occurs in most families.
+However, it is important to note that *all* within-family cross-lagged effects had significant variance components, highlighting the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.
+
+We also found partial support for the hypothesized positive associations of parental consistency with emotional adjustment.
+These associations occurred primarily at the between-family level, where both maternal and paternal consistency were positively correlated with children's emotional adjustment.
+This means that, in families where parents tend to display high consistency, children tend to be more emotionally well-adjusted.
+In terms of within-family correlations, we found the predicted links for mothers', but not fathers', consistency.
+On average, we found no support for within-family predictive links between consistency and adjustment.
+It is interesting that we found only between-family links and no over-time predictive effects.
+There are several reasons why between-family links might be observed: For instance, these can reflect the accumulated effect of within-family causal effects at an earlier age.
+However, our study included children from the age of six, and - although it is possible that consistency did, on average, affect children's emotional development prior to the preschool years, it seems unlikely that this effect would have dissipated entirely by the time our study commenced.
+Alternatively, associations at the between-family level can reflect stable differences between families that impact both constructs; for example, it has been proposed that the economic hardship associated with low SES diminishes parents' capacity for consistent parenting, while at the same time, undermining children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+The fact that we observed no within-family predictive effects, despite including children from a very young age, might suggest that previously found associations between consistency and emotional adjustment might be less attributable to causal processes than prior research had concluded, based on methods that did not control for stable between-family differences [@brandPerceivedParentalInconsistency1990].
+
+With regard to inductive reasoning, we found mixed support for the hypothesized positive associations with emotional adjustment: Associations were opposite in sign at the between- versus within-family levels.
+At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment.
+This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior, although a constructive one that stimulates children's internalization of rules and expectations [@grusecImpactParentalDiscipline1994].
+Parents' reports of engaging in inductive reasoning might thus be confounded with how often they discipline their children.
+Disciplining children might result in, be precipitated by, or share a common cause with emotional maladjustment.
+Paradoxically, at the within-family level, emotional adjustment was positively correlated with paternal inductive reasoning, and emotional adjustment had a positive prospective effect on maternal inductive reasoning.
+In other words, increased maladjustment goes hand in hand with increased inductive reasoning.
+The fact that between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment might suggest that between-family differences in inductive reasoning tap more strongly into the amount of discipline parents engage in, whereas the within-family fluctuations might more strongly reflect changes in the way parents discipline their children.
 
@@
 This means that, if children demonstrated better emotional adjustment than usual in one year, their mothers tended to display more anger than usual in the next year.
-The literature on expectation-based loss aversion offers a possible explanation for this puzzling finding [@KoszegiModelReferenceDependentPreferences].
+The literature on expectation-based loss aversion offers a possible explanation for this puzzling finding [@koszegiModelReferenceDependentPreferences].
 Note that, on average, emotional adjustment showed no-to-negative stability coefficients in all models.
@@
 
-We found partial but straightforward support for the hypothesized positive associations of parental consistency with emotional adjustment; again primarily at the between-family level, where both maternal and paternal consistency were positively correlated with children's emotional adjustment.
-This means that, in families where parents tend to display high consistency, children tend to be more emotionally well-adjusted.
-In terms of within-family correlations, we found the predicted links for mothers', but not fathers', consistency.
-On average, we found no support for within-family predictive links between consistency and adjustment.
-It is interesting that we found only between-family links and no over-time predictive effects.
-There are several reasons why between-family links might be observed: For instance, these can reflect the accumulated effect of within-family causal effects at an earlier age.
-However, our study included children from the age of six, and - although it is possible that consistency did, on average, affect children's emotional development prior to the preschool years, it seems unlikely that this effect would have dissipated entirely by the time our study commenced.
-Alternatively, associations at the between-family level can reflect stable differences between families that impact both constructs; for example, it has been proposed that the economic hardship associated with low SES diminishes parents' capacity for consistent parenting, while at the same time, undermining children's socio-emotional adjustment [@McLoydImpactEconomicHardship1990].
-The fact that we observed no within-family predictive effects, despite including children from a very young age, might suggest that the "causal" role of consistency in children's emotional adjustment is more limited than prior research has concluded based on methods that did not control for stable between-family differences [REF].
+## Quantitative involvement and qualitative parenting behaviors
 
-With regard to inductive reasoning, we found mixed support for the hypothesized positive associations with emotional adjustment: Associations were opposite in sign at the between- versus within-family levels.
-At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment.
-This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior, although a constructive one that stimulates children's internalization of rules and expectations **(better ref here)**.
-Parents' reports of engaging in inductive reasoning might thus be confounded with how often they discipline their children.
-Disciplining children might result in, be precipitated by, or share a common cause with emotional maladjustment.
-Paradoxically, at the within-family level, emotional adjustment was positively correlated with paternal inductive reasoning, and emotional adjustment had a positive prospective effect on maternal inductive reasoning.
-In other words, increased maladjustment goes hand in hand with increased inductive reasoning.
-The fact that between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment might suggest that between-family differences in inductive reasoning tap more strongly into the amount of discipline parents engage in, whereas the within-family fluctuations might more strongly reflect changes in the way parents discipline their children.
-
-Finally, we found partial, but unequivocal, support for the hypothesized positive associations between warmth and emotional adjustment.
-These links were absent at the between-family level, but abundant at the within-family level.
-Specifically, we found within-family correlations for both parents' warmth; an average prospective effect of maternal warmth on children's emotional adjustment, and child effects of adjustment on both parents' warmth.
-In other words, well-adjusted children evoke more parental warmth.
-The only socialization effect that was significant on average, which suggests that it is applicable to most families, was this predictive effect from maternal warmth to children's emotional adjustment.
-However, it is important to note that *all* within-family cross-lagged effects had significant variance components, highlighting the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.
+Associations between quantitative involvement and qualitative parenting behaviors were in line with the notion that greater involvement is associated with more desireable parenting behaviors [@sabattiniRelationMothersFathers2004].
+When both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning.
+Fathers, in particular, reported substantially greater parental warmth when both parents were more involved.
+When childcare was more unequally distributed, with mothers performing more care tasks than fathers, we found that mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning.
+The fact that mothers displayed more warmth when fathers were relatively less involved might reflect an attempt to compensate for his absence [@shinnFatherAbsenceChildren1978].
+In sum, these results suggest that fathers engage in less desirable qualitative parenting behaviors, in families where childcare is unequally distributed [@sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011].
 
@@
 For total joint care, we found that, counter to our hypothesis, joint care was *negatively* associated with between-family differences in children's emotional adjustment.
-One potential explanation of this finding is suggested by self-determination: Perhaps excessive levels of involvement undermine children fundamental needs for autonomy and a sense of competence, thereby leading to increased emotional maladjustment [@DeciFacilitatingoptimalmotivation2008].
-Research on "helicopter parenting" provides some support for this interpretation, as psychological wellbeing has been found to be lower in children of overly involved parents [@SchiffrinHelpingHoveringEffects2014].
+One potential explanation of this finding is suggested by self-determination: Perhaps excessive levels of involvement undermine children fundamental needs for autonomy and a sense of competence, thereby leading to increased emotional maladjustment [@deciFacilitatingOptimalMotivation2008].
+Research on "helicopter parenting" provides some support for this interpretation, as psychological wellbeing has been found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014].
 It should be noted, however, that this literature is mostly cross-sectional, and focused on adolescents.
@@
 Mothers' warmth predicted children's adjustment more strongly if mothers were relatively more involved than fathers.
-This suggests that having a warm mother is of greater developmental importance with regard to children's emotional adjustment if fathers are less involved with the day-to-day minutiae of childcare.
-Conversely, we also found that mothers' warmth was less strong of a predictor of adjustment development when both parents were more involved.
-This suggests that maternal warmth is less essential when both parents are highly involved.
-Together, these findings are reminiscent of a "buffering" effect, where mothers' warmth can make up for any disadvantages incurred by having a relatively uninvolved father.
+Conversely, mothers' warmth was less strong of a predictor of adjustment development when both parents were more involved.
+Together, these findings suggest that having a warm mother is of greater developmental importance with regard to children's emotional adjustment if fathers are less involved with the day-to-day minutiae of childcare, and is less essential when both parents are highly involved.
+This suggests a compensatory effect, whereby mothers' warmth can make up for any disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].
 
@@
 For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can be expressed in one-off interactions, such as when disciplining a child (inductive reasoning), or during play time (warmth).
-Moreover, the fact that the association between inductive reasoning and emotional adjustment was stronger for fathers than mothers is in line with the classic view of the father as disciplinarian [REF].
+Moreover, the fact that the association between inductive reasoning and emotional adjustment was stronger for fathers than mothers is in line with the classic view of the father as disciplinarian [@ferrariImpactCultureChild2002].
 
 With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers.
-This is in line with prior research, which suggested that warmth is an important socialization mechanism for mothers [REF mezelf].
+This is in line with prior research, which suggested that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].
 
@@
 The present study casts doubt on this interpretation, as we found that child effects outweighed parenting effects, even at a young age.
-Although still preliminary, the prominence of child effects across two large longitudinal studies, conducted on samples from different continents and at different age ranges, invites the question: Are parents shaping children's development, or are they responding to it?
+Although still preliminary, the prominence of child effects across two large longitudinal studies, conducted on samples from different continents and at different age ranges, invites the question: Are parents really shaping children's development, or are they, instead, responding to it?
 An extreme interpretation of the findings from our study and this preceding work might be that parents' influence is limited to "setting the stage" for children's development, through their stable differences from other parents, and from there on, are merely adjusting to children's developmental needs.
@@
 In other words, parents relax their verbal discipline more in response to positive emotional adjustment when children are male, than when they are female.
-When viewed in the context of the fact that parents are generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.
+When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.
 
@@
 A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment.
-Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are known to be poor judges of others' emotional states [REF Ickes].
+Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are known to be poor judges of others' emotional states [@ickesEmpathicAccuracy1997].
 Consequently, it is important to consider the fact that our effect sizes might be attenuated by noisy measures of emotional adjustment.
@@
 
-We set out to assess whether the quantity of parental care interacted with within-family links between qualitative parenting behaviors and children's emotional adjustment.
 Our findings suggest that, in families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development.
 In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment.
-This suggests that, when differential involvement was higher, mothers more strongly adopted the classic role of a warm "attachment figure", and fathers that of an authority figure.
+This suggests that, when differential involvement was higher, parents adopted more differentiated roles: Mothers' role more strongly resembled that of the classic warm "attachment figure", and fathers that of an authority figure.
 When both parents were more involved, mothers' warmth was less strong of a predictor for emotional adjustment, and mothers' inductive reasoning also became more contingent on adjustment, suggesting that mothers in such families played more similar roles to fathers.
@@
 Children's emotional adjustment, too, was lower when mothers were more involved than fathers.
-These findings suggest that all family members might benefit from a more equal distribution of childcare tasks.
+These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.
 
@@
 Evidence for fathers' unique role was relatively limited; significant within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role in the capacity of disciplinarian.
-The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - fathers' role is more context dependent.
+The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - the role of fathers is more context dependent than that of mothers.
 

### 2019-05-01 LSAC.scales.list.R

@@ -1,0 +1,41 @@
+scales.list <-
+  list(
+    emo1 = grep("dse03t3\\w", names(kdata), value = TRUE),
+    emo2 = grep("ese03t3\\w", names(kdata), value = TRUE),
+    emo3 = grep("fse03t3\\w", names(kdata), value = TRUE),
+    emo4 = grep("gse03t3\\w", names(kdata), value = TRUE),
+    Mwa1 = grep("dpa03m\\d", names(kdata), value = TRUE),
+    Mwa2 = grep("epa03m\\d", names(kdata), value = TRUE),
+    Mwa3 = grep("fpa03m\\d", names(kdata), value = TRUE),
+    Mwa4 = grep("gpa03m\\d", names(kdata), value = TRUE),
+    Fwa1 = grep("dpa03f\\d", names(kdata), value = TRUE),
+    Fwa2 = grep("epa03f\\d", names(kdata), value = TRUE),
+    Fwa3 = grep("fpa03f\\d", names(kdata), value = TRUE),
+    Fwa4 = grep("gpa03f\\d", names(kdata), value = TRUE),
+    Min1 = grep("dpa10m\\d", names(kdata), value = TRUE),
+    Fin1 = grep("dpa10f\\d", names(kdata), value = TRUE),
+    Mco1 = grep("dpa11m\\d", names(kdata), value = TRUE),
+    Mco2 = grep("epa11m\\d", names(kdata), value = TRUE),
+    Mco3 = grep("fpa11m\\d", names(kdata), value = TRUE),
+    Mco4 = grep("gpa11m\\d", names(kdata), value = TRUE),
+    Fco1 = grep("dpa11f\\d", names(kdata), value = TRUE),
+    Fco2 = grep("epa11f\\d", names(kdata), value = TRUE),
+    Fco3 = grep("fpa11f\\d", names(kdata), value = TRUE),
+    Fco4 = grep("gpa11f\\d", names(kdata), value = TRUE),
+    Man1 = grep("dpa13m\\d", names(kdata), value = TRUE),
+    Man2 = grep("epa13m\\d", names(kdata), value = TRUE),
+    Man3 = grep("fpa13m\\d", names(kdata), value = TRUE),
+    Man4 = grep("gpa13m\\d", names(kdata), value = TRUE),
+    Fan1 = grep("dpa13f\\d", names(kdata), value = TRUE),
+    Fan2 = grep("epa13f\\d", names(kdata), value = TRUE),
[…]

### 2019-05-01 invariance/table_coefficients.R

@@ -1,0 +1,113 @@
+library(MplusAutomation)
+library(motley)
+source("rfunctions.R")
+if(!file.exists("results.RData")){
+  results <- readModels("c:/git_repositories/involvement_quant_qual/Mplus")
+  save(results, file = "results.RData")
+} else {
+  load("results.RData")
+}
+
+order_rows <- c("Between-family level correlations",
+                "MOTH.WITH.EMO", "FATH.WITH.EMO", "MOTH.WITH.FATH",
+                "Within-family level correlations",
+                "EMO.WITH.MOTH.Within", "FATH.WITH.EMO.Within", "FATH.WITH.MOTH.Within",
+                "Intercepts of parenting effects",
+                "Intercepts.SEMOMOTH", "Intercepts.SEMOFATH",
+                "Intercepts of child effects",
+                "Intercepts.SMOTHEMO", "Intercepts.SFATHEMO",
+                "Effects of quantitative involvement on parenting and adjustment",
+                "MOTH.ON.MEAN", "FATH.ON.MEAN", "EMO.ON.MEAN", 
+                "MOTH.ON.MMORE", "FATH.ON.MMORE", "EMO.ON.MMORE", 
+                "Moderating effect of quantitative involvement on parenting effects",
+                "SEMOMOTH.ON.MEAN", "SEMOFATH.ON.MEAN",
+                "SEMOMOTH.ON.MMORE", "SEMOFATH.ON.MMORE",
+                "Moderating effect of quantitative involvement on child effects",
+                "SMOTHEMO.ON.MEAN", "SFATHEMO.ON.MEAN",
+                "SMOTHEMO.ON.MMORE", "SFATHEMO.ON.MMORE",
+                "Sex differences in parenting and adjustment",
+                "MOTH.ON.MALE", "FATH.ON.MALE", "EMO.ON.MALE", 
+                "Sex differences in cross-lagged effects",
+                "SEMOMOTH.ON.MALE", "SEMOFATH.ON.MALE",
+                "SMOTHEMO.ON.MALE", "SFATHEMO.ON.MALE",
+                "Within-family level residual variances",
+                "Residual.Variances.MOTH.Within", "Residual.Variances.FATH.Within", "Residual.Variances.EMO.Within",
+                "Between-family level residual variances",
+                "Residual.Variances.MOTH", "Residual.Variances.FATH", "Residual.Variances.EMO", 
[…]

### 2019-05-01 rfunctions.R

@@ -1,0 +1,53 @@
+rename_pars <- function(x){
+  new_labels <- x
+  new_labels <- gsub("\\.Within", "", new_labels)
+  new_labels <- gsub("S(EMO|MOTH|FATH)(EMO|MOTH|FATH)", "(\\1 on \\2)", new_labels)
+  new_labels <- gsub("MEAN", "Joint", new_labels)
+  new_labels <- gsub("MMORE", "Relative", new_labels)
+  new_labels <- gsub("^(Residual\\.Variances|Residual|Variances|Intercepts|Means)\\.", "", new_labels)
+  new_labels <- gsub("FATH", "Father", new_labels)
+  new_labels <- gsub("MOTH", "Mother", new_labels)
+  new_labels <- gsub("EMO", "Emo.", new_labels)
+  new_labels <- gsub("MALE", "sex", new_labels)
+  new_labels <- gsub("\\.ON\\.", " on ", new_labels)
+  new_labels <- gsub("\\.WITH\\.", " with ", new_labels)
+  new_labels
+}
+
+order_by_name <- function(x){
+  x[sapply(c("warmth", "consistency", "reason", "anger"), grep, names(x))]
+}
+
+colspan_fun <- function(x, skip = 0){
+  
+  num_cols <- ncol(x)-skip
+  out <- mapply(
+    function(starts, ends) {
+      c(starts, ends)
+    },
+    starts = seq(
+      from = 1,
+      to = (num_cols - (num_cols / 4) + 1),
+      length.out = 4
+    ) + skip,
+    ends = seq(
+      from = (num_cols / 4),
+      to = num_cols,
+      length.out = 4
+    ) + skip
+  , SIMPLIFY = FALSE)
+  
+  
+  names(out) <- c("Warmth", "Consistency", "Reason", "Anger")
+  out
+}
+
+reg_table <- function(x, ...){
+  rownames(x) <- NULL
+  names(x) <- gsub("^.+est_sig$", "Est.", names(x))
+  names(x) <- gsub("^.+confint$", "95\\% HPDI", names(x))
+  tab <- capture.output(apa_table(x, format = "latex", 
+                                  col_spanners = colspan_fun(x, skip = 1), ...))
+  tab <- gsub("^(.+?) & NA & NA & NA & NA & NA & NA & NA & NA\\\\\\\\$", "\\\\multicolumn\\{9\\}\\{l\\}\\{\\1\\}\\\\\\\\", tab)
+  cat(tab, sep = "\n")
+}

### 2019-10-29 Paper_shortened.Rmd
DP submission

---
title             : "Quantitative involvement moderates longitudinal associations between mothers' and fathers' qualitative parenting and children's emotional adjustment"
shorttitle        : "INVOLVEMENT AND EMOTIONAL ADJUSTMENT"

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Renske Keizer"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Erasmus University Rotterdam Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"

abstract: |
  The question whether fathers and mothers play similar roles in children’s
  emotional adjustment has been approached differently in different research
  traditions: Developmental psychologists have focused on qualitative parenting,
  and sociologists have focused on quantitative child care involvement. This
  longitudinal study (N = 1884, age 6-12) bridges these approaches, 
  investigating qualitative and quantitative involvement, and their interaction,
  in relation to emotional adjustment. On average, evidence for mothers'
  and fathers' unique roles was limited.
  When both parents were more involved, both engaged
  more in desirable qualitative parenting. When mothers were disproportionally
  involved, however, parenting roles were more differentiated:
  Fathers displayed less
  desirable qualitative parenting, and children were less well-adjusted.
  This suggests that all family members benefit from a more equal
  childcare distribution.

keywords          : "involvement; emotional adjustment; fathers; parenting; longitudinal"
wordcount         : "7361"

bibliography      : ["DSEM_2017.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
classoption       : "noextraspace"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
library(motley)
source("rfunctions.R")
load("results.RData")
load("table_coefficients.RData")
mask <- TRUE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

\setlength{\parskip}{0em}

\vspace{\baselineskip}
Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Emotional adjustment development is a domain in which fathers might play a unique role, because fathers encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000]. The attachment relationships between children and parents are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities.

## Fatherhood theory

The past decades have seen major strides in the consolidation of theories to guide fatherhood research. Inspired by increasing divorce rates and maternal employment, early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Over time, the notion of the father as a coparent gained ground [@pleckFatherhoodIdealsUnited1997], according to which fathers should share childrearing responsibilities with mothers, adopting roles historically fulfilled by mothers. To keep pace with changing fatherhood norms, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].

Although it is now widely acknowledged that involvement quality and quantity are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together. Historically, developmental psychology has emphasized how qualitative parenting behaviors shape child development, whereas sociology and gender studies have focused on involvement quantity and its implications for maternal well-being and gender equality [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

## Qualitative parenting and emotional adjustment

### Parenting effects

Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2018a]. Qualitative parenting behaviors have been operationalized in terms of the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983]. Substantial evidence indicates that responsiveness, which involves warm and supportive parenting, is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior. Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. The present study investigates two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Inductive reasoning is thought to promote the internalization of parental expectations, including display rules for emotions, and might thus play a role in emotional adjustment development as well [@grusecImpactParentalDiscipline1994].

Negative parenting is an additional qualitative parenting dimension that appears to be negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. We operationalized negative parenting as parental anger. Parental anger is thought to model anger in children, and undermine their emotion regulation [@scaramellaParentingChildrenDistress2008].

Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010]. Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers. For various reasons, mothers and fathers sometimes appear to fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011]. In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage, and respond more supportively to, emotional expression, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for adolescents' emotional adjustment, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is some evidence that emotional adjustment is more negatively impacted by negative parenting from mothers, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship. Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.

### Child effects

A view of the family as a dynamic system, whose members reciprocally influence one another, has become commonplace [@kerrPerceivedParentingStyle2012]. Nevertheless, bi-directional effects are not routinely investigated in the childhood literature [@flouriFathersBehaviorsChildren2010]. The available evidence suggests that maladjustment elicits less parental warmth, more demandingness, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @lenguaBidirectionalAssociationsTemperament2005]. 
The adolescent literature provides additional support that emotional maladjustment predicts decreased parental support, and greater control [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014; @wijsbroekDirectionEffectsPerceived2011]. Thus, parents' display punitive responses to maladjustment, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness. We might therefore expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.

### Direction of effects

Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if stable between-family differences are not controlled [@hamakerCritiqueCrosslaggedPanel2015]. Innovative methods provide better estimates of dynamic reciprocal processes within families, by controlling for stable differences: The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].

Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples. One such study found that effects of parenting on adjustment that emerged using classical methods, disappeared when stable between-family differences in parenting and adjustment were included in the model [@keijsersParentalMonitoringAdolescent2016]. Another study similarly found that most associations between maternal and paternal qualitative parenting and adolescents' emotional adjustment, were explained by between-family level correlations. The remaining predictive effects consisted primarily of child effects, not parenting effects [@vanlissaRoleFathersMothers2018a]. The authors theorized that this finding was reflective of the relative equality of parent-child relationships in adolescence. If this is correct, we would expect to find more evidence of parenting effects in the present, younger, sample.

## Quantitative involvement

### Operationalizing quantitative involvement

Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the **division** of childcare tasks is more relevant for children's adjustment [@krollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003]. Indeed, the equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001]. Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006]. This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality.

A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001]. Examining the childcare distribution instead, and controlling for total care, mitigates this confound. Some prior research has attempted to measure the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016]. When quantitative involvement has been measured in terms of absolute involvement, however, a "latent mean and difference" approach can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].

## Main effects of quantitative involvement

### Effects on qualitative involvement

The amount of time parents spend with children is associated with the quality of parenting behaviors  [@sabattiniRelationMothersFathers2004]. Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a]. Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014]. Consequently, we might expect greater total joint care to positively predict both parents' parenting quality.

With regard to differential involvement, we might also expect the more involved parent to display more desirable qualitative parenting behaviors, due to the effects of experience [@sabattiniRelationMothersFathers2004]. However, the benefit of parents monitoring each other's childrearing should be diminished when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Indeed, retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. Additionally, greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child wellbeing [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008]. We might thus expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.

### Effects on emotional adjustment

Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. This suggests that the total amount of care provided should be positively associated with children's emotional adjustment. Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their wellbeing [@deutschPaternalParticipationChild2001].

## Interactions between involvement quality and quantity

Interactions between involvement quality and quantity have largely been overlooked in theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@jaffeeLifeFatherBenefits2003]. In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the role of parent. Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families. According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of differential involvement, affects qualitative childrearing behaviors and children's adjustment directly, and moderates the link between these two constructs.

There is some empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012]. In adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014]. Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.

## Sex differences

Sex differences are a relevant factor in the interplay between parenting and emotional adjustment. Parents tend to raise boys and girls differently; expressing more control with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more negative parenting with boys [@strausCorporalPunishmentAmerican1999]. Furthermore, social learning theory posits that the same-sex parent plays a stronger role in socialization; particularly for sex-typed behavior, such as emotion display rules [@busseySocialCognitiveTheory1999]. Finally, mounting evidence suggests that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a]. We might thus expect parents to display more responsiveness, structure, and inductive reasoning with girls, and more negative parenting with boys. Furthermore, we might expect associations between qualitative parenting and emotional adjustment to be stronger between children and the same-sex parent.

Although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a]. We consequently did not expect to find sex differences in children's average levels of emotional adjustment.

## The present study

The present study aims to make several contributions. First, we include both mothers and fathers, in order to tease apart their common and unique roles in children's emotional adjustment development. Second, we incorporate both qualitative and quantitative measures of involvement. Third, we disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative involvement and emotional adjustment. Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.

# Methods

## Participants

We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.

Our target sample consisted of 1884 native `r masked("Australian")`, non-minority, intact, biological families. At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$). Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$; 669 father-only, 41 mother-only). The modal annual income for fathers was between \$26,000-\$51,999 (N = 663), and for mothers, \$25,999 or less per year (N = 1152).

## Measures

### Qualitative involvement.

Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001]. Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000]. These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".

### Quantitative involvement.

Measures of quantitative involvement varied from wave to wave, due to children's changing care needs [@wangATUSFathersInvolvement2009]. We used the first wave reports as an indicator of between-family differences in quantitative involvement. Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].

### Emotional adjustment.

Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].

## Strategy of analyses

We calculated mean scores for all of the study variables. Scale descriptive statistics are provided in Supplementary Table 1, and scale intercorrelations are provided in Supplementary Table 2. We deemed it defensible to assume measurement invariance, as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).

To address our research questions, we used multilevel, autoregressive structural equation modelling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015]. Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.

All analyses were conducted using Mplus 8.0 and the R-package MplusAutomation. This type of multilevel autoregressive model can only be estimated using Bayesian methods. We used two Gibbs-sampler chains of 50000 iterations each. We applied mildly informative inverse Wishart priors to the variance/covariance matrices, and normal priors centered on zero for all regression coefficients. We assessed convergence through visual inspections of the chains and posterior distributions. We report the posterior median, and a 95% highest posterior density interval (HPDI). This is the "narrowest" possible interval that contains 95% of samples from the posterior distribution. It represents the most likely population values of the parameter in question. Results are best interpreted by examining the effect size (standardized regression coefficients and correlations), and the width of the HPDI. Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate. All standardized and unstandardized model parameters are reported in Supplementary Table 4.

# Results

## Correlations between qualitative parenting and emotional adjustment

We hypothesized that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment. At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)). Random intercepts reflect the average level of each variable across all waves of the study. Counter to our hypotheses, we found no significant correlations between parental warmth and children's emotional adjustment. In support of hypotheses, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment. Contrary to hypotheses, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment. This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment. Finally, as hypothesized, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.

```{r tabcors, results = "asis"}
cor_tables <- lapply(results, corTable)

table_between_cor <- lapply(cor_tables, function(x){x[["Between"]][3:5, 3:5]})
table_between_cor <- order_by_name(table_between_cor)
table_between_cor <- do.call(cbind, table_between_cor)
table_between_cor <- cbind(rename_pars(rownames(table_between_cor)), table_between_cor)
table_between_cor <- as.matrix(table_between_cor)
colnames(table_between_cor) <- c("Parameter", rep(table_between_cor[,1], 4))
rownames(table_between_cor) <- NULL
table_between_cor <- gsub("\\*+", "*", table_between_cor)
table_between_cor <- gsub("^1$", "", table_between_cor)
table_between_cor <- table_between_cor[-3, -seq(2, 11, by = 3)]
table_between_cor <- rbind(c("Between-family level correlations", rep(NA, ncol(table_between_cor)-1)), table_between_cor)

table_within_cor <- lapply(cor_tables, `[[`, "Within.Std.Averaged.Over.Clusters")
table_within_cor <- order_by_name(table_within_cor)
table_within_cor <- do.call(cbind, table_within_cor)
table_within_cor <- cbind(rename_pars(rownames(table_within_cor)), table_within_cor)
table_within_cor <- as.matrix(table_within_cor)
colnames(table_within_cor) <- c("Parameter", rep(table_within_cor[,1], 4))
rownames(table_within_cor) <- NULL
table_within_cor <- gsub("\\*+", "*", table_within_cor)
table_within_cor <- gsub("^1$", "", table_within_cor)
table_within_cor <- table_within_cor[-3, -seq(2, 11, by = 3)]
table_within_cor <- rbind(c("Within-families level correlations", rep(NA, ncol(table_within_cor)-1)), table_within_cor)

table_cors <- rbind(table_between_cor, table_within_cor)
write.csv(table_cors, "table_cors.csv", row.names = FALSE)

reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)). In support of hypotheses, we found significant positive within-family level associations between emotional adjustment and fathers' and mothers' warmth. Notably, these associations were nearly twice as strong for fathers' warmth, as compared to mothers' warmth. In partial support of hypotheses, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning. Finally, contrary to our expectations, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.

## Within-family predictive effects

At the within-family level, we examined over-time cross-lagged effects between qualitative parenting and emotional adjustment. The intercepts of these random effects reflect the average cross-lagged effect across families. Their variances indicate whether the effect differs in strength (and sign) between families.

With regard to within-families parenting effects, none of the intercepts differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment (Table \@ref(tab:tabcrosspath)). Thus, on average, mothers' warmth had a positive significant predictive effect on children's emotional adjustment. The variances of all parenting effects were significant, however, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families. Moreover, these variances were notably larger for fathers than for mothers, indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.

```{r tabcrosspath, results = "asis"}
tabparenteff <- rbind(tab_std_sections$`Intercepts of parenting effects`, 
                      tab_std_sections$`Between-family level residual variances`[c(1, 5:6),])

tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
tabparenteff$Parameter[c(1,4)] <- c("Random intercepts of parenting effects of:", "Variance of parenting effects of:")

tabchildeff <- rbind(tab_std_sections$`Intercepts of child effects`,
                      tab_std_sections$`Between-family level residual variances`[c(1, 7:8),])

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")

reg_table(rbind(tabparenteff, tabchildeff), caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

With regard to child effects, in line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)). Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year. Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant. As hypothesized, the intercept for maternal inductive reasoning was positive. Finally, counter to hypotheses, the intercept for maternal anger was positive, not negative. In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant. None of the other intercepts for child effects on paternal parenting were significant. The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' childrearing behaviors differs between families.

## Effects of quantitative involvement on parenting and adjustment

We examined regression coefficients from total joint involvement and relative involvement to the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning. Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint total involvement on parental anger. Counter to our hypothesis, we found that greater total joint involvement was negatively associated with emotional adjustment. Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.

With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict greater warmth, consistency, and inductive reasoning, and reduced anger, only for the more involved parent. For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers. For fathers, we found that, in line with this hypothesis, fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers. Finally, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
reg_table(rbind(tabinv, tabinvmod), caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

## Interactions between involvement quality and quantity

To examine interactions between involvement quality and quantity, we tested whether between-family differences in quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth. Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when both parents' total joint involvement was higher. This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers. We found no effects for consistency. For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when both parents' joint involvement was higher. In other words, when both parents were  more involved, the cross-lagged effect of children's adjustment on maternal discipline was more negative. Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

## Effects of child sex

We found sex differences in the random intercepts of qualitative parenting behaviors: Both mothers and fathers displayed less warmth, and more inductive reasoning and anger, towards sons as compared to daughters  (Table \@ref(tab:tabsex)). No significant differences were found for maternal warmth, and for both parents' consistency. We found no sex differences in the random intercept of emotional adjustment. The only cross-lagged effects significantly moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning. Specifically, emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls. In other words, if children display better adjustment than usual in one year, their parents decrease inductive reasoning more strongly in the next year when children are male.

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

# Discussion

## Associations between qualitative parenting and emotional adjustment

There was partial but unequivocal support for the hypothesized positive associations between warmth and emotional adjustment. These links were absent at the between-family level, but abundant at the within-family level - for instance, in the form of within-family correlations with both parents' warmth. The random intercept for the parenting effect of maternal warmth was significant, suggesting that this parenting effect occurs in most families. Finally, child effects on both parents' warmth indicated that well-adjusted children evoke greater warmth.

We found partial support for positive associations of parental consistency with emotional adjustment, primarily at the between-family level: children in families with highly consistent parents tended to be more emotionally well-adjusted. We found the predicted within-family correlations for mothers', but not fathers', consistency. On average, we found no support for within-family predictive links between consistency and adjustment. It is interesting that we found only between-family links and no predictive effects for consistency: Between-family level correlations can reflect the consolidated influence of predictive effects at an earlier age, or alternatively, the influence of third variables that impact both constructs. Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990]. The fact that we observed no within-family predictive effects, despite including children from a very young age, casts doubt on the former explanation, and on causal interpretations of prior research that did not control for stable between-family differences [@brandPerceivedParentalInconsistency1990].

With regard to inductive reasoning, we found mixed support for the hypothesized positive associations with emotional adjustment: Associations were opposite in sign at the between- versus within-family levels. At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment. This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior - albeit a constructive one [@grusecImpactParentalDiscipline1994]. Parents' reports of engaging in inductive reasoning might thus be confounded with how often they discipline their children. At the within-family level, by contrast, paternal inductive reasoning was positively correlated with emotional adjustment, and emotional adjustment had a positive predictive effect on maternal inductive reasoning. The fact that between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment might suggest that between-family differences in inductive reasoning tap more strongly into the amount of discipline parents engage in, whereas within-family fluctuations reflect changes in the way parents discipline their children.

With regard to parental anger, we found mixed evidence for the predicted negative associations with emotional adjustment. These associations existed at the between-family level only. This means that, in families where mothers or fathers display low anger, children tend to be more emotionally well-adjusted. Counter to our expectations, however, we found no evidence for within-family correlations or parenting effects. This suggests that the brunt of negative associations between anger and adjustment are explained by differences between families. One potential explanation for the absence of within-family associations might be a ceiling effect; it might be the case that, once the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger no longer make much of a difference. Another explanation might be that any further effects of fluctuations in anger play out on a different time scale than the one captured by our study; for example, if such effects primarily play out immediately, in dyadic interactions between parents and children. Finally, we found that, on average, emotional adjustment showed a paradoxical positive predictive effect on maternal anger. This means that, if children demonstrated better emotional adjustment than usual in one year, their mothers tended to display more anger than usual in the next year. A possible explanation might be regression to the mean: On average, emotional adjustment showed no-to-negative stability coefficients in all models. That means that better emotional adjustment in one year, was likely to be followed worse adjustment in the next year. The literature on expectation-based loss aversion suggests that mothers might be more angered by a child who unexpectedly displays poor adjustment, than they would be appeased by a child who displays unexpected positive adjustment [@koszegiModelReferenceDependentPreferences].

In sum, we found mixed evidence for our hypotheses that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment. Importantly, all within-family predictive effects had significant variance components, highlighting the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.

## Quantitative involvement and qualitative parenting behaviors

When both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning. Fathers, in particular, reported substantially greater parental warmth when both parents were more involved. We found that, when mothers performed more care tasks than fathers, mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning. This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978]. These results are in line with our hypothesis that greater involvement would be associated with more desireable parenting behaviors [@sabattiniRelationMothersFathers2004]. Additionally, they suggest that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [@craigHowMothersFathers2011].

## Quantitative involvement and emotional adjustment

Our operationalization of quantitative involvement, in terms of total joint care and the childcare distribution, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. However, counter to our hypothesis, joint care was *negatively* associated with between-family differences in emotional adjustment. One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014]. Research on "helicopter parenting" supports this interpretation, as psychological wellbeing was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were more involved than fathers - even after controlling for the total amount of care provided. Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001]. These correlational findings at the between-family level should not be interpreted causally, however. Future research should examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their wellbeing, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.

## Interactions between quantitative and qualitative involvement

We found limited evidence that quantitative involvement moderated within-family parenting effects. One such effect is that the effect of maternal warmth on adjustment was contingent on both total joint care and the childcare distribution. Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers. These findings suggests a compensatory effect, whereby mothers' warmth can make up for disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].

The evidence that quantitative moderates child effects was somewhat stronger. Firstly, when total joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth. These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth. In families where mothers' involvement was high - either in terms of joint, or relative involvement - this positive effect was attenuated somewhat. Secondly, total joint involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning. This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved. When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.

## Differences between fathers and mothers

Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development. These differences all played out at the within-family level, however. Stable between-family differences in either parent's qualitative parenting were all similarly associated with children's adjustment.

At the within-family level we found stronger associations with emotional adjustment for maternal consistency, and for paternal reasoning and warmth. As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample. This might help explain the stronger associations for maternal consistency - a behavior which implies, by definition, consistent repeated interactions with the child. For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can also be expressed in one-off interactions, such as when  disciplining a child (inductive reasoning), or during play time (warmth). The stronger association between inductive reasoning and emotional adjustment for fathers is in line with the classic view of the father as disciplinarian [@flouriFathersBehaviorsChildren2010].

With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers. This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a]. With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting. Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.

## Direction of effects

Historically, developmental theory has focused on how parents cultivate their children's development. Nowadays, it is commonly acknowledged that influences are reciprocal. However, commonly used developmental methods have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015]. One recent study found that, in adolescence, child effects substantially outnumbered parenting effects [@vanlissaRoleFathersMothers2018a]. The authors speculated that this might be due to the greater equality of parent-child relationships in adolescence, compared to earlier childhood. The present study casts doubt on this interpretation, as we found that child effects outweighed parenting effects, even at a young age. The prominence of child effects across two large longitudinal studies, with samples from different continents and at different age ranges, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it? An extreme interpretation of our findings might be that parents "set the stage" for children's development, through their stable differences from other parents, and from there on, are merely responding to children's changing developmental needs. This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations. For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time. Consequently, stable parent characteristics can cause change in children. Future research might examine alternative theoretical conceptualizations of parental influences in child development. Another point to consider is that parent effects on children might play out on a different time scale than was captured by these macro-longitudinal studies. If parents shape their children through specific behaviors in dyadic interactions, then we might fail to capture parenting effects when using annual reflective self-reports.

## Sex differences

In line with prior research, we found that parents raise boys and girls differently. Mothers and fathers both displayed more anger and inductive reasoning with sons than with daughters, and fathers displayed less warmth towards sons. These findings suggest that parents are stricter towards their sons. Parenting effects on children were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors. Child effects were moderated by sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male. When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.

## Strengths and Limitations

Several strengths allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated both quantitative and qualitative involvement. This allowed us to provide a more nuanced overview of the links between parental involvement and emotional adjustment development than prior work, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001]. We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately. Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. As teachers are likely to be "blind" to parents' childrearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.

A potential limitation of the present study is that teachers' reports of emotional adjustment might be more noisy than self- or parent-reports. A teacher is responsible for many pupils, and might not be aware of each pupil's emotional adjustment. Moreover, although the instrument we used assessed *overt* symptoms of emotional maladjustment, people are generally poor judges of others' emotional states [@ickesEmpathicAccuracy1997]. Consequently, our effect sizes might be attenuated by noisy measures of emotional adjustment. Teachers' reports do have clear benefits as well: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies. Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983]. We focused on specific parenting behaviors for two reasons. First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction. This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in childrearing. Moreover, the longitudinal models we used are not suitable for categorical variables, such as parenting style. Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation. Finally, one potential limitation is that we only included quantitative involvement in the first year of the study. This was done for pragmatic reasons: Children's developmental needs change rapidly, especially in the first years of life. It would be impossible to establish measurement invariance for quantitative involvement from age 6 to 12. Our approach assumes that a "snapshot" of quantitative involvement at age 6 taps into parents' general level of involvement. Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact. Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range.

## Conclusions

Our findings suggest that, in families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development. In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment. This suggests that, when differential involvement was higher, parents adopted more differentiated roles: Mothers' role more strongly resembled that of the classic warm "attachment figure", and fathers that of an authority figure. When both parents were more involved, mothers' warmth was less strong of a predictor for emotional adjustment, and mothers' inductive reasoning also became more contingent on adjustment, suggesting that mothers in such families played more similar roles to fathers.

We also found substantial evidence that both parents showed more desirable patterns of qualitative parenting behaviors - behaviors that are positively associated with children's adjustment - when total joint involvement was higher. When mothers were more involved than fathers, however, mothers' expressed more warmth, whereas fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.

With regard to mothers' versus fathers' unique roles in children's emotional development, we found that, in line with the notion of the mother as a warm, supportive attachment figure, maternal warmth and consistency were associated with children's emotional adjustment. Evidence for fathers' unique role was relatively limited; significant within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role in the capacity of disciplinarian. The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - the role of fathers is more context dependent than that of mothers.

\newpage

# References
```{r create_r-references}
#r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\setlength{\parskip}{0em}

<div id = "refs"></div>
\endgroup

### 2019-10-29 Paper_shortened_doc.Rmd
DP submission

--- 2019-10-29 Paper_shortened.Rmd
@@
 Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013].
-The present study investigates two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
+We examined two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
 Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990].
@@
 Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
-Finally, there is some evidence that emotional adjustment is more negatively impacted by negative parenting from mothers, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
-Based on these findings, we might expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.
+Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
+We might therefore expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.
 
@@
 Some prior research has attempted to measure the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016].
-When quantitative involvement has been measured in terms of absolute involvement, however, a "latent mean and difference" approach can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].
+When quantitative involvement has been measured in terms of absolute involvement, latent means and differences can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].
 
@@
 write.csv(table_cors, "table_cors.csv", row.names = FALSE)
-
+table_cors[is.na(table_cors)] <- ""
+rownames(table_cors) <- NULL
+#reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
 reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
@@
 tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
-
-reg_table(rbind(tabparenteff, tabchildeff), caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+tab <- rbind(tabparenteff, tabchildeff)
+tab[is.na(tab)] <- ""
+rownames(tab) <- NULL
+reg_table(tab, caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
 ```
@@
                               "Moderating effects of relative involvement on child effects on:")
-reg_table(rbind(tabinv, tabinvmod), caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+tab <- rbind(tabinv, tabinvmod)
+tab[is.na(tab)] <- ""
+rownames(tab) <- NULL
+reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
 ```
@@
                               "Sex differences in parenting effects on:")
+tabsex[is.na(tabsex)] <- ""
+rownames(tabsex) <- NULL
 reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")

### 2019-10-29 Paper_shortened_pdf.Rmd
DP submission

--- 2019-10-29 Paper_shortened_doc.Rmd
@@
 keywords          : "involvement; emotional adjustment; fathers; parenting; longitudinal"
-wordcount         : "7361"
+wordcount         : "7356"
 
@@
 lineno            : yes
-mask              : no
+mask              : yes
 
@@
 We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012.
+Each wave was reviewed and approved by the
+`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
+Ethics Committee registered with the `r masked("National Health and Medical Research Council")`.
+Documentation of the ethics approval is available online at `r masked("https://growingupinaustralia.gov.au/sites/default/files/certification_of_ethics_approval_for_lsac.pdf")`.
 `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population.
@@
 write.csv(table_cors, "table_cors.csv", row.names = FALSE)
-table_cors[is.na(table_cors)] <- ""
+#table_cors[is.na(table_cors)] <- ""
 rownames(table_cors) <- NULL
@@
 tab <- rbind(tabparenteff, tabchildeff)
-tab[is.na(tab)] <- ""
+#tab[is.na(tab)] <- ""
 rownames(tab) <- NULL
@@
 tab <- rbind(tabinv, tabinvmod)
-tab[is.na(tab)] <- ""
+#tab[is.na(tab)] <- ""
 rownames(tab) <- NULL
@@
                               "Sex differences in parenting effects on:")
-tabsex[is.na(tabsex)] <- ""
+#tabsex[is.na(tabsex)] <- ""
 rownames(tabsex) <- NULL
@@
 
-With regard to parenting effects, the only difference between mothers and fathers we found was for the predictive effect of warmth on emotional adjustment, which was found only for mothers.
+With regard to parenting effects, the only difference between mothers and fathers was for the predictive effect of warmth on emotional adjustment, which was evident only for mothers.
 This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].
@@
 These findings suggest that parents are stricter towards their sons.
-Parenting effects on children were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors.
+Parenting effects were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors.
 Child effects were moderated by sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male.
@@
 
-We also found substantial evidence that both parents showed more desirable patterns of qualitative parenting behaviors - behaviors that are positively associated with children's adjustment - when total joint involvement was higher.
-When mothers were more involved than fathers, however, mothers' expressed more warmth, whereas fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth).
+We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when total joint involvement was higher.
+Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth).
 Children's emotional adjustment, too, was lower when mothers were more involved than fathers.

### 2019-10-29 rfunctions.R

@@ -51,3 +51,20 @@ reg_table <- function(x, ...){
   tab <- gsub("^(.+?) & NA & NA & NA & NA & NA & NA & NA & NA\\\\\\\\$", "\\\\multicolumn\\{9\\}\\{l\\}\\{\\1\\}\\\\\\\\", tab)
   cat(tab, sep = "\n")
 }
+
+
+reg_table_html <- function(x, ...){
+  rownames(x) <- NULL
+  names(x) <- gsub("^.+est_sig$", "Est.", names(x))
+  names(x) <- gsub("^.+confint$", "95\\% HPDI", names(x))
+  multiline_rows <- rowSums(is.na(x)) == 8
+  multiline_text <- x[multiline_rows, 1]
+  x[multiline_rows, 1] <- ""
+  x[is.na(x)] <- ""
+  tab <- capture.output(apa_table(x, format = "html", 
+                                   ...))
+  browser()
+  div_line <- grep("^-[ -]+$", tab)
+  tab[which(multiline_rows)+div_line] <- sprintf(paste0("%-", nchar(tab[which(multiline_rows)[1]+div_line]), "s"), multiline_text)
+  cat(tab, sep = "\n")
+}

### 2019-10-29 Action letter.Rmd
Create revision letter

---
title       : "Quantitative involvement moderates longitudinal associations between mothers' and fathers' qualitative parenting and children's emotional adjustment"
authors     : "Caspar J. van Lissa & Renske Keizer"
journal     : "Developmental Psychology"
manuscript  : "DEV-2019-2824"
bibliography: ["DSEM_2017.bib"]  
class       : "draft"

output      : papaja::revision_letter_pdf
---

```{r, echo = FALSE}
editor = c("Dr.", "Sara", "Jaffee")
journal = "Developmental Psychology"
title = "Quantitative involvement moderates longitudinal associations between mothers' and fathers' qualitative parenting and children's emotional adjustment"
```

Dear Dr. `r editor[3]`,

Thank you for considering our manuscript for publication in `r journal`. We hereby submit a major revision of the manuscript.

We thank you and the Reviewers for your thorough comments, for the helpful summary you've provided. We hope that you will find that our responses adequately address the issues raised, and that the quality of the work has been improved.

<!--The manuscript (including title page, abstract, and references) comprises 35 pages, or 8666 words. Four supplementary tables are included, whose size prohibits them from being included in the manuscript: A full correlation table (total sample and paneled by adolescent sex); the full matrix of parameter estimates; detailed scale descriptives and intercorrelations, and a full matrix of parameter estimates for an alternative analysis suggested by a Reviewer. 
Should the manuscript be submitted, we are prepared to make all syntax and results of the analyses conducted available on a public code repository referenced in the paper, to safeguard reproducibility. 
Treatment of participants was in accordance with the APA ethical principles. On behalf of the authors, I declare that I have no conflicts of interest concerning this article. I will serve as the corresponding author for this manuscript.-->

Yours sincerely,
Caspar van Lissa

thank you very much for taking the time to consider our manuscript for publication at the _Journal XYZ_.
In the following we address your and each reviewer's concerns point-by-point.

### Actual comments

DEV-2019-2824
Quantitative involvement moderates longitudinal associations between mothers’ and fathers’ qualitative parenting and children’s emotional adjustment
Developmental Psychology

Dear Dr. van Lissa,

Thank you very much for submitting your manuscript "Quantitative involvement moderates longitudinal associations between mothers’ and fathers’ qualitative parenting and children’s emotional adjustment" for review and consideration for publication in Developmental Psychology. I sincerely appreciate the opportunity to review the manuscript. I have now received two reviews of the manuscript, and I have read the manuscript carefully myself. The reviews are appended to this email. As you will note, the reviewers agreed that your manuscript pertains to topics of importance—(1) how mothers and fathers influence their children's emotional adjustment and (2) the interplay between how much time and how parents spend caring for children. The strengths of the study include the sophisticated methodological approach, the inclusion of mothers and fathers, and the effort to include quantitative and qualitative measures of parenting. However, at the same time, the reviewers also identified a number of significant conceptual problems with the manuscript. Thus, I cannot accept the manuscript for publication in its present form. However, the reviewers and I agree that these issues could be adequately addressed in a revision. I encourage you therefore to revise and resubmit the manuscript. Although I believe you will find the reviewers' comments clear, it may be useful for me to summarize the major points made in the reviews.

The reviewers were very consistent in their comments about the manuscript and I agree with them. One of the biggest concerns they raised was about the (lack of) clarity of the writing. Thus, my comments are organized to reflect my recommendations for improving clarity in each section of the manuscript.
1. Introduction
1a. Both reviewers (and I) think you are addressing an important set of questions, but the organization of the Introduction is very difficult to follow. I think it's fine to start with a description of emotional adjustment and to explain that the quality of the parent-child relationship is associated with children's emotional adjustment, but then I think you need to revise the text dramatically. Readers want to know (1) what is already known about the association between the parent-child relationship and children's adjustment? (2) what is not known about the association between the parent-child relationship and children's adjustment? (3) how will your study fill that gap in knowledge?
2. Method
2a. Both reviewers found it hard to evaluate the findings without additional information about the sample, including how representative it is, how missing data were handled, and how much attrition there was across waves of the study. It is possible that reviewers and I have misinterpreted a sentence in the Discussion section in which you mention two longitudinal studies across two continents. At least one of the reviewers understood this to mean that your study involves samples from two continents, so please clarify if this is not the case.
2b. I was confused as to how joint and relative quantitative involvement were defined. Is joint quantitative involvement the total amount of time mothers or fathers spend caring for children and relative involvement the ratio of one parent's involvement to both parents' involvement? Please also use consistent language throughout. In the Results section, you describe total and relative involvement, but in the Intro and Discussion, you describe total joint care and the childcare distribution. 
3. Results
3a. Although I suspect a growing number of Developmental Psychology readers are aware of statistical advances on the standard cross-lagged model, many of them probably have not estimated those models. You would help readers understand your findings by presenting and explaining a conceptual figure of the statistical model. That way, readers might be better able to visualize it when you describe various random intercepts or bidirectional effects.
3b. More extensive table notes would also be helpful. These could be used to remind readers of how some of the constructs are defined (e.g., joint and relative quantitative involvement).
3c. You could educate readers by helping them interpret the values in the tables. You do this nicely in some places; for example, on p. 14 where you explain, "Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year." Even there, however, it's not clear what a value of .14 means in the association of quant. involvement on maternal warmth.  
3d. The results section is very dense. You could make it less dense by articulating clear hypotheses at the end of the Introduction and then reporting whether those specific hypotheses were supported. As just one example, you state on p. 9, "Applying this insight from the coparenting literature, we might thus expect parents’ quantitative involvement to be directly associated with their qualitative parenting behaviors and children’s emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment." This is not a directional hypothesis, as stated. It would be helpful if you could be more specific. Are you predicting that the more time mothers and fathers spend caring for children, the more warm, consistent, etc. their parenting will be? Do you expect this to be equally true for mothers and fathers? How do you expect the quantitative measure of parenting to moderate the association between qualitative measures and children's emotional adjustment? 
4. Discussion
4a. The Discussion, as it is currently written, raises a number of issues that are not necessarily anticipated in the Introduction. For example, the Introduction would lead me to think that your analyses would focus primarily on how patterns of association between dimensions of parenting and children's adjustment differ for fathers versus mothers. The first several paragraphs of your discussion, however, focus on how patterns of results differ across parenting dimensions (e.g., results for inconsistent parenting are different from results for parental warmth). I understand that the data show these differences, but the presentation of the results and discussion would be more streamlined if you focused on those that were directly relevant to your hypothesis.  
4b. There are a number of issues you raise in the Discussion that don't have a parallel in the Introduction. Thus, I think readers are not well prepared to understand the significance of those analyses or findings. For example, you make special note of it when results are observed at the within-family, but not the between-family level (e.g., mother vs. father differences) or when they are observed at the between-family, but not within-family level (e.g., some of the parenting dimension-emotional adjustment associations). Because you haven't really explained clearly when you expect to see between-family effects and when you expect to see within-family effects, I think the significance of much of the discussion will be lost on readers. 
4c. I have two comments about your discussion of child effects (p. 23). First, although you note that you are measuring the parent-child dyad early in development, the children in your sample are already 6 years old when they are first assessed and they are 12 by the last measurement occasion. Although there is no question in my mind that child effects are present from the earliest stages in development, I think the likelihood of observing parenting effects is greater at still younger ages (e.g., infancy and toddlerhood). Second, there are pronounced individual differences in how parents react to children. The same level of emotionally dysregulated behavior might elicit anger from one parent and inductive reasoning from another. There is a great deal of evidence from parenting interventions that how parents react to difficult behavior affects whether their children remain noncompliant and dysregulated or whether their behavior improves. So I think you're selling parents short if your claim is that the only way they influence their children is through the changing expression of the genes they have bequeathed them.   
So, we encourage you to submit a revision. To do so, please go to https://www.editorialmanager.com/dvl/ and log in as an Author. You will see a menu item called "Submissions Needing Revision". You will find your submission record there. Please resubmit within 30 days. If you need more time to complete additional analyses, this time period is negotiable, but let us know beforehand. Please include an 'Authors Response to Reviewers' with your revision that includes details on how you addressed the various issues that I highlighted above, as well as the other issues raised by the reviewers. I will return the revision to the original reviewers for additional comments .

Thank you for considering Developmental Psychology as an outlet for your work.

Sincerely,
Sara Jaffee
Associate Editor
Developmental Psychology
EDITOR'S NOTE: Please make sure your revision is concise. Our guidelines for length of manuscripts are 35 pages inclusive of everything except tables and figures. Because your manuscript reports on a longitudinal study, we could allow a few extra pages, but please keep close to this guideline.

NOTE FROM EDITORIAL OFFICE:  We ask that you upload either Word or RTF versions of both your title page and manuscript,  rather than PDFs, when resubmitting.

Reviewers' comments:

Reviewer 1: This manuscript addresses an important research area: understanding the role of mothers and fathers in the emotional adjustment of children.  Strengths of this paper include (a) the inclusion of quantitative and qualitative measures of parenting from both parents, (b) a within- and between-person design, and (c) a longitudinal study with repeated measures of children and parenting. Below, I describe my suggestions to improve the manuscript as currently written.

Introduction

The first paragraph of the introduction introduces many different ideas and comes across as disjointed. Revision should make it clear to readers that the focus is on emotional adjustment among children (as well as the specific developmental age of focus), describe a more complex, well-cited definition of emotional adjustment (the current citation and definition does not seem accurate), introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers. Finally, it would help the reader to provide the research goals in the first 1-2 pages of the introduction to get a sense where the paper is headed.

The authors make the argument that developmental psychology has emphasized qualitative parenting behaviors whereas sociology and gender studies have focused on involvement quantity. The authors need to defend this point with more citations and nuance, but more broadly, perhaps the focus should move away from disciplinary differences in order to explain and justify why including both types of measures is important. It's seems this later idea is what the authors are intending to put forward to the reader but it gets lost a bit when describing disciplinary differences that do not seem precise.

The remaining introduction is thorough in its review, but needs more synthesis and direction for the reader. Providing the research goals toward the beginning of the introduction and then carefully walking the reader through the relevant existing literature may help to do this.

While the possibility of sex differences is brought up in the introduction, it is not brought forth in the research questions so it is unclear if the authors are going to investigate it, despite their justification for doing so.

Finally, hypotheses based upon the existing literature and theory would be useful in the introduction.

Methods
The annual income of the parents seems quite low, and the decision to focus only on non-minority families raises questions, however perhaps this is due to my lack of knowledge of the context in which this study is placed. It would be helpful to specify the geographic area of the study to provide some context. Similarly, more information on the sample is needed to understand how these results may or may not generalize to broader populations.

More information on all of the scales should be provided. Specifically, the number of items in each scale, reliability statistics, any information on external validity, and the reporter. It was unclear to me that the measure of emotional adjustment was teacher reported until the limitations section.

There is also more information needed on study recruitment, who were the participants, when each round of data collection took place, how the assessments were completed, ect.

Relatedly, how did the authors handle missing data across the study waves?

Results

Please include effect sizes when describing the results. What are the values in the tables? Please include this in a table note and describe in the text.

How were sex differences tested in the analytic models?

Showing the results in the form of figures and tables would be very helpful for readers.

When the authors describe effects as being larger for fathers in comparison to mothers, did they test this statistically?

Discussion

I am unsure of the context of this study, but there seem to be issues in terms of the representativeness of the sample and therefore these issues should be references in the limitations section.

Minor points:
* In the first sentence of the abstract, I believe it should be "The question of whether…" rather than "The question whether…"
* Extra spaces between paragraphs should be removed
* Miss spelling on page 12: agerage

Reviewer 2: Dear Authors,

Thank you for the opportunity to review this article. Overall I think it is an important study that makes contributions to the literature. However, I found it hard to follow most of the time with some inconsistencies that need to be addressed before it is published. The more substantial comments that highlight confusing sections and/or inconsistencies are bolded:

1. The opening paragraph does not convince me that fathers may uniquely contribute to children's emotional adjustment. I can see the argument that fathers may stimulate emotion by encouraging children outside of their comfort zones, but warmth and security (what the authors argue mothers do more of) should also make emotion safe and commonplace.
2. I don't think "incursions" is the right word (line 65).
3. Please do not capitalize the word after a colon within a sentence. In general there are some grammatical problems throughout that should be addressed before publication. There are also some typos (e.g., line 236 "agerage instead of average").
4. Can you say more about why parental anger is thought to undermine emotion regulation?
5. Why would emotional adjustment be negatively associated with consistency and inductive reasoning (line 105)? It doesn't follow to me that because maladjustment predicts less warmth, more demandingness, and more negative parenting that adjustment predicts more warmth and less demandingness. Shouldn't we expect to see demandingness, as you define it, along with other positive aspects of parenting? They are talked about as the more positive dimensions of demandingness that are not overly restrictive and applicable to children despite their emotional adjustment. This is also inconsistent with what you say your hypotheses are in the results section.
6. It would be helpful in the background to specify when you're talking about child influences (e.g., label the section "sex differences" as "child sex differences")
7. If you have income from ~1700 dads, why are you only reporting it for ~700?
8. I understand the need to maintain a blind document, but with so much information about the sample removed it is very hard to interpret the results. I did not realize the sample spanned two continents/cultures until the discussion - there is no mention of potential site differences before that point, or a discussion of how parenting is the same or different in different cultures.
9. Please provide examples of the structure and negative parenting items used.
10. Please define measurement invariance (parentheses is fine - e.g., measurement equivalence) for the less statistically trained reader.
11. The authors need to walk the reader through the tables better. The results section and tables need to be easier to navigate and understand. It is clear you did a tremendous amount of work on these analyses, and I want all readers to be able to follow the results very easily.
12. The findings on line 336 don't make sense given the previous line. The authors need to introduce all of the outcomes discussed in line 336: With regard to child effects, in line with hypotheses, the intercept of the cross-lagged
335 effect of emotional adjustment on maternal warmth was positive and significant (Table 2).
336 Thus, when children displayed better emotional adjustment than usual in one year, their
337 mothers displayed more anger and inductive reasoning, but also more warmth, than usual in
338 the next year.
13. Line 383 you use "discipline" for the first time. Do you mean demandingness?
14. I don't understand your argument in the discussion around the within family significant association between consistency and adjustment. You're saying that because you found tween family level correlations but not within family correlations that this suggests other correlated variables like economic hardship do not explain this association? What do your findings mean?
15. Regarding the direction of effects, is your conclusion that child effects are stronger than parenting effects driven entirely by the size of the coefficient? Did you conduct any tests to compare the two models and their coefficients? I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?
16. Can you test for non-linear effects of quantitative involvement on emotional adjustment? Your explanation that excessive involvement leads to maladjustment makes sense, but one would expect a u-shaped curve in which adjustment is highest at an optimal middle ground of involvement. You could also test for mean differences in emotional adjustment for children with highly involved parents versus less involved parents. It could be, as you argue in the next section of the discussion, that the effect of involvement on emotional adjustment is stronger when only one parent is involved because there is less variability in children's adjustment when both parents are highly involved. This would help test your two suggested explanations.

### end actual comment

# Reviewer \#1

\RC
This is a reviewer comment

This is our response

\Assignment{First Author}
\WorkInProgress
\Easy

\RC{This is the first paragraph of a long reviewer comment

This is the second paragraph}

This is our response.

> This is a section quoted from the revised manuscript to illustrate the change.

\Assignment{Second Author}
\AlmostDone
\Medium

\RC
This is a reviewer comment

This is our response

\Assignment{First Author}
\Done
\TimeConsuming
\Hard
\NotEstimated

# References

\setlength{\leftskip}{0.4in}
\setlength{\parindent}{-0.4in}

### 2019-10-29 Paper_shortened_pdf.Rmd
Fix spelling

@@
   When both parents were more involved, both engaged
-  more in desirable qualitative parenting. When mothers were disproportionally
+  more in desirable qualitative parenting. When mothers were disproportionately
   involved, however, parenting roles were more differentiated:
@@
 It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
-Emotional adjustment development is a domain in which fathers might play a unique role, because fathers encourage children to challenge themselves, in play and in life [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
-The attachment relationships between children and parents are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998].
-Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
-By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities.
+Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of the quantitative amount of care parents provide (who does what), and in terms of qualitative parenting behaviors (how do parents behave towards children).
+Nevertheless, the existing literature has two major shortcomings.
+Firstly, quantitative involvement and qualitative parenting behaviors are rarely jointly investigated.
+Therefore, we currently don't know how the interplay between quantitative and qualitative involvement relates to emotional development.
+Secondly, although fathers have been theorized to play a unique role in emotional adjustment development, empirical work has rarely addressed the joint and relative contributions of fathers' and mothers' involvement.
+The present study addresses overcomes these limitations by incorporating both quantitative and quantitative measures of mothers' and fathers' involvement.
+We use sophisticated methods to examine how joint and relative quantitative involvement moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment.
+We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment.
 
@@
 
+Attachment relationships are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+Consequently, fathers and mothers are theorized influence emotional adjustment development in somewhat different ways.
+
 The past decades have seen major strides in the consolidation of theories to guide fatherhood research.
-Inspired by increasing divorce rates and maternal employment, early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987].
-Over time, the notion of the father as a coparent gained ground [@pleckFatherhoodIdealsUnited1997], according to which fathers should share childrearing responsibilities with mothers, adopting roles historically fulfilled by mothers.
-To keep pace with changing fatherhood norms, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
+Early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987].
+Over time, the notion the fathers should share child-rearing responsibilities with mothers, adopting roles historically fulfilled by mothers, gained ground [@pleckFatherhoodIdealsUnited1997].
+Consequently, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
 
-Although it is now widely acknowledged that involvement quality and quantity are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together.
-Historically, developmental psychology has emphasized how qualitative parenting behaviors shape child development, whereas sociology and gender studies have focused on involvement quantity and its implications for maternal well-being and gender equality [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-The present study attempts to bridge these approaches by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
+It is now widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004].
+Nevertheless, these dimensions are rarely investigated together [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
@@
 Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013].
-We examined two aspects of demandingness: Consistency; the extent to which consequences for incursions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
+We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
 Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990].
@@
 With regard to differential involvement, we might also expect the more involved parent to display more desirable qualitative parenting behaviors, due to the effects of experience [@sabattiniRelationMothersFathers2004].
-However, the benefit of parents monitoring each other's childrearing should be diminished when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014].
+However, the benefit of parents monitoring each other's child-rearing should be diminished when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014].
 Indeed, retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
@@
 Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001].
-An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child wellbeing [@davies2006interparental].
+An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental].
 When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious.
@@
 This suggests that the total amount of care provided should be positively associated with children's emotional adjustment.
-Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their wellbeing [@deutschPaternalParticipationChild2001].
+Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
 
@@
 Coparenting refers to the ways parents relate to each other in the role of parent.
-Although the coparenting literature is typically invoked to describe childrearing in divorced families, it also holds insights for intact families.
-According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of differential involvement, affects qualitative childrearing behaviors and children's adjustment directly, and moderates the link between these two constructs.
+Although the coparenting literature is typically invoked to describe child-rearing in divorced families, it also holds insights for intact families.
+According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of differential involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs.
 
@@
 
-Emotional adjusmtment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].
+Emotional adjustment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].
 
@@
 We calculated mean scores for all of the study variables.
-Scale descriptive statistics are provided in Supplementary Table 1, and scale intercorrelations are provided in Supplementary Table 2.
+Scale descriptive statistics are provided in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
 We deemed it defensible to assume measurement invariance, as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
 
-To address our research questions, we used multilevel, autoregressive structural equation modelling [@asparouhovDynamicStructuralEquation2018].
+To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018].
 In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
@@
 None of the other intercepts for child effects on paternal parenting were significant.
-The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' childrearing behaviors differs between families.
+The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs between families.
 
@@
 This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978].
-These results are in line with our hypothesis that greater involvement would be associated with more desireable parenting behaviors [@sabattiniRelationMothersFathers2004].
+These results are in line with our hypothesis that greater involvement would be associated with more desirable parenting behaviors [@sabattiniRelationMothersFathers2004].
 Additionally, they suggest that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [@craigHowMothersFathers2011].
@@
 One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014].
-Research on "helicopter parenting" supports this interpretation, as psychological wellbeing was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014].
+Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014].
 Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs.
@@
 These correlational findings at the between-family level should not be interpreted causally, however.
-Future research should examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their wellbeing, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.
+Future research should examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their well-being, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.
 
@@
 This avoids common method bias and single reporter bias.
-As teachers are likely to be "blind" to parents' childrearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.
+As teachers are likely to be "blind" to parents' child-rearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.
 
@@
 First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction.
-This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in childrearing.
+This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in child-rearing.
 Moreover, the longitudinal models we used are not suitable for categorical variables, such as parenting style.

### 2019-10-29 Paper_shortened_pdf.Rmd
Fix income

@@
 At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$).
-Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$; 669 father-only, 41 mother-only).
+Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only).
 The modal annual income for fathers was between \$26,000-\$51,999 (N = 663), and for mothers, \$25,999 or less per year (N = 1152).

### 2019-10-29 Paper_shortened_pdf.Rmd
Reviewer 2 comment 8

@@
 Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only).
-The modal annual income for fathers was between \$26,000-\$51,999 (N = 663), and for mothers, \$25,999 or less per year (N = 1152).
+The modal annual income for fathers was between \$26,000-\$51,999 ($N = 663$ fathers reported this value), and for mothers, \$25,999 or less per year ($N = 1152$ reported this value).
 
@@
 However, commonly used developmental methods have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015].
-One recent study found that, in adolescence, child effects substantially outnumbered parenting effects [@vanlissaRoleFathersMothers2018a].
-The authors speculated that this might be due to the greater equality of parent-child relationships in adolescence, compared to earlier childhood.
-The present study casts doubt on this interpretation, as we found that child effects outweighed parenting effects, even at a young age.
-The prominence of child effects across two large longitudinal studies, with samples from different continents and at different age ranges, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it?
+One recent study with a similar design to the present study found that child effects substantially outnumbered parenting effects, and speculated that this might be due to the greater equality of parent-adolescent relationships, compared to earlier childhood [@vanlissaRoleFathersMothers2018a].
+The present study casts doubt on this speculation, as we found that child effects outweighed parenting effects, even at a young age.
+The fact that both the present study and this previous work found a preponderance of child effects, despite the different age ranges of the samples, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it?
 An extreme interpretation of our findings might be that parents "set the stage" for children's development, through their stable differences from other parents, and from there on, are merely responding to children's changing developmental needs.

### 2019-10-29 Action letter.Rmd
Reviewer 2 comment 9

@@
 
-thank you very much for taking the time to consider our manuscript for publication at the _Journal XYZ_.
-In the following we address your and each reviewer's concerns point-by-point.
+# Editor comments
 
-### Actual comments
+\rc{Introduction
 
-DEV-2019-2824
-Quantitative involvement moderates longitudinal associations between mothers’ and fathers’ qualitative parenting and children’s emotional adjustment
-Developmental Psychology
-
-Dear Dr. van Lissa,
-
-Thank you very much for submitting your manuscript "Quantitative involvement moderates longitudinal associations between mothers’ and fathers’ qualitative parenting and children’s emotional adjustment" for review and consideration for publication in Developmental Psychology.
-I sincerely appreciate the opportunity to review the manuscript.
-I have now received two reviews of the manuscript, and I have read the manuscript carefully myself.
-The reviews are appended to this email.
-As you will note, the reviewers agreed that your manuscript pertains to topics of importance—(1) how mothers and fathers influence their children's emotional adjustment and (2) the interplay between how much time and how parents spend caring for children.
-The strengths of the study include the sophisticated methodological approach, the inclusion of mothers and fathers, and the effort to include quantitative and qualitative measures of parenting.
-However, at the same time, the reviewers also identified a number of significant conceptual problems with the manuscript.
-Thus, I cannot accept the manuscript for publication in its present form.
-However, the reviewers and I agree that these issues could be adequately addressed in a revision.
-I encourage you therefore to revise and resubmit the manuscript.
-Although I believe you will find the reviewers' comments clear, it may be useful for me to summarize the major points made in the reviews.
-
-The reviewers were very consistent in their comments about the manuscript and I agree with them.
-One of the biggest concerns they raised was about the (lack of) clarity of the writing.
-Thus, my comments are organized to reflect my recommendations for improving clarity in each section of the manuscript.
-1. Introduction
-1a.
-Both reviewers (and I) think you are addressing an important set of questions, but the organization of the Introduction is very difficult to follow.
+The organization of the Introduction is very difficult to follow.
 I think it's fine to start with a description of emotional adjustment and to explain that the quality of the parent-child relationship is associated with children's emotional adjustment, but then I think you need to revise the text dramatically.
@@
 (2) what is not known about the association between the parent-child relationship and children's adjustment?
-(3) how will your study fill that gap in knowledge?
+(3) how will your study fill that gap in knowledge?}
+
 2. Method
@@
 
-Sincerely,
-Sara Jaffee
-Associate Editor
-Developmental Psychology
-EDITOR'S NOTE: Please make sure your revision is concise.
+\RC{
+
+}
+This is a reviewer comment
+
+This is our response
+
+\Assignment{First Author}
+\WorkInProgress
+\Easy
+
+\RC{This is the first paragraph of a long reviewer comment
+
+This is the second paragraph}
+
+This is our response.
+
+> This is a section quoted from the revised manuscript to illustrate the change.
+
+\Assignment{Second Author}
+\AlmostDone
+\Medium
+
+\RC
+This is a reviewer comment
+
+This is our response
+
+\Assignment{First Author}
+\Done
+\TimeConsuming
+\Hard
+\NotEstimated
+
+### Actual comments
+
+Dear Dr. van Lissa,
+
+Thank you very much for submitting your manuscript "Quantitative involvement moderates longitudinal associations between mothers’ and fathers’ qualitative parenting and children’s emotional adjustment" for review and consideration for publication in Developmental Psychology.
+I sincerely appreciate the opportunity to review the manuscript.
+I have now received two reviews of the manuscript, and I have read the manuscript carefully myself.
+The reviews are appended to this email.
+As you will note, the reviewers agreed that your manuscript pertains to topics of importance—(1) how mothers and fathers influence their children's emotional adjustment and (2) the interplay between how much time and how parents spend caring for children.
+The strengths of the study include the sophisticated methodological approach, the inclusion of mothers and fathers, and the effort to include quantitative and qualitative measures of parenting.
+However, at the same time, the reviewers also identified a number of significant conceptual problems with the manuscript.
+Thus, I cannot accept the manuscript for publication in its present form.
+However, the reviewers and I agree that these issues could be adequately addressed in a revision.
+I encourage you therefore to revise and resubmit the manuscript.
+Although I believe you will find the reviewers' comments clear, it may be useful for me to summarize the major points made in the reviews.
+
+## EDITOR'S NOTE: Please make sure your revision is concise.
 Our guidelines for length of manuscripts are 35 pages inclusive of everything except tables and figures.
@@
 
-NOTE FROM EDITORIAL OFFICE:  We ask that you upload either Word or RTF versions of both your title page and manuscript,  rather than PDFs, when resubmitting.
+## NOTE FROM EDITORIAL OFFICE:  We ask that you upload either Word or RTF versions of both your title page and manuscript,  rather than PDFs, when resubmitting.
 
@@
 
-1.
+\rc{1.
 The opening paragraph does not convince me that fathers may uniquely contribute to children's emotional adjustment.
-I can see the argument that fathers may stimulate emotion by encouraging children outside of their comfort zones, but warmth and security (what the authors argue mothers do more of) should also make emotion safe and commonplace.
-2. I don't think "incursions" is the right word (line 65).
-3.
+I can see the argument that fathers may stimulate emotion by encouraging children outside of their comfort zones, but warmth and security (what the authors argue mothers do more of) should also make emotion safe and commonplace.}
+
+We completely agree with the Reviewer that warmth and security, provided by either parent, are positively related to emotional adjustment development.
+We discuss this in REFER TO PARAGRAPH.
+It appears that the Reviewer interpreted the word *uniquely* to mean that *only* fathers contribute to emotional adjustment.
+We intended to say that fathers contribute in a *different* way from mothers.
+We merely intended to summarize fatherhood theory regarding the distinction between the attachment vs activation relationship: A widely endorsed theory that posits that demandingness is a more important socialization mechanism for fathers than for mothers, and that warmth is a more important socialization mechanism for mothers than for fathers.
+
+In the Revision, we have attempted to address the Reviewer's comment by disambiguating this section, which now reads:
+
+\AlmostDone
+
+\rc{2. I don't think "incursions" is the right word (line 65).}
+
+We appreciate the correction; we meant to say *transgressions* and have corrected this in the Revision.
+
+\Done
+
+\rc{3.
 Please do not capitalize the word after a colon within a sentence.
 In general there are some grammatical problems throughout that should be addressed before publication.
-There are also some typos (e.g., line 236 "agerage instead of average").
-4. Can you say more about why parental anger is thought to undermine emotion regulation?
+There are also some typos (e.g., line 236 "agerage instead of average").}
+
+We have capitalized the word after a colon if it is the beginning of an independent clause, in line with the the sixth edition of the APA Publication Manual.
+We would generally to prefer to adhere to the Publication Manual in these matters, but we defer to the Editor as to whether further changes are required.
+To address this typo, and other spelling errors, we have checked the manuscript with the R-package `hunspell` and fixed about 10 other errors.
+
+\Done
+
+\rc{4. Can you say more about why parental anger is thought to undermine emotion regulation?}
+
 5.
@@
 This is also inconsistent with what you say your hypotheses are in the results section.
+
 6. It would be helpful in the background to specify when you're talking about child influences (e.g., label the section "sex differences" as "child sex differences")
-7. If you have income from ~1700 dads, why are you only reporting it for ~700?
-8.
+
+\rc{7. If you have income from ~1700 dads, why are you only reporting it for ~700?}
+
+We are reporting the modal income for fathers and mothers, which is the appropriate measure of central tendency for an ordinal variable.
+Of all fathers, 663 reported this modal value.
+Based on this Reviewer comment, we understand that the way we reported income was ambiguous.
+In the revision, we have attempted to clarify this section as follows:
+
+> Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only).
+The modal annual income for fathers was between \$26,000-\$51,999 ($N = 663$ fathers reported this value), and for mothers, \$25,999 or less per year ($N = 1152$ reported this value).
+
+\rc{8.
 I understand the need to maintain a blind document, but with so much information about the sample removed it is very hard to interpret the results.
-I did not realize the sample spanned two continents/cultures until the discussion - there is no mention of potential site differences before that point, or a discussion of how parenting is the same or different in different cultures.
-9. Please provide examples of the structure and negative parenting items used.
+I did not realize the sample spanned two continents/cultures until the discussion - there is no mention of potential site differences before that point, or a discussion of how parenting is the same or different in different cultures.}
+
+There appears to be a misunderstanding here, engendered by a paragraph in the Discussion that relates our findings to the findings of a published study on a similar topic with a European adolescent sample.
+The published study found that child effects outnumbered parenting effects in adolescence, and speculated that this was due to the relative equality of parent-adolescent relationships.
+The present study also found a preponderance of child effects, at a younger age, which discredits this speculation.
+Thus, this is nothing more than a discussion point.
+We don't think that an extensive discussion of cultural differences is relevant (especially because the findings are similar).
+
+To address this comment, we have made two changes in the revision:
+
+1. We have clarified this section in the discussion.
+2.
+We no longer mention the different continents.
+Although it is certainly relevant that these studies were conducted on different populations, that information is orthogonal to the point we are trying to make about different age ranges.
+3. We have removed masking of sample details.
+
+The Discussion section in question now reads:
+
+> One recent study with a similar design to the present study found that child effects substantially outnumbered parenting effects, and speculated that this might be due to the greater equality of parent-adolescent relationships, compared to earlier childhood [@vanlissaRoleFathersMothers2018a].
+The present study casts doubt on this speculation, as we found that child effects outweighed parenting effects, even at a young age.
+The fact that both the present study and this previous work found a preponderance of child effects, despite the different age ranges of the samples, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it?
+
+\Done
+
+\rc{9. Please provide examples of the structure and negative parenting items used.}
+
+As requested, we have added these examples:
+
+> Structure and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
+
 10. Please define measurement invariance (parentheses is fine - e.g., measurement equivalence) for the less statistically trained reader.
+
 11.
@@
 
-# Reviewer \#1
-
-\RC
-This is a reviewer comment
-
-This is our response
-
-\Assignment{First Author}
-\WorkInProgress
-\Easy
-
-\RC{This is the first paragraph of a long reviewer comment
-
-This is the second paragraph}
-
-This is our response.
-
-> This is a section quoted from the revised manuscript to illustrate the change.
-
-\Assignment{Second Author}
-\AlmostDone
-\Medium
-
-\RC
-This is a reviewer comment
-
-This is our response
-
-\Assignment{First Author}
-\Done
-\TimeConsuming
-\Hard
-\NotEstimated
-
 # References

### 2019-10-29 Paper_shortened_pdf.Rmd
Reviewer 2 comment 9

@@
 Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001].
-Structure and negative parenting were measured using the consistency and anger scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
+Structure and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
 These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".

### 2019-10-29 Action letter.Rmd
Reviewer 2 Comment 15

@@
 
-10. Please define measurement invariance (parentheses is fine - e.g., measurement equivalence) for the less statistically trained reader.
+\rc{10. Please define measurement invariance (parentheses is fine - e.g., measurement equivalence) for the less statistically trained reader.}
+
+We have addressed this comment by including the following explanation:
+
+> (i.e., to assume that the factor loadings of each construct are the same across time points)
 
@@
 It is clear you did a tremendous amount of work on these analyses, and I want all readers to be able to follow the results very easily.
+
 12.
@@
 The authors need to introduce all of the outcomes discussed in line 336: With regard to child effects, in line with hypotheses, the intercept of the cross-lagged
+
 335 effect of emotional adjustment on maternal warmth was positive and significant (Table 2).
@@
 338 the next year.
-13. Line 383 you use "discipline" for the first time. Do you mean demandingness?
-14.
+
+\rc{13. Line 383 you use "discipline" for the first time. Do you mean demandingness?}
+
+Apologies for the oversight; we simply meant to say Inductive Reasoning. We have corrected this typo in the revision.
+
+\rc{14.
 I don't understand your argument in the discussion around the within family significant association between consistency and adjustment.
 You're saying that because you found tween family level correlations but not within family correlations that this suggests other correlated variables like economic hardship do not explain this association?
-What do your findings mean?
+What do your findings mean?}
+
+With all due respect, we do not entirely understand the Reviewer's comment, but we take it to mean that our argumentation in this paragraph was unclear.
+In the revision, we have rewritten this section as follows:
+
+> On average, we found no support for within-family predictive links between consistency and adjustment, but we did find support for between-family level correlations.
+Between-family level correlations can have different causes: They can reflect the consolidated influence of predictive effects that occurred prior to the study, or alternatively, the influence of third variables that impact both constructs.
+The fact that our study included children from a very young age casts doubt on the former explanation.
+It seems more likely that third variables explain these associations.
+Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+This finding means that we should reflect critically on causal interpretations of prior research that did not control for stable between-family differences [e.g., @brandPerceivedParentalInconsistency1990].
+
 15.

### 2019-10-29 Paper_shortened_pdf.Rmd
Reviewer 2 Comment 15

@@
 Scale descriptive statistics are provided in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
-We deemed it defensible to assume measurement invariance, as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
+We deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points), as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
 
@@
 We found no effects for consistency.
-For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when both parents' joint involvement was higher.
-In other words, when both parents were  more involved, the cross-lagged effect of children's adjustment on maternal discipline was more negative.
+For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when joint involvement was higher.
 Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
@@
 We found the predicted within-family correlations for mothers', but not fathers', consistency.
-On average, we found no support for within-family predictive links between consistency and adjustment.
-It is interesting that we found only between-family links and no predictive effects for consistency: Between-family level correlations can reflect the consolidated influence of predictive effects at an earlier age, or alternatively, the influence of third variables that impact both constructs.
+On average, we found no support for within-family predictive links between consistency and adjustment, but we did find support for between-family level correlations.
+Between-family level correlations can have different causes: They can reflect the consolidated influence of predictive effects that occurred prior to the study, or alternatively, the influence of third variables that impact both constructs.
+The fact that our study included children from a very young age casts doubt on the former explanation.
+It seems more likely that third variables explain these associations.
 Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
-The fact that we observed no within-family predictive effects, despite including children from a very young age, casts doubt on the former explanation, and on causal interpretations of prior research that did not control for stable between-family differences [@brandPerceivedParentalInconsistency1990].
+This finding means that we should reflect critically on causal interpretations of prior research that did not control for stable between-family differences [e.g., @brandPerceivedParentalInconsistency1990].
 

### 2019-10-29 Action letter.Rmd
Reviewer 2 Comment 16 17

@@
 
-15.
+\rc{15.
 Regarding the direction of effects, is your conclusion that child effects are stronger than parenting effects driven entirely by the size of the coefficient?
-Did you conduct any tests to compare the two models and their coefficients?
-I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?
-16.
+Did you conduct any tests to compare the two models and their coefficients?}
+
+We appreciate the suggestion; we are indeed only interpreting the pattern of effects (i.e., which effects differ 'significantly' from zero), and it would be better to explicitly test the difference between these parameters (i.e., which parent and child effects differ significantly from one another).
+In the revision, we have attempted to address this comment by explicitly acknowledging this in the Discussion section.
+
+We would prefer not to incorporate these tests for two reasons:
+
+1.
+We have no hypotheses about the difference between parent and child effects.
+Including difference tests would therefore introduce an additional and unnecessary layer of complexity to the manuscript - which is already very complex.
+2. For practical reasons: Each of the four models we conducted takes approximately a week to run (MCMC sampling is very computationally expensive).
+
+If the Editor deems it necessary to include these tests nonetheless, we will do so of course, but we'd need a longer turnaround time for the next revision.
+
+\Done
+
+\rc{I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?}
+
+This is not the case; parent and child effects refer to the reciprocal effects between the parenting variables and child adjustment; both variables must be included in one model to examine these reciprocal effects.
+
+\Done
+
+\rc{16.
 Can you test for non-linear effects of quantitative involvement on emotional adjustment?
-Your explanation that excessive involvement leads to maladjustment makes sense, but one would expect a u-shaped curve in which adjustment is highest at an optimal middle ground of involvement.
-You could also test for mean differences in emotional adjustment for children with highly involved parents versus less involved parents.
+Your explanation that excessive involvement leads to maladjustment makes sense, but one would expect a u-shaped curve in which adjustment is highest at an optimal middle ground of involvement.}
+
+We appreciate the Reviewer's suggestion.
+There are probably good grounds to question the linearity of *all* processes within the family system.
+However, it is not within the scope of our paper to make that point.
+We would be happy to conduct the exploratory analysis proposed by the Reviewer, but we are limited by the possibilities of contemporary statistical software.
+Estimating this non-linear effect involves a latent variable interaction of quantitative involvement with itself, to obtain the quadratic term.
+Latent variable interactions cannot currently be estimated within the Bayesian framework (they require Monte Carlo integration).
+Thus, we must choose: We can either test our theoretical model, which requires Bayesian estimation using Gibbs sampling, or we can conduct some kind of exploratory analysis of this proposed non-linear effect but with a much simpler model that can be estimated using numerical integration.
+Of course, we would prefer to analyze our theoretical model.
+
+\Done
+
+\rc{You could also test for mean differences in emotional adjustment for children with highly involved parents versus less involved parents.
 It could be, as you argue in the next section of the discussion, that the effect of involvement on emotional adjustment is stronger when only one parent is involved because there is less variability in children's adjustment when both parents are highly involved.
-This would help test your two suggested explanations.
+This would help test your two suggested explanations.}
 
-### end actual comment
+We appreciate the suggestion, but we believe that this comment might be due to an oversight on the Reviewer's part.
+We do investiage these effects, and they are reported in the section titled **Effects of quantitative involvement on parenting and adjustment**.
+That is why we suggested those explanations in the Discussion.
+Here are the relevant fragments from the Results section:
+
+> Counter to our hypothesis, we found that greater total joint involvement was negatively associated with emotional adjustment.
+Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
+
+And:
+
+> Finally, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.
+
+\Done
 

### 2019-10-30 Action letter.Rmd
Updates

--- 2019-10-29 Action letter.Rmd
@@
 title = "Quantitative involvement moderates longitudinal associations between mothers' and fathers' qualitative parenting and children's emotional adjustment"
+mask <- FALSE
+masked <- function(text){
+  ifelse(mask, "[masked]", text)
+}
 ```
+
+\subsection*{EDITOR'S NOTE: Please make sure your revision is concise.
+Our guidelines for length of manuscripts are 35 pages inclusive of everything except tables and figures.
+Because your manuscript reports on a longitudinal study, we could allow a few extra pages, but please keep close to this guideline.}
+
+\subsection*{NOTE FROM EDITORIAL OFFICE:  We ask that you upload either Word or RTF versions of both your title page and manuscript,  rather than PDFs, when resubmitting.}
 
@@
 
-# Editor comments
+\section*{Editor comments}
 
-\rc{Introduction
+\RC{Introduction
 
@@
 
-2. Method
-2a.
-Both reviewers found it hard to evaluate the findings without additional information about the sample, including how representative it is, how missing data were handled, and how much attrition there was across waves of the study.
-It is possible that reviewers and I have misinterpreted a sentence in the Discussion section in which you mention two longitudinal studies across two continents.
-At least one of the reviewers understood this to mean that your study involves samples from two continents, so please clarify if this is not the case.
-2b.
+\WorkInProgress
+
+\RC{Both reviewers found it hard to evaluate the findings without additional information about the sample, including how representative it is, how missing data were handled, and how much attrition there was across waves of the study.}
+
+We have addressed each of these issues in our detailed responses to the Reviewers below. Here, we provide an overview of changes made:
+
+1. We have unmasked all sections in the revised manuscript that describe the sample, so the Reviewers can assess its quality and generalizability.
+2. Bayesian estimation handles missing data intrinsically, in a similar way to FIML. We have explained this in the revised manuscript and in this letter.
+3.
+Attrition is minor, and is covered in Supplementary Table S1.
+In the revised manuscript, we now explicitly refer to all information covered in S1 (scale descriptive statistics, number of items, number of complete cases, and reliability).
+
+\Done
+
+\RC{It is possible that reviewers and I have misinterpreted a sentence in the Discussion section in which you mention two longitudinal studies across two continents.
+At least one of the reviewers understood this to mean that your study involves samples from two continents, so please clarify if this is not the case.}
+
+This is indeed a misunderstanding, and we apologize for the lack of clarity. Please see our response to Reviewer 2, comment 8.
+
+\Done
+
+\RC{2b.
 I was confused as to how joint and relative quantitative involvement were defined.
@@
 Please also use consistent language throughout.
-In the Results section, you describe total and relative involvement, but in the Intro and Discussion, you describe total joint care and the childcare distribution.
-3. Results
-3a.
+In the Results section, you describe total and relative involvement, but in the Intro and Discussion, you describe total joint care and the childcare distribution.}
+
+We thank the Editor for pointing out the lack of consistency across sections, and we have edited the revised manuscript to resolve this issue.
+We now consistently use the terms "joint and relative quantitative involvement".
+Joint quantitative involvement is defined as the average involvement of two parents.
+Relative involvement is defined as the difference between mothers and fathers (not a ratio).
+These variables are obtained by re-expressing the variance of the two observed variables, absolute maternal and paternal involvement, into two latent variables, reflecting the mean difference.
+We have included a visual representation of the model, which should clarify this technique.
+
+\WorkInProgress
+
+\RC{3a.
 Although I suspect a growing number of Developmental Psychology readers are aware of statistical advances on the standard cross-lagged model, many of them probably have not estimated those models.
 You would help readers understand your findings by presenting and explaining a conceptual figure of the statistical model.
-That way, readers might be better able to visualize it when you describe various random intercepts or bidirectional effects.
-3b. More extensive table notes would also be helpful. These could be used to remind readers of how some of the constructs are defined (e.g., joint and relative quantitative involvement).
-3c.
+That way, readers might be better able to visualize it when you describe various random intercepts or bidirectional effects.}
+
+We appreciate the suggestion, and have incorporated such a figure.
+
+\WorkInProgress
+
+\RC{3b. More extensive table notes would also be helpful. These could be used to remind readers of how some of the constructs are defined (e.g., joint and relative quantitative involvement).}
+
+\WorkInProgress
+
+\RC{3c.
 You could educate readers by helping them interpret the values in the tables.
@@
 14 where you explain, "Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year." Even there, however, it's not clear what a value of .14 means in the association of quant.
-involvement on maternal warmth.
-3d.
+involvement on maternal warmth.}
+
+\WorkInProgress
+
+\RC{3d.
 The results section is very dense.
@@
 How do you expect the quantitative measure of parenting to moderate the association between qualitative measures and children's emotional adjustment?
-4. Discussion
-4a.
+}
+
+\WorkInProgress
+
+\RC{4a.
 The Discussion, as it is currently written, raises a number of issues that are not necessarily anticipated in the Introduction.
@@
 The first several paragraphs of your discussion, however, focus on how patterns of results differ across parenting dimensions (e.g., results for inconsistent parenting are different from results for parental warmth).
-I understand that the data show these differences, but the presentation of the results and discussion would be more streamlined if you focused on those that were directly relevant to your hypothesis.
-4b.
+I understand that the data show these differences, but the presentation of the results and discussion would be more streamlined if you focused on those that were directly relevant to your hypothesis.}
+
+\WorkInProgress
+
+\RC{4b.
 There are a number of issues you raise in the Discussion that don't have a parallel in the Introduction.
@@
 father differences) or when they are observed at the between-family, but not within-family level (e.g., some of the parenting dimension-emotional adjustment associations).
-Because you haven't really explained clearly when you expect to see between-family effects and when you expect to see within-family effects, I think the significance of much of the discussion will be lost on readers.
-4c.
+Because you haven't really explained clearly when you expect to see between-family effects and when you expect to see within-family effects, I think the significance of much of the discussion will be lost on readers.}
+
+\WorkInProgress
+
+\RC{4c.
 I have two comments about your discussion of child effects (p.
@@
 First, although you note that you are measuring the parent-child dyad early in development, the children in your sample are already 6 years old when they are first assessed and they are 12 by the last measurement occasion.
-Although there is no question in my mind that child effects are present from the earliest stages in development, I think the likelihood of observing parenting effects is greater at still younger ages (e.g., infancy and toddlerhood).
-Second, there are pronounced individual differences in how parents react to children.
+Although there is no question in my mind that child effects are present from the earliest stages in development, I think the likelihood of observing parenting effects is greater at still younger ages (e.g., infancy and toddlerhood).}
+
+\WorkInProgress
+
+\RC{Second, there are pronounced individual differences in how parents react to children.
 The same level of emotionally dysregulated behavior might elicit anger from one parent and inductive reasoning from another.
 There is a great deal of evidence from parenting interventions that how parents react to difficult behavior affects whether their children remain noncompliant and dysregulated or whether their behavior improves.
-So I think you're selling parents short if your claim is that the only way they influence their children is through the changing expression of the genes they have bequeathed them.
-So, we encourage you to submit a revision.
-To do so, please go to https://www.editorialmanager.com/dvl/ and log in as an Author.
-You will see a menu item called "Submissions Needing Revision".
-You will find your submission record there.
-Please resubmit within 30 days.
-If you need more time to complete additional analyses, this time period is negotiable, but let us know beforehand.
-Please include an 'Authors Response to Reviewers' with your revision that includes details on how you addressed the various issues that I highlighted above, as well as the other issues raised by the reviewers.
-I will return the revision to the original reviewers for additional comments .
+So I think you're selling parents short if your claim is that the only way they influence their children is through the changing expression of the genes they have bequeathed them.}
 
-Thank you for considering Developmental Psychology as an outlet for your work.
+\WorkInProgress
 
-\RC{
+# Reviewer 1:
 
-}
-This is a reviewer comment
+\RC{This manuscript addresses an important research area: understanding the role of mothers and fathers in the emotional adjustment of children.
+Strengths of this paper include (a) the inclusion of quantitative and qualitative measures of parenting from both parents, (b) a within- and between-person design, and (c) a longitudinal study with repeated measures of children and parenting.
+Below, I describe my suggestions to improve the manuscript as currently written.}
 
-This is our response
+We thank the Reviewer for their time, and for the helpful suggestions. We have attempted to address them thoroughly and in good faith.
 
-\Assignment{First Author}
+\RC{The first paragraph of the introduction introduces many different ideas and comes across as disjointed.
+Revision should make it clear to readers that the focus is on emotional adjustment among children (as well as the specific developmental age of focus), describe a more complex, well-cited definition of emotional adjustment (the current citation and definition does not seem accurate), introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers.
+Finally, it would help the reader to provide the research goals in the first 1-2 pages of the introduction to get a sense where the paper is headed.}
+
 \WorkInProgress
-\Easy
 
-\RC{This is the first paragraph of a long reviewer comment
+\RC{The authors make the argument that developmental psychology has emphasized qualitative parenting behaviors whereas sociology and gender studies have focused on involvement quantity.
+The authors need to defend this point with more citations and nuance, but more broadly, perhaps the focus should move away from disciplinary differences in order to explain and justify why including both types of measures is important.
+It's seems this later idea is what the authors are intending to put forward to the reader but it gets lost a bit when describing disciplinary differences that do not seem precise.}
 
-This is the second paragraph}
+\WorkInProgress
 
-This is our response.
+\RC{The remaining introduction is thorough in its review, but needs more synthesis and direction for the reader.
+Providing the research goals toward the beginning of the introduction and then carefully walking the reader through the relevant existing literature may help to do this.}
 
-> This is a section quoted from the revised manuscript to illustrate the change.
+\WorkInProgress
 
-\Assignment{Second Author}
-\AlmostDone
-\Medium
+\RC{While the possibility of sex differences is brought up in the introduction, it is not brought forth in the research questions so it is unclear if the authors are going to investigate it, despite their justification for doing so.}
 
-\RC
-This is a reviewer comment
+\WorkInProgress
 
-This is our response
+\RC{Finally, hypotheses based upon the existing literature and theory would be useful in the introduction.}
 
-\Assignment{First Author}
+We agree wholeheartedly with this point, but there is a challenge: Because nobody has previously modeled the interplay between parental involvement and emotional adjustment in the way we have, controlling for stable between-family differences, and examining quantitative involvement in terms of the joint and relative contribution of mothers and fathers, there is no precedent upon which to base such hypotheses.
+Therefore, the best we can do is to present *conceptual* hypotheses.
+
+\WorkInProgress
+
+\RC{The annual income of the parents seems quite low, and the decision to focus only on non-minority families raises questions, however perhaps this is due to my lack of knowledge of the context in which this study is placed.
+It would be helpful to specify the geographic area of the study to provide some context.
+Similarly, more information on the sample is needed to understand how these results may or may not generalize to broader populations.}
+
+To facilitate evaluation of the sample, we have un-masked this information in the Revision.
+
+Median household income in Australia was $\$58708$ in 2006, the first year of data collection.
+We only have information on income categories, but in our sample, most families were dual earners, with most fathers reporting between \$26,000-\$51,999, and most mothers reporting \$25,999 of income.
+This seems to be very comfortably around the median income in Australia at the time.
+That makes sense, because the study contains a representative sample.
+We did decide to focus only on white urbanized Australians, because the amount of heterogeneity introduced by including rural and nonwhite/aboriginal Australians led to model non-convergence.
+
+\WorkInProgress
+
+\RC{More information on all of the scales should be provided. Specifically, the number of items in each scale, reliability statistics...}
+
+All of this information requested is provided Supplementary Table 1, as stated in the text (*Scale descriptive statistics are provided in Supplementary Table 1*).
+Perhaps these supplementary Tables are not automatically shared with reviewers?
+
+To address this comment, we've clarified this sentence in the manuscript, which now reads: *Scale descriptive statistics, number of items, number of complete cases, and reliability are reported in Supplementary Table 1*.
+We also reproduce Table S1 here, in case the Reviewer does not have access:
+
+```{r, echo = FALSE, results = 'asis'}
+library(knitr)
+library(kableExtra)
+kable(read.csv("Supplementary.Table.1.item.descriptives.csv"), booktabs = TRUE) %>%
+  kable_styling(font_size = 8)
+```
+
 \Done
-\TimeConsuming
-\Hard
-\NotEstimated
 
-### Actual comments
+\RC{...any information on external validity...}
 
-Dear Dr. van Lissa,
+We have followed the convention of 1) providing example items as an indication of face validity, and 2) citing the scale validation papers for more information on predictive validity.
+We would like to defer to the Editor as to whether further actions need to be taken (we are also not sure what these actions would be).
 
-Thank you very much for submitting your manuscript "Quantitative involvement moderates longitudinal associations between mothers’ and fathers’ qualitative parenting and children’s emotional adjustment" for review and consideration for publication in Developmental Psychology.
-I sincerely appreciate the opportunity to review the manuscript.
-I have now received two reviews of the manuscript, and I have read the manuscript carefully myself.
-The reviews are appended to this email.
-As you will note, the reviewers agreed that your manuscript pertains to topics of importance—(1) how mothers and fathers influence their children's emotional adjustment and (2) the interplay between how much time and how parents spend caring for children.
-The strengths of the study include the sophisticated methodological approach, the inclusion of mothers and fathers, and the effort to include quantitative and qualitative measures of parenting.
-However, at the same time, the reviewers also identified a number of significant conceptual problems with the manuscript.
-Thus, I cannot accept the manuscript for publication in its present form.
-However, the reviewers and I agree that these issues could be adequately addressed in a revision.
-I encourage you therefore to revise and resubmit the manuscript.
-Although I believe you will find the reviewers' comments clear, it may be useful for me to summarize the major points made in the reviews.
+\Done
 
-## EDITOR'S NOTE: Please make sure your revision is concise.
-Our guidelines for length of manuscripts are 35 pages inclusive of everything except tables and figures.
-Because your manuscript reports on a longitudinal study, we could allow a few extra pages, but please keep close to this guideline.
+\RC{...and the reporter. It was unclear to me that the measure of emotional adjustment was teacher reported until the limitations section.}
 
-## NOTE FROM EDITORIAL OFFICE:  We ask that you upload either Word or RTF versions of both your title page and manuscript,  rather than PDFs, when resubmitting.
+This information is provided in the Measures section, as it should be:
 
-Reviewers' comments:
+> Emotional adjustment was measured using the reverse-coded emotional distress subscale of the __teacher-reported__ Strengths and Difficulties Questionnaire
 
-Reviewer 1: This manuscript addresses an important research area: understanding the role of mothers and fathers in the emotional adjustment of children.
-Strengths of this paper include (a) the inclusion of quantitative and qualitative measures of parenting from both parents, (b) a within- and between-person design, and (c) a longitudinal study with repeated measures of children and parenting.
-Below, I describe my suggestions to improve the manuscript as currently written.
+\Done
 
-Introduction
+\RC{There is also more information needed on study recruitment, who were the participants, when each round of data collection took place, how the assessments were completed, ect.}
 
-The first paragraph of the introduction introduces many different ideas and comes across as disjointed.
-Revision should make it clear to readers that the focus is on emotional adjustment among children (as well as the specific developmental age of focus), describe a more complex, well-cited definition of emotional adjustment (the current citation and definition does not seem accurate), introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers.
-Finally, it would help the reader to provide the research goals in the first 1-2 pages of the introduction to get a sense where the paper is headed.
+Several working papers have been written on the recruitment, participants, data collection, etc.
+for this sample.
+However, this information was masked in the original manuscript, to facilitate blind peer review.
+We feel that it is unnecessary to reproduce this information for a well-documented large cohort study, as we can simply cite the original material.
+However, as we also understand that the Reviewer needs access to these references to evaluate the study quality, we have now revealed the masked sections:
 
-The authors make the argument that developmental psychology has emphasized qualitative parenting behaviors whereas sociology and gender studies have focused on involvement quantity.
-The authors need to defend this point with more citations and nuance, but more broadly, perhaps the focus should move away from disciplinary differences in order to explain and justify why including both types of measures is important.
-It's seems this later idea is what the authors are intending to put forward to the reader but it gets lost a bit when describing disciplinary differences that do not seem precise.
+> Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.
 
-The remaining introduction is thorough in its review, but needs more synthesis and direction for the reader.
-Providing the research goals toward the beginning of the introduction and then carefully walking the reader through the relevant existing literature may help to do this.
+\Done
 
-While the possibility of sex differences is brought up in the introduction, it is not brought forth in the research questions so it is unclear if the authors are going to investigate it, despite their justification for doing so.
+\RC{Relatedly, how did the authors handle missing data across the study waves?}
 
-Finally, hypotheses based upon the existing literature and theory would be useful in the introduction.
+One property of the Bayesian estimation algorithm used, is that it intrinsically handles missing values without the need to impute them.
+It is comparable in that sense to full-information maximum likelihood estimation.
+Under the assumption of MAR, this produces consistent estimates.
+We have now clarified this in the revised manuscript by including the following sentence in the Strategy of Analyses:
 
-Methods
-The annual income of the parents seems quite low, and the decision to focus only on non-minority families raises questions, however perhaps this is due to my lack of knowledge of the context in which this study is placed.
-It would be helpful to specify the geographic area of the study to provide some context.
-Similarly, more information on the sample is needed to understand how these results may or may not generalize to broader populations.
+> Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
 
-More information on all of the scales should be provided.
-Specifically, the number of items in each scale, reliability statistics, any information on external validity, and the reporter.
-It was unclear to me that the measure of emotional adjustment was teacher reported until the limitations section.
-
-There is also more information needed on study recruitment, who were the participants, when each round of data collection took place, how the assessments were completed, ect.
-
-Relatedly, how did the authors handle missing data across the study waves?
+\Done
 
@@
 
-Please include effect sizes when describing the results. What are the values in the tables? Please include this in a table note and describe in the text.
+\RC{Please include effect sizes when describing the results.}
 
-How were sex differences tested in the analytic models?
+All of our results are reported in terms of correlations (which, when squared, gives the proportion of explained variance shared by two variables) and standardized regression coefficients (a one SD increase in X predicts a $\beta$ SD increase in Y, conditional on the other predictors), both of which are already effect sizes [@nieminenStandardisedRegressionCoefficient2013].
 
-Showing the results in the form of figures and tables would be very helpful for readers.
+\RC{What are the values in the tables? Please include this in a table note and describe in the text.}
 
-When the authors describe effects as being larger for fathers in comparison to mothers, did they test this statistically?
+\RC{How were sex differences tested in the analytic models?}
 
-Discussion
+This is explained in the strategy of analyses:
 
-I am unsure of the context of this study, but there seem to be issues in terms of the representativeness of the sample and therefore these issues should be references in the limitations section.
+> Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
+In this way, we included total joint childcare, the childcare distribution, __and child sex__ as moderators.
 
-Minor points:
-* In the first sentence of the abstract, I believe it should be "The question of whether…" rather than "The question whether…"
-* Extra spaces between paragraphs should be removed
-* Miss spelling on page 12: agerage
+\RC{Showing the results in the form of figures and tables would be very helpful for readers.}
 
-Reviewer 2: Dear Authors,
+We do currently show the results in the form of Tables, but we have also included conceptual Figures to further clarify the model.
 
-Thank you for the opportunity to review this article.
-Overall I think it is an important study that makes contributions to the literature.
-However, I found it hard to follow most of the time with some inconsistencies that need to be addressed before it is published.
-The more substantial comments that highlight confusing sections and/or inconsistencies are bolded:
+\WorkInProgress
 
-\rc{1.
+\RC{When the authors describe effects as being larger for fathers in comparison to mothers, did they test this statistically?}
+
+We appreciate this suggestion!
+Strictly speaking, since Bayesian analyses focus on estimation rather than inference, it is correct to say that the (maximum a-posteriori) estimate is different between mothers and fathers, without further substantiation - because in the Bayesian framework, the maximum a-posteriori estimate can be interpreted as the most likely population value for the parameter, conditional on the data and prior.
+In practice, since we use Bayesian estimation for pragmatic reasons, we understand that there is added value in knowing the posterior distribution of the difference between mothers and fathers.
+
+Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between mothers and fathers.
+We have reported this information in-text whenever making a statement about mother/father differences, and have added it to the relevant Supplementary Tables with the full model parameters.
+We have omitted it from the smaller Tables in the manuscript, because their function is to help make it easier for readers to digest the results, and we feel that adding four columns with difference tests would interfere with that goal.
+
+\WorkInProgress
+
+\RC{Discussion: I am unsure of the context of this study, but there seem to be issues in terms of the representativeness of the sample and therefore these issues should be references in the limitations section.}
+
+\WorkInProgress
+
+\RC{Minor points:
+* In the first sentence of the abstract, I believe it should be "The question of whether…" rather than "The question whether…"}
+
+We appreciate the correction.
+
+\Done
+
+\RC{* Extra spaces between paragraphs should be removed}
+
+\WorkInProgress
+
+\RC{* Miss spelling on page 12: agerage}
+
+We appreciate the correction. To address this typo, and other spelling errors, we have checked the manuscript with the R-package `hunspell` and fixed about 10 other errors.
+
+\Done
+
+# Reviewer 2:
+
+\RC{Dear Authors, Thank you for the opportunity to review this article. Overall I think it is an important study that makes contributions to the literature.}
+
+Thank you sincerely for the time you have spent reviewing this manuscript, and for your kind words about its potential contribution.
+Your comments were very helpful in improving the quality of the manuscript, and we have attempted to address all of them in good faith, as detailed below.
+
+Sincerely,
+the authors
+
+\RC{1.
 The opening paragraph does not convince me that fathers may uniquely contribute to children's emotional adjustment.
@@
 
-\rc{2. I don't think "incursions" is the right word (line 65).}
+\RC{2. I don't think "incursions" is the right word (line 65).}
 
@@
 
-\rc{3.
+\RC{3.
 Please do not capitalize the word after a colon within a sentence.
@@
 
-\rc{4. Can you say more about why parental anger is thought to undermine emotion regulation?}
+\RC{4. Can you say more about why parental anger is thought to undermine emotion regulation?}
 
-5.
+\WorkInProgress
+
+\RC{5.
 Why would emotional adjustment be negatively associated with consistency and inductive reasoning (line 105)?
@@
 They are talked about as the more positive dimensions of demandingness that are not overly restrictive and applicable to children despite their emotional adjustment.
-This is also inconsistent with what you say your hypotheses are in the results section.
+This is also inconsistent with what you say your hypotheses are in the results section.}
 
-6. It would be helpful in the background to specify when you're talking about child influences (e.g., label the section "sex differences" as "child sex differences")
+\WorkInProgress
 
-\rc{7. If you have income from ~1700 dads, why are you only reporting it for ~700?}
+\RC{6. It would be helpful in the background to specify when you're talking about child influences (e.g., label the section "sex differences" as "child sex differences")}
 
-We are reporting the modal income for fathers and mothers, which is the appropriate measure of central tendency for an ordinal variable.
+\WorkInProgress
+
+\RC{7. If you have income from ~1700 dads, why are you only reporting it for ~700?}
+
+We believe that this comment might be due to a misunderstanding, that can be resolved by clearer writing on our part.
+We reported the modal income for fathers and mothers, which is the appropriate measure of central tendency for an ordinal variable (the LSAC recorded income in ordinal categories, with relatively little dispersion).
 Of all fathers, 663 reported this modal value.
-Based on this Reviewer comment, we understand that the way we reported income was ambiguous.
+That is where the number ~700 comes from.
 In the revision, we have attempted to clarify this section as follows:
@@
 > Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only).
+Income was recorded in ordered categories.
 The modal annual income for fathers was between \$26,000-\$51,999 ($N = 663$ fathers reported this value), and for mothers, \$25,999 or less per year ($N = 1152$ reported this value).
 
-\rc{8.
+\Done
+
+\RC{8.
 I understand the need to maintain a blind document, but with so much information about the sample removed it is very hard to interpret the results.
@@
 
-There appears to be a misunderstanding here, engendered by a paragraph in the Discussion that relates our findings to the findings of a published study on a similar topic with a European adolescent sample.
+We understand that all Reviewers and the Editor feel that the masked manuscript interferes with proper assessment of the sample.
+Therefore, we have decided to unmask the revised manuscript.
+With regard to the point about two continents: This appears to be a misunderstanding, engendered by a paragraph in the Discussion that relates our findings to the findings of a published study on a similar topic with a European adolescent sample.
 The published study found that child effects outnumbered parenting effects in adolescence, and speculated that this was due to the relative equality of parent-adolescent relationships.
-The present study also found a preponderance of child effects, at a younger age, which discredits this speculation.
-Thus, this is nothing more than a discussion point.
-We don't think that an extensive discussion of cultural differences is relevant (especially because the findings are similar).
+The present study also found a preponderance of child effects, at a younger age, which discredits that speculation.
+Thus, there are no two continents/cultures within the present study.
 
@@
 1. We have clarified this section in the discussion.
-2.
-We no longer mention the different continents.
-Although it is certainly relevant that these studies were conducted on different populations, that information is orthogonal to the point we are trying to make about different age ranges.
-3. We have removed masking of sample details.
+2. We have removed masking of sample details.
 
@@
 
-\rc{9. Please provide examples of the structure and negative parenting items used.}
+\RC{9. Please provide examples of the structure and negative parenting items used.}
 
@@
 
-\rc{10. Please define measurement invariance (parentheses is fine - e.g., measurement equivalence) for the less statistically trained reader.}
+\Done
 
-We have addressed this comment by including the following explanation:
+\RC{10. Please define measurement invariance (parentheses is fine - e.g., measurement equivalence) for the less statistically trained reader.}
 
-> (i.e., to assume that the factor loadings of each construct are the same across time points)
+We have done as requested, and included the following explanation:
 
-11.
+> (i.e., to assume that the factor loadings of each construct are the same across time points and respondents)
+
+\RC{11.
 The authors need to walk the reader through the tables better.
 The results section and tables need to be easier to navigate and understand.
-It is clear you did a tremendous amount of work on these analyses, and I want all readers to be able to follow the results very easily.
+It is clear you did a tremendous amount of work on these analyses, and I want all readers to be able to follow the results very easily.}
 
-12.
+\WorkInProgress
+
+\RC{12.
 The findings on line 336 don't make sense given the previous line.
-The authors need to introduce all of the outcomes discussed in line 336: With regard to child effects, in line with hypotheses, the intercept of the cross-lagged
+The authors need to introduce all of the outcomes discussed in line 336: With regard to child effects, in line with hypotheses, the intercept of the cross-lagged}
 
@@
 
-\rc{13. Line 383 you use "discipline" for the first time. Do you mean demandingness?}
+\WorkInProgress
+
+\RC{13. Line 383 you use "discipline" for the first time. Do you mean demandingness?}
 
@@
 
-\rc{14.
+\Done
+
+\RC{14.
 I don't understand your argument in the discussion around the within family significant association between consistency and adjustment.
@@
 
-With all due respect, we do not entirely understand the Reviewer's comment, but we take it to mean that our argumentation in this paragraph was unclear.
+We are not sure that we fully understand the Reviewer's comment, but we take it to mean that our argumentation in this paragraph was unclear.
 In the revision, we have rewritten this section as follows:
@@
 
-15.
+\Done
+
+\RC{15.
 Regarding the direction of effects, is your conclusion that child effects are stronger than parenting effects driven entirely by the size of the coefficient?
-Did you conduct any tests to compare the two models and their coefficients?
-I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?
-16.
+Did you conduct any tests to compare the two models and their coefficients?}
+
+We appreciate this suggestion!
+Strictly speaking, since Bayesian analyses focus on estimation rather than inference, it is correct to say that the (maximum a-posteriori) estimate is different between mothers and fathers, without further substantiation - because in the Bayesian framework, the maximum a-posteriori estimate can be interpreted as the most likely population value for the parameter, conditional on the data and prior.
+In practice, since we use Bayesian estimation for pragmatic reasons, we understand that there is added value in knowing the posterior distribution of the difference between mothers and fathers.
+
+Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between mothers and fathers.
+We have reported this information in-text whenever making a statement about mother/father differences, and have added it to the relevant Supplementary Tables with the full model parameters.
+We have omitted it from the smaller Tables in the manuscript, because their function is to help make it easier for readers to digest the results, and we feel that adding four columns with difference tests would interfere with that goal.
+
+\WorkInProgress
+
+<!--We appreciate the suggestion; we are indeed only interpreting the pattern of effects (i.e., which effects differ 'significantly' from zero), and it would be better to explicitly test the difference between these parameters (i.e., which parent and child effects differ significantly from one another).
+In the revision, we have attempted to address this comment by explicitly acknowledging this in the Discussion section.
+
+We would prefer not to incorporate these tests for two reasons:
+
+1.
+We have no hypotheses about the difference between parent and child effects.
+Including difference tests would therefore introduce an additional and unnecessary layer of complexity to the manuscript - which is already very complex.
+2. For practical reasons: Each of the four models we conducted takes approximately a week to run (MCMC sampling is very computationally expensive).
+
+If the Editor deems it necessary to include these tests nonetheless, we will do so of course, but we'd need a longer turnaround time for the next revision.
+
+\Done-->
+
+\RC{I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?}
+
+This assumption is incorrect; parent and child effects refer to the reciprocal effects between the parenting variables and child adjustment; both variables must be included in one model to examine these reciprocal effects.
+This is outlined in the paragraph below:
+
+> To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018].
+In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
+This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015].
+Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts.
+Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
+In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.
+We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
+
+\Done
+
+\RC{16.
 Can you test for non-linear effects of quantitative involvement on emotional adjustment?
-Your explanation that excessive involvement leads to maladjustment makes sense, but one would expect a u-shaped curve in which adjustment is highest at an optimal middle ground of involvement.
+Your explanation that excessive involvement leads to maladjustment makes sense, but one would expect a u-shaped curve in which adjustment is highest at an optimal middle ground of involvement.}
+
+We appreciate the Reviewer's suggestion, and agree that there are probably good grounds to question the linearity of *all* processes within the family system.
+However, it is not within the scope of the present paper to make that point.
+In principle, we would be happy to conduct the exploratory analysis proposed by the Reviewer, but we are stopped by the limitations of contemporary statistical software.
+Including a quadratic term for this non-linear effect requires estimating a latent variable interaction of quantitative involvement with itself.
+Such latent variable interactions cannot currently be estimated within the Bayesian framework; they require Monte Carlo integration.
+Thus, we must choose: We can either test our theoretical model, which requires Bayesian estimation using Gibbs sampling, or we analyze this proposed non-linear effect, but within a much simpler model that can be estimated using numerical integration.
+Of course, we would prefer to analyze our theoretical model.
+
+\Done
+
+\RC{17.
 You could also test for mean differences in emotional adjustment for children with highly involved parents versus less involved parents.
 It could be, as you argue in the next section of the discussion, that the effect of involvement on emotional adjustment is stronger when only one parent is involved because there is less variability in children's adjustment when both parents are highly involved.
-This would help test your two suggested explanations.
+This would help test your two suggested explanations.}
 
-### end actual comment
+We agree with the Reviewer that mean differences in emotional adjustment based on low vs high involvement are important.
+That is why we do investiage these effects, and they are reported in the section titled **Effects of quantitative involvement on parenting _and adjustment_**.
+That is why we suggested those explanations in the Discussion.
+Here are the relevant fragments from the Results section:
 
-# References
+> Counter to our hypothesis, we found that greater total joint involvement was negatively associated with emotional adjustment.
+Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
+*[...]*  
+Finally, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.
+
+\Done
+
+\section*{References}
 

### 2019-10-30 Paper_shortened_pdf.Rmd
Reviewer 2 Comment 16 17
Updates

@@
 abstract: |
-  The question whether fathers and mothers play similar roles in children’s
+  The question of whether fathers and mothers play similar roles in children’s
   emotional adjustment has been approached differently in different research
@@
 load("table_coefficients.RData")
-mask <- TRUE
+mask <- FALSE
 masked <- function(text){
@@
 Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only).
+Income was recorded in ordered categories.
 The modal annual income for fathers was between \$26,000-\$51,999 ($N = 663$ fathers reported this value), and for mothers, \$25,999 or less per year ($N = 1152$ reported this value).
@@
 We calculated mean scores for all of the study variables.
-Scale descriptive statistics are provided in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
-We deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points), as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
+Scale descriptive statistics, number of items, number of complete cases, and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
+We deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents), as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
 
@@
 In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.
+We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
 
@@
 We used two Gibbs-sampler chains of 50000 iterations each.
+Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
+The missing data handling is likelihood based and guarantees consistent estimation as long as the missing data is missing at random (MAR).
 We applied mildly informative inverse Wishart priors to the variance/covariance matrices, and normal priors centered on zero for all regression coefficients.

### 2019-10-30 Paper_shortened_pdf.Rmd
Create figure 1, fix tables

@@
 In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.
+Figure \@ref(fig:figure1) shows a graphical representation of the resulting model.
 We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
+
+```{r figure1, out.width="100%", fig.cap="Abbreviations: `T-1' (the previous measurement occasion), `ri' (random intercept), `s' random slope, `M' (Mother's qualitative involvement), `E' (Emotional adjustment), `F' (Father's qualitative involvement), `inv' (quantitative involvement).
+Thus, `sME' represents the random slope of Mothers' qualitative involvement on emotional adjustment."}
+knitr::include_graphics("Figure1.png")
+```
 
@@
 rownames(tab) <- NULL
-reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+cat(as.character(reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")))
 ```

### 2019-10-30 Figure1.R

@@ -1,0 +1,8 @@
+library(gridExtra)
+source("Figure1between.R")
+source("Figure1within.R")
+p <- ggplot()+geom_vline(xintercept = 1, linetype = 3)+theme_void()
+p <- grid.arrange(p_within, p, p_between, nrow = 1, widths = c(10,1,10))
+
+Fig1 <- p
+ggsave("Figure1.png", p)

### 2019-10-30 Figure1between.R

@@ -1,0 +1,148 @@
+library(motley)
+library(ggplot2)
+library(ggforce)
+
+layout <- get_layout(read.csv("fig1_between_layout.csv", header = FALSE, stringsAsFactors = FALSE))
+
+df_nodes <-  data.frame(node_id = 1:length(layout$param), param = layout$param, stringsAsFactors = FALSE)
+df_nodes$shape <- "rect"
+df_nodes$shape[grepl("^(ri|s|Joint|Relative)", df_nodes$param)] <- "oval"
+df_nodes$label <- df_nodes$param
+#df_nodes$param <- df_nodes$node_id
+df_edges <- data.frame(matrix(c(
+  3, 1, "last", "1", 
+  3, 2, "last", "1", 
+  
+  5, 1, "last", "+1", 
+  5, 2, "last", "-1", 
+  
+  3, 6, "last", "", 
+  3, 7, "last", "", 
+  3, 8, "last", "", 
+  3, 9, "last", "", 
+  3, 10, "last", "", 
+  3, 11, "last", "", 
+  3, 12, "last", "",   
+  
+  4, 6, "last", "", 
+  4, 7, "last", "", 
+  4, 8, "last", "", 
+  4, 9, "last", "", 
+  4, 10, "last", "", 
+  4, 11, "last", "", 
+  4, 12, "last", "",   
+  
+  5, 6, "last", "", 
+  5, 7, "last", "", 
+  5, 8, "last", "", 
+  5, 9, "last", "", 
+  5, 10, "last", "", 
+  5, 11, "last", "", 
+  5, 12, "last", ""
+  ), ncol = 4, byrow = TRUE), stringsAsFactors = FALSE)
+names(df_edges) <- c("from", "to", "arrow", "label")
+
+df_edges$from <- df_nodes$param[as.numeric(df_edges$from)]
+df_edges$to <- df_nodes$param[as.numeric(df_edges$to)]
+df_edges$label[grepl("^ri", df_edges$from)] <- "1"
+prep <- prepare_plot_sem(nodes = df_nodes, layout = layout, edges = df_edges)
+
+
+# Plot --------------------------------------------------------------------
+x <- prep
+df_nodes <- x$nodes
+df_edges <- x$edges
+rect_width <- x$rect_width
+rect_height <- x$rect_height
+ellipses_a <- x$ellipses_a
+ellipses_b <- x$ellipses_b
+spacing_x <- x$spacing_x
[…]

### 2019-10-30 Figure1within.R

@@ -1,0 +1,126 @@
+library(motley)
+library(ggplot2)
+library(ggforce)
+
+layout <- get_layout(read.csv("fig1_within_layout.csv", header = FALSE, stringsAsFactors = FALSE))
+
+df_nodes <-  data.frame(node_id = 1:length(layout$param), param = layout$param, stringsAsFactors = FALSE)
+df_nodes$shape <- "rect"
+df_nodes$shape[grepl("^(ri|s)", df_nodes$param)] <- "oval"
+df_nodes$label <- df_nodes$param
+
+df_edges <- data.frame(matrix(c(
+  1, 11, "last", "", 
+  1, 12, "last", "", 
+  #1, 13, "last", "", 
+  
+  2, 11, "last", "", 
+  2, 12, "last", "", 
+  2, 13, "last", "", 
+  
+  #3, 11, "last", "", 
+  3, 12, "last", "", 
+  3, 13, "last", "", 
+  
+  14, 11, "last", "",
+  15, 12, "last", "",
+  16, 13, "last", ""), ncol = 4, byrow = TRUE), stringsAsFactors = FALSE)
+names(df_edges) <- c("from", "to", "arrow", "label")
+
+df_edges$from <- df_nodes$param[as.numeric(df_edges$from)]
+df_edges$to <- df_nodes$param[as.numeric(df_edges$to)]
+df_edges$label[grepl("^ri", df_edges$from)] <- "1"
+prep <- prepare_plot_sem(nodes = df_nodes, layout = layout, edges = df_edges)
+
+# Plot --------------------------------------------------------------------
+x <- prep
+df_nodes <- x$nodes
+df_edges <- x$edges
+rect_width <- x$rect_width
+rect_height <- x$rect_height
+ellipses_a <- x$ellipses_a
+ellipses_b <- x$ellipses_b
+spacing_x <- x$spacing_x
+spacing_y <- x$spacing_y
+text_size <- x$text_size
+
+df_nodes[grepl("^s(EM|ME|EF|FE)", df_nodes$param), c("x", "node_xmin", "node_xmax")] <- df_nodes[grepl("^s(EM|ME|EF|FE)", df_nodes$param), c("x", "node_xmin", "node_xmax")]+.25
+
+
+connect_points <- setNames(data.frame(t(
+  mapply(function(from, to, startpoint, endpoint){
+    c(
+      switch(startpoint,
+             right = df_nodes$node_xmax[which(df_nodes$node_id == from)],
[…]

### 2019-10-30 rfunctions.R

@@ -46,10 +46,18 @@ reg_table <- function(x, ...){
   rownames(x) <- NULL
   names(x) <- gsub("^.+est_sig$", "Est.", names(x))
   names(x) <- gsub("^.+confint$", "95\\% HPDI", names(x))
-  tab <- capture.output(apa_table(x, format = "latex", 
-                                  col_spanners = colspan_fun(x, skip = 1), ...))
-  tab <- gsub("^(.+?) & NA & NA & NA & NA & NA & NA & NA & NA\\\\\\\\$", "\\\\multicolumn\\{9\\}\\{l\\}\\{\\1\\}\\\\\\\\", tab)
-  cat(tab, sep = "\n")
+  tab <- apa_table(x, format = "latex", col_spanners = colspan_fun(x, skip = 1), ...)
+  tab <- gsub('(^"|"$)', '', tab)
+  tab <- gsub("^\\[1\\] ", "", tab)
+  
+  tab <- gsub("\\n(.+?) & NA & NA & NA & NA & NA & NA & NA & NA\\\\\\\\\\n", "\\\\multicolumn\\{9\\}\\{l\\}\\{\\1\\}\\\\\\\\", tab, perl = TRUE)
+  #tab <- gsub("\\\\", "\\", tab)
+  #tab <- gsub("\\\\begin", "\\begin", tab)
+  #tab <- gsub("\\\\n(?!=orm)", "\\\\", tab, perl = TRUE)
+  #tab <- gsub("%", "\\%", tab)
+  #cat(tab, sep = "\n")
+  #knitr::raw_latex(tab)
+  tab
 }
 
 
@@ -62,7 +70,7 @@ reg_table_html <- function(x, ...){
   x[multiline_rows, 1] <- ""
   x[is.na(x)] <- ""
   tab <- capture.output(apa_table(x, format = "html", 
-                                   ...))
+                                   ...)[1])
   browser()
   div_line <- grep("^-[ -]+$", tab)
   tab[which(multiline_rows)+div_line] <- sprintf(paste0("%-", nchar(tab[which(multiline_rows)[1]+div_line]), "s"), multiline_text)

### 2019-11-01 Action letter.Rmd
Add DSEM results and update action letter

@@
 
-We have addressed each of these issues in our detailed responses to the Reviewers below. Here, we provide an overview of changes made:
+We have addressed these four issues in our detailed responses to the Reviewers below. Here, we provide an overview of changes made:
 
@@
 
-This is indeed a misunderstanding, and we apologize for the lack of clarity. Please see our response to Reviewer 2, comment 8.
+This is indeed a misunderstanding, and we apologize for the lack of clarity.
+Please see our response to Reviewer 2, comment 8.
+In this part of the Discussion section, we relate our findings to those of a previous study with a similar design.
 
@@
 We now consistently use the terms "joint and relative quantitative involvement".
-Joint quantitative involvement is defined as the average involvement of two parents.
-Relative involvement is defined as the difference between mothers and fathers (not a ratio).
-These variables are obtained by re-expressing the variance of the two observed variables, absolute maternal and paternal involvement, into two latent variables, reflecting the mean difference.
-We have included a visual representation of the model, which should clarify this technique.
+Joint quantitative involvement is defined as the average involvement of both parents (a latent mean).
+Relative involvement is defined as the difference between mothers and fathers (not a ratio; a latent difference).
+These variables are obtained by re-expressing the variance of the two observed variables, absolute amount maternal and paternal involvement, into two latent variables, reflecting the mean and difference.
+We have included a Figure that presents the model visually, which should clarify this technique.
+We also explicitly point out, once in the Introduction and once in the Discussion, that joint involvement represents the total amount of care provided to children, whereas relative involvement captures the childcare distribution who does what).
 
-\WorkInProgress
+\Done
 
@@
 
-\WorkInProgress
+\Done
 
 \RC{3b. More extensive table notes would also be helpful. These could be used to remind readers of how some of the constructs are defined (e.g., joint and relative quantitative involvement).}
+
+We have elaborated on the Table notes.
+We think that the definition of joint and relative involvement is adequately addressed by the increased consistency with regard to terminology, in conjunction with the new Figure - but we are happy to take further action if necessary.
 
@@
 
+To address this comment, we've attempted to ensure that every issue in the Introduction is paralleled in the Discussion.
+We have also attempted to clarify the between/within distinction.
+A major contribution of the present study is that we are (one of?) the first to make this distinction in this particular research area.
+Consequently, there is no prior literature available to base specific hypotheses about between/within effects on.
+The best we can do is develop exploratory hypotheses based on a conflicting literature that has lumped together between/within effects.
+Thus, we cannot make more specific predictions than we currently do.
+
 \WorkInProgress
@@
 
+We appreciate this suggestion.
+We agree that parent effects could very well be present at even younger ages, and explicitly acknowledge this in the revision.
+At the same time, we think that it is interesting and potentially meaningful that parent effects - but not child effects - tend to become smaller/non significant when stable between-family differences are accounted for.
+We have attempted to clarify our discussion of the potential theoretical/methodological implications of this finding.
+
+> Nowadays, it is commonly acknowledged that parents and children reciprocally influence one another.
+However, the methods used to demonstrate bi-directional effects have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015].
+We are aware of only one prior study that controlled for stable between-family differences when examining the interplay between parenting and emotional adjustment.
+This study found that child effects substantially outnumbered parenting effects in adolescence [@vanlissaRoleFathersMothers2018a].
+The authors explained this finding in light of the relative equality of parent-adolescent relationships.
+The present study casts doubt on that interpretation, as we found a preponderance of child effects between the ages of 6-12.
+Although it is, of course, possible that parent effects are in effect at an even younger age, we can also consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
+The finding that parent effects become smaller or non-significant once stable between-family differences are accounted for has also been demonstrated in other areas of developmental research (REF).
+As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
+Such findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
+Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
+At the same time, we should consider the limitations of these methods.
+The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
+Stable factors can also exert changing influences over time.
+As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time.
+It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
+Another important limitation is that within-family cross-lagged effects are time scale dependent.
+If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
+Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
+
 \WorkInProgress
@@
 
-\WorkInProgress
+We agree, and as we mention in the manuscript, we also find these between-family differences in parents' responses to childen's emotional adjustment in the present study.
+In the rewritten paragraphs above, we have attempted to clarify that we are not making any claims about genes, but instead, we are questioning the assumption of the RI-CLPM model that all reciprocal influences are captured by within-family fluctuation.
+We then introduce genes as an illustrative example of something very stable that can nonetheless have changing influences over the life course.
+In the revision, we have emphasized that this is an illustrative example, and we then draw a parallel to how stable psychosocial traits might similarly have changing effects over the course of development.
+
+\Done
 
@@
 
-\WorkInProgress
+Thank you for pointing out that sex differences are omitted from the list of contributions mentioned in *The Present Study*.
+We have attempted to address this issue by adding sex differences to the list, and we have also clarified our specific hypotheses regarding sex differences in the preceding paragraph.
+
+> We therefore hypothesize that parents will display more responsiveness, structure, and inductive reasoning towards girls, and more negative parenting towards boys.
+Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and the same-sex parent.
+*[...]*
+Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.  
+*[...]*
+Finally, we examined sex as a moderator of between-family differences in all study variables and within-family predictive effects
+
+\Done
 
@@
 
-We agree wholeheartedly with this point, but there is a challenge: Because nobody has previously modeled the interplay between parental involvement and emotional adjustment in the way we have, controlling for stable between-family differences, and examining quantitative involvement in terms of the joint and relative contribution of mothers and fathers, there is no precedent upon which to base such hypotheses.
+We agree wholeheartedly with this point, and we had formulated hypotheses for all patterns studied, but we understand that our colloquial use of language made it hard to identify our specific hypotheses.
+For example, in the original manuscript, we wrote:
+
+> Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.
+
+In the revision, we have attempted to address this comment by being more specific. Now, this section reads:
+
+> Applying this insight from the coparenting literature, we thus hypothesize that  parents' quantitative involvement will be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and will moderate effects between qualitative parenting behaviors and emotional adjustment.
+
+There is a bit of a challenge, however: Because there is no precedent for using the random-intercept cross-lagged panel model or its multilevel DSEM extension in the literature on parental involvement and emotional adjustment, there is no literature upon which to base such hypotheses.
 Therefore, the best we can do is to present *conceptual* hypotheses.
@@
 
-Median household income in Australia was $\$58708$ in 2006, the first year of data collection.
+Median household income in Australia was $\$58.708$ in 2006, the first year of data collection.
 We only have information on income categories, but in our sample, most families were dual earners, with most fathers reporting between \$26,000-\$51,999, and most mothers reporting \$25,999 of income.
-This seems to be very comfortably around the median income in Australia at the time.
+This seems to be very close to the median income in Australia at the time.
 That makes sense, because the study contains a representative sample.
-We did decide to focus only on white urbanized Australians, because the amount of heterogeneity introduced by including rural and nonwhite/aboriginal Australians led to model non-convergence.
+We did decide to select only white urbanized Australians from that sample, because initial exploratory analyses indicated that the amount of heterogeneity introduced by including rural and nonwhite/aboriginal Australians led to model non-convergence.
 
-\WorkInProgress
+\Done
 
@@
 
-All of this information requested is provided Supplementary Table 1, as stated in the text (*Scale descriptive statistics are provided in Supplementary Table 1*).
+This might be an oversight, as all of this information requested is provided in Supplementary Table 1 (*Scale descriptive statistics are provided in Supplementary Table 1*).
 Perhaps these supplementary Tables are not automatically shared with reviewers?
@@
 
-This information is provided in the Measures section, as it should be:
+This information is provided in the Measures section:
 
@@
 for this sample.
-However, this information was masked in the original manuscript, to facilitate blind peer review.
-We feel that it is unnecessary to reproduce this information for a well-documented large cohort study, as we can simply cite the original material.
-However, as we also understand that the Reviewer needs access to these references to evaluate the study quality, we have now revealed the masked sections:
+However, this information was masked in the original manuscript, to enable blind peer review.
+However, as both Reviewers and the Editor requested access to these references to evaluate the study quality, we have now unmasked these sections in the revised manuscript:
 
@@
 
-Results
-
 \RC{Please include effect sizes when describing the results.}
@@
 
+*Probably, we should cite the values in the text, as the Editor also suggested.*
+
+\WorkInProgress
+
 \RC{What are the values in the tables? Please include this in a table note and describe in the text.}
+
+We have attempted to address this comment by labeling and annotating the Tables more extensively.
+
+\WorkInProgress
 
@@
 
-This is explained in the strategy of analyses:
+In the original manuscript, we explained how sex differences were tested in the strategy of analyses:
 
@@
 
+In the revision, we have attempted to further clarify this by including a visual representation of the model, and by labeling Tables more clearly.
+
+\Done
+
 \RC{Showing the results in the form of figures and tables would be very helpful for readers.}
 
-We do currently show the results in the form of Tables, but we have also included conceptual Figures to further clarify the model.
+In the original manuscript, we included four Tables with the results reported in the Results section. In the revision, we have additionally included a Figure to further clarify the model.
 
-\WorkInProgress
+\Done
 

### 2019-11-01 Paper_shortened_pdf.Rmd
Add DSEM results and update action letter

--- 2019-10-30 Paper_shortened_pdf.Rmd
@@
 A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001].
-Examining the childcare distribution instead, and controlling for total care, mitigates this confound.
-Some prior research has attempted to measure the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016].
-When quantitative involvement has been measured in terms of absolute involvement, latent means and differences can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].
+This confound can be mitigated by examining the effect of the distribution of childcare, while controlling for total care.
+Some prior research has attempted to achieve this by assessing the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016; @keizerInfluenceFathersMothers2019].
+Alternatively, when quantitative involvement has been measured in terms of absolute involvement, one can estimate a latent mean and difference to statistically disentangle joint involvement from relative involvement [@mcardleLatentVariableModeling2009].
+Joint involvement reflects the total amount of care provided to children, and relative involvement reflects the childcare distribution.
 
@@
 Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014].
-Consequently, we might expect greater total joint care to positively predict both parents' parenting quality.
+Consequently, we might expect greater joint involvement to positively predict both parents' parenting quality.
 
@@
 Finally, mounting evidence suggests that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
-We might thus expect parents to display more responsiveness, structure, and inductive reasoning with girls, and more negative parenting with boys.
-Furthermore, we might expect associations between qualitative parenting and emotional adjustment to be stronger between children and the same-sex parent.
+We therefore hypothesize that parents will display more responsiveness, structure, and inductive reasoning towards girls, and more negative parenting towards boys.
+Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and the same-sex parent.
 
 Although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a].
-We consequently did not expect to find sex differences in children's average levels of emotional adjustment.
+Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.
 
@@
 Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
+Finally, we examined sex as a moderator of between-family differences in all study variables and within-family predictive effects.
 
@@
 Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
-In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.
+In this way, we included joint and relative quantitative involvement, and child sex as moderators.
+Figure \@ref(fig:figure1) shows a graphical representation of the resulting model.
 We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
+
+```{r figure1, out.width="100%", fig.cap="T-1: previous measurement occasion.
+ri: random intercept.
+s: random slope.
+M: Mother's qualitative involvement.
+E: Emotional adjustment.
+F: Father's qualitative involvement.
+inv: quantitative involvement.
+Thus, sME represents the random slope of Mothers' qualitative involvement on emotional adjustment."}
+knitr::include_graphics("Figure1.png")
+```
 
@@
 rownames(tab) <- NULL
-reg_table(tab, caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+reg_table(tab, caption = "Between-family level intercepts and variances of within-family predictive effects", note = "Est.: Estimate.
+HDPI: Highest posterior density interval.
+Emo.: Emotional adjustment.
+*: <5% of posterior has opposite sign from estimate.")
 ```
@@
 
-We examined regression coefficients from total joint involvement and relative involvement to the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+We examined regression coefficients from joint and relative quantitative involvement to the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
 With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning.
-Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers.
-We did not find the hypothesized negative effects of joint total involvement on parental anger.
-Counter to our hypothesis, we found that greater total joint involvement was negatively associated with emotional adjustment.
+Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers.
+We did not find the hypothesized negative effects of joint involvement on parental anger.
+Counter to our hypothesis, we found that greater joint involvement was negatively associated with emotional adjustment.
 Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
@@
                               "Effects of relative quantitative involvement on random intercepts of:")
-tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
+tabinv$Parameter <- gsub("^(.+) on (?!random).*$", "\\1", tabinv$Parameter, perl= TRUE)
 #reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
@@
 rownames(tab) <- NULL
-reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of between-family differences in quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
+Est.: Estimate.
+HDPI: Highest posterior density interval.
+Emo.: Emotional adjustment.
+*: <5% of posterior has opposite sign from estimate.")))
 ```
@@
 With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
-Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when both parents' total joint involvement was higher.
+Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when both parents' joint quantitative involvement was higher.
 This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement.
@@
 
-With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers.
+With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint quantitative involvement was higher, and also when mothers' relative involvement was greater compared to fathers.
 We found no effects for consistency.
@@
 rownames(tabsex) <- NULL
-reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Sex is a binary predictor (-1 = female, 1 = male).
+The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
+Est.: Estimate.
+HDPI: Highest posterior density interval.
+Emo.: Emotional adjustment.
+*: <5% of posterior has opposite sign from estimate.")
 ```
@@
 
-Our operationalization of quantitative involvement, in terms of total joint care and the childcare distribution, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001].
-However, counter to our hypothesis, joint care was *negatively* associated with between-family differences in emotional adjustment.
+Our operationalization of quantitative involvement, in terms of joint involvement and relative involvement, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001].
+However, counter to our hypothesis, joint involvement was *negatively* associated with between-family differences in emotional adjustment.
 One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014].
@@
 We found limited evidence that quantitative involvement moderated within-family parenting effects.
-One such effect is that the effect of maternal warmth on adjustment was contingent on both total joint care and the childcare distribution.
+One such effect is that the effect of maternal warmth on adjustment was contingent on both joint- and relative quantitative involvement.
 Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers.
@@
 The evidence that quantitative moderates child effects was somewhat stronger.
-Firstly, when total joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth.
+Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth.
 These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth.
 In families where mothers' involvement was high - either in terms of joint, or relative involvement - this positive effect was attenuated somewhat.
-Secondly, total joint involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning.
+Secondly, joint quantitative involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning.
 This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved.
@@
 
-Historically, developmental theory has focused on how parents cultivate their children's development.
-Nowadays, it is commonly acknowledged that influences are reciprocal.
-However, commonly used developmental methods have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015].
-One recent study with a similar design to the present study found that child effects substantially outnumbered parenting effects, and speculated that this might be due to the greater equality of parent-adolescent relationships, compared to earlier childhood [@vanlissaRoleFathersMothers2018a].
-The present study casts doubt on this speculation, as we found that child effects outweighed parenting effects, even at a young age.
-The fact that both the present study and this previous work found a preponderance of child effects, despite the different age ranges of the samples, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it?
-An extreme interpretation of our findings might be that parents "set the stage" for children's development, through their stable differences from other parents, and from there on, are merely responding to children's changing developmental needs.
-This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations.
-For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time.
-Consequently, stable parent characteristics can cause change in children.
-Future research might examine alternative theoretical conceptualizations of parental influences in child development.
-Another point to consider is that parent effects on children might play out on a different time scale than was captured by these macro-longitudinal studies.
-If parents shape their children through specific behaviors in dyadic interactions, then we might fail to capture parenting effects when using annual reflective self-reports.
+Nowadays, it is commonly acknowledged that parents and children reciprocally influence one another.
+However, the methods used to demonstrate bi-directional effects have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015].
+We are aware of only one prior study that controlled for stable between-family differences when examining the interplay between parenting and emotional adjustment.
+This study found that child effects substantially outnumbered parenting effects in adolescence [@vanlissaRoleFathersMothers2018a].
+The authors explained this finding in light of the relative equality of parent-adolescent relationships.
+The present study casts doubt on that interpretation, as we found a preponderance of child effects between the ages of 6-12.
+Although it is, of course, possible that parent effects are in effect at an even younger age, we can also consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
+
+The finding that parent effects become smaller or non-significant once stable between-family differences are accounted for has also been demonstrated in other areas of developmental research (REF).
+As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
+Such findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
+Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
+
+At the same time, we should consider the limitations of these methods.
+The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
+Stable factors can also exert changing influences over time.
+As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time.
+It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
+Another important limitation is that within-family cross-lagged effects are time scale dependent.
+If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
+Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
 
@@
 Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001].
-We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement.
+We addressed this limitation by estimating the latent mean and difference of absolute maternal and paternal quantitative involvement.
+This allows us to disentangle the effect of joint involvement, which reflects the total amount of care children receive, from relative involvement, which reflects the childcare distribution.
 Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level.
@@
 
+<!--  -->
+
 ## Conclusions
@@
 
-We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when total joint involvement was higher.
+We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when joint involvement was higher.
 Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth).

### 2019-11-07 Action letter.Rmd
Sent to Renske

@@
 
-We thank you and the Reviewers for your thorough comments, for the helpful summary you've provided.
+We thank you and the Reviewers for your thorough comments, and for the helpful summary you've provided.
 We hope that you will find that our responses adequately address the issues raised, and that the quality of the work has been improved.
 
-<!--The manuscript (including title page, abstract, and references) comprises 35 pages, or 8666 words.
+<!--The revised manuscript (including title page, abstract, and references) comprises 35 pages, or 8666 words.
 Four supplementary tables are included, whose size prohibits them from being included in the manuscript: A full correlation table (total sample and paneled by adolescent sex); the full matrix of parameter estimates; detailed scale descriptives and intercorrelations, and a full matrix of parameter estimates for an alternative analysis suggested by a Reviewer.
-Should the manuscript be submitted, we are prepared to make all syntax and results of the analyses conducted available on a public code repository referenced in the paper, to safeguard reproducibility.
+Should the manuscript be accepted, we are prepared to make all syntax and results of the analyses conducted available on a public code repository referenced in the paper, to safeguard reproducibility.
 Treatment of participants was in accordance with the APA ethical principles.
@@
 
-1. We have unmasked all sections in the revised manuscript that describe the sample, so the Reviewers can assess its quality and generalizability.
-2. Bayesian estimation handles missing data intrinsically, in a similar way to FIML. We have explained this in the revised manuscript and in this letter.
+1.
+We have unmasked all sections in the revised manuscript that describe the sample, so the Reviewers can assess its quality and generalizability.
+We also address generalizability in the *Limitations* section.
+
+The *Participants* section now reads:
+
+> We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012.
+Each wave was reviewed and approved by the
+`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
+Ethics Committee registered with the `r masked("National Health and Medical Research Council")`.
+Documentation of the ethics approval is available online at `r masked("https://growingupinaustralia.gov.au/sites/default/files/certification_of_ethics_approval_for_lsac.pdf")`.
+`r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population.
+Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.
+
+The *Limitations* section now reads:
+
+> A final potential limitation pertains to the generalizability of the findings.
+Our target sample consisted of native `r masked("Australian")`, non-minority, intact, biological families.
+It is possible, but beyond the scope of this study, that the findings might have limited generalizability for Australian minorities, or for families in other industrialized nations.
+
+2.
+Bayesian estimation handles missing data intrinsically, in a similar way to FIML.
+We have explained this in the revised manuscript (below) and in this letter (see our response to Reviewer 1, comment 1.10).
+
+> Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
+The missing data handling is likelihood based and guarantees consistent estimation as long as the missing data is missing at random (MAR).
+
 3.
@@
 
-\RC{This manuscript addresses an important research area: understanding the role of mothers and fathers in the emotional adjustment of children.
+\RC{1.1 This manuscript addresses an important research area: understanding the role of mothers and fathers in the emotional adjustment of children.
 Strengths of this paper include (a) the inclusion of quantitative and qualitative measures of parenting from both parents, (b) a within- and between-person design, and (c) a longitudinal study with repeated measures of children and parenting.
@@
 
-\RC{The first paragraph of the introduction introduces many different ideas and comes across as disjointed.
-Revision should make it clear to readers that the focus is on emotional adjustment among children (as well as the specific developmental age of focus), describe a more complex, well-cited definition of emotional adjustment (the current citation and definition does not seem accurate), introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers.
+\RC{1.2a The first paragraph of the introduction introduces many different ideas and comes across as disjointed.
+Revision should make it clear to readers that the focus is on emotional adjustment among children (as well as the specific developmental age of focus)...}
+
+\RC{1.2b ...describe a more complex, well-cited definition of emotional adjustment (the current citation and definition does not seem accurate)...}
+
+We operationalized emotional adjustment in terms of the emotional adjustment subscale of the SDQ.
+Consequently, we cite the scale validation paper of the SDQ here.
+It is not evident to us what other reference would be more appropriate than the one we used to operationalize this construct, but if the Reviewer feels that a particular citation should be included, we are happy to do so.
+
+\RC{1.2c ...introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers.
 Finally, it would help the reader to provide the research goals in the first 1-2 pages of the introduction to get a sense where the paper is headed.}
@@
 
-\RC{The authors make the argument that developmental psychology has emphasized qualitative parenting behaviors whereas sociology and gender studies have focused on involvement quantity.
+\RC{1.3 The authors make the argument that developmental psychology has emphasized qualitative parenting behaviors whereas sociology and gender studies have focused on involvement quantity.
 The authors need to defend this point with more citations and nuance, but more broadly, perhaps the focus should move away from disciplinary differences in order to explain and justify why including both types of measures is important.
@@
 
-\RC{The remaining introduction is thorough in its review, but needs more synthesis and direction for the reader.
+\RC{1.4 The remaining introduction is thorough in its review, but needs more synthesis and direction for the reader.
 Providing the research goals toward the beginning of the introduction and then carefully walking the reader through the relevant existing literature may help to do this.}
@@
 
-\RC{While the possibility of sex differences is brought up in the introduction, it is not brought forth in the research questions so it is unclear if the authors are going to investigate it, despite their justification for doing so.}
+\RC{1.5 While the possibility of sex differences is brought up in the introduction, it is not brought forth in the research questions so it is unclear if the authors are going to investigate it, despite their justification for doing so.}
 
@@
 
-\RC{Finally, hypotheses based upon the existing literature and theory would be useful in the introduction.}
+\RC{1.6 Finally, hypotheses based upon the existing literature and theory would be useful in the introduction.}
 
@@
 
-\RC{The annual income of the parents seems quite low, and the decision to focus only on non-minority families raises questions, however perhaps this is due to my lack of knowledge of the context in which this study is placed.
+\RC{1.7 The annual income of the parents seems quite low, and the decision to focus only on non-minority families raises questions, however perhaps this is due to my lack of knowledge of the context in which this study is placed.
 It would be helpful to specify the geographic area of the study to provide some context.
@@
 
+Furthermore, we now address generalizability in the *Limitations* section:
+
+> A final potential limitation pertains to the generalizability of the findings.
+Our target sample consisted of native `r masked("Australian")`, non-minority, intact, biological families.
+It is possible, but beyond the scope of this study, that the findings might have limited generalizability for Australian minorities, or for families in other industrialized nations.
+
 \Done
 
-\RC{More information on all of the scales should be provided. Specifically, the number of items in each scale, reliability statistics...}
+\RC{1.8a More information on all of the scales should be provided. Specifically, the number of items in each scale, reliability statistics...}
 
@@
 
-\RC{...any information on external validity...}
+\RC{1.8b ...any information on external validity...}
 
@@
 
-\RC{...and the reporter. It was unclear to me that the measure of emotional adjustment was teacher reported until the limitations section.}
+\RC{1.8c ...and the reporter. It was unclear to me that the measure of emotional adjustment was teacher reported until the limitations section.}
 
@@
 
-\RC{There is also more information needed on study recruitment, who were the participants, when each round of data collection took place, how the assessments were completed, ect.}
+\RC{1.9 There is also more information needed on study recruitment, who were the participants, when each round of data collection took place, how the assessments were completed, ect.}
 
@@
 
-\RC{Relatedly, how did the authors handle missing data across the study waves?}
+\RC{1.10 Relatedly, how did the authors handle missing data across the study waves?}
 
@@
 
-\RC{Please include effect sizes when describing the results.}
+\RC{1.11 Please include effect sizes when describing the results.}
 
@@
 
-\RC{What are the values in the tables? Please include this in a table note and describe in the text.}
+\RC{1.12 What are the values in the tables? Please include this in a table note and describe in the text.}
 
@@
 
-\RC{How were sex differences tested in the analytic models?}
+\RC{1.13 How were sex differences tested in the analytic models?}
 
@@
 
-\RC{Showing the results in the form of figures and tables would be very helpful for readers.}
+\RC{1.14 Showing the results in the form of figures and tables would be very helpful for readers.}
 
@@
 
-\RC{When the authors describe effects as being larger for fathers in comparison to mothers, did they test this statistically?}
+\RC{1.15 When the authors describe effects as being larger for fathers in comparison to mothers, did they test this statistically?}
 
@@
 
-\RC{Discussion: I am unsure of the context of this study, but there seem to be issues in terms of the representativeness of the sample and therefore these issues should be references in the limitations section.}
+\RC{1.16 Discussion: I am unsure of the context of this study, but there seem to be issues in terms of the representativeness of the sample and therefore these issues should be references in the limitations section.}
 
-\WorkInProgress
+We have attempted to address this comment by unmasking the *Participants* section (see below), and discussing generalizability in the *Limitations* section (see our response to 1.7).
 
-\RC{Minor points:
+The *Participants* section now reads:
+
+> We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012.
+Each wave was reviewed and approved by the
+`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
+Ethics Committee registered with the `r masked("National Health and Medical Research Council")`.
+Documentation of the ethics approval is available online at `r masked("https://growingupinaustralia.gov.au/sites/default/files/certification_of_ethics_approval_for_lsac.pdf")`.
+`r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population.
+Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.
+
+\Done
+
+\RC{1.17 Minor points:
 * In the first sentence of the abstract, I believe it should be "The question of whether…" rather than "The question whether…"}
@@
 
-\RC{* Extra spaces between paragraphs should be removed}
+\RC{1.18 * Extra spaces between paragraphs should be removed}
 
@@
 
-\RC{* Miss spelling on page 12: agerage}
+\RC{1.19 * Miss spelling on page 12: agerage}
 
@@
 We intended to say that fathers contribute in a *different* way from mothers.
-We merely intended to summarize fatherhood theory regarding the distinction between the attachment vs activation relationship: A widely endorsed theory that posits that demandingness is a more important socialization mechanism for fathers than for mothers, and that warmth is a more important socialization mechanism for mothers than for fathers.
+We merely intended to summarize fatherhood theory regarding the distinction between the attachment vs activation relationship: An often-cited theory that implies that demandingness is a more important socialization mechanism for fathers than for mothers, and that warmth is a more important socialization mechanism for mothers than for fathers.
 
@@
 
+<!--[RK14]Socialization; children model their parents and see that their parents are also not able to regulate their emotions??-->
+
 \WorkInProgress
@@
 
+<!-- [RK15]The worse children fare, the more parents want to correct them/put them in line? -->
 \WorkInProgress
@@
 
-We appreciate this suggestion!
+<!-- WRONG ANSWER!! [RK17]Let op: dit is niet het goede antwoord op de vraag!!! -->
+
+<!--We appreciate this suggestion!
 Strictly speaking, since Bayesian analyses focus on estimation rather than inference, it is correct to say that the (maximum a-posteriori) estimate is different between mothers and fathers, without further substantiation - because in the Bayesian framework, the maximum a-posteriori estimate can be interpreted as the most likely population value for the parameter, conditional on the data and prior.
@@
 We have omitted it from the smaller Tables in the manuscript, because their function is to help make it easier for readers to digest the results, and we feel that adding four columns with difference tests would interfere with that goal.
+-->
 
@@
 Thus, we must choose: We can either test our theoretical model, which requires Bayesian estimation using Gibbs sampling, or we analyze this proposed non-linear effect, but within a much simpler model that can be estimated using numerical integration.
-Of course, we would prefer to analyze our theoretical model.
+Our choice to analyze our theoretical model also means deciding *against* testing alternative possible models, such as the one proposed by the Reviewer.
 
@@
 We agree with the Reviewer that mean differences in emotional adjustment based on low vs high involvement are important.
-That is why we do investiage these effects, and they are reported in the section titled **Effects of quantitative involvement on parenting _and adjustment_**.
+That is why we do investigate these effects, and they are reported in the section titled **Effects of quantitative involvement on parenting _and adjustment_**.
 That is why we suggested those explanations in the Discussion.

### 2019-11-07 Paper_shortened_pdf.Rmd
Sent to Renske

@@
   The question of whether fathers and mothers play similar roles in children’s
-  emotional adjustment has been approached differently in different research
-  traditions: Developmental psychologists have focused on qualitative parenting,
-  and sociologists have focused on quantitative child care involvement. This
+  emotional adjustment has been addressed using either qualitative parenting 
+  measures or  quantitative child care involvement. This
   longitudinal study (N = 1884, age 6-12) bridges these approaches, 
@@
 It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
-Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of the quantitative amount of care parents provide (who does what), and in terms of qualitative parenting behaviors (how do parents behave towards children).
+Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for the child), and in terms of qualitative parenting behaviors (how do parents behave towards children).
 Nevertheless, the existing literature has two major shortcomings.
-Firstly, quantitative involvement and qualitative parenting behaviors are rarely jointly investigated.
-Therefore, we currently don't know how the interplay between quantitative and qualitative involvement relates to emotional development.
+Firstly, in assessing linkages with children’s emotional adjustment, quantitative involvement and qualitative parenting behaviors are rarely jointly investigated.
+Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
+Children’s emotional adjustment might be differently impacted by parents’  control if that parent is generally highly involved with the day-to-day minutiae of childrearing, than when (s)he is generally absent.
+Conversely, children’s emotional adjustment might elicit different qualitative parenting behaviors from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be sanctioned.
+The present study set out to address this question.
+
 Secondly, although fathers have been theorized to play a unique role in emotional adjustment development, empirical work has rarely addressed the joint and relative contributions of fathers' and mothers' involvement.
-The present study addresses overcomes these limitations by incorporating both quantitative and quantitative measures of mothers' and fathers' involvement.
-We use sophisticated methods to examine how joint and relative quantitative involvement moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment.
-We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment.
+The present study overcomes these limitations by incorporating both quantitative and quantitative measures of mothers' and fathers' involvement.
+We use sophisticated methods to examine how joint and relative quantitative involvement moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
+We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.
 
@@
 By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
-Consequently, fathers and mothers are theorized influence emotional adjustment development in somewhat different ways.
+Consequently, fathers and mothers are theorized to influence emotional adjustment development in somewhat different ways.
 
@@
 It is now widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004].
-Nevertheless, these dimensions are rarely investigated together [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+Nevertheless, these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
 The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
@@
 
-### Parenting effects
-
 Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2018a].
 Qualitative parenting behaviors have been operationalized in terms of the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983].
-Substantial evidence indicates that responsiveness, which involves warm and supportive parenting, is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].
+Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].
 
@@
 Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990].
-Inductive reasoning is thought to promote the internalization of parental expectations, including display rules for emotions, and might thus play a role in emotional adjustment development as well [@grusecImpactParentalDiscipline1994].
+There is less empirical evidence for linkages between inductive reasoning and children’s emotional adjustment.
+Given that inductive inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would be positively related to emotional adjustment.
 
-Negative parenting is an additional qualitative parenting dimension that appears to be negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
-We operationalized negative parenting as parental anger.
-Parental anger is thought to model anger in children, and undermine their emotion regulation [@scaramellaParentingChildrenDistress2008].
+Negative parenting, which we operationalized as parental anger, is an additional qualitative parenting dimension that is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+Parental anger directly undermines children's emotional adjustment [@scaramellaParentingChildrenDistress2008], and models emotional dysregulation [@changHarshParentingRelation2003].
+We thus hypothesized that it would be negatively related to emotional adjustment.
 
-Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' influence [@flouriFathersBehaviorsChildren2010].
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
 Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
@@
 In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage, and respond more supportively to, emotional expression, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-Moreover, mothers' supportive behavior was found to be a more salient socialization mechanism than control for adolescents' emotional adjustment, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
 Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
 Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
-We might therefore expect emotional adjustment to be more strongly associated with maternal warmth, and with paternal demanding behaviors, such as consistency and inductive reasoning.
+We might therefore expect emotional adjustment to be associated more strongly with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning.
 
-### Child effects
+### Parent or child effects?
 
@@
 The available evidence suggests that maladjustment elicits less parental warmth, more demandingness, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @lenguaBidirectionalAssociationsTemperament2005].
-The adolescent literature provides additional support that emotional maladjustment predicts decreased parental support, and greater control [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014; @wijsbroekDirectionEffectsPerceived2011].
+The adolescent literature provides additional support that emotional maladjustment predicts decreased parental support, and greater demandingness [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014; @wijsbroekDirectionEffectsPerceived2011].
 Thus, parents' display punitive responses to maladjustment, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness.
-We might therefore expect emotional adjustment to positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.
+We therefore hypothesize that emotional adjustment will positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.
 
-### Direction of effects
+### Analyzing the direction of effects
 
@@
 
+<!-- Hier iets over de hypotheses; dat we niet weten wat er within en wat between speelt -->
+
 ## Quantitative involvement
@@
 Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006].
-This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality.
+This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].
 
-A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001].
+There are also methodological advantages to focusing on the distribution of childcare.
+Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001].
 This confound can be mitigated by examining the effect of the distribution of childcare, while controlling for total care.
 Some prior research has attempted to achieve this by assessing the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016; @keizerInfluenceFathersMothers2019].
+
 Alternatively, when quantitative involvement has been measured in terms of absolute involvement, one can estimate a latent mean and difference to statistically disentangle joint involvement from relative involvement [@mcardleLatentVariableModeling2009].
-Joint involvement reflects the total amount of care provided to children, and relative involvement reflects the childcare distribution.
+Joint involvement reflects the total (average) amount of care provided to children, and relative involvement reflects the extent to which the mother is more involved than the father (i.e., the childcare distribution).
 
@@
 Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014].
-Consequently, we might expect greater joint involvement to positively predict both parents' parenting quality.
+Consequently, we expect greater joint involvement to positively predict both parents' parenting quality.
 
@@
 Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].
-We might thus expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
+We thus expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
 
@@
 Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
-This suggests that the total amount of care provided should be positively associated with children's emotional adjustment.
-Differential involvement, on the other hand, might be negatively linked to emotional adjustment: When children perceive that one of their parents is relatively uninvolved, they likely feel less supported than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
+This implies that joint care should be positively associated with children's emotional adjustment.
+Relative involvement, on the other hand, might be negatively linked to emotional adjustment, because children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
 
@@
 
-Interactions between involvement quality and quantity have largely been overlooked in theoretical models of father involvement; perhaps in part because these two dimensions have rarely been jointly investigated [@jaffeeLifeFatherBenefits2003].
-In this respect, theories of father involvement can be complemented by Feinberg's coparenting model [@feinbergInternalStructureEcological2003].
-Coparenting refers to the ways parents relate to each other in the role of parent.
+Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003].
+One notable exception is Feinberg's coparenting model [@feinbergInternalStructureEcological2003].
+Coparenting refers to the ways parents relate to each other in the parenting role.
 Although the coparenting literature is typically invoked to describe child-rearing in divorced families, it also holds insights for intact families.
-According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of differential involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs.
+According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of relative involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs.
 
@@
 Sex differences are a relevant factor in the interplay between parenting and emotional adjustment.
-Parents tend to raise boys and girls differently; expressing more control with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more negative parenting with boys [@strausCorporalPunishmentAmerican1999].
+Parents tend to raise boys and girls differently; expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999].
 Furthermore, social learning theory posits that the same-sex parent plays a stronger role in socialization; particularly for sex-typed behavior, such as emotion display rules [@busseySocialCognitiveTheory1999].
-Finally, mounting evidence suggests that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
-We therefore hypothesize that parents will display more responsiveness, structure, and inductive reasoning towards girls, and more negative parenting towards boys.
-Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and the same-sex parent.
+Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
+We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning towards girls, and more anger towards boys.
+Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent.
 
-Although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a].
+Furthermore, although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a].
 Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.
@@
 First, we include both mothers and fathers, in order to tease apart their common and unique roles in children's emotional adjustment development.
-Second, we incorporate both qualitative and quantitative measures of involvement.
+Second, we incorporate both qualitative and quantitative measures of involvement, and examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
 Third, we disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative involvement and emotional adjustment.
-Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
 Finally, we examined sex as a moderator of between-family differences in all study variables and within-family predictive effects.
@@
 Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001].
-Structure and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
+Consistency and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
 These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".
@@
 Thus, on average, mothers' warmth had a positive significant predictive effect on children's emotional adjustment.
-The variances of all parenting effects were significant, however, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
+The variances of all parenting effects were all significant, however, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
 Moreover, these variances were notably larger for fathers than for mothers, indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
@@
 tabparenteff <- rbind(tab_std_sections$`Intercepts of parenting effects`, 
-                      tab_std_sections$`Between-family level residual variances`[c(1, 5:6),])
+                      tab_unst_sections$`Between-family level residual variances`[c(1, 5:6),])
 
@@
 tabchildeff <- rbind(tab_std_sections$`Intercepts of child effects`,
-                      tab_std_sections$`Between-family level residual variances`[c(1, 7:8),])
+                      tab_unst_sections$`Between-family level residual variances`[c(1, 7:8),])
 
@@
 rownames(tab) <- NULL
-reg_table(tab, caption = "Between-family level intercepts and variances of within-family predictive effects", note = "Est.: Estimate.
+tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
+reg_table(tab, caption = "Standardized between-family level intercepts and unstandardized variances of within-family predictive effects", note = "Est.: Estimate.
 HDPI: Highest posterior density interval.
@@
 With regard to child effects, in line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)).
-Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger and inductive reasoning, but also more warmth, than usual in the next year.
+Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more warmth than usual in the next year.
 Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant.
 As hypothesized, the intercept for maternal inductive reasoning was positive.
-Finally, counter to hypotheses, the intercept for maternal anger was positive, not negative.
+Counter to hypotheses, the intercept for maternal anger was positive, not negative.
+Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger than usual in the next year.
 In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant.
 None of the other intercepts for child effects on paternal parenting were significant.
-The variances of all child effects were significant, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs between families.
+The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
 
@@
 With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
-Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when both parents' joint quantitative involvement was higher.
-This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement.
+Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when joint involvement was higher, and positively attenuated when mothers were relatively more involved than fathers.
 We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.
@@
 For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when joint involvement was higher.
-Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
+Emotional adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
 Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning.
@@
 
-We found sex differences in the random intercepts of qualitative parenting behaviors: Both mothers and fathers displayed less warmth, and more inductive reasoning and anger, towards sons as compared to daughters  (Table \@ref(tab:tabsex)).
+We found sex differences in the random intercepts of qualitative parenting behaviors: Fathers displayed less warmth, and both parents displayed more inductive reasoning and anger, towards sons as compared to daughters  (Table \@ref(tab:tabsex)).
 No significant differences were found for maternal warmth, and for both parents' consistency.
@@
 Specifically, emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls.
-In other words, if children display better adjustment than usual in one year, their parents decrease inductive reasoning more strongly in the next year when children are male.
+In other words, if children display worse emotional adjustment than usual in one year, parents increase inductive reasoning more strongly for boys than for girls.
 
@@
 
-There was partial but unequivocal support for the hypothesized positive associations between warmth and emotional adjustment.
-These links were absent at the between-family level, but abundant at the within-family level - for instance, in the form of within-family correlations with both parents' warmth.
-The random intercept for the parenting effect of maternal warmth was significant, suggesting that this parenting effect occurs in most families.
-Finally, child effects on both parents' warmth indicated that well-adjusted children evoke greater warmth.
+In terms of between-family level correlations, we found partial support for our hypotheses.
+We did not find the hypothesized positive associations between parental warmth and emotional adjustment.
+The predicted positive associations of parental consistency with emotional adjustment did emerge, indicating that children in families with highly consistent parents tended to be more emotionally well-adjusted.
+Counter to hypotheses, parents who engaged in more inductive reasoning had children with poorer emotional adjustment.
+Finally, for anger, we found the predicted negative associations, which means that, in families where mothers or fathers display high anger, children tend to be less emotionally well-adjusted.
 
-We found partial support for positive associations of parental consistency with emotional adjustment, primarily at the between-family level: children in families with highly consistent parents tended to be more emotionally well-adjusted.
-We found the predicted within-family correlations for mothers', but not fathers', consistency.
-On average, we found no support for within-family predictive links between consistency and adjustment, but we did find support for between-family level correlations.
-Between-family level correlations can have different causes: They can reflect the consolidated influence of predictive effects that occurred prior to the study, or alternatively, the influence of third variables that impact both constructs.
-The fact that our study included children from a very young age casts doubt on the former explanation.
-It seems more likely that third variables explain these associations.
-Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
-This finding means that we should reflect critically on causal interpretations of prior research that did not control for stable between-family differences [e.g., @brandPerceivedParentalInconsistency1990].
+In terms of parenting effects, we found only one effect that was significant on average, i.e., its random intercept was significant and positive: Maternal warmth positively predicted emotional adjustment in most families.
+However, all within-family predictive effects had significant variance components, which suggests that the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.
+Variability was notably greater for fathers, indicating that fathers' influence differs most across families.
 
-With regard to inductive reasoning, we found mixed support for the hypothesized positive associations with emotional adjustment: Associations were opposite in sign at the between- versus within-family levels.
-At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment.
+Regarding child effects, we found that, on average, well-adjusted children evoked greater warmth from both parents, as hypothesized.
+Counter to predictions, we did not find average child effects for consistency.
+In partial support of our hypothesis, we found that children who displayed better emotional adjustment in one year evoked greater inductive reasoning from mothers in the next year.
+But paradoxically, and counter to oru hypothesis, we also found that children who displayed better emotional adjustment in one year evoked greater maternal anger in the next year.
+
+In sum, we found mixed evidence for our hypotheses that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
+
+## Direction of effects
+
+We set out to investigate the direction of effects between parents and children, using innovative methods that control for stable between-family differences, in order to obtain more accurate estimates of within-family effects [@hamakerCritiqueCrosslaggedPanel2015].
+We found ample evidence for within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
+We found no between-family level correlations for warmth.
+This suggests that warmth has an ongoing positive within-family effect on emotional adjustment.
+For consistency, however, we found no within-family predictive effects, but only between-family level associations.
+This suggests that the benefits for emotional adjustment associated with greater consistency might be primarily due to stable between-family differences.
+
+For inductive reasoning, paradoxically, between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment.
 This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior - albeit a constructive one [@grusecImpactParentalDiscipline1994].
-Parents' reports of engaging in inductive reasoning might thus be confounded with how often they discipline their children.
-At the within-family level, by contrast, paternal inductive reasoning was positively correlated with emotional adjustment, and emotional adjustment had a positive predictive effect on maternal inductive reasoning.
-The fact that between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment might suggest that between-family differences in inductive reasoning tap more strongly into the amount of discipline parents engage in, whereas within-family fluctuations reflect changes in the way parents discipline their children.
+When parents report engaging in more inductive reasoning, this introduces a confound between *how often* they discipline their children, and *how* they discipline them.
+The fact that between-family correlations between inductive reasoning and emotional adjustment were negative suggests that between-family differences in inductive reasoning might tap more strongly into how often parents engage in discipline.
+The within-family fluctuations, by contrast, might reflect changes in how parents discipline their children.
 
-With regard to parental anger, we found mixed evidence for the predicted negative associations with emotional adjustment.
-These associations existed at the between-family level only.
-This means that, in families where mothers or fathers display low anger, children tend to be more emotionally well-adjusted.
+For anger, we found moderately strong between-family level correlations for both parents.
 Counter to our expectations, however, we found no evidence for within-family correlations or parenting effects.
-This suggests that the brunt of negative associations between anger and adjustment are explained by differences between families.
-One potential explanation for the absence of within-family associations might be a ceiling effect; it might be the case that, once the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger no longer make much of a difference.
-Another explanation might be that any further effects of fluctuations in anger play out on a different time scale than the one captured by our study; for example, if such effects primarily play out immediately, in dyadic interactions between parents and children.
-Finally, we found that, on average, emotional adjustment showed a paradoxical positive predictive effect on maternal anger.
+This suggests that the brunt of negative associations between anger and adjustment is explained by between-family differences.
+One potential explanation for the absence of within-family associations might be a ceiling effect: If the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger might no longer make much of a difference.
+We also found a paradoxical positive predictive effect of emotional adjustment on maternal anger.
 This means that, if children demonstrated better emotional adjustment than usual in one year, their mothers tended to display more anger than usual in the next year.
@@
 
-In sum, we found mixed evidence for our hypotheses that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
-Importantly, all within-family predictive effects had significant variance components, highlighting the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.
+To recapitulate, we set out to examine the direction of within-family effects by controlling for stable between-family differences.
+We found only limited evidence for within-family effects, most of which were child effects on parents.
+Instead, we mostly found evidence for between-family level correlations.
+We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016].
+In the area of parenting and emotional adjustment development, we are aware of only one prior study that also controlled for stable between-family differences.
+That study also reported mostly child effects and between-family level corelations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
+They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
+The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of children from age 6-12 suggests that either parent effects occurred at an even younger age, or that something else is at play.
+
+As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
+We should consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
+Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
+Our findings also serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general.
+
+Future research might also devote greater attention to between-family level correlations as an outcome of interest.
+Between-family level correlations might reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they might reflect the influence of third variables that impact both constructs.
+Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors such as economic hardship and neighborhood effects [@dengFamilyProcessesMediating2006], which are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.
+
+Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
+At the same time, we should consider the limitations of these methods.
+The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
+Stable factors can also exert changing influences over time.
+As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time.
+It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
+Another important limitation is that within-family cross-lagged effects are time scale dependent.
+If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
+Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
+
+## Differences between fathers and mothers
+
+Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development.
+These differences all played out at the within-family level, however.
+Stable between-family differences in either parent's qualitative parenting were all similarly associated with children's adjustment.
+
+At the within-family level we found stronger associations with emotional adjustment for maternal consistency, and for paternal reasoning and warmth.
+As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample.
+This might help explain the stronger associations for maternal consistency - a behavior which implies, by definition, consistent repeated interactions with the child.
+For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can also be expressed in one-off interactions, such as when  disciplining a child (inductive reasoning), or during play time (warmth).
+The stronger association between inductive reasoning and emotional adjustment for fathers is in line with the classic view of the father as disciplinarian [@flouriFathersBehaviorsChildren2010].
+
+With regard to parenting effects, the only difference between mothers and fathers was for the predictive effect of warmth on emotional adjustment, which was evident only for mothers.
+This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].
+With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting.
+Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.
 
@@
 Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs.
-Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.
+Factors related to children's objective needs - such as developmental delay, or atypical development - might drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, resulting in a negative correlation between the two.
 
@@
 Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.
-
-## Differences between fathers and mothers
-
-Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development.
-These differences all played out at the within-family level, however.
-Stable between-family differences in either parent's qualitative parenting were all similarly associated with children's adjustment.
-
-At the within-family level we found stronger associations with emotional adjustment for maternal consistency, and for paternal reasoning and warmth.
-As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample.
-This might help explain the stronger associations for maternal consistency - a behavior which implies, by definition, consistent repeated interactions with the child.
-For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can also be expressed in one-off interactions, such as when  disciplining a child (inductive reasoning), or during play time (warmth).
-The stronger association between inductive reasoning and emotional adjustment for fathers is in line with the classic view of the father as disciplinarian [@flouriFathersBehaviorsChildren2010].
-
-With regard to parenting effects, the only difference between mothers and fathers was for the predictive effect of warmth on emotional adjustment, which was evident only for mothers.
-This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].
-With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting.
-Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.
-
-## Direction of effects
-
-Nowadays, it is commonly acknowledged that parents and children reciprocally influence one another.
-However, the methods used to demonstrate bi-directional effects have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015].
-We are aware of only one prior study that controlled for stable between-family differences when examining the interplay between parenting and emotional adjustment.
-This study found that child effects substantially outnumbered parenting effects in adolescence [@vanlissaRoleFathersMothers2018a].
-The authors explained this finding in light of the relative equality of parent-adolescent relationships.
-The present study casts doubt on that interpretation, as we found a preponderance of child effects between the ages of 6-12.
-Although it is, of course, possible that parent effects are in effect at an even younger age, we can also consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
-
-The finding that parent effects become smaller or non-significant once stable between-family differences are accounted for has also been demonstrated in other areas of developmental research (REF).
-As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
-Such findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
-Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
-
-At the same time, we should consider the limitations of these methods.
-The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
-Stable factors can also exert changing influences over time.
-As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time.
-It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
-Another important limitation is that within-family cross-lagged effects are time scale dependent.
-If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
-Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
 
@@
 Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range.
-
-<!--  -->
+A final potential limitation pertains to the generalizability of the findings.
+Our target sample consisted of native `r masked("Australian")`, non-minority, intact, biological families.
+It is possible, but beyond the scope of this study, that the findings might have limited generalizability for Australian minorities, or for families in other industrialized nations.
 

### 2019-11-07 invariance/table_coefficients.R

@@ -110,4 +110,15 @@ starts = sections[-length(sections)],
 ends = sections[-1], SIMPLIFY = FALSE)
 names(tab_std_sections) <- tab_std$Parameter[sections][-length(sections)]
 
-save(tab_std, tab_unst, tab_std_sections, file = "table_coefficients.RData")
+
+# Split UNST --------------------------------------------------------------
+
+sections <- c(which(rowSums(is.na(tab_unst)) == 8), (nrow(tab_unst)+1))
+tab_unst_sections <- mapply(function(starts, ends) {
+  tab_unst[starts:(ends-1),]
+},
+starts = sections[-length(sections)],
+ends = sections[-1], SIMPLIFY = FALSE)
+names(tab_unst_sections) <- tab_unst$Parameter[sections][-length(sections)]
+
+save(tab_std, tab_unst, tab_std_sections, tab_unst_sections, file = "table_coefficients.RData")

### 2019-11-08 table_coefficients.R
Add additional tests and re-make coefficient tables

@@ -1,0 +1,149 @@
+library(MplusAutomation)
+library(motley)
+source("rfunctions.R")
+if(!file.exists("results.RData")){
+  results <- readModels("./Mplus", filefilter = "revision")
+  save(results, file = "results.RData")
+} else {
+  load("results.RData")
+}
+
+order_rows <- c("Between-family level correlations",
+                "MOTH.WITH.EMO", "FATH.WITH.EMO", "MOTH.WITH.FATH",
+                "Within-family level correlations",
+                "EMO.WITH.MOTH.Within", "FATH.WITH.EMO.Within", "FATH.WITH.MOTH.Within",
+                "Intercepts of parenting effects",
+                "Intercepts.SEMOMOTH", "Intercepts.SEMOFATH",
+                "Intercepts of child effects",
+                "Intercepts.SMOTHEMO", "Intercepts.SFATHEMO",
+                "Effects of quantitative involvement on parenting and adjustment",
+                "MOTH.ON.MEAN", "FATH.ON.MEAN", "EMO.ON.MEAN", 
+                "MOTH.ON.MMORE", "FATH.ON.MMORE", "EMO.ON.MMORE", 
+                "Moderating effect of quantitative involvement on parenting effects",
+                "SEMOMOTH.ON.MEAN", "SEMOFATH.ON.MEAN",
+                "SEMOMOTH.ON.MMORE", "SEMOFATH.ON.MMORE",
+                "Moderating effect of quantitative involvement on child effects",
+                "SMOTHEMO.ON.MEAN", "SFATHEMO.ON.MEAN",
+                "SMOTHEMO.ON.MMORE", "SFATHEMO.ON.MMORE",
+                "Sex differences in parenting and adjustment",
+                "MOTH.ON.MALE", "FATH.ON.MALE", "EMO.ON.MALE", 
+                "Sex differences in cross-lagged effects",
+                "SEMOMOTH.ON.MALE", "SEMOFATH.ON.MALE",
+                "SMOTHEMO.ON.MALE", "SFATHEMO.ON.MALE",
+                "Within-family level residual variances",
+                "Residual.Variances.MOTH.Within", "Residual.Variances.FATH.Within", "Residual.Variances.EMO.Within",
+                "Between-family level residual variances",
+                "Residual.Variances.MOTH", "Residual.Variances.FATH", "Residual.Variances.EMO", 
[…]

### 2019-11-08 Action letter.Rmd
Contrasts in table

@@
 We hope that you will find that our responses adequately address the issues raised, and that the quality of the work has been improved.
+
+With regard to the Editorial Office's request that we upload a Word or RTF file: In line with the Submission Guidelines on the Developmental Psychology website, we have provided \latex as a `zip`-file, with accompanying `.pdf`.
+We do not have a Word version of this manuscript available.
 
@@
 
-Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between mothers and fathers.
-We have reported this information in-text whenever making a statement about mother/father differences, and have added it to the relevant Supplementary Tables with the full model parameters.
-We have omitted it from the smaller Tables in the manuscript, because their function is to help make it easier for readers to digest the results, and we feel that adding four columns with difference tests would interfere with that goal.
+Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between corresponding parent- and child effects.
+We have explained this procedure in the text (fragment cited below), and have added it to Table 2 in the form of labels that indicate which parameters differed significantly, and we have reported the full difference tests in Supplementary Table S5.
+
+<!--The interpretation of the difference tests is straightforward, because they exactly mirror the pattern of significant/non-significant parent and child effects: All effects that are significant in one direction, but non-significant in the reverse direction, have significant difference tests.
+All effects that are significant in both directions, or non-significant in both directions, have non-significant difference tests.
+It is therefore justifiable to simply interpret the pattern of significant parent- and child effects.
+
+Section in Results:
+
+> We established direction of effects by testing the significance of the difference between corresponding parent- and child effects (see Table \@ref(tab:tabcrosspath) and Supplementary Table S5).
+All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests.
+All effects that were significant (or non-significant) in both directions had non-significant difference tests.
+
+Table note:
+
+> Matching letters indicate that parameters differed significantly, see Supplementary Table S5.-->
 
@@
 
+We appreciate this suggestion!
+Strictly speaking, since Bayesian analyses focus on estimation rather than inference, it is correct to say that the (maximum a-posteriori) estimate is different between mothers and fathers, without further substantiation - because in the Bayesian framework, the maximum a-posteriori estimate can be interpreted as the most likely population value for the parameter, conditional on the data and prior.
+In practice, since we use Bayesian estimation for pragmatic reasons, we understand that there is added value in knowing the posterior distribution of the difference between mothers and fathers.
+
+Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between corresponding parent- and child effects.
+We have explained this procedure in the text (fragment cited below), and have added it to Table 2 in the form of labels that indicate which parameters differed significantly, and we have reported the full difference tests in Supplementary Table S5.
+
+The interpretation of the difference tests is straightforward, because they exactly mirror the pattern of significant/non-significant parent and child effects: All effects that are significant in one direction, but non-significant in the reverse direction, have significant difference tests.
+All effects that are significant in both directions, or non-significant in both directions, have non-significant difference tests.
+It is therefore justifiable to simply interpret the pattern of significant parent- and child effects.
+
+Section in Results:
+
+> We established direction of effects by testing the significance of the difference between corresponding parent- and child effects (see Table \@ref(tab:tabcrosspath) and Supplementary Table S5).
+All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests.
+All effects that were significant (or non-significant) in both directions had non-significant difference tests.
+
+Table note:
+
+> Matching letters indicate that parameters differed significantly, see Supplementary Table S5.
+
 <!-- WRONG ANSWER!! [RK17]Let op: dit is niet het goede antwoord op de vraag!!! -->
@@
 
-\WorkInProgress
+\Done
 

### 2019-11-08 Paper_shortened_pdf.Rmd
Contrasts in table

@@
 
-At the within-family level, we examined over-time cross-lagged effects between qualitative parenting and emotional adjustment.
-The intercepts of these random effects reflect the average cross-lagged effect across families.
+At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment.
+The intercepts of these random effects reflect the average effect across families.
 Their variances indicate whether the effect differs in strength (and sign) between families.
-
 With regard to within-families parenting effects, none of the intercepts differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment (Table \@ref(tab:tabcrosspath)).
@@
 rownames(tab) <- NULL
+
+# Add letters for contrast tests
+
+stars <- tab_unst_sections$`Defined parameters`[tab_unst_sections$`Defined parameters`$Parameter %in% c("Mother effect - child effect on mother", "Father effect - child effect on father"), ]
+letter <- 1
+tabrows <- list(c(2,8),
+                c(3, 9))
+
+for(i in 1:length(tabrows)){
+  if(any(grepl("\\*", stars[i, ]))){
+    for(c in which(grepl("\\*", stars[i, ]))){
+      # c = 6
+      tab[tabrows[[i]], c] <- paste0(tab[tabrows[[i]], c], letters[letter])
+      letter <<- letter + 1
+    }
+  }
+}
+
 tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
@@
 Emo.: Emotional adjustment.
-*: <5% of posterior has opposite sign from estimate.")
+*: <5% of posterior has opposite sign from estimate.
+Matching letters indicate that parameters differed significantly, see Supplementary Table S5.")
 ```
@@
 The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
+
+We established direction of effects by testing the significance of the difference between corresponding parent- and child effects (see Table \@ref(tab:tabcrosspath) and Supplementary Table S5).
+All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests.
+All effects that were significant (or non-significant) in both directions had non-significant difference tests.
+
+We examined differences between fathers and mothers the same way. <!-- NOG DOEN -->
 

### 2019-11-12 Action letter.Rmd
Final changes before introduction

--- 2019-11-07 Action letter.Rmd
@@
 
-\subsection*{NOTE FROM EDITORIAL OFFICE:  We ask that you upload either Word or RTF versions of both your title page and manuscript,  rather than PDFs, when resubmitting.}
-
 Dear Dr. `r editor[3]`,
@@
 
-We thank you and the Reviewers for your thorough comments, and for the helpful summary you've provided.
-We hope that you will find that our responses adequately address the issues raised, and that the quality of the work has been improved.
+We thank you and the Reviewers for your thorough comments, and for your helpful summary of these comments.
+We have attempted to address all comments, and hope that you will find that the quality of the work has been improved.
 
-<!--The revised manuscript (including title page, abstract, and references) comprises 35 pages, or 8666 words.
-Four supplementary tables are included, whose size prohibits them from being included in the manuscript: A full correlation table (total sample and paneled by adolescent sex); the full matrix of parameter estimates; detailed scale descriptives and intercorrelations, and a full matrix of parameter estimates for an alternative analysis suggested by a Reviewer.
-Should the manuscript be accepted, we are prepared to make all syntax and results of the analyses conducted available on a public code repository referenced in the paper, to safeguard reproducibility.
-Treatment of participants was in accordance with the APA ethical principles.
-On behalf of the authors, I declare that I have no conflicts of interest concerning this article.
-I will serve as the corresponding author for this manuscript.-->
+With regard to the Editorial Office's request that we upload a Word or RTF file: In line with the Submission Guidelines on the Developmental Psychology website, we have provided \latex as a `zip`-file, with accompanying `.pdf`.
+We do not have a Word version of this manuscript available.
+We could manually convert this to Word, but the typographical advantages of using Latex would then be lost.
+Please let us know how to proceed.
 
@@
 
-We thank the Editor for pointing out the lack of consistency across sections, and we have edited the revised manuscript to resolve this issue.
+We thank the Editor for pointing out the lack of consistent terminology across sections, and we have edited the revised manuscript to resolve this issue.
 We now consistently use the terms "joint and relative quantitative involvement".
@@
 
-We appreciate the suggestion, and have incorporated such a figure.
+We appreciate the suggestion, and have incorporated this Figure, see below:
+
+```{r figure1, out.width="100%", fig.cap="T-1: previous measurement occasion.
+ri: random intercept.
+s: random slope.
+M: Mother's qualitative involvement.
+E: Emotional adjustment.
+F: Father's qualitative involvement.
+inv: quantitative involvement.
+Thus, sME represents the random slope of Mothers' qualitative involvement on emotional adjustment."}
+knitr::include_graphics("Figure1.png")
+```
 
@@
 
-We have elaborated on the Table notes.
-We think that the definition of joint and relative involvement is adequately addressed by the increased consistency with regard to terminology, in conjunction with the new Figure - but we are happy to take further action if necessary.
+As requested, we have elaborated and clarified the Table notes in the revision.
+Regarding the definition of joint and relative involvement - we believe that this is adequately clarified by the consistent use of terminology in the revised manuscript, and the new Figure describing the model specification - but we are happy to take further action if necessary.
 
-\WorkInProgress
+\Done
 
@@
 
-\WorkInProgress
+We have attempted to address this comment by adding extensive Table notes and clarifying in-table text throughout the *Results* section.
+We assume that readers are familiar with the interpretation of correlations, and have therefore not further clarified the interpretation of correlations.
+With regard to the standardized regression coefficients (and their variance), we have included an elaborate explanation of the interpretation of the first standardized regression coefficient reported for readers not familiar with standardized regression coefficients.
+For the sake of saving space, we do not interpret each coefficient so elaborately (we prefer to stick with the APA publication manual, and report results either in-text or in a table, but not both).
+We explain the basic random effects once, and the moderated effects once.
+
+The added sections for the random effects reads:
+
+> At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment.
+The intercepts of these random effects reflect the average standardized regression coefficient across families.
+The variances in Table 2 indicate whether this coefficient differs in strength (and sign) between families.
+We report unstandardized variances, because the standardized variances are, by definition, set to 1.
+With regard to within-families parenting effects, we found no evidence for hypotheses $H_{1b-d}$, as none of the intercepts differed significantly from zero.
+In line with hypothesis $H_{1a}$, however, on average, a maternal warmth positively predicted children's emotional adjustment in the next wave.
+The standardized regression coefficient of this effect, and others, is interpreted as follows: A $.99$ standard deviation increase in children's emotional adjustment in the next wave (see Table 2).
+
+The added section for the moderation effects reads:
+
+> To examine interactions between involvement quality and quantity, we examined standardized regression coefficients of between-family differences in quantitative involvement on within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table 3).
+With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
+Specifically, a one-standard deviation increase in joint involvement predicted a decrease in the within-families predictive effect of maternal warmth on emotional adjustment of $-.14$ of a standard deviation of the effect (see Table 2).
+
+\Done
 
@@
 their parenting will be?
-Do you expect this to be equally true for mothers and fathers?
-How do you expect the quantitative measure of parenting to moderate the association between qualitative measures and children's emotional adjustment?
-}
+Do you expect this to be equally true for mothers and fathers?}
 
-\WorkInProgress
+We have attempted to address this comment by clarifying and labelling our hypotheses, and re-structuring the *Introduction*, *Results*, and *Discussion* sections to follow the order of these hypotheses.
+The hypotheses are:
+
+* $H_{1a-d}$ Warmth, consistency, and inductive reasoning will be positively, and anger negatively, associated with children's emotional adjustment
+* $H_{2a-d}$ Emotional adjustment will positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger
+* $H_{3}$ Emotional adjustment will be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning
+* $H_{4a-d}$ Greater joint involvement will positively attenuate both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively attenuate parental anger
+* $H_{5a-d} Greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively
+* $H_{6a}$ Greater joint care should positively predict children's emotional adjustment
+* $H_{6b}$ Greater relative involvement will negatively predict emotional adjustment
+* $H_{7a-d}$ Parents will display more warmth, consistency, and inductive reasoning towards girls than boys, and more anger towards boys than girls
+* $H_{8a-d}$ Associations between qualitative parenting and emotional adjustment will be stronger between children and their same-sex parent
+
+\Done
+
+\RC{3e How do you expect the quantitative measure of parenting to moderate the association between qualitative measures and children's emotional adjustment? }
+
+Because such interactions have not been examined before, any hypothesis would be unsubstantiated. Thus, our work in this area is exploratory. We now acknowledge this explicitly:
+
+> Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003].
+Consequently, it is not possible to substantiate specific hypotheses for such interactions.
+[...] We thus explored the moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.
+
+\Done
 
@@
 
-\WorkInProgress
+As mentioned in our response to the Editor's comment 3d, we have now re-written several sections of the paper, including the *Discussion*, to follow the same structure, guided by the (numbered) hypotheses.
+
+\Done
 
@@
 
-To address this comment, we've attempted to ensure that every issue in the Introduction is paralleled in the Discussion.
-We have also attempted to clarify the between/within distinction.
+This issue is also partly addressed by our efforts to ensure that the structures of the *Introduction*, *Results* section, and *Discussion* are parallel.
+We have also attempted to clarify why it is important to use methods that disentangle between/within variance:
+
+> Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if stable between-family differences are not controlled [@hamakerCritiqueCrosslaggedPanel2015].
+Innovative methods provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].
+
 A major contribution of the present study is that we are (one of?) the first to make this distinction in this particular research area.
 Consequently, there is no prior literature available to base specific hypotheses about between/within effects on.
-The best we can do is develop exploratory hypotheses based on a conflicting literature that has lumped together between/within effects.
-Thus, we cannot make more specific predictions than we currently do.
+The best we can do is develop exploratory hypotheses based on literature that has lumped together between/within effects.
+In the revision, we have attempted to explain why we cannot make (more) specific predictions about when we expect to see between-family effects or within-family effects:
 
-\WorkInProgress
+> Several studies have re-examined the direction of effects between parents and children using these methods [...] These two studies also illustrate that applying these new methods changes the pattern of results in unexpected ways.
+Consequently, it is impossible to make specific predictions about the pattern of between-family correlations and within-family predictive effects.
+For now, we will assume that $H_{1a-d}$ and $H_{2a-d}$ apply at both the between- and within-family level.
+
+\Done
 
@@
 
-We appreciate this suggestion.
-We agree that parent effects could very well be present at even younger ages, and explicitly acknowledge this in the revision.
-At the same time, we think that it is interesting and potentially meaningful that parent effects - but not child effects - tend to become smaller/non significant when stable between-family differences are accounted for.
-We have attempted to clarify our discussion of the potential theoretical/methodological implications of this finding.
+We appreciate this suggestion. In the revision, we explicitly acknowledge that parent effects could very well be present at even younger ages:
 
-> Nowadays, it is commonly acknowledged that parents and children reciprocally influence one another.
-However, the methods used to demonstrate bi-directional effects have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015].
-We are aware of only one prior study that controlled for stable between-family differences when examining the interplay between parenting and emotional adjustment.
-This study found that child effects substantially outnumbered parenting effects in adolescence [@vanlissaRoleFathersMothers2018a].
-The authors explained this finding in light of the relative equality of parent-adolescent relationships.
-The present study casts doubt on that interpretation, as we found a preponderance of child effects between the ages of 6-12.
-Although it is, of course, possible that parent effects are in effect at an even younger age, we can also consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
-The finding that parent effects become smaller or non-significant once stable between-family differences are accounted for has also been demonstrated in other areas of developmental research (REF).
-As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
-Such findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
-Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
-At the same time, we should consider the limitations of these methods.
+> In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
+They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
+The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that __either parent effects occurred at an even younger age__, or that something else is at play.
+
+\Done
+
+\RC{4d.
+Second, there are pronounced individual differences in how parents react to children.
+The same level of emotionally dysregulated behavior might elicit anger from one parent and inductive reasoning from another.
+There is a great deal of evidence from parenting interventions that how parents react to difficult behavior affects whether their children remain noncompliant and dysregulated or whether their behavior improves.}
+
+We agree, and as mentioned in the manuscript, we also find these between-family differences in parents' responses to childen's emotional adjustment in the present study:
+
+> The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
+[...] The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
+
+\Done
+
+\RC{4e. So I think you're selling parents short if your claim is that the only way they influence their children is through the changing expression of the genes they have bequeathed them.}
+
+This is certainly not a claim we intend to make, and we have rewritten the paragraph in question to clarify that we are not making any claims about genes, but instead, we are questioning the assumption of the RI-CLPM model that all reciprocal influences are captured by within-family fluctuation.
+We then introduce genes as an illustrative example of something very stable that can nonetheless have changing influences over the life course.
+In the revision, we have emphasized that this is an illustrative example, and we then draw a parallel to how stable psychosocial traits might similarly have changing effects over the course of development:
+
+> At the same time, we should consider the limitations of these methods.
 The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
@@
 It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
-Another important limitation is that within-family cross-lagged effects are time scale dependent.
-If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
-Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
-
-\WorkInProgress
-
-\RC{Second, there are pronounced individual differences in how parents react to children.
-The same level of emotionally dysregulated behavior might elicit anger from one parent and inductive reasoning from another.
-There is a great deal of evidence from parenting interventions that how parents react to difficult behavior affects whether their children remain noncompliant and dysregulated or whether their behavior improves.
-So I think you're selling parents short if your claim is that the only way they influence their children is through the changing expression of the genes they have bequeathed them.}
-
-We agree, and as we mention in the manuscript, we also find these between-family differences in parents' responses to childen's emotional adjustment in the present study.
-In the rewritten paragraphs above, we have attempted to clarify that we are not making any claims about genes, but instead, we are questioning the assumption of the RI-CLPM model that all reciprocal influences are captured by within-family fluctuation.
-We then introduce genes as an illustrative example of something very stable that can nonetheless have changing influences over the life course.
-In the revision, we have emphasized that this is an illustrative example, and we then draw a parallel to how stable psychosocial traits might similarly have changing effects over the course of development.
 
@@
 
+\WorkInProgress
+
 \RC{1.2b ...describe a more complex, well-cited definition of emotional adjustment (the current citation and definition does not seem accurate)...}
 
-We operationalized emotional adjustment in terms of the emotional adjustment subscale of the SDQ.
-Consequently, we cite the scale validation paper of the SDQ here.
-It is not evident to us what other reference would be more appropriate than the one we used to operationalize this construct, but if the Reviewer feels that a particular citation should be included, we are happy to do so.
+We operationalized emotional adjustment in terms of the emotional adjustment subscale of the SDQ, and cite the SDQ validation paper.
+It is not evident to us what other reference would be more appropriate than the one we used to operationalize this construct, but if the Reviewer feels that specific literature should be cited, we are happy to do so.
 
-\RC{1.2c ...introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers.
-Finally, it would help the reader to provide the research goals in the first 1-2 pages of the introduction to get a sense where the paper is headed.}
+\Done
+
+\RC{1.2c ...introduce the idea that mothers and fathers may plan a unique role in the development of emotional adjustment, before focusing in on the role of fathers...}
+
+\WorkInProgress
+
+\RC{1.2d ...Finally, it would help the reader to provide the research goals in the first 1-2 pages of the introduction to get a sense where the paper is headed.}
 
@@
 
-\WorkInProgress
+We appreciate the suggestion; indeed, the idea that both quantitative and qualitative involvement are important (and might interact) is most important.
+The rough division along disciplinary lines is interesting, but not central to the argument.
+We have therefore edited our argumentation to focus on the importance of jointly investigating quantitative and qualitative involvement, and removed all references to disciplinary differences.
+
+\Done
 
@@
 
-Thank you for pointing out that sex differences are omitted from the list of contributions mentioned in *The Present Study*.
-We have attempted to address this issue by adding sex differences to the list, and we have also clarified our specific hypotheses regarding sex differences in the preceding paragraph.
+This issue is partly addressed by our efforts to ensure that the structures of the *Introduction*, *Results* section, and *Discussion* are parallel.
+Furthermore, we have added sex differences to the list of contributions in *The present study* (*"we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects"*), and we have also formulated our specific hypotheses regarding sex differences in the preceding paragraph.
 
-> We therefore hypothesize that parents will display more responsiveness, structure, and inductive reasoning towards girls, and more negative parenting towards boys.
-Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and the same-sex parent.
+> We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning towards girls than boys, and more anger towards boys than girls, $H_{7a-d}$.
+Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent, $H_{8a-d}$.
 *[...]*
 Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.
-*[...]*
-Finally, we examined sex as a moderator of between-family differences in all study variables and within-family predictive effects
 
@@
 
-We agree wholeheartedly with this point, and we had formulated hypotheses for all patterns studied, but we understand that our colloquial use of language made it hard to identify our specific hypotheses.
-For example, in the original manuscript, we wrote:
+We have attempted to address this comment by clarifying and labelling our hypotheses, and re-structuring the *Introduction*, *Results*, and *Discussion* sections to follow the order of these hypotheses.
+The hypotheses are:
 
-> Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.
+* $H_{1a-d}$ Warmth, consistency, and inductive reasoning will be positively, and anger negatively, associated with children's emotional adjustment
+* $H_{2a-d}$ Emotional adjustment will positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger
+* $H_{3}$ Emotional adjustment will be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning
+* $H_{4a-d}$ Greater joint involvement will positively attenuate both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively attenuate parental anger
+* $H_{5a-d} Greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively
+* $H_{6a}$ Greater joint care should positively predict children's emotional adjustment
+* $H_{6b}$ Greater relative involvement will negatively predict emotional adjustment
+* $H_{7a-d}$ Parents will display more warmth, consistency, and inductive reasoning towards girls than boys, and more anger towards boys than girls
+* $H_{8a-d}$ Associations between qualitative parenting and emotional adjustment will be stronger between children and their same-sex parent
 
-In the revision, we have attempted to address this comment by being more specific. Now, this section reads:
-
-> Applying this insight from the coparenting literature, we thus hypothesize that  parents' quantitative involvement will be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and will moderate effects between qualitative parenting behaviors and emotional adjustment.
-
-There is a bit of a challenge, however: Because there is no precedent for using the random-intercept cross-lagged panel model or its multilevel DSEM extension in the literature on parental involvement and emotional adjustment, there is no literature upon which to base such hypotheses.
-Therefore, the best we can do is to present *conceptual* hypotheses.
-
-\WorkInProgress
+\Done
 
@@
 
-This might be an oversight, as all of this information requested is provided in Supplementary Table 1 (*Scale descriptive statistics are provided in Supplementary Table 1*).
+We respectfully believe that this comment might be the result of an oversight, or of a misunderstanding on our part with regard to what information is relayed to Reviewers.
+All of this information was provided in Supplementary Table 1, as mentioned in the manuscript: *Scale descriptive statistics are provided in Supplementary Table 1*.
 Perhaps these supplementary Tables are not automatically shared with reviewers?
@@
 To address this comment, we've clarified this sentence in the manuscript, which now reads: *Scale descriptive statistics, number of items, number of complete cases, and reliability are reported in Supplementary Table 1*.
-We also reproduce Table S1 here, in case the Reviewer does not have access:
+We also reproduce Table S1 here, in case the Reviewer does not have access.
+Hopefully this addresses the issue.
 
@@
 
-We have followed the convention of 1) providing example items as an indication of face validity, and 2) citing the scale validation papers for more information on predictive validity.
+We have followed the convention of 1) providing example items as an indicator of face validity, and 2) citing the scale validation papers for more information on predictive validity.
 We would like to defer to the Editor as to whether further actions need to be taken (we are also not sure what these actions would be).
@@
 
-Several working papers have been written on the recruitment, participants, data collection, etc.
-for this sample.
-However, this information was masked in the original manuscript, to enable blind peer review.
+The recruitment, participants, data collection, etc.
+for this sample are well-documented and these protocols are public.
+This information was masked in the original manuscript, to enable blind peer review.
 However, as both Reviewers and the Editor requested access to these references to evaluate the study quality, we have now unmasked these sections in the revised manuscript:
@@
 
-All of our results are reported in terms of correlations (which, when squared, gives the proportion of explained variance shared by two variables) and standardized regression coefficients (a one SD increase in X predicts a $\beta$ SD increase in Y, conditional on the other predictors), both of which are already effect sizes [@nieminenStandardisedRegressionCoefficient2013].
+All of our results are reported in terms of correlations and standardized regression coefficients, both of which are already effect sizes [@nieminenStandardisedRegressionCoefficient2013].
+We agree with the Reviewer that this was not communicated clearly enough, however, so we have attempted to address this comment by clarifying in the text and in the Tables that these results are reported as correlations and standardized regression coefficients.
+We assume that all readers are familiar with the interpretation of correlation coefficients, but we have included two detailed examples of the interpretation of standardized regression coefficients in the context of basic random effects, and moderated effects.
 
-*Probably, we should cite the values in the text, as the Editor also suggested.*
+The added sections for the random effects reads:
 
-\WorkInProgress
+> At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment.
+The intercepts of these random effects reflect the average standardized regression coefficient across families.
+The variances in Table 2 indicate whether this coefficient differs in strength (and sign) between families.
+We report unstandardized variances, because the standardized variances are, by definition, set to 1.
+With regard to within-families parenting effects, we found no evidence for hypotheses $H_{1b-d}$, as none of the intercepts differed significantly from zero.
+In line with hypothesis $H_{1a}$, however, on average, a maternal warmth positively predicted children's emotional adjustment in the next wave.
+The standardized regression coefficient of this effect, and others, is interpreted as follows: A $.99$ standard deviation increase in children's emotional adjustment in the next wave (see Table 2).
+
+The added section for the moderation effects reads:
+
+> To examine interactions between involvement quality and quantity, we examined standardized regression coefficients of between-family differences in quantitative involvement on within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table 3).
+With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
+Specifically, a one-standard deviation increase in joint involvement predicted a decrease in the within-families predictive effect of maternal warmth on emotional adjustment of $-.14$ of a standard deviation of the effect (see Table 2).
+
+\Done
 
@@
 
-We have attempted to address this comment by labeling and annotating the Tables more extensively.
+We have attempted to address this comment by labeling and annotating the Tables more extensively, and by explaining the coefficients in-text, as illustrated in our response to the previous question.
 
-\WorkInProgress
+\Done
 
@@
 
-In the original manuscript, we explained how sex differences were tested in the strategy of analyses:
+In the original manuscript, we explained how sex differences were tested in the strategy of analyses (see fragment below), and in the revision, we have attempted to further clarify this by including a visual representation of the model (Figure 1), and by labeling Tables more clearly.
 
@@
 In this way, we included total joint childcare, the childcare distribution, __and child sex__ as moderators.
-
-In the revision, we have attempted to further clarify this by including a visual representation of the model, and by labeling Tables more clearly.
 
@@
 We appreciate this suggestion!
-Strictly speaking, since Bayesian analyses focus on estimation rather than inference, it is correct to say that the (maximum a-posteriori) estimate is different between mothers and fathers, without further substantiation - because in the Bayesian framework, the maximum a-posteriori estimate can be interpreted as the most likely population value for the parameter, conditional on the data and prior.
-In practice, since we use Bayesian estimation for pragmatic reasons, we understand that there is added value in knowing the posterior distribution of the difference between mothers and fathers.
+We appreciate this suggestion!
+To address this comment, we have re-run all analyses to obtain the posterior for the differences between corresponding parent- and child effects.
+We have explained this procedure in the text (fragment cited below), and have added it to Table 2 in the form of labels that indicate which parameters differed significantly, and we have reported the full difference tests in Supplementary Table S5.
+These difference tests do not alter the interpretation of most results, because they closely mirror the pattern of significant/non-significant effects for fathers in comparison to mothers: All but one of the effects that are significant for one parent, but non-significant for the other parent, have significant difference tests.
+The exception is the child effect on parental anger.
 
-Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between mothers and fathers.
-We have reported this information in-text whenever making a statement about mother/father differences, and have added it to the relevant Supplementary Tables with the full model parameters.
-We have omitted it from the smaller Tables in the manuscript, because their function is to help make it easier for readers to digest the results, and we feel that adding four columns with difference tests would interfere with that goal.
+Section in Results:
 
-\WorkInProgress
+> We examined differences between fathers and mothers in parent- and child effects the same way [by testing the significance of the difference between corresponding mother- and father effects (see Table 2 and Supplementary Table S5)].
+All effects that were significant for one parent but not significant for the other parent had significant difference tests, except for the child effect of emotional adjustment on parental anger, which did not differ significantly between parents.
+
+Table note:
+
+> Matching letters indicate that parameters differed significantly, see Supplementary Table S5.
+
+\Done
 
@@
 
-\WorkInProgress
+This is automatically done by the typesetting program, to space the text evenly, but we have attempted to override the behavior.
+
+\Done
 
@@
 We completely agree with the Reviewer that warmth and security, provided by either parent, are positively related to emotional adjustment development.
-We discuss this in REFER TO PARAGRAPH.
 It appears that the Reviewer interpreted the word *uniquely* to mean that *only* fathers contribute to emotional adjustment.
 We intended to say that fathers contribute in a *different* way from mothers.
-We merely intended to summarize fatherhood theory regarding the distinction between the attachment vs activation relationship: An often-cited theory that implies that demandingness is a more important socialization mechanism for fathers than for mothers, and that warmth is a more important socialization mechanism for mothers than for fathers.
-
+Specifically, the effect of specific parenting behaviors might be a bit stronger for mothers versus fathers.
 In the Revision, we have attempted to address the Reviewer's comment by disambiguating this section, which now reads:
 
-\AlmostDone
+> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
+Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
+Some have theorized that mothers and fathers might fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011].
+In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage and respond supportively to emotional expression to a greater extent than fathers, whereas fathers structure and sanction emotion expressions more than mothers [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
+We might therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning, $H_{3}$.
+
+\Done
 
@@
 
-We have capitalized the word after a colon if it is the beginning of an independent clause, in line with the the sixth edition of the APA Publication Manual.
-We would generally to prefer to adhere to the Publication Manual in these matters, but we defer to the Editor as to whether further changes are required.
+We have capitalized the word after a colon if it is the beginning of an independent clause, in line with the APA Publication Manual (sixth edition).
+We would generally prefer to adhere to the Publication Manual in these matters, but we defer to the Editor as to whether further changes are required.
 To address this typo, and other spelling errors, we have checked the manuscript with the R-package `hunspell` and fixed about 10 other errors.
@@
 
-<!--[RK14]Socialization; children model their parents and see that their parents are also not able to regulate their emotions??-->
+Parental anger undermines emotional adjustment because anger is upsetting to children, and because parental anger reflects poor emotional adjustment on the parent's side, which children copy.
+We have rewritten the section to clarify this point:
 
-\WorkInProgress
+> Negative parenting, which we operationalized as parental anger, is an additional qualitative parenting dimension that is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+Parental anger expressions negatively impact children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because parental anger is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
+
+\Done
 
@@
 
-<!-- [RK15]The worse children fare, the more parents want to correct them/put them in line? -->
-\WorkInProgress
+We have completely rewritten this section, as we agree that it was confusing.
+Basically, the literature supports the notion that warmth, consistency, and inductive reasoning are positively associated, and anger is negatively associated, with emotional adjustment.
+However, research focusing on *behavioral control*, a different aspect of demandingness that we did not measure, is more equivocal.
+Some studies have found positive associations with emotional adjustment, and others have found negative associations - particularly in terms of child effects.
+We had tried to convey this subtlety in the original manuscript, but we have now removed it, because it is confusing, and because it is only tangentially relevant, because we did not measure behavioral control.
+The section now reads:
+
+> [...] bi-directional effects are not routinely investigated in the childhood literature [@flouriFathersBehaviorsChildren2010].
+Consequently, most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could be taken as support for child effects as well as parent effects.
+We therefore hypothesize that child effects are similar in direction, and that warmth, consistency, and inductive reasoning positively predict, and anger negatively predicts, emotional adjustment, $H_{2a-d}$.
+This is supported by studies that investigated child effects specifically, and found that children's maladjustment elicits less parental warmth, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
+Bi-directional positive links between parental consistency and emotional adjustment have also been reported [@lenguaBidirectionalAssociationsTemperament2005].
+The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011].
+Thus, in a way, parents respond positively to emotional adjustment, and negatively to maladjustment.
+
+\Done
 
@@
 
-\WorkInProgress
+We have done as asked. Nearly every instance of "sex" has been replaced by "child sex", except where the fact that we are discussing child sex is made clear in the rest of the sentence.
+
+\Done
 
@@
 
-1. We have clarified this section in the discussion.
-2. We have removed masking of sample details.
+1. We have removed masking of sample details.
+2. We have clarified this section in the discussion, which now reads:
 
-The Discussion section in question now reads:
-
-> One recent study with a similar design to the present study found that child effects substantially outnumbered parenting effects, and speculated that this might be due to the greater equality of parent-adolescent relationships, compared to earlier childhood [@vanlissaRoleFathersMothers2018a].
-The present study casts doubt on this speculation, as we found that child effects outweighed parenting effects, even at a young age.
-The fact that both the present study and this previous work found a preponderance of child effects, despite the different age ranges of the samples, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it?
+> we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
+They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
+The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.
 
@@
 
-\WorkInProgress
+We have attempted to address this comment by adding extensive Table notes and clarifying in-table text throughout the *Results* section.
+Some of the results are presented as correlation coefficients.
+We assume that readers are familiar with the interpretation of correlations, and have therefore not further clarified these sections.
+The remaining results are presented as standardized regression coefficients.
+Some of these vary between families, and thus have a random intercept and a variance.
+We have included an elaborate explanation of the interpretation of the first random intercept for the within-family standardized regression coefficients, and one elaborate example for the interpretation of the moderated effects.
+
+The added sections for the random effects reads:
+
+> At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment.
+The intercepts of these random effects reflect the average standardized regression coefficient across families.
+The variances in Table 2 indicate whether this coefficient differs in strength (and sign) between families.
+We report unstandardized variances, because the standardized variances are, by definition, set to 1.
+With regard to within-families parenting effects, we found no evidence for hypotheses $H_{1b-d}$, as none of the intercepts differed significantly from zero.
+In line with hypothesis $H_{1a}$, however, on average, a maternal warmth positively predicted children's emotional adjustment in the next wave.
+The standardized regression coefficient of this effect, and others, is interpreted as follows: A $.99$ standard deviation increase in children's emotional adjustment in the next wave (see Table 2).
+
+The added section for the moderation effects reads:
+
+> To examine interactions between involvement quality and quantity, we examined standardized regression coefficients of between-family differences in quantitative involvement on within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table 3).
+With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
+Specifically, a one-standard deviation increase in joint involvement predicted a decrease in the within-families predictive effect of maternal warmth on emotional adjustment of $-.14$ of a standard deviation of the effect (see Table 2).
+
+\Done
 
@@
 
-335 effect of emotional adjustment on maternal warmth was positive and significant (Table 2).
-336 Thus, when children displayed better emotional adjustment than usual in one year, their
-337 mothers displayed more anger and inductive reasoning, but also more warmth, than usual in
-338 the next year.
+We were referring back to another finding, but we recognize that this was confusing and did not serve any particular purpose. The section now reads:
 
-\WorkInProgress
+> With regard to child effects, in line with hypothesis $H_{2a}$, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table 2).
+Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave.
+
+\Done
 
@@
 
-Apologies for the oversight; we simply meant to say Inductive Reasoning. We have corrected this typo in the revision.
+Apologies for the oversight; we meant to say "inductive reasoning". We have corrected this typo in the revision.
 
@@
 We are not sure that we fully understand the Reviewer's comment, but we take it to mean that our argumentation in this paragraph was unclear.
-In the revision, we have rewritten this section as follows:
+In the revision, we have separated our discussion of specific findings (such as for consistency) from a more general discussion of the implications of the fact that we, and two other studies, have found that once stable between-family level correlations are taken into account, within-family effects tend to mostly disappear.
+This section now reads:
 
-> On average, we found no support for within-family predictive links between consistency and adjustment, but we did find support for between-family level correlations.
-Between-family level correlations can have different causes: They can reflect the consolidated influence of predictive effects that occurred prior to the study, or alternatively, the influence of third variables that impact both constructs.
-The fact that our study included children from a very young age casts doubt on the former explanation.
-It seems more likely that third variables explain these associations.
-Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
-This finding means that we should reflect critically on causal interpretations of prior research that did not control for stable between-family differences [e.g., @brandPerceivedParentalInconsistency1990].
+> We found only limited evidence for within-family predictive effects, most of which were child effects on parents.
+Instead, we mostly found evidence for between-family level correlations.
+We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016].
+In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
+They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
+The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.
+  As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
+Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
+Our findings serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general.
+We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
+  Future research might also devote greater attention to between-family level correlations as an outcome of interest.
+These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs.
+We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors.
+For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
 
@@
 
-<!-- WRONG ANSWER!! [RK17]Let op: dit is niet het goede antwoord op de vraag!!! -->
+We appreciate this suggestion!
+To address this comment, we have re-run all analyses to obtain the posterior for the differences between corresponding parent- and child effects.
+We have explained this procedure in the text (fragment cited below), and have added it to Table 2 in the form of labels that indicate which parameters differed significantly, and we have reported the full difference tests in Supplementary Table S5.
+These difference tests do not alter the pattern of results, because they exactly mirror the pattern of significant/non-significant parent and child effects: All effects that are significant in one direction, but non-significant in the reverse direction, have significant difference tests.
+All effects that are significant in both directions, or non-significant in both directions, have non-significant difference tests.
 
-<!--We appreciate this suggestion!
-Strictly speaking, since Bayesian analyses focus on estimation rather than inference, it is correct to say that the (maximum a-posteriori) estimate is different between mothers and fathers, without further substantiation - because in the Bayesian framework, the maximum a-posteriori estimate can be interpreted as the most likely population value for the parameter, conditional on the data and prior.
-In practice, since we use Bayesian estimation for pragmatic reasons, we understand that there is added value in knowing the posterior distribution of the difference between mothers and fathers.
+Section in Results:
 
-Therefore, to address this comment, we have re-run all analyses to obtain the posterior for the differences between mothers and fathers.
-We have reported this information in-text whenever making a statement about mother/father differences, and have added it to the relevant Supplementary Tables with the full model parameters.
-We have omitted it from the smaller Tables in the manuscript, because their function is to help make it easier for readers to digest the results, and we feel that adding four columns with difference tests would interfere with that goal.
--->
+> We established direction of effects by testing the significance of the difference between corresponding parent- and child effects (see Table 2 and Supplementary Table S5).
+All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests.
+All effects that were significant (or non-significant) in both directions had non-significant difference tests.
 
-\WorkInProgress
+Table note:
 
-<!--We appreciate the suggestion; we are indeed only interpreting the pattern of effects (i.e., which effects differ 'significantly' from zero), and it would be better to explicitly test the difference between these parameters (i.e., which parent and child effects differ significantly from one another).
-In the revision, we have attempted to address this comment by explicitly acknowledging this in the Discussion section.
+> Matching letters indicate that parameters differed significantly, see Supplementary Table S5.
 
-We would prefer not to incorporate these tests for two reasons:
+\Done
 
-1.
-We have no hypotheses about the difference between parent and child effects.
-Including difference tests would therefore introduce an additional and unnecessary layer of complexity to the manuscript - which is already very complex.
-2. For practical reasons: Each of the four models we conducted takes approximately a week to run (MCMC sampling is very computationally expensive).
-
-If the Editor deems it necessary to include these tests nonetheless, we will do so of course, but we'd need a longer turnaround time for the next revision.
-
-\Done-->
-
-\RC{I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?}
+\RC{16. I assume parenting and child effects were run separately considering they are modeled as predicting different outcomes?}
 
 This assumption is incorrect; parent and child effects refer to the reciprocal effects between the parenting variables and child adjustment; both variables must be included in one model to examine these reciprocal effects.
-This is outlined in the paragraph below:
+This is outlined in the *Strategy of Analyses*, and visually displayed in Figure 1:
 
@@
 Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
-In this way, we included total joint childcare, the childcare distribution, and child sex as moderators.
+In this way, we included joint and relative quantitative involvement, and child sex as moderators.
+Figure \@ref(fig:figure1) shows a graphical representation of the resulting model.
 We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
@@
 
-\RC{16.
+\RC{17.
 Can you test for non-linear effects of quantitative involvement on emotional adjustment?
@@
 Thus, we must choose: We can either test our theoretical model, which requires Bayesian estimation using Gibbs sampling, or we analyze this proposed non-linear effect, but within a much simpler model that can be estimated using numerical integration.
-Our choice to analyze our theoretical model also means deciding *against* testing alternative possible models, such as the one proposed by the Reviewer.
+Our choice to analyze our theoretical model also means deciding *against* testing certain alternative models, such as the one proposed by the Reviewer.
 
@@
 
-\RC{17.
+\RC{18.
 You could also test for mean differences in emotional adjustment for children with highly involved parents versus less involved parents.
@@
 
-> Counter to our hypothesis, we found that greater total joint involvement was negatively associated with emotional adjustment.
+> Counter to our hypothesis $H_{6a}$, we found that greater joint involvement was negatively associated with emotional adjustment.
 Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
-*[...]*  
-Finally, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.
+However, hypothesis $H_{6b}$, that children's emotional adjustment would be lower in families where mothers were relatively more involved than fathers, was supported.
 

### 2019-11-12 Paper_shortened_pdf.Rmd
Final changes before introduction

@@
 output            : papaja::apa6_pdf
+header-includes   :
+    - \raggedbottom
 ---
@@
 It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
-Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for the child), and in terms of qualitative parenting behaviors (how do parents behave towards children).
+Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their children).
 Nevertheless, the existing literature has two major shortcomings.
@@
 Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
-Children’s emotional adjustment might be differently impacted by parents’  control if that parent is generally highly involved with the day-to-day minutiae of childrearing, than when (s)he is generally absent.
+Children’s emotional adjustment might, for example, be differently impacted by a parent's warmth if that parent is generally highly involved with the day-to-day minutiae of childrearing, than when (s)he is generally absent.
 Conversely, children’s emotional adjustment might elicit different qualitative parenting behaviors from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be sanctioned.
@@
 Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].
+Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.
 
@@
 Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990].
+Therefore we hypothesized that consistency would be positively associated with emotional adjustment, $H_{1b}$.
 There is less empirical evidence for linkages between inductive reasoning and children’s emotional adjustment.
-Given that inductive inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would be positively related to emotional adjustment.
+Given that inductive inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.
 
 Negative parenting, which we operationalized as parental anger, is an additional qualitative parenting dimension that is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
-Parental anger directly undermines children's emotional adjustment [@scaramellaParentingChildrenDistress2008], and models emotional dysregulation [@changHarshParentingRelation2003].
-We thus hypothesized that it would be negatively related to emotional adjustment.
-
-Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
-Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
-For various reasons, mothers and fathers sometimes appear to fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011].
-In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage, and respond more supportively to, emotional expression, whereas fathers structure and sanction emotion expressions to a greater extent [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
-Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
-Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
-We might therefore expect emotional adjustment to be associated more strongly with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning.
+Parental anger expressions negatively impact children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because parental anger is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
+We thus hypothesized that it would be negatively related to emotional adjustment, $H_{1d}$.
 
 ### Parent or child effects?
-
+<!-- Hier iets doen 
+Was wel echt meer control, maar rigid control oid?-->
 A view of the family as a dynamic system, whose members reciprocally influence one another, has become commonplace [@kerrPerceivedParentingStyle2012].
 Nevertheless, bi-directional effects are not routinely investigated in the childhood literature [@flouriFathersBehaviorsChildren2010].
-The available evidence suggests that maladjustment elicits less parental warmth, more demandingness, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004; @nepplIntergenerationalContinuityParenting2009; @lenguaBidirectionalAssociationsTemperament2005].
-The adolescent literature provides additional support that emotional maladjustment predicts decreased parental support, and greater demandingness [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014; @wijsbroekDirectionEffectsPerceived2011].
-Thus, parents' display punitive responses to maladjustment, including decreased support and increased negative parenting, and corrective responses, such as increased demandingness.
-We therefore hypothesize that emotional adjustment will positively predict parental warmth, and negatively predict consistency, inductive reasoning, and anger.
+Consequently, most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could be taken as support for child effects as well as parent effects.
+We therefore hypothesize that child effects are similar in direction, and that warmth, consistency, and inductive reasoning positively predict, and anger negatively predicts, emotional adjustment, $H_{2a-d}$.
+This is supported by studies that investigated child effects specifically, and found that children's maladjustment elicits less parental warmth, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
+Bi-directional positive links between parental consistency and emotional adjustment have also been reported [@lenguaBidirectionalAssociationsTemperament2005].
+The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011].
+Thus, in a way, parents respond positively to emotional adjustment, and negatively to maladjustment.
+<!-- Neppl 2009: Adolescents' harsh behavior towards parents predicts harsh parenting, and positive behavior predicts positive parenting -->
 
@@
 Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if stable between-family differences are not controlled [@hamakerCritiqueCrosslaggedPanel2015].
-Innovative methods provide better estimates of dynamic reciprocal processes within families, by controlling for stable differences: The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].
+Innovative methods provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].
 
-Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples.
+Several studies have re-examined the direction of effects between parents and children using these methods, but (to our knowledge) only in adolescent samples.
 One such study found that effects of parenting on adjustment that emerged using classical methods, disappeared when stable between-family differences in parenting and adjustment were included in the model [@keijsersParentalMonitoringAdolescent2016].
@@
 If this is correct, we would expect to find more evidence of parenting effects in the present, younger, sample.
+These two studies also illustrate that applying these new methods changes the pattern of results in unexpected ways.
+Consequently, it is impossible to make specific predictions about the pattern of between-family correlations and within-family predictive effects.
+For now, we will assume that $H_{1a-d}$ and $H_{2a-d}$ apply at both the between- and within-family level.
+<!-- Hier iets over de hypotheses; dat we niet weten wat er within en wat between speelt 
+We are, however, a bit cautious about formulating more focused hypotheses about within- versus between-family effects than the one formulated above.
+Because there is no precedent for using the random-intercept cross-lagged panel model or its multilevel DSEM extension in the literature on parental involvement and emotional adjustment, there is no literature upon which to base such hypotheses.
+Our examination of between and within-family effects will therefore mainly be exploratory.
+-->
 
-<!-- Hier iets over de hypotheses; dat we niet weten wat er within en wat between speelt -->
+### Differences between mothers and fathers
+
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
+Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
+Some have theorized that mothers and fathers might fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011].
+In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage and respond supportively to emotional expression to a greater extent than fathers, whereas fathers structure and sanction emotion expressions more than mothers [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
+We might therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning, $H_{3}$.
 
@@
 Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014].
-Consequently, we expect greater joint involvement to positively predict both parents' parenting quality.
+Consequently, we hypothesize that greater joint involvement will positively attenuate both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively attenuate parental anger, $H_{4a-d}$.
 
-With regard to differential involvement, we might also expect the more involved parent to display more desirable qualitative parenting behaviors, due to the effects of experience [@sabattiniRelationMothersFathers2004].
-However, the benefit of parents monitoring each other's child-rearing should be diminished when parents are not equally involved [@mclanahanConsequencesFatherAbsence2014].
-Indeed, retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
-Additionally, greater differential involvement might undermine both parents' involvement quality, through its effect on the marital relationship.
+By the same reasoning, we might hypothesize that greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004].
+However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
+Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
+One reason might be that parents are less able to monitor each other's child-rearing behavior when they are not equally involved [@mclanahanConsequencesFatherAbsence2014].
+Additionally, differential involvement might negatively impact the marital relationship, and thereby undermine both parents' involvement quality.
 Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001].
@@
 Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].
-We thus expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
 
@@
 Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
-This implies that joint care should be positively associated with children's emotional adjustment.
-Relative involvement, on the other hand, might be negatively linked to emotional adjustment, because children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
+We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$.
+By contrast, we predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$, because children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
 
@@
 Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003].
+Consequently, it is not possible to substantiate specific hypotheses for such interactions.
 One notable exception is Feinberg's coparenting model [@feinbergInternalStructureEcological2003].
@@
 According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of relative involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs.
+Applying this insight from the coparenting literature, we might expect parents' quantitative involvement to moderate effects between qualitative parenting behaviors and emotional adjustment.
+Due to the limited literature, it is not possible to formulate specific hypotheses about the nature of these moderating effects.
+There is, however, preliminary evidence to indicate that interactions between involvement quality and quantity are relevant.
+For instance, fathers' playful and caring involvement positively predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012].
+Furthermore, the quantity of paternal school-related involvement interacted with paternal warmth to positively predict adolescents' academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014].
+We thus explored the moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.
 
-There is some empirical support for interactions between involvement quality and quantity; for instance, fathers' playful and caring involvement predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012].
-In adolescence, the quantity of paternal education-related structure, an age-appropriate form of involvement, interacted with paternal warmth to predict both academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014].
-Applying this insight from the coparenting literature, we might thus expect parents' quantitative involvement to be directly associated with their qualitative parenting behaviors and children's emotional adjustment, and to moderate effects between qualitative parenting behaviors and emotional adjustment.
+## Child sex differences
 
-## Sex differences
-
-Sex differences are a relevant factor in the interplay between parenting and emotional adjustment.
+Child sex differences are a relevant factor in the interplay between parenting and emotional adjustment.
 Parents tend to raise boys and girls differently; expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999].
@@
 Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
-We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning towards girls, and more anger towards boys.
-Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent.
+We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning towards girls than boys, and more anger towards boys than girls, $H_{7a-d}$.
+Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent, $H_{8a-d}$.
 
@@
 Third, we disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative involvement and emotional adjustment.
-Finally, we examined sex as a moderator of between-family differences in all study variables and within-family predictive effects.
+Finally, we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects.
 
@@
 
-We hypothesized that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
+We hypothesized that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
 At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)).
 Random intercepts reflect the average level of each variable across all waves of the study.
-Counter to our hypotheses, we found no significant correlations between parental warmth and children's emotional adjustment.
-In support of hypotheses, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment.
-Contrary to hypotheses, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment.
+Counter to hypothesis $H_{1a}$, we found no significant correlations between parental warmth and children's emotional adjustment.
+In support of hypothesis $H_{1b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment.
+Contrary to hypothesis $H_{1c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment.
 This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment.
-Finally, as hypothesized, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
+Finally, as hypothesized in $H_{1d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
 
@@
 rownames(table_cors) <- NULL
+colnames(table_cors)[1] <- "Param."
 #reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
-reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
+reg_table(table_cors, caption = "Correlations between qualitative parenting and emotional adjustment", note = "Param.: Parameter.
+Emo.: Emotional adjustment.
+*: <5% of posterior has opposite sign from estimate.")
 ```
@@
 We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)).
-In support of hypotheses, we found significant positive within-family level associations between emotional adjustment and fathers' and mothers' warmth.
-Notably, these associations were nearly twice as strong for fathers' warmth, as compared to mothers' warmth.
-In partial support of hypotheses, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
-Finally, contrary to our expectations, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.
+In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between emotional adjustment and fathers' and mothers' warmth.
+In partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
+Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.
+Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S5).
+This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.
 
@@
 At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment.
-The intercepts of these random effects reflect the average effect across families.
-Their variances indicate whether the effect differs in strength (and sign) between families.
-With regard to within-families parenting effects, none of the intercepts differed significantly from zero, excepting the effect of maternal warmth on children's emotional adjustment (Table \@ref(tab:tabcrosspath)).
-Thus, on average, mothers' warmth had a positive significant predictive effect on children's emotional adjustment.
-The variances of all parenting effects were all significant, however, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
-Moreover, these variances were notably larger for fathers than for mothers, indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
+The intercepts of these random effects reflect the average standardized regression coefficient across families.
+The variances in Table \@ref(tab:tabcrosspath) indicate whether this coefficient differs in strength (and sign) between families.
+We report unstandardized variances, because the standardized variances are, by definition, set to 1.
+With regard to within-families parenting effects, we found no evidence for hypotheses $H_{1b-d}$, as none of the intercepts differed significantly from zero.
+In line with hypothesis $H_{1a}$, however, on average, a maternal warmth positively predicted children's emotional adjustment in the next wave.
+The standardized regression coefficient of this effect, and others, is interpreted as follows: A $.99$ standard deviation increase in children's emotional adjustment in the next wave (see Table \@ref(tab:tabcrosspath)).
+The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
+Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S5), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
 
@@
 # Add letters for contrast tests
+letter <- 1
+tab <- add_letters(list("Mother effect - child effect on mother" = c(2, 8),
+                 "Father effect - child effect on father" = c(3, 9)),
+            tab, tab_unst_sections)
 
-stars <- tab_unst_sections$`Defined parameters`[tab_unst_sections$`Defined parameters`$Parameter %in% c("Mother effect - child effect on mother", "Father effect - child effect on father"), ]
-letter <- 1
-tabrows <- list(c(2,8),
-                c(3, 9))
-
-for(i in 1:length(tabrows)){
-  if(any(grepl("\\*", stars[i, ]))){
-    for(c in which(grepl("\\*", stars[i, ]))){
-      # c = 6
-      tab[tabrows[[i]], c] <- paste0(tab[tabrows[[i]], c], letters[letter])
-      letter <<- letter + 1
-    }
-  }
-}
+tab <- add_letters(list("F-M Parent effects on emo." = c(2, 3),
+                        "F-M Variance of parenting effect" = c(5,6),
+                 "F-M Child effects on parenting" = c(8, 9),
+                 "F-M Variance of child effect" = c(11, 12)),
+            tab, tab_unst_sections)
 
 tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
-reg_table(tab, caption = "Standardized between-family level intercepts and unstandardized variances of within-family predictive effects", note = "Est.: Estimate.
+names(tab)[1] <- "Param."
+reg_table(tab, caption = "Standardized between-family level intercepts and unstandardized variances of within-family predictive effects", note = "Param.: Parameter.
+Est.: Estimate.
 HDPI: Highest posterior density interval.
@@
 *: <5% of posterior has opposite sign from estimate.
-Matching letters indicate that parameters differed significantly, see Supplementary Table S5.")
+Matching letters indicate that parameters differed significantly; see Supplementary Table S5 for all difference tests.")
 ```
 
-With regard to child effects, in line with hypotheses, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)).
-Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more warmth than usual in the next year.
-Contrary to hypotheses, the intercept of the effect of emotional adjustment on maternal consistency was non-significant.
-As hypothesized, the intercept for maternal inductive reasoning was positive.
-Counter to hypotheses, the intercept for maternal anger was positive, not negative.
-Thus, when children displayed better emotional adjustment than usual in one year, their mothers displayed more anger than usual in the next year.
-In partial support of hypotheses, the intercept for the child effect of adjustment on paternal warmth was positive and significant.
-None of the other intercepts for child effects on paternal parenting were significant.
+With regard to child effects, in line with hypothesis $H_{2a}$, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)).
+Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave.
+Contrary to hypothesis $H_{2b}$, the intercept of the effect of emotional adjustment on maternal consistency was non-significant.
+As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive.
+Counter to hypothesis $H_{2d}$, the intercept for maternal anger was positive, not negative.
+Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more anger than usual in the next wave.
+In support of hypothesis $H_{2a}$, the intercept for the child effect of adjustment on paternal warmth was positive and significant.
+None of the other intercepts ($H_{2b-d}$)for child effects on paternal parenting were significant.
 The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
@@
 All effects that were significant (or non-significant) in both directions had non-significant difference tests.
-
-We examined differences between fathers and mothers the same way. <!-- NOG DOEN -->
+We examined differences between fathers and mothers in parent- and child effects the same way.
+All effects that were significant for one parent but not significant for the other parent had significant difference tests, except for the child effect of emotional adjustment on parental anger, which did not differ significantly between parents.
 
@@
 
-We examined regression coefficients from joint and relative quantitative involvement to the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
-With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning.
+We examined standardized regression coefficients of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+With regard to parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning.
 Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers.
-We did not find the hypothesized negative effects of joint involvement on parental anger.
-Counter to our hypothesis, we found that greater joint involvement was negatively associated with emotional adjustment.
+We did not find the hypothesized negative effects of joint involvement on parental anger, $H_{4d}$.
+With regard to relative involvement, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively.
+For mothers, the only significant association we found was that mothers displayed more warmth when they were relatively more involved than fathers.
+We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.
+
+Counter to our hypothesis $H_{6a}$, we found that greater joint involvement was negatively associated with emotional adjustment.
 Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
-
-With regard to relative involvement, we found partial support for our hypothesis that more skewed relative involvement should predict greater warmth, consistency, and inductive reasoning, and reduced anger, only for the more involved parent.
-For mothers, the only significant association we found was that, in line with this hypothesis, mothers displayed more warmth when they were relatively more involved than fathers.
-For fathers, we found that, in line with this hypothesis, fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.
-Finally, our hypothesis that children's emotional adjustment would be poorer in families where mothers were relatively more involved than fathers was supported.
+However, hypothesis $H_{6b}$, that children's emotional adjustment would be lower in families where mothers were relatively more involved than fathers, was supported.
 
@@
 rownames(tab) <- NULL
-cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of between-family differences in quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
+names(tab)[1] <- "Param."
+cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
+Param.: Parameter.
 Est.: Estimate.
@@
 
-To examine interactions between involvement quality and quantity, we tested whether between-family differences in quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+To examine interactions between involvement quality and quantity, we examined standardized regression coefficients of between-family differences in quantitative involvement on within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
 With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
-Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when joint involvement was higher, and positively attenuated when mothers were relatively more involved than fathers.
+Specifically, a one-standard deviation increase in joint involvement predicted a decrease in the within-families predictive effect of maternal warmth on emotional adjustment of $-.14$ of a standard deviation of the effect (see Table \@ref(tab:tabcrosspath)).
+This same effect was positively attenuated by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
 We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.
@@
 
-We found sex differences in the random intercepts of qualitative parenting behaviors: Fathers displayed less warmth, and both parents displayed more inductive reasoning and anger, towards sons as compared to daughters  (Table \@ref(tab:tabsex)).
-No significant differences were found for maternal warmth, and for both parents' consistency.
+We found sex differences in the random intercepts of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)).
+Counter to hypothesis $H_{7c}$, both parents also displayed more inductive reasoning towards sons.
 We found no sex differences in the random intercept of emotional adjustment.
-The only cross-lagged effects significantly moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning.
-Specifically, emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls.
-In other words, if children display worse emotional adjustment than usual in one year, parents increase inductive reasoning more strongly for boys than for girls.
+Hypotheses $H_{8a-d}$ were rejected: There was no evidence for stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent.
+The only predictive effects moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning.
+Emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls.
+In other words, if children display worse emotional adjustment than usual in one wave, parents increase inductive reasoning more strongly in the next wave for boys than for girls.
 
@@
 tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
-tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
-                              "Sex differences in parenting effects of:",
-                              "Sex differences in parenting effects on:")
+tabsex$Parameter[c(1,5,8)] <- c("Sex differences in random intercepts of parenting and adjustment",
+                              "Moderation of parenting effects by sex",
+                              "Moderation of child effects by sex")
 #tabsex[is.na(tabsex)] <- ""
 rownames(tabsex) <- NULL
-reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Sex is a binary predictor (-1 = female, 1 = male).
+names(tab)[1] <- "Param."
+reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Child sex is a binary predictor (-1 = female, 1 = male).
 The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
+Param.: Parameter.
 Est.: Estimate.
@@
 
+This study set out to obtain a more comprehensive and nuanced understanding of how fathers' and mothers' quantitative involvement relate to children's emotional adjustment development.
+We used innovative methods that control for stable between-family differences, in order to obtain more accurate estimates of within-family effects and investigate the direction of effects between parents and children.
+Furthermore, we examined how joint and relative quantitative involvement predict parenting and emotional adjustment, and moderate the longitudinal interplay between qualitative parenting and emotional adjustment.
+
 ## Associations between qualitative parenting and emotional adjustment
 
-In terms of between-family level correlations, we found partial support for our hypotheses.
-We did not find the hypothesized positive associations between parental warmth and emotional adjustment.
-The predicted positive associations of parental consistency with emotional adjustment did emerge, indicating that children in families with highly consistent parents tended to be more emotionally well-adjusted.
-Counter to hypotheses, parents who engaged in more inductive reasoning had children with poorer emotional adjustment.
-Finally, for anger, we found the predicted negative associations, which means that, in families where mothers or fathers display high anger, children tend to be less emotionally well-adjusted.
+We found mixed evidence, at both the within- and between-family levels, for our hypotheses that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
+Parental warmth was not significantly correlated with adjustment at the between-family level, but changes in warmth were correlated with changes in adjustment at the within-family level.
+Consistency was positively correlated with adjustment for both parents at the between-family level, indicating that children in families with highly consistent parents tended to be more emotionally well-adjusted, and for mothers at the within-level.
+Counter to hypotheses, inductive reasoning was negatively correlated with adjustment at the between-family level, indicating that parents who engage in more inductive reasoning had children with poorer emotional adjustment.
+At the within-family level, increases in fathers' inductive reasoning were positively correlated with adjustment.
+Finally, for anger, we found the predicted negative associations at the between-family level only, which means that, in families where mothers or fathers display high anger, children tend to be less emotionally well-adjusted.
 
-In terms of parenting effects, we found only one effect that was significant on average, i.e., its random intercept was significant and positive: Maternal warmth positively predicted emotional adjustment in most families.
-However, all within-family predictive effects had significant variance components, which suggests that the fact that the findings discussed here only reflect the average across all families, and that dynamic reciprocal processes vary from one family to the next.
-Variability was notably greater for fathers, indicating that fathers' influence differs most across families.
+In terms of parenting effects, the only effect that was significant on average was that maternal warmth positively predicted emotional adjustment in most families.
+However, all within-family predictive effects significantly varied across families, and this variability was notably greater for fathers, indicating that fathers' influence differs most across families.
+Our results thus emphasize the fact that dynamic reciprocal processes vary from one family to the next.
 
@@
 Counter to predictions, we did not find average child effects for consistency.
-In partial support of our hypothesis, we found that children who displayed better emotional adjustment in one year evoked greater inductive reasoning from mothers in the next year.
-But paradoxically, and counter to oru hypothesis, we also found that children who displayed better emotional adjustment in one year evoked greater maternal anger in the next year.
-
-In sum, we found mixed evidence for our hypotheses that, at both the within- and between-family levels, warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
+In partial support of our hypothesis, we found that children who displayed better emotional adjustment in one wave evoked greater inductive reasoning from mothers in the next wave.
+But paradoxically, and counter to oru hypothesis, we also found that children who displayed better emotional adjustment in one wave evoked greater maternal anger in the next wave.
 
@@
 
-We set out to investigate the direction of effects between parents and children, using innovative methods that control for stable between-family differences, in order to obtain more accurate estimates of within-family effects [@hamakerCritiqueCrosslaggedPanel2015].
 We found ample evidence for within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
-We found no between-family level correlations for warmth.
-This suggests that warmth has an ongoing positive within-family effect on emotional adjustment.
+This suggests that warmth and emotional adjustment reciprocally influence one another within families; potentially through causal mechanisms.
 For consistency, however, we found no within-family predictive effects, but only between-family level associations.
 This suggests that the benefits for emotional adjustment associated with greater consistency might be primarily due to stable between-family differences.
-
 For inductive reasoning, paradoxically, between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment.
-This apparently paradoxical finding might be explained by considering that inductive reasoning is a disciplinary behavior - albeit a constructive one [@grusecImpactParentalDiscipline1994].
-When parents report engaging in more inductive reasoning, this introduces a confound between *how often* they discipline their children, and *how* they discipline them.
-The fact that between-family correlations between inductive reasoning and emotional adjustment were negative suggests that between-family differences in inductive reasoning might tap more strongly into how often parents engage in discipline.
-The within-family fluctuations, by contrast, might reflect changes in how parents discipline their children.
+One potential explanation for this finding might be that the construct of inductive reasoning confounds *how* parents discipline their children (sample item: "How often do you talk it over and reason with this child..."), with *how often* their child misbehaves ("...when he/she misbehaved?").
+The fact that between-family correlations between inductive reasoning and emotional adjustment were negative suggests that between-family differences in inductive reasoning primarily capture variance related to the frequency of misbehavior.
+The within-family fluctuations, by contrast, might tap into changes in how parents discipline their children to a greater extent.
 
@@
 We also found a paradoxical positive predictive effect of emotional adjustment on maternal anger.
-This means that, if children demonstrated better emotional adjustment than usual in one year, their mothers tended to display more anger than usual in the next year.
+This means that, if children demonstrated better emotional adjustment than usual in one wave, their mothers tended to display more anger than usual in the next wave.
 A possible explanation might be regression to the mean: On average, emotional adjustment showed no-to-negative stability coefficients in all models.
-That means that better emotional adjustment in one year, was likely to be followed worse adjustment in the next year.
+That means that better emotional adjustment in one wave, was likely to be followed worse adjustment in the next wave.
 The literature on expectation-based loss aversion suggests that mothers might be more angered by a child who unexpectedly displays poor adjustment, than they would be appeased by a child who displays unexpected positive adjustment [@koszegiModelReferenceDependentPreferences].
-
-To recapitulate, we set out to examine the direction of within-family effects by controlling for stable between-family differences.
-We found only limited evidence for within-family effects, most of which were child effects on parents.
-Instead, we mostly found evidence for between-family level correlations.
-We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016].
-In the area of parenting and emotional adjustment development, we are aware of only one prior study that also controlled for stable between-family differences.
-That study also reported mostly child effects and between-family level corelations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
-The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
-They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
-The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of children from age 6-12 suggests that either parent effects occurred at an even younger age, or that something else is at play.
-
-As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
-We should consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
-Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
-Our findings also serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general.
-
-Future research might also devote greater attention to between-family level correlations as an outcome of interest.
-Between-family level correlations might reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they might reflect the influence of third variables that impact both constructs.
-Economic hardship, for instance, undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
-We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors such as economic hardship and neighborhood effects [@dengFamilyProcessesMediating2006], which are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.
-
-Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
-At the same time, we should consider the limitations of these methods.
-The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
-Stable factors can also exert changing influences over time.
-As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time.
-It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
-Another important limitation is that within-family cross-lagged effects are time scale dependent.
-If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
-Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
 
@@
 
+<!-- To do!! --> 
 Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development.
@@
 
-When both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning.
+Turning to our findings from linkages between quantitative involvement and qualitative parenting behaviors, our results were in line with our hypothesis that greater involvement would be associated with more desirable parenting behaviors [@sabattiniRelationMothersFathers2004].
+Our results revealed that, when both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning.
 Fathers, in particular, reported substantially greater parental warmth when both parents were more involved.
@@
 This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978].
-These results are in line with our hypothesis that greater involvement would be associated with more desirable parenting behaviors [@sabattiniRelationMothersFathers2004].
-Additionally, they suggest that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [@craigHowMothersFathers2011].
+These results indicate that, when childcare is unequally distributed, fathers engage in less desirable qualitative parenting behaviors [see also @craigHowMothersFathers2011].
 
@@
 
-In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were more involved than fathers - even after controlling for the total amount of care provided.
+In line with predictions derived from coparenting theory, we found that children's emotional adjustment was lower in families where mothers were relatively more involved than fathers - even after controlling for the joint amount of care provided.
 Although substantial empirical work has indicated that the marital relationship suffers when childcare is unequally distributed [@shockleyItNotWhat2018], our research is one of few to examine the implications for children's adjustment [see also @deutschPaternalParticipationChild2001].
-These correlational findings at the between-family level should not be interpreted causally, however.
-Future research should examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their well-being, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.
+We caution against undue causal interpretation of these findings, however.
+Future research is required to examine potential pathways, such as whether children's adjustment suffers when they feel like both parents are not equally invested in their well-being, or whether fathers are more likely to withdraw from childcare responsibilities when children display poor emotional adjustment, or whether differential involvement strains the marital relationship, and this strain spills over to children's emotional adjustment.
 
@@
 
-The evidence that quantitative moderates child effects was somewhat stronger.
+Evidence that quantitative moderates child effects was somewhat stronger.
 Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth.
@@
 
-## Sex differences
+## Child sex differences
 
@@
 Parenting effects were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors.
-Child effects were moderated by sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male.
+Child effects were moderated by child sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male.
 When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.
+
+## Summary
+
+To recapitulate, we set out to examine the direction of within-family effects by controlling for stable between-family differences.
+We found only limited evidence for within-family predictive effects, most of which were child effects on parents.
+Instead, we mostly found evidence for between-family level correlations.
+We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016].
+In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
+They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
+The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.
+
+As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
+Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
+Our findings serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general.
+We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
+
+Future research might also devote greater attention to between-family level correlations as an outcome of interest.
+These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs.
+We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors.
+For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.
+
+Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.
+At the same time, we should consider the limitations of these methods.
+The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable.
+Stable factors can also exert changing influences over time.
+As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time.
+It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points.
+Another important limitation is that within-family cross-lagged effects are time scale dependent.
+If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects.
+Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.
 
@@
 
-Our findings suggest that, in families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development.
+Our findings suggest that between-family differences in qualitative parenting behaviors explain a substantial amount of between-family differences in children's emotional adjustment.
+We found limited evidence for within-family parent- or child effects that apply to most families; instead, we found that these effects vary between families.
+Some of this variance was explained by between-family differences in joint and relative quantitative involvement.
+In families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development.
 In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment.
@@
 
-We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when joint involvement was higher.
+We found substantial evidence that, when joint involvement was higher, both parents reported engaging more in qualitative parenting behaviors that are positively associated with children's adjustment.
 Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth).
-Children's emotional adjustment, too, was lower when mothers were more involved than fathers.
+Children's emotional adjustment, too, was lower when mothers were relatively more involved than fathers.
 These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.

### 2019-11-12 rfunctions.R

@@ -1,3 +1,25 @@
+add_letters <- function(contrast_list, tab, tab_unst_sections, superscript = FALSE){
+  stars <- tab_unst_sections$`Defined parameters`[match(names(contrast_list),
+                                                        tab_unst_sections$`Defined parameters`$Parameter), ]
+  
+  for(i in 1:nrow(stars)){
+    if(any(grepl("\\*", stars[i, ]))){
+      for(c in which(grepl("\\*", stars[i, ]))){
+        if(superscript){
+          tab[contrast_list[[i]], c] <- paste0(tab[contrast_list[[i]], c], "$^{", letters[letter], "}$")
+          tab[contrast_list[[i]], c] <- gsub("\\}\\$\\$\\^\\{", "", tab[contrast_list[[i]], c])
+        } else {
+          tab[contrast_list[[i]], c] <- paste0(tab[contrast_list[[i]], c], letters[letter])
+          tab[contrast_list[[i]], c] <- gsub("\\}\\$\\$\\^\\{", "", tab[contrast_list[[i]], c])
+        }
+      
+        letter <<- letter + 1
+      }
+    }
+  }
+  tab
+}
+
 rename_pars <- function(x){
   new_labels <- x
   new_labels <- gsub("\\.Within", "", new_labels)
@@ -60,6 +82,24 @@ reg_table <- function(x, ...){
   tab
 }
 
+
+reg_table_escape_false <- function(x, ...){
+  rownames(x) <- NULL
+  names(x) <- gsub("^.+est_sig$", "Est.", names(x))
+  names(x) <- gsub("^.+confint$", "95\\% HPDI", names(x))
+  tab <- apa_table(x, format = "latex", col_spanners = colspan_fun(x, skip = 1), escape = FALSE, ...)
+  browser()
+  tab <- gsub('(^"|"$)', '', tab)
+  tab <- gsub("^\\[1\\] ", "", tab)
+  tab <- gsub("\\n(.+?) & NA & NA & NA & NA & NA & NA & NA & NA\\\\\\\\\\n", "\\\\multicolumn\\{9\\}\\{l\\}\\{\\1\\}\\\\\\\\", tab, perl = TRUE)
+  #tab <- gsub("\\\\", "\\", tab)
+  #tab <- gsub("\\\\begin", "\\begin", tab)
+  #tab <- gsub("\\\\n(?!=orm)", "\\\\", tab, perl = TRUE)
[…]

### 2019-11-12 table_coefficients.R

@@ -47,8 +47,7 @@ order_rows <- c("Between-family level correlations",
                 "Random intercept variances",
                 "Variances.SMOTHMOTH", "Variances.SFATHFATH", "Variances.SEMOEMO",
                 "Defined parameters",
-                paste0("New.Additional.Parameters.P", 1:18)
-                
+                paste0("New.Additional.Parameters.P", 1:20)
 )
 
 label_defined_pars <- c(p1 = "F-M Within-family correlations with emo.",
@@ -68,7 +67,9 @@ label_defined_pars <- c(p1 = "F-M Within-family correlations with emo.",
                          p15 = "F-M Variance of parenting effect",
                          p16 = "F-M Variance of child effect",
                          p17 = "Mother effect - child effect on mother",
-                         p18 = "Father effect - child effect on father")
+                         p18 = "Father effect - child effect on father",
+                        p19 = "F-M Parent effects on emo.",
+                        p20 = "F-M Child effects on parenting")
 
 tab_fun <- function(x, ...){
   tables <- lapply(results, printResultsTable, keepCols = c("label", "est_sig",
@@ -111,7 +112,6 @@ tab_fun <- function(x, ...){
   names(tables)[1] <- "Parameter"
   
   tables$Parameter <- rename_pars(tables$Parameter)
-  
   tables$Parameter[match(paste0("New.Additional.Parameters.", toupper(names(label_defined_pars))), tables$Parameter)] <- label_defined_pars
   tables
 }
@@ -120,7 +120,15 @@ tab_std <- suppressWarnings(tab_fun(results, parameters = "stdyx.standardized"))
 tab_std <- tab_std[!tab_std$Parameter %in% c("Defined parameters", label_defined_pars),]
 tab_unst <- suppressMessages(tab_fun(results, parameters = "unstandardized"))
 
+# Replace minus because I accidentally reversed F and M for p19/p20 -------
 
+replace_minus <- 
+sapply(tab_unst[tab_unst$Parameter %in% c("F-M Parent effects on emo.",
+                                          "F-M Child effects on parenting"), ], gsub, pattern = "(^|\\s)0", replacement = "X0", perl = TRUE)
+replace_minus <- gsub(pattern = "-0", replacement = " 0", replace_minus)
+replace_minus <- gsub(pattern = "X0", replacement = "-0", replace_minus)
+tab_unst[tab_unst$Parameter %in% c("F-M Parent effects on emo.",
[…]

### 2019-11-13 Figure1.R
Submit revision

@@ -6,3 +6,6 @@ p <- grid.arrange(p_within, p, p_between, nrow = 1, widths = c(10,1,10))
 
 Fig1 <- p
 ggsave("Figure1.png", p)
+
+ggsave("Figure1.pdf", p)
+ggsave("Figure1.eps", p, device = "eps")

### 2019-11-13 Figure1between.R

@@ -129,7 +129,14 @@ if(any(df_nodes$shape == "oval")){
                                                                            b = .5*ellipses_b, angle = 0), fill = "white", colour = "black")
 }
 p <- p + geom_text(data = df_nodes, aes(x = x, y = y, label = label), size = text_size) 
-
+move_by = -.24
+df_cors <- data.frame(t(mapply(function(from, to){
+  c(unlist(df_nodes[df_nodes$param == from, c("y", "x")])+c(0,move_by),
+    unlist(df_nodes[df_nodes$param == to, c("y", "x")])+c(0,move_by))
+}, from = c("riM", "riM", "riE"), to = c("riE", "riF", "riF"), USE.NAMES = FALSE)))
+names(df_cors) <- c("ystart", "xstart", "ymax", "xmax")
+p <- p + geom_curve(data = df_cors, aes(x = xstart, y = ystart, xend = xmax, yend = ymax), curvature = .5, linetype = 2) 
+p <- p + scale_x_continuous(expand = expand_scale(mult = c(0, .05)))
 df_edges$text_x[c(2:3)] <- 1.5
 df_edges$text_y[c(2:3)] <- c(6.2,7.8)
 p <- p + geom_label(data = df_edges[!df_edges$label == "", ],

### 2019-11-13 Figure1within.R

@@ -109,6 +109,14 @@ if(any(df_nodes$shape == "rect")){
 if(any(df_nodes$shape == "oval")){
   p <- p + geom_point(data = df_nodes[df_nodes$shape == "oval", ], aes(x = x, y = y), fill = "white", colour = "black", size = 14, shape = 21)
 }
+move_by <- -.34
+df_cors <- data.frame(t(mapply(function(from, to){
+  c(unlist(df_nodes[df_nodes$param == from, c("y", "x")])+c(0,move_by),
+    unlist(df_nodes[df_nodes$param == to, c("y", "x")])+c(0,move_by))
+}, from = c("riM", "riM", "riE"), to = c("riE", "riF", "riF"), USE.NAMES = FALSE)))
+names(df_cors) <- c("ystart", "xstart", "ymax", "xmax")
+p <- p + geom_curve(data = df_cors, aes(x = xstart, y = ystart, xend = xmax, yend = ymax), curvature = .2, linetype = 2) 
+p <- p + scale_x_continuous(expand = expand_scale(mult = c(0, .055)))
 p <- p + geom_text(data = df_nodes, aes(x = x, y = y, label = label), size = text_size) 
 p <- p + geom_label(data = df_edges[df_edges$label == "1", ],
            aes(x = text_x, y = text_y, label = label), size = text_size, fill = "white", label.size = NA)+ theme(axis.line = element_blank(),

### 2020-03-23 revision.Rmd
Submitted Revision 1

--- 2019-11-12 Paper_shortened_pdf.Rmd
@@
 ---
-title             : "Quantitative involvement moderates longitudinal associations between mothers' and fathers' qualitative parenting and children's emotional adjustment"
+title             : "Quantitative involvement moderates the within-family interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment"
 shorttitle        : "INVOLVEMENT AND EMOTIONAL ADJUSTMENT"
@@
 abstract: |
-  The question of whether fathers and mothers play similar roles in children’s
-  emotional adjustment has been addressed using either qualitative parenting 
-  measures or  quantitative child care involvement. This
-  longitudinal study (N = 1884, age 6-12) bridges these approaches, 
-  investigating qualitative and quantitative involvement, and their interaction,
-  in relation to emotional adjustment. On average, evidence for mothers'
-  and fathers' unique roles was limited.
-  When both parents were more involved, both engaged
-  more in desirable qualitative parenting. When mothers were disproportionately
-  involved, however, parenting roles were more differentiated:
-  Fathers displayed less
-  desirable qualitative parenting, and children were less well-adjusted.
-  This suggests that all family members benefit from a more equal
-  childcare distribution.
+  Ample research has shown that children's emotional adjustment and parenting
+  are related. Several important questions remain unaddressed. First, both
+  quantitative involvement (time spent on childcare), and qualitative
+  involvement (specific parenting behaviors) have been implicated in emotional
+  adjustment development. Yet these involvement dimensions are rarely
+  investigated jointly or in interaction. Secondly, although fathers have been
+  theorized to contribute uniquely to emotional adjustment development, they
+  are not routinely included in research. Consequently, the potential
+  implications of joint- (total) and relative (who does what) quantitative
+  involvement for the interplay between qualitative parenting and emotional
+  adjustment have not been adequately researched. This longitudinal study of
+  Australian families (N = 1884, from age 6-12) addressed these questions,
+  using dynamic structural equation modeling to examine the within-family
+  interplay between qualitative parenting and emotion regulation, while
+  controlling for stable between-family differences in these constructs. Joint
+  and relative quantitative involvement, and child sex, were included as
+  between-family moderators of within-family processes. We primarily found
+  between-family correlations between qualitative involvement and emotional
+  adjustment. Evidence for within-family effects was limited: On average,
+  maternal warmth positively predicted emotional adjustment. Several child
+  effects indicated that emotional adjustment evokes changes in parenting.
+  Within-family effects varied substantially across families however,
+  especially for fathers. When joint quantitative involvement was higher, both
+  parents engaged in more desirable qualitative parenting. When mothers were
+  relatively more involved than fathers, parental influence was more
+  differentiated, fathers displayed less desirable qualitative parenting, and
+  children were less well-adjusted. This suggests that all family members
+  benefit from a more equal childcare distribution.
 
 keywords          : "involvement; emotional adjustment; fathers; parenting; longitudinal"
-wordcount         : "7356"
+wordcount         : "8634"
 
@@
 \vspace{\baselineskip}
-Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
+Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
 It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
-Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their children).
+Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child).
 Nevertheless, the existing literature has two major shortcomings.
-Firstly, in assessing linkages with children’s emotional adjustment, quantitative involvement and qualitative parenting behaviors are rarely jointly investigated.
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
 Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
-Children’s emotional adjustment might, for example, be differently impacted by a parent's warmth if that parent is generally highly involved with the day-to-day minutiae of childrearing, than when (s)he is generally absent.
-Conversely, children’s emotional adjustment might elicit different qualitative parenting behaviors from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be sanctioned.
-The present study set out to address this question.
+For example, children’s emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
+Conversely, children’s poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning.
+The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
-Secondly, although fathers have been theorized to play a unique role in emotional adjustment development, empirical work has rarely addressed the joint and relative contributions of fathers' and mothers' involvement.
-The present study overcomes these limitations by incorporating both quantitative and quantitative measures of mothers' and fathers' involvement.
-We use sophisticated methods to examine how joint and relative quantitative involvement moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
+Secondly, although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014].
+Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development.
+The present study overcomes this limitation by incorporating both parents.
+We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
 We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.
-
-## Fatherhood theory
-
-Attachment relationships are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998].
-Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
-By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
-Consequently, fathers and mothers are theorized to influence emotional adjustment development in somewhat different ways.
-
-The past decades have seen major strides in the consolidation of theories to guide fatherhood research.
-Early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987].
-Over time, the notion the fathers should share child-rearing responsibilities with mothers, adopting roles historically fulfilled by mothers, gained ground [@pleckFatherhoodIdealsUnited1997].
-Consequently, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
-
-It is now widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004].
-Nevertheless, these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
@@
 Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2018a].
-Qualitative parenting behaviors have been operationalized in terms of the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983].
-Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's socio-emotional adjustment [@khalequePerceivedParentalAcceptanceRejection2002; @morrisRoleFamilyContext2007].
-Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.
+Qualitative parenting behaviors are related to the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983].
+Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's emotional adjustment [@morrisRoleFamilyContext2007].
+Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.<!--@khalequePerceivedParentalAcceptanceRejection2002; -->
 
 The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007].
-Demandingness encompasses several parenting practices related to setting and enforcing expectations for children's behavior.
-Some of these might help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013].
+Demandingness encompasses parenting behaviors related to setting and enforcing expectations for children.
+Some of these behaviors help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013].
 We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
-Consistency has been found to correlate strongly with emotional adjustment [@brandPerceivedParentalInconsistency1990].
-Therefore we hypothesized that consistency would be positively associated with emotional adjustment, $H_{1b}$.
-There is less empirical evidence for linkages between inductive reasoning and children’s emotional adjustment.
-Given that inductive inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.
+Consistency correlates positively with emotional adjustment [@brandPerceivedParentalInconsistency1990].
+Therefore we also hypothesized that consistency would be positively associated with emotional adjustment, $H_{1b}$.
+There is less empirical evidence for links between inductive reasoning and emotional adjustment.
+Given that inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.
 
-Negative parenting, which we operationalized as parental anger, is an additional qualitative parenting dimension that is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
-Parental anger expressions negatively impact children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because parental anger is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
-We thus hypothesized that it would be negatively related to emotional adjustment, $H_{1d}$.
+Negative parenting, which we operationalized as parental anger, is a qualitative parenting dimension not originally subsumed in the parenting styles typology.
+It is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
+We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.
 
 ### Parent or child effects?
-<!-- Hier iets doen 
-Was wel echt meer control, maar rigid control oid?-->
+
 A view of the family as a dynamic system, whose members reciprocally influence one another, has become commonplace [@kerrPerceivedParentingStyle2012].
@@
 Consequently, most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could be taken as support for child effects as well as parent effects.
-We therefore hypothesize that child effects are similar in direction, and that warmth, consistency, and inductive reasoning positively predict, and anger negatively predicts, emotional adjustment, $H_{2a-d}$.
+We therefore hypothesize that child effects are similar in direction, and that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$.
 This is supported by studies that investigated child effects specifically, and found that children's maladjustment elicits less parental warmth, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
@@
 For now, we will assume that $H_{1a-d}$ and $H_{2a-d}$ apply at both the between- and within-family level.
-<!-- Hier iets over de hypotheses; dat we niet weten wat er within en wat between speelt 
-We are, however, a bit cautious about formulating more focused hypotheses about within- versus between-family effects than the one formulated above.
-Because there is no precedent for using the random-intercept cross-lagged panel model or its multilevel DSEM extension in the literature on parental involvement and emotional adjustment, there is no literature upon which to base such hypotheses.
-Our examination of between and within-family effects will therefore mainly be exploratory.
--->
 
@@
 
-Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
-Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
-Some have theorized that mothers and fathers might fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011].
-In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage and respond supportively to emotional expression to a greater extent than fathers, whereas fathers structure and sanction emotion expressions more than mothers [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
-Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
-Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
-We might therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning, $H_{3}$.
+The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997].
+Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010].
+Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014].
+Others have argued that mothers and fathers might contribute through somewhat different mechanisms.
+Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+
+In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers.
+Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
 
@@
 
+Early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987].
 Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004].
-However, some have argued that the **division** of childcare tasks is more relevant for children's adjustment [@krollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003].
+However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [@krollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003].
 Indeed, the equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001].
@@
 There are also methodological advantages to focusing on the distribution of childcare.
-Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001].
+Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving *more* total care [see @deutschPaternalParticipationChild2001].
 This confound can be mitigated by examining the effect of the distribution of childcare, while controlling for total care.
@@
 We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$.
-By contrast, we predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$, because children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
+Further, it has been argued that children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
+We therefore predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$.
 
@@
 
-Child sex differences are a relevant factor in the interplay between parenting and emotional adjustment.
+Child sex differences are relevant in the interplay between parenting and emotional adjustment.
 Parents tend to raise boys and girls differently; expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999].
@@
 Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent, $H_{8a-d}$.
-
-Furthermore, although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a].
-Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.
+Sex differences in emotional adjustment have been found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a].
+However, a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a].
+Consequently, we have no hypothesis regarding sex differences in average levels of emotional adjustment.
 
@@
 
-All analyses were conducted using Mplus 8.0 and the R-package MplusAutomation.
+Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation.
 This type of multilevel autoregressive model can only be estimated using Bayesian methods.
 We used two Gibbs-sampler chains of 50000 iterations each.
-Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
-The missing data handling is likelihood based and guarantees consistent estimation as long as the missing data is missing at random (MAR).
-We applied mildly informative inverse Wishart priors to the variance/covariance matrices, and normal priors centered on zero for all regression coefficients.
-We assessed convergence through visual inspections of the chains and posterior distributions.
+Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parameteric MCAR test resulted in $p > .05$).
+We used mildly informative inverse Wishart priors for covariance matrices, and normal priors centered on zero for regression coefficients.
+We assessed convergence through visual inspection of the chains and posterior distributions.
 We report the posterior median, and a 95% highest posterior density interval (HPDI).
@@
 It represents the most likely population values of the parameter in question.
-Results are best interpreted by examining the effect size (standardized regression coefficients and correlations), and the width of the HPDI.
+Results are best interpreted by examining the effect size (correlations and standardized regression coefficients), and the width of the HPDI.
 Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate.
-All standardized and unstandardized model parameters are reported in Supplementary Table 4.
+Supplementary Table S4 reports all standardized and unstandardized model parameters.
+All code and output are available online at [masked].
 
@@
 Random intercepts reflect the average level of each variable across all waves of the study.
-Counter to hypothesis $H_{1a}$, we found no significant correlations between parental warmth and children's emotional adjustment.
-In support of hypothesis $H_{1b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment.
-Contrary to hypothesis $H_{1c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment.
+Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment.
+In support of hypotheses $H_{1,2b}$, we found significant positive correlations between maternal and paternal consistency, and emotional adjustment.
+Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment.
 This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment.
-Finally, as hypothesized in $H_{1d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
+Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
 
@@
 Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.
-Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S5).
+Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4).
 This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.
@@
 The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
-Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S5), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
+Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
 
@@
 *: <5% of posterior has opposite sign from estimate.
-Matching letters indicate that parameters differed significantly; see Supplementary Table S5 for all difference tests.")
+Matching letters indicate that parameters differed significantly; see Supplementary Table S4 for all difference tests.")
 ```
@@
 In support of hypothesis $H_{2a}$, the intercept for the child effect of adjustment on paternal warmth was positive and significant.
-None of the other intercepts ($H_{2b-d}$)for child effects on paternal parenting were significant.
+None of the other intercepts ($H_{2b-d}$) for child effects on paternal parenting were significant.
 The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
 
-We established direction of effects by testing the significance of the difference between corresponding parent- and child effects (see Table \@ref(tab:tabcrosspath) and Supplementary Table S5).
+We established direction of effects by testing the difference between corresponding parent- and child effects (see Table \@ref(tab:tabcrosspath) and Supplementary Table S4).
 All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests.
@@
 We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.
-
-Counter to our hypothesis $H_{6a}$, we found that greater joint involvement was negatively associated with emotional adjustment.
-Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
-However, hypothesis $H_{6b}$, that children's emotional adjustment would be lower in families where mothers were relatively more involved than fathers, was supported.
+Counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment.
+Thus, children in families with highly involved parents generally displayed poorer emotional adjustment.
+Hypothesis $H_{6b}$ was supported: Children's emotional adjustment was lower in families where mothers were relatively more involved than fathers.
 
@@
 We found sex differences in the random intercepts of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)).
-Counter to hypothesis $H_{7c}$, both parents also displayed more inductive reasoning towards sons.
+Counter to hypotheses $H_{7b,c}$, there were no sex differences in consistency, and both parents displayed more inductive reasoning towards sons.
 We found no sex differences in the random intercept of emotional adjustment.
@@
 
-In terms of parenting effects, the only effect that was significant on average was that maternal warmth positively predicted emotional adjustment in most families.
-However, all within-family predictive effects significantly varied across families, and this variability was notably greater for fathers, indicating that fathers' influence differs most across families.
-Our results thus emphasize the fact that dynamic reciprocal processes vary from one family to the next.
+The only parenting effect that was significant on average, was for maternal warmth, which positively predicted emotional adjustment in most families.
+All within-family predictive effects varied significantly across families, however, which highlights the importance of capturing and explaining these differences.
+This variability was notably greater for fathers than mothers, which suggests that fathers' influence is more context-dependent.
 
-Regarding child effects, we found that, on average, well-adjusted children evoked greater warmth from both parents, as hypothesized.
-Counter to predictions, we did not find average child effects for consistency.
+Regarding child effects, as hypothesized, well-adjusted children evoked greater warmth from both parents on average.
+Counter to predictions, we did not find child effects for consistency.
 In partial support of our hypothesis, we found that children who displayed better emotional adjustment in one wave evoked greater inductive reasoning from mothers in the next wave.
-But paradoxically, and counter to oru hypothesis, we also found that children who displayed better emotional adjustment in one wave evoked greater maternal anger in the next wave.
+But paradoxically, and counter to our hypothesis, we also found that children who displayed better emotional adjustment in one wave evoked greater maternal anger in the next wave.
+Perhaps the two go hand in hand: When mothers get angry, they (eventually) also explain what provoked their anger, through inductive reasoning.
 
@@
 That means that better emotional adjustment in one wave, was likely to be followed worse adjustment in the next wave.
-The literature on expectation-based loss aversion suggests that mothers might be more angered by a child who unexpectedly displays poor adjustment, than they would be appeased by a child who displays unexpected positive adjustment [@koszegiModelReferenceDependentPreferences].
+The literature on expectation-based loss aversion suggests that mothers might be more angered by a child who unexpectedly displays poor adjustment, than they would be appeased by a child who displays unexpected positive adjustment [@koszegiModelReferenceDependentPreferences2006].
 
@@
 
-<!-- To do!! --> 
 Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development.
@@
 
-At the within-family level we found stronger associations with emotional adjustment for maternal consistency, and for paternal reasoning and warmth.
+At the within-family level, associations with emotional adjustment were stronger for maternal consistency, and for paternal reasoning and warmth.
 As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample.
@@
 For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can also be expressed in one-off interactions, such as when  disciplining a child (inductive reasoning), or during play time (warmth).
-The stronger association between inductive reasoning and emotional adjustment for fathers is in line with the classic view of the father as disciplinarian [@flouriFathersBehaviorsChildren2010].
+The stronger association with inductive reasoning for fathers is in line with the classic view of the father as disciplinarian [@flouriFathersBehaviorsChildren2010].
 
@@
 
-We found limited evidence that quantitative involvement moderated within-family parenting effects.
-One such effect is that the effect of maternal warmth on adjustment was contingent on both joint- and relative quantitative involvement.
+We found limited evidence that quantitative involvement moderates within-family parenting effects.
+We did find both joint- and relative quantitative involvement moderated the effect of maternal warmth on adjustment.
 Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers.
@@
 With regard to mothers' versus fathers' unique roles in children's emotional development, we found that, in line with the notion of the mother as a warm, supportive attachment figure, maternal warmth and consistency were associated with children's emotional adjustment.
-Evidence for fathers' unique role was relatively limited; significant within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role in the capacity of disciplinarian.
-The aforementioned interaction effects between qualitative and quantitative involvement, however, seem to suggest that - above all - the role of fathers is more context dependent than that of mothers.
+Evidence for a consistent unique role for fathers was limited, although within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role as disciplinarian.
+Effects of fathers' parenting varied significantly more than mothers', however.
+Combined with the aforementioned interaction effects between qualitative and quantitative involvement, this suggests that - above all - the role of fathers is more context dependent than that of mothers.
 

### 2020-03-23 Paper_shortened_pdf.Rmd
Address Reviewer 1 comments

@@
 \vspace{\baselineskip}
-Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
+Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
 It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
-Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their children).
+Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child).
 Nevertheless, the existing literature has two major shortcomings.
-Firstly, in assessing linkages with children’s emotional adjustment, quantitative involvement and qualitative parenting behaviors are rarely jointly investigated.
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
 Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
-Children’s emotional adjustment might, for example, be differently impacted by a parent's warmth if that parent is generally highly involved with the day-to-day minutiae of childrearing, than when (s)he is generally absent.
-Conversely, children’s emotional adjustment might elicit different qualitative parenting behaviors from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be sanctioned.
-The present study set out to address this question.
+For example, children's emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
+Conversely, children's poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children's emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning.
+The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
-Secondly, although fathers have been theorized to play a unique role in emotional adjustment development, empirical work has rarely addressed the joint and relative contributions of fathers' and mothers' involvement.
-The present study overcomes these limitations by incorporating both quantitative and quantitative measures of mothers' and fathers' involvement.
-We use sophisticated methods to examine how joint and relative quantitative involvement moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
+Secondly, although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014].
+Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development.
+The present study overcomes this limitation by incorporating both parents.
+We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
 We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.
-
-## Fatherhood theory
-
-Attachment relationships are grounded in a balance between security and exploration [@grossmannDeveloppementAttachementAdaptation1998].
-Fatherhood researchers have posited that warmth and security are more central to mother-child attachment relationships, whereas father-child relationships have been construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
-By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
-Consequently, fathers and mothers are theorized to influence emotional adjustment development in somewhat different ways.
-
-The past decades have seen major strides in the consolidation of theories to guide fatherhood research.
-Early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987].
-Over time, the notion the fathers should share child-rearing responsibilities with mothers, adopting roles historically fulfilled by mothers, gained ground [@pleckFatherhoodIdealsUnited1997].
-Consequently, definitions of father involvement evolved to incorporate *qualitative* dimensions, such as warmth and control [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
-
-It is now widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004].
-Nevertheless, these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
@@
 
-The present study aims to make several contributions.
-First, we include both mothers and fathers, in order to tease apart their common and unique roles in children's emotional adjustment development.
-Second, we incorporate both qualitative and quantitative measures of involvement, and examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
-Third, we disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative involvement and emotional adjustment.
+The present study examined whether specific qualitative parenting behaviors are associated with children's emotional adjustment, and whether these associations are moderated by quantitative involvement.
+We aimed to make several contributions.
+First, by including both mothers and fathers, we can tease apart their common and unique roles in children's emotional adjustment development.
+Second, by incorporating both qualitative and quantitative measures of involvement, we can examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
+Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can more accurately assess the direction of effects between qualitative involvement and emotional adjustment.
 Finally, we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects.
@@
 Ethics Committee registered with the `r masked("National Health and Medical Research Council")`.
-Documentation of the ethics approval is available online at `r masked("https://growingupinaustralia.gov.au/sites/default/files/certification_of_ethics_approval_for_lsac.pdf")`.
+Documentation of the ethics approval is available online [@australianinstituteoffamilystudiesGrowingAustralia2020].
 `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population.
-Complete details regarding the study design, sample, and procedure are available online at `r masked("http://data.growingupinaustralia.gov.au/")`.
+Complete details regarding the study design, sample, and procedure are available online [@australianinstituteoffamilystudiesGrowingAustralia2020].
 
-Our target sample consisted of 1884 native `r masked("Australian")`, non-minority, intact, biological families.
-At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$).
+From the total sample (4.983 focal children at the first wave), we selected the 1884 native Australian, non-minority, intact biological families that constituted our target population.
+Although minority- or immigration status and family composition are important factors, they are beyond the scope of this paper.
+At the first wave, children in our sample were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on average, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$).
 Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only).
-Income was recorded in ordered categories.
-The modal annual income for fathers was between \$26,000-\$51,999 ($N = 663$ fathers reported this value), and for mothers, \$25,999 or less per year ($N = 1152$ reported this value).
+The LSAC did not measure income as a continuous variable, but as an ordered categorical variable.
+For fathers, the modal category was "\$26,000-\$51,999" ($N = 663$ fathers reported this value), and for mothers, "\$25,999 or less per year" ($N = 1152$ reported this value).
 
 ## Measures
+
+We calculated mean scores for all study variables with multiple indicators.
+Scale descriptive statistics, number of items, attrition (indicated by the number of complete cases), and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
+The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
+We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents).
 
@@
 ## Strategy of analyses
-
-We calculated mean scores for all of the study variables.
-Scale descriptive statistics, number of items, number of complete cases, and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
-We deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents), as the difference in fit between structurally invariant models, and metrically invariant models, was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
 
@@
 
-All analyses were conducted using Mplus 8.0 and the R-package MplusAutomation.
+Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation.
 This type of multilevel autoregressive model can only be estimated using Bayesian methods.
 We used two Gibbs-sampler chains of 50000 iterations each.
-Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
-The missing data handling is likelihood based and guarantees consistent estimation as long as the missing data is missing at random (MAR).
-We applied mildly informative inverse Wishart priors to the variance/covariance matrices, and normal priors centered on zero for all regression coefficients.
-We assessed convergence through visual inspections of the chains and posterior distributions.
+Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parametric MCAR test resulted in $p > .05$).
+We used mildly informative inverse Wishart priors for covariance matrices, and normal priors centered on zero for regression coefficients.
+We assessed convergence through visual inspection of the chains and posterior distributions.
 We report the posterior median, and a 95% highest posterior density interval (HPDI).

### 2020-03-24 Paper_shortened_pdf.Rmd
Editor comments 1 and 2

@@
 Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014].
-Consequently, we hypothesize that greater joint involvement will positively attenuate both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively attenuate parental anger, $H_{4a-d}$.
+Consequently, we hypothesize that greater joint involvement will positively predict both parents' desirable parenting behaviors (warmth, consistency, and inductive reasoning), and negatively predict parental anger, $H_{4a-d}$.
 
-By the same reasoning, we might hypothesize that greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004].
+By the same reasoning, we might hypothesize that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004].
 However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
@@
 We calculated mean scores for all study variables with multiple indicators.
-Scale descriptive statistics, number of items, attrition (indicated by the number of complete cases), and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
+Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
 The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
@@
 
-## Correlations between qualitative parenting and emotional adjustment
+## Is qualitative parenting correlated with emotional adjustment?
 
@@
 At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)).
-Random intercepts reflect the average level of each variable across all waves of the study.
+These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment.
 Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment.
@@
 Contrary to hypotheses $H_{1,2c}$, we found that maternal and paternal inductive reasoning were negatively correlated with emotional adjustment.
-This means that, in families where parents use more inductive reasoning, children displayed poorer emotional adjustment.
 Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
+This means that, in families where parents displayed more consistent parenting, children had better emotional adjustment - and in families where parents engaged in more inductive reasoning or anger, children had poorer emotional adjustment.
 
@@
 We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)).
-In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between emotional adjustment and fathers' and mothers' warmth.
-In partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
+These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in children's emotional adjustment.
+In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment.
+This means that if, in one particular wave, parents displayed more warmth than usual, their children tended to display better emotional adjustment than usual.
+Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
 Finally, contrary to hypothesis $H_{1d}$, we found no significant within-family associations between children's emotional adjustment and maternal or paternal anger.
@@
 
-## Within-family predictive effects
+## Predictive effects between within-family changes in parenting and emotional adjustment
 
-At the within-family level, we examined over-time predictive effects between qualitative parenting and emotional adjustment.
-The intercepts of these random effects reflect the average standardized regression coefficient across families.
-The variances in Table \@ref(tab:tabcrosspath) indicate whether this coefficient differs in strength (and sign) between families.
-We report unstandardized variances, because the standardized variances are, by definition, set to 1.
-With regard to within-families parenting effects, we found no evidence for hypotheses $H_{1b-d}$, as none of the intercepts differed significantly from zero.
-In line with hypothesis $H_{1a}$, however, on average, a maternal warmth positively predicted children's emotional adjustment in the next wave.
-The standardized regression coefficient of this effect, and others, is interpreted as follows: A $.99$ standard deviation increase in children's emotional adjustment in the next wave (see Table \@ref(tab:tabcrosspath)).
+In our analyses, each family has family-specific regression coefficients for the cross-lagged effects between wave-to-wave fluctuations in qualitative parenting, and emotional adjustment.
+Our hypotheses $H_{1a-d}$ also apply to these within-family predictive effects.
+To test these hypotheses, we examine the average standardized regression coefficients across families (see Table \@ref(tab:tabcrosspath)).
+In technical terms, these are random intercepts of the within-family predictive effects.
+Their variances indicate whether the family-specific regression coefficient differs in strength (and sign) across families.
+With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave.
+The standardized regression coefficient of this effect is interpreted as follows: On average across families, a one standard deviation increase in maternal warmth in one wave predicts a $.99$ standard deviation increase in children's emotional adjustment in the next wave.
+We found no evidence for hypotheses $H_{1b-d}$ in terms of parenting effects.
 The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
@@
 tabparenteff$Parameter <- gsub("^\\(.+on (.+)\\)$", "\\1", tabparenteff$Parameter)
-tabparenteff$Parameter[c(1,4)] <- c("Random intercepts of parenting effects of:", "Variance of parenting effects of:")
+tabparenteff$Parameter[c(1,4)] <- c("Average parenting effect of:", "Variance of parenting effect of:")
 
@@
 tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
-tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
+tabchildeff$Parameter[c(1,4)] <- c("Average child effect on:", "Variance of child effect on:")
 tab <- rbind(tabparenteff, tabchildeff)
@@
 names(tab)[1] <- "Param."
-reg_table(tab, caption = "Standardized between-family level intercepts and unstandardized variances of within-family predictive effects", note = "Param.: Parameter.
+reg_table(tab, caption = "Standardized between-family level intercepts (average effect) and unstandardized variances of within-family predictive effects.", note = "Param.: Parameter.
 Est.: Estimate.
@@
 
-With regard to child effects, in line with hypothesis $H_{2a}$, the intercept of the cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)).
-Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave.
-Contrary to hypothesis $H_{2b}$, the intercept of the effect of emotional adjustment on maternal consistency was non-significant.
+With regard to child effects, in line with hypothesis $H_{2a}$, the average cross-lagged effect of emotional adjustment on maternal warmth was positive and significant (Table \@ref(tab:tabcrosspath)).
+Thus, on average, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more warmth than usual in the next wave.
+Contrary to hypothesis $H_{2b}$, the average effect of emotional adjustment on maternal consistency was non-significant.
 As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive.
@@
 Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more anger than usual in the next wave.
-In support of hypothesis $H_{2a}$, the intercept for the child effect of adjustment on paternal warmth was positive and significant.
+In support of hypothesis $H_{2a}$, the average child effect of adjustment on paternal warmth was positive and significant.
 None of the other intercepts ($H_{2b-d}$) for child effects on paternal parenting were significant.
-The variances of all child effects were all significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
+The variances of all child effects were significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
 
@@
 
-## Effects of quantitative involvement on parenting and adjustment
+## Does quantitative involvement predict the average level of parenting and adjustment?
 
-We examined standardized regression coefficients of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
-With regard to parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning.
+To investigate our hypotheses $H_{4-6}$, we examined the predictive effects of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+With regard to effects of joint quantitative involvement on parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, greater joint involvement predicted increased warmth, consistency, and inductive reasoning.
 Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers.
 We did not find the hypothesized negative effects of joint involvement on parental anger, $H_{4d}$.
-With regard to relative involvement, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would attenuate the desirable parenting behaviors of the more involved parent positively, and those of the less involved parent negatively.
+With regard to effects of relative quantitative involvement on parenting, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent.
 For mothers, the only significant association we found was that mothers displayed more warmth when they were relatively more involved than fathers.
 We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.
-Counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment.
-Thus, children in families with highly involved parents generally displayed poorer emotional adjustment.
+With regard to effects of quantitative involvement on emotional adjustment, counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment.
+Thus, children of highly involved parents generally displayed poorer emotional adjustment.
 Hypothesis $H_{6b}$ was supported: Children's emotional adjustment was lower in families where mothers were relatively more involved than fathers.
@@
 
-To examine interactions between involvement quality and quantity, we examined standardized regression coefficients of between-family differences in quantitative involvement on within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+To examine interactions between involvement quality and quantity, we investigated whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
 With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
-Specifically, a one-standard deviation increase in joint involvement predicted a decrease in the within-families predictive effect of maternal warmth on emotional adjustment of $-.14$ of a standard deviation of the effect (see Table \@ref(tab:tabcrosspath)).
-This same effect was positively attenuated by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
+Specifically, a one-standard deviation increase in joint involvement predicted a $-.14$ standard deviation decrease in the within-families predictive effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
+This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
 We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.
 
-With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint quantitative involvement was higher, and also when mothers' relative involvement was greater compared to fathers.
+With regard to the moderation of child effects by quantitative involvement, we found that parents' greater joint quantitative involvement, and mothers' greater relative involvement, both negatively predicted the within-family effect of adjustment on maternal warmth.
 We found no effects for consistency.
-For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when joint involvement was higher.
+For inductive reasoning, we found that greater joint involvement negatively predicted the within-family effect of adjustment on maternal inductive reasoning.
 Emotional adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
@@
 These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth.
-In families where mothers' involvement was high - either in terms of joint, or relative involvement - this positive effect was attenuated somewhat.
-Secondly, joint quantitative involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning.
+This positive effect was somewhat smaller in families where mothers' involvement was high - either in terms of joint, or relative involvement.
+Secondly, joint quantitative involvement positively moderated the link from emotional adjustment to maternal inductive reasoning.
 This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved.

### 2020-03-25 Paper_shortened_pdf.Rmd
Editor comment 6

@@
 
-We hypothesized that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
-At the between-family level, we examined correlations between the random intercepts of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)).
+We hypothesized that warmth, consistency, and inductive reasoning would correlate positively, and anger negatively, with emotional adjustment.
+At the between-family level, we examined correlations between families' average levels of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)).
 These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment.
@@
 Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
-This means that, in families where parents displayed more consistent parenting, children had better emotional adjustment - and in families where parents engaged in more inductive reasoning or anger, children had poorer emotional adjustment.
+This means that, in families where parents reported more consistency, children displayed better emotional adjustment - and in families where parents reported more inductive reasoning or anger, children showed poorer emotional adjustment.
 
@@
 
-We also examined within-family correlations, which are based on the residuals of the study variables, across all waves and families (Table \@ref(tab:tabcors)).
-These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in children's emotional adjustment.
+We also examined within-family correlations, which are based on within-family fluctuations in the study variables (Table \@ref(tab:tabcors)).
+These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in their children's emotional adjustment.
 In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment.
-This means that if, in one particular wave, parents displayed more warmth than usual, their children tended to display better emotional adjustment than usual.
+This means that if, in any particular wave, parents report more warmth than usual, their children displayed better emotional adjustment than usual.
 Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
@@
 
-In our analyses, each family has family-specific regression coefficients for the cross-lagged effects between wave-to-wave fluctuations in qualitative parenting, and emotional adjustment.
-Our hypotheses $H_{1a-d}$ also apply to these within-family predictive effects.
-To test these hypotheses, we examine the average standardized regression coefficients across families (see Table \@ref(tab:tabcrosspath)).
-In technical terms, these are random intercepts of the within-family predictive effects.
-Their variances indicate whether the family-specific regression coefficient differs in strength (and sign) across families.
+In addition to the aforementioned within-family correlations, we also estimated over-time predictive effects between within-family fluctuations in qualitative parenting, and emotional adjustment.
+Each family has family-specific regression coefficients, and across families, we estimate the intercept and variance of these cross-lagged effects.
+The intercept indicates the average of the within-family predictive effects across families, and the variance indicate whether the family-specific regression coefficient varies in strength (and sign) across families.
+Hypotheses $H_{1a-d}$ also apply to these within-family predictive effects.
+To test these hypotheses, we examine the average standardized regression coefficients (see Table \@ref(tab:tabcrosspath)).
 With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave.
@@
 
-To examine interactions between involvement quality and quantity, we investigated whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
-With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth.
-Specifically, a one-standard deviation increase in joint involvement predicted a $-.14$ standard deviation decrease in the within-families predictive effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
+We conducted exploratory analyses of interactions between involvement quality and quantity, by investigating whether between-family differences in quantitative involvement had significant effects on the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+The novelty of this research area precluded us from formulating specific hypotheses.
+With regard to parenting effects, the only significant moderation we found was for maternal warmth.
+Specifically, in families where both parents were more involved, the cross-lagged effect of maternal warmth on emotional adjustment was *lower* than in families with less-involved parents.
+Specifically, $+1SD$ in joint involvement predicted a $-.14SD$ decrease in the within-families cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
 This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
@@
 
-With regard to the moderation of child effects by quantitative involvement, we found that parents' greater joint quantitative involvement, and mothers' greater relative involvement, both negatively predicted the within-family effect of adjustment on maternal warmth.
+With regard to the moderation of child effects by quantitative involvement, we found that greater joint quantitative involvement, and mothers' greater relative involvement, both negatively moderated the within-family effect of emotional adjustment on maternal warmth.
 We found no effects for consistency.
-For inductive reasoning, we found that greater joint involvement negatively predicted the within-family effect of adjustment on maternal inductive reasoning.
-Emotional adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers.
-Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning.
+For inductive reasoning, we found that greater joint involvement negatively moderated the within-family effect of adjustment on maternal inductive reasoning.
+Greater relative involvement also negatively moderated the effect of emotional adjustment on fathers' inductive reasoning.
+Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment predicted a stronger decrease in paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning.
 We found no moderating effects for anger.

### 2020-03-25 Paper_shortened_pdf.Rmd
Editor sex comment

@@
 
-Child sex differences are relevant in the interplay between parenting and emotional adjustment.
-Parents tend to raise boys and girls differently; expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999].
-Furthermore, social learning theory posits that the same-sex parent plays a stronger role in socialization; particularly for sex-typed behavior, such as emotion display rules [@busseySocialCognitiveTheory1999].
+Child sex differences are considered to be an important covariate in the literature on parenting and emotional adjustment.
+Parents raise boys and girls differently, expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999].
+Furthermore, the same-sex parent is thought to play a stronger role in socialization [@busseySocialCognitiveTheory1999].
 Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
-We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning towards girls than boys, and more anger towards boys than girls, $H_{7a-d}$.
-Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent, $H_{8a-d}$.
-Sex differences in emotional adjustment have been found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a].
-However, a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a].
-Consequently, we have no hypothesis regarding sex differences in average levels of emotional adjustment.
+We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning, and less anger, towards girls than boys, $H_{7a-d}$.
+We also predict associations between qualitative parenting and emotional adjustment to be stronger between children and their same-sex parent, $H_{8a-d}$.
+Finally, although sex differences in emotional adjustment have been found in older children [@vanlissaRoleFathersMothers2018a], there is little evidence for such differences in young children [@else-questGenderDifferencesTemperament2006a].
 
@@
 
-We found sex differences in the random intercepts of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)).
+We found sex differences in the average levels of qualitative parenting behaviors: In partial support of $H_{7a,d}$, fathers displayed more warmth towards daughters, and both parents displayed more anger towards sons (Table \@ref(tab:tabsex)).
 Counter to hypotheses $H_{7b,c}$, there were no sex differences in consistency, and both parents displayed more inductive reasoning towards sons.
-We found no sex differences in the random intercept of emotional adjustment.
+We found no sex differences in the average level of emotional adjustment.
 Hypotheses $H_{8a-d}$ were rejected: There was no evidence for stronger associations between qualitative parenting and emotional adjustment between children and their same-sex parent.
-The only predictive effects moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning.
-Emotional adjustment was more negatively predictive of parents' inductive reasoning for boys, as compared to girls.
-In other words, if children display worse emotional adjustment than usual in one wave, parents increase inductive reasoning more strongly in the next wave for boys than for girls.
+The only predictive effects moderated by child sex were child effects of emotional adjustment on maternal and paternal inductive reasoning: When children displayed worse emotional adjustment than usual in one wave, their parents reported stronger increases in inductive reasoning in the next wave for boys than for girls.
 
@@
 
-In line with prior research, we found that parents raise boys and girls differently.
-Mothers and fathers both displayed more anger and inductive reasoning with sons than with daughters, and fathers displayed less warmth towards sons.
-These findings suggest that parents are stricter towards their sons.
-Parenting effects were not moderated by child sex, suggesting that boys and girls are similarly affected by qualitative parenting behaviors.
-Child effects were moderated by child sex, however: Emotional adjustment was more negatively predictive of both parents' inductive reasoning when children were male.
-When viewed in the context of the finding that parents were generally stricter with boys, these findings suggest that some of this greater strictness is conditional on sons demonstrating positive adjustment.
+In line with prior research, we found that both parents displayed more anger and inductive reasoning, and less warmth, towards sons than towards daughters.
+These findings suggest that parents are stricter with sons.
+Sex did not moderate parenting effects, which suggests that boys and girls are similarly affected by qualitative parenting behaviors.
+Sex did moderate child effects, however: Emotional adjustment predicted both parents' inductive reasoning more negatively when children were male.
+This finding suggests that some of parents' greater strictness toward boys is alleviated when boys are well-adjusted.
 

### 2020-03-26 Paper_shortened_pdf.Rmd
Improve discussion

@@
 # Discussion
-
-This study set out to obtain a more comprehensive and nuanced understanding of how fathers' and mothers' quantitative involvement relate to children's emotional adjustment development.
-We used innovative methods that control for stable between-family differences, in order to obtain more accurate estimates of within-family effects and investigate the direction of effects between parents and children.
+<!-- To recapitulate, we set out to investigate how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development.
+We used innovative methods to disentangle stable between-family differences from within-family fluctuations.
+-->
+This study set out to obtain a more comprehensive and nuanced understanding of how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development.
+We used innovative methods to disentangle stable between-family differences from within-family fluctuations in the study variables, allowing us to investigate the direction of effects between parents and children.
 Furthermore, we examined how joint and relative quantitative involvement predict parenting and emotional adjustment, and moderate the longitudinal interplay between qualitative parenting and emotional adjustment.
 
-## Associations between qualitative parenting and emotional adjustment
+## Correlations between qualitative parenting and emotional adjustment
 
-We found mixed evidence, at both the within- and between-family levels, for our hypotheses that warmth, consistency, and inductive reasoning would be positively associated, and anger would be negatively associated, with emotional adjustment.
-Parental warmth was not significantly correlated with adjustment at the between-family level, but changes in warmth were correlated with changes in adjustment at the within-family level.
-Consistency was positively correlated with adjustment for both parents at the between-family level, indicating that children in families with highly consistent parents tended to be more emotionally well-adjusted, and for mothers at the within-level.
-Counter to hypotheses, inductive reasoning was negatively correlated with adjustment at the between-family level, indicating that parents who engage in more inductive reasoning had children with poorer emotional adjustment.
-At the within-family level, increases in fathers' inductive reasoning were positively correlated with adjustment.
+We found mixed evidence for our hypotheses that warmth, consistency, and inductive reasoning would correlate positively, and anger negatively, with emotional adjustment.
+Parental warmth was only correlated with emotional adjustment at the within-family level.
+Consistency was positively correlated with adjustment at the between-family level for both parents, and at the within-family level for mothers.
+Correlations with inductive reasoning were, paradoxically, opposite in sign at the between- and within-family levels.
+At the between-family level, parents who engaged in more inductive reasoning had children with poorer emotional adjustment.
+At the within-family level, fathers' inductive reasoning was positively correlated with adjustment.
 Finally, for anger, we found the predicted negative associations at the between-family level only, which means that, in families where mothers or fathers display high anger, children tend to be less emotionally well-adjusted.
+Although most of these findings are in line with hypotheses based on prior research that did not disentangle between- and within-family variance, the few deviations from the expected pattern illustrate an important insight: Correlations do not necessarily occur at both of these levels, and can even be opposite in sign; a phenomenon known as the ecological fallacy [see @keijsersParentalMonitoringAdolescent2016].
 
-The only parenting effect that was significant on average, was for maternal warmth, which positively predicted emotional adjustment in most families.
-All within-family predictive effects varied significantly across families, however, which highlights the importance of capturing and explaining these differences.
-This variability was notably greater for fathers than mothers, which suggests that fathers' influence is more context-dependent.
-
-Regarding child effects, as hypothesized, well-adjusted children evoked greater warmth from both parents on average.
-Counter to predictions, we did not find child effects for consistency.
-In partial support of our hypothesis, we found that children who displayed better emotional adjustment in one wave evoked greater inductive reasoning from mothers in the next wave.
-But paradoxically, and counter to our hypothesis, we also found that children who displayed better emotional adjustment in one wave evoked greater maternal anger in the next wave.
-Perhaps the two go hand in hand: When mothers get angry, they (eventually) also explain what provoked their anger, through inductive reasoning.
+A potential explanation for the paradoxical findings for inductive reasoning might be that the construct of inductive reasoning confounds *how* parents discipline their children (sample item: "How often do you talk it over and reason with this child..."), with *how often* their child misbehaves ("...when he/she misbehaved?").
+Stable between-family differences in inductive reasoning might tap into *how often* the child misbehaves.
+The negative between-family correlations between inductive reasoning and emotional adjustment suggest that poorly adjusted children are disciplined more often.
+Within-family fluctuations, by contrast, might tap into context-specific changes in *how* parents discipline their children to a greater extent [@critchleyParentDisciplinaryBehavior2006].
+When parents use more inductive reasoning than they usually do, talking over and reasoning about misbehavior, this might benefit emotional adjustment.
 
@@
 
-We found ample evidence for within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
-This suggests that warmth and emotional adjustment reciprocally influence one another within families; potentially through causal mechanisms.
-For consistency, however, we found no within-family predictive effects, but only between-family level associations.
-This suggests that the benefits for emotional adjustment associated with greater consistency might be primarily due to stable between-family differences.
-For inductive reasoning, paradoxically, between-family differences and within-family fluctuations showed opposing patterns of associations with adjustment.
-One potential explanation for this finding might be that the construct of inductive reasoning confounds *how* parents discipline their children (sample item: "How often do you talk it over and reason with this child..."), with *how often* their child misbehaves ("...when he/she misbehaved?").
-The fact that between-family correlations between inductive reasoning and emotional adjustment were negative suggests that between-family differences in inductive reasoning primarily capture variance related to the frequency of misbehavior.
-The within-family fluctuations, by contrast, might tap into changes in how parents discipline their children to a greater extent.
+In terms of the direction of within-family predictive effects, we found ample evidence for within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
+This suggests that parental warmth and children's emotional adjustment reciprocally influence one another within families; potentially through causal mechanisms.
+There was no other evidence for parenting effects however.
+Child effects outweighed parenting effects both in number and in size.
+The direction of effects thus appears to flow primarily from children to parents.
 
-For anger, we found moderately strong between-family level correlations for both parents.
-Counter to our expectations, however, we found no evidence for within-family correlations or parenting effects.
-This suggests that the brunt of negative associations between anger and adjustment is explained by between-family differences.
-One potential explanation for the absence of within-family associations might be a ceiling effect: If the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger might no longer make much of a difference.
-We also found a paradoxical positive predictive effect of emotional adjustment on maternal anger.
-This means that, if children demonstrated better emotional adjustment than usual in one wave, their mothers tended to display more anger than usual in the next wave.
-A possible explanation might be regression to the mean: On average, emotional adjustment showed no-to-negative stability coefficients in all models.
+With regard to these child effects, on average, children who displayed better emotional adjustment in one wave evoked more warmth from both parents, and more inductive reasoning from mothers in the next wave.
+This latter finding is in line with the aforementioned within-family correlations for inductive reasoning.
+Counter to our hypothesis, we found that children who displayed better emotional adjustment in one wave evoked greater maternal anger in the next wave.
+One possible explanation is regression to the mean: On average, emotional adjustment showed no-to-negative stability coefficients in all models.
 That means that better emotional adjustment in one wave, was likely to be followed worse adjustment in the next wave.
-The literature on expectation-based loss aversion suggests that mothers might be more angered by a child who unexpectedly displays poor adjustment, than they would be appeased by a child who displays unexpected positive adjustment [@koszegiModelReferenceDependentPreferences2006].
+Looking beyond direction of effects, it is worth noting that the brunt of negative associations between anger and adjustment appears to be due to stable between-family differences.
+The sparsity of within-family effects of anger might be due to a ceiling effect: If the stable level of parental anger in a family is high enough to negatively impact children's emotional adjustment, over-time fluctuations of that anger might no longer make much of a difference.
 
-## Differences between fathers and mothers
+## Differences between mothers and fathers
+<!-- All within-family predictive effects varied significantly across families, however, which highlights the importance of capturing and explaining these differences.
+This variability was notably greater for fathers than mothers, which suggests that fathers' influence is more context-dependent.
+-->
 
-Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development.
-These differences all played out at the within-family level, however.
-Stable between-family differences in either parent's qualitative parenting were all similarly associated with children's adjustment.
+Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development at the within-family level.
+These findings are in line with existing theory [@faganShouldResearchersConceptualize2014; @paquetteTheorizingFatherchildRelationship2004].
+Stable between-family differences in mothers' and fathers' parenting behaviors were all similarly associated with children's adjustment, however.
 
-At the within-family level, associations with emotional adjustment were stronger for maternal consistency, and for paternal reasoning and warmth.
+Within-family associations of emotional adjustment with consistency were stronger for mothers, and associations with warmth and reasoning were stronger for fathers.
 As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample.
-This might help explain the stronger associations for maternal consistency - a behavior which implies, by definition, consistent repeated interactions with the child.
+This might explain why associations for maternal consistency were stronger: Spending more time with the child provides greater opportunity for consistent behavior.
 For fathers, associations were stronger for inductive reasoning and warmth - behaviors which can also be expressed in one-off interactions, such as when  disciplining a child (inductive reasoning), or during play time (warmth).
@@
 Specifically, mothers' anger, inductive reasoning, and warmth were predicted by adjustment, whereas these effects were absent for fathers.
+This is in line with some prior studies, which found that mothers are more sensitive to children's needs [see @faganShouldResearchersConceptualize2014, p.
+395].
 
-## Quantitative involvement and qualitative parenting behaviors
+## Does quantitative involvement predict qualitative parenting behaviors?
 
@@
 Fathers, in particular, reported substantially greater parental warmth when both parents were more involved.
-We found that, when mothers performed more care tasks than fathers, mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning.
+We also found that, when mothers performed more care tasks than fathers, mothers displayed more warmth, and fathers displayed less warmth, consistency, and inductive reasoning.
 This increase in maternal warmth might reflect an attempt to compensate for fathers' absence [@shinnFatherAbsenceChildren1978].
@@
 Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014].
-Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs.
-Factors related to children's objective needs - such as developmental delay, or atypical development - might drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, resulting in a negative correlation between the two.
+Another potential explanation is that the amount of quantitative care provided is driven, in part, by children's objective needs, such as atypical or delayed development.
+Such factors require increased quantitative care, and are also associated with children's emotional maladjustment, which could result in a negative correlation between the two.
 
@@
 
-## Interactions between quantitative and qualitative involvement
+## Does quantitative involvement moderate links between qualitative parenting and emotional adjustment?
 
-We found limited evidence that quantitative involvement moderates within-family parenting effects.
-We did find both joint- and relative quantitative involvement moderated the effect of maternal warmth on adjustment.
-Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers.
+We found limited evidence that quantitative involvement moderates parenting effects on emotional adjustment.
+Specifically, joint- and relative quantitative involvement moderated the effect of maternal warmth on adjustment: Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers.
 These findings suggests a compensatory effect, whereby mothers' warmth can make up for disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].
 
-Evidence that quantitative moderates child effects was somewhat stronger.
+Evidence that quantitative involvement moderates child effects was somewhat stronger.
 Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth.
@@
 
-In line with prior research, we found that both parents displayed more anger and inductive reasoning, and less warmth, towards sons than towards daughters.
-These findings suggest that parents are stricter with sons.
-Sex did not moderate parenting effects, which suggests that boys and girls are similarly affected by qualitative parenting behaviors.
-Sex did moderate child effects, however: Emotional adjustment predicted both parents' inductive reasoning more negatively when children were male.
+In line with prior research, we found indications that parents are stricter with sons than with daughters; specifically, parents displayed more anger and inductive reasoning, and less warmth, towards sons.
+Sex also moderated child effects: Emotional adjustment predicted both parents' inductive reasoning more negatively when children were male.
 This finding suggests that some of parents' greater strictness toward boys is alleviated when boys are well-adjusted.
 
-## Summary
+## Theoretical implications of disentangling within- and between-family variance
 
-To recapitulate, we set out to examine the direction of within-family effects by controlling for stable between-family differences.
-We found only limited evidence for within-family predictive effects, most of which were child effects on parents.
-Instead, we mostly found evidence for between-family level correlations.
-We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016].
-In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
-The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
-They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age.
-The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.
+The present study mostly found evidence for correlations at the between-family level; evidence for within-family predictive effects was limited, and revealed mostly child effects.
+In a previous study with a comparable design, conducted in an adolescent sample, we similarly found mostly between-family level correlations and child effects [@vanlissaRoleFathersMothers2018a; see also @keijsersParentalMonitoringAdolescent2016].
+We explained that finding in light of the relative equality of parent-adolescent relationships.
+We theorized that the between-family level correlations might reflect the consolidated influence of within-family parent effects that took place at an earlier age, and hypothesized that parenting effects would be more prevalent, and between-family correlations weaker, in younger children.
+The present findings refute these conjectures.
+The preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.
 
@@
 These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs.
-We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors.
+We should seek to understand why between-family level correlations are so prominent by comparing between-family correlations across children of different ages, including very young children, and by examining the influence of between-family covariates, including sociological factors.
 For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].

### 2020-04-10 Paper_shortened_pdf.Rmd
Final changes for resubmission

@@
 ---
-title             : "Quantitative involvement moderates the within-family interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment"
+title             : "Mothers' and fathers' quantitative and qualitative parenting in relation to children's emotional adjustment: A between- and within-family investigation"
 shorttitle        : "INVOLVEMENT AND EMOTIONAL ADJUSTMENT"
@@
 abstract: |
-  Ample research has shown that children's emotional adjustment and parenting
-  are related. Several important questions remain unaddressed. First, both
-  quantitative involvement (time spent on childcare), and qualitative
-  involvement (specific parenting behaviors) have been implicated in emotional
-  adjustment development. Yet these involvement dimensions are rarely
-  investigated jointly or in interaction. Secondly, although fathers have been
-  theorized to contribute uniquely to emotional adjustment development, they
-  are not routinely included in research. Consequently, the potential
-  implications of joint- (total) and relative (who does what) quantitative
-  involvement for the interplay between qualitative parenting and emotional
-  adjustment have not been adequately researched. This longitudinal study of
-  Australian families (N = 1884, from age 6-12) addressed these questions,
-  using dynamic structural equation modeling to examine the within-family
-  interplay between qualitative parenting and emotion regulation, while
-  controlling for stable between-family differences in these constructs. Joint
-  and relative quantitative involvement, and child sex, were included as
-  between-family moderators of within-family processes. We primarily found
-  between-family correlations between qualitative involvement and emotional
-  adjustment. Evidence for within-family effects was limited: On average,
-  maternal warmth positively predicted emotional adjustment. Several child
-  effects indicated that emotional adjustment evokes changes in parenting.
-  Within-family effects varied substantially across families however,
-  especially for fathers. When joint quantitative involvement was higher, both
-  parents engaged in more desirable qualitative parenting. When mothers were
-  relatively more involved than fathers, parental influence was more
-  differentiated, fathers displayed less desirable qualitative parenting, and
-  children were less well-adjusted. This suggests that all family members
-  benefit from a more equal childcare distribution.
+  This longitudinal study of Australian families (n = 1884, from age 6-12)
+  examined how fathers' and mothers' quantitative involvement (time spent on
+  childcare) and qualitative involvement (specific parenting behaviors) relate
+  to children's emotional adjustment development. We used dynamic structural
+  equation modeling to disentangle stable between-family differences from
+  within-family fluctuations in qualitative parenting and emotional adjustment,
+  allowing us to investigate the direction of effects between parents and
+  children. Because fathers have been theorized to contribute uniquely to
+  emotional adjustment development, we examined differences between mothers and
+  fathers, and the interplay between parent- and child sex. We further examined
+  whether between-family differences in quantitative involvement,
+  operationalized as joint (total) and relative (one parent does more)
+  involvement, predict qualitative parenting behaviors and emotional
+  adjustment, and moderate the within-family interplay between them. Our
+  results primarily revealed between-family correlations of qualitative
+  involvement with emotional adjustment. Evidence for within-family effects was
+  limited to a parenting effect of maternal warmth, and several child effects,
+  which indicated that emotional adjustment evokes changes in parenting.
+  Parenting effects varied substantially across families however, especially
+  for fathers. Between-family differences in quantitative involvement explained
+  some of these differences. When joint quantitative involvement was higher,
+  both parents engaged in more desirable qualitative parenting. When mothers
+  were relatively more involved than fathers, fathers displayed less desirable
+  qualitative parenting, and children were less well-adjusted. There was some
+  limited evidence that within-family predictive effects were moderated by
+  quantitative involvement, particularly relative involvement. These findings
+  suggest that all family members benefit from a more equal childcare
+  distribution.
 
@@
 Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
+Yet it is plausible that quantitative involvement might moderate predictive effects between qualitative parenting and emotional adjustment.
 For example, children's emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
@@
 
-Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2018a].
+Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2019].
 Qualitative parenting behaviors are related to the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983].
@@
 
-A view of the family as a dynamic system, whose members reciprocally influence one another, has become commonplace [@kerrPerceivedParentingStyle2012].
-Nevertheless, bi-directional effects are not routinely investigated in the childhood literature [@flouriFathersBehaviorsChildren2010].
-Consequently, most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could be taken as support for child effects as well as parent effects.
-We therefore hypothesize that child effects are similar in direction, and that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$.
-This is supported by studies that investigated child effects specifically, and found that children's maladjustment elicits less parental warmth, and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
-Bi-directional positive links between parental consistency and emotional adjustment have also been reported [@lenguaBidirectionalAssociationsTemperament2005].
+The family is commonly viewed as a dynamic system, whose members reciprocally influence one another [@kerrPerceivedParentingStyle2012].
+Nevertheless, bi-directional effects are not routinely investigated [@flouriFathersBehaviorsChildren2010].
+Most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could also support child effects.
+We thus hypothesize that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$.
+These predictions are in line with studies that investigated child effects specifically, and found that maladjustment elicits less warmth and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
+There is also evidence for bi-directional positive links between parental consistency and emotional adjustment [@lenguaBidirectionalAssociationsTemperament2005].
 The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011].
@@
 
-Methods commonly used to examine direction of effects between parents and children have been criticized for yielding biased results if stable between-family differences are not controlled [@hamakerCritiqueCrosslaggedPanel2015].
-Innovative methods provide better estimates of dynamic reciprocal processes within families, by controlling for stable between-family differences: The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].
+Methods commonly used to examine direction of effects between parents and children have been criticized for failing to disentangle between-family differences from within-family predictive effects [@hamakerCritiqueCrosslaggedPanel2015].
+The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018], overcome this limitation by disentangling the two levels.
 
-Several studies have re-examined the direction of effects between parents and children using these methods, but (to our knowledge) only in adolescent samples.
-One such study found that effects of parenting on adjustment that emerged using classical methods, disappeared when stable between-family differences in parenting and adjustment were included in the model [@keijsersParentalMonitoringAdolescent2016].
-Another study similarly found that most associations between maternal and paternal qualitative parenting and adolescents' emotional adjustment, were explained by between-family level correlations.
-The remaining predictive effects consisted primarily of child effects, not parenting effects [@vanlissaRoleFathersMothers2018a].
+Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples.
+One such study found that effects of parenting on adjustment that emerged using classical methods became non-significant when stable between-family differences were taken into account [@keijsersParentalMonitoringAdolescent2016].
+Another study similarly found that associations between maternal and paternal qualitative parenting and adolescents' emotional adjustment, were mostly explained by between-family level correlations.
+The remaining predictive effects consisted primarily of child effects, not parenting effects [@vanlissaRoleFathersMothers2019].
 The authors theorized that this finding was reflective of the relative equality of parent-child relationships in adolescence.
@@
 In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019].
 Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
@@
 
-Early fatherhood research primarily operationalized involvement in *quantitative* terms: Presence versus absence in divorced families, and the sharing of childcare duties in dual-earner families [@lambBiosocialPerspectivePaternal1987].
+Early fatherhood research primarily operationalized involvement in *quantitative* terms: Father absence in divorced families, and childcare sharing in dual-earner families [@lambBiosocialPerspectivePaternal1987].
 Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004].
-However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [@krollEarlyFatherInvolvement2016; @sabattiniRelationMothersFathers2004; @craigHowMothersFathers2011; @feinbergInternalStructureEcological2003].
+However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011].
 Indeed, the equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001].
@@
 Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving *more* total care [see @deutschPaternalParticipationChild2001].
-This confound can be mitigated by examining the effect of the distribution of childcare, while controlling for total care.
-Some prior research has attempted to achieve this by assessing the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016; @keizerInfluenceFathersMothers2019].
+This confound can be mitigated by estimating a latent mean and difference [@mcardleLatentVariableModeling2009].
+The latent mean reflects joint involvement; the total (average) amount of care provided to children, and tge latent difference reflects relative involvement; the extent to which the mother is more involved than the father (i.e., the childcare distribution).
 
-Alternatively, when quantitative involvement has been measured in terms of absolute involvement, one can estimate a latent mean and difference to statistically disentangle joint involvement from relative involvement [@mcardleLatentVariableModeling2009].
-Joint involvement reflects the total (average) amount of care provided to children, and relative involvement reflects the extent to which the mother is more involved than the father (i.e., the childcare distribution).
+<!-- examining the effect of the distribution of childcare, while controlling for total care.
+Some prior research has attempted to achieve this by assessing the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016; ].
+
+Alternatively, when quantitative involvement has been measured in terms of absolute involvement, one can-->
 
@@
 
-The amount of time parents spend with children is associated with the quality of parenting behaviors [@sabattiniRelationMothersFathers2004].
+Time spent with children is associated with the quality of parenting behavior [@sabattiniRelationMothersFathers2004].
 Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a].
@@
 
-Bioecological theories of development indicate that children require a certain amount of care, and who provides this care is, to some extent, flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
+Bioecological theories argue that children require a certain amount of care; who provides it is somewhat flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
 We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$.
@@
 Child sex differences are considered to be an important covariate in the literature on parenting and emotional adjustment.
-Parents raise boys and girls differently, expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more anger with boys [@strausCorporalPunishmentAmerican1999].
+Parents raise boys and girls differently, expressing more demandingness with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2019], and more anger with boys [@strausCorporalPunishmentAmerican1999].
 Furthermore, the same-sex parent is thought to play a stronger role in socialization [@busseySocialCognitiveTheory1999].
-Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a].
+Finally, mounting evidence suggests that parental warmth is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2019].
 We therefore hypothesize that parents will display more warmth, consistency, and inductive reasoning, and less anger, towards girls than boys, $H_{7a-d}$.
 We also predict associations between qualitative parenting and emotional adjustment to be stronger between children and their same-sex parent, $H_{8a-d}$.
-Finally, although sex differences in emotional adjustment have been found in older children [@vanlissaRoleFathersMothers2018a], there is little evidence for such differences in young children [@else-questGenderDifferencesTemperament2006a].
+Finally, although sex differences in emotional adjustment have been found in older children [@vanlissaRoleFathersMothers2019], there is little evidence for such differences in young children [@else-questGenderDifferencesTemperament2006a].
 
@@
 Second, by incorporating both qualitative and quantitative measures of involvement, we can examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.
-Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can more accurately assess the direction of effects between qualitative involvement and emotional adjustment.
+Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can establish at which level associations occur, and assess the within-family direction of effects between qualitative involvement and emotional adjustment.
 Finally, we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects.
@@
 We calculated mean scores for all study variables with multiple indicators.
-Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), and reliability are reported in Supplementary Table 1, and scale correlations are provided in Supplementary Table 2.
+Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), reliability, and intraclass correlations (ICCs) are reported in Supplementary Table 1; scale correlations are provided in Supplementary Table 2.
 The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
@@
 In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
-This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015].
-Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts.
+The ICCs (supplementary Table S1) indicate the proportion of variance at the between- versus within-family level.
+For warmth, consistency, and anger, the ICCs were high, ranging from [.60, .64].
+This indicates that about 60% of the variance in these parenting behaviors is due to between-family differences.
+For inductive reasoning, they were notably lower, [.46, .50].
+For emotional adjustment, they were even lower, $r = .25$.
+
+Disentangling between- and within-family level variance allows us to determine at which of these levels associations occur, and to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015].
+Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between random intercepts.
 Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
@@
 Finally, as hypothesized in $H_{1,2d}$, we found significant negative correlations between maternal and paternal anger, and emotional adjustment.
-This means that, in families where parents reported more consistency, children displayed better emotional adjustment - and in families where parents reported more inductive reasoning or anger, children showed poorer emotional adjustment.
+To sum up, in families where parents reported more consistency, children displayed better emotional adjustment - and in families where parents reported more inductive reasoning or anger, children showed poorer emotional adjustment.
 
@@
 Counter to hypothesis $H_{3}$, none of the differences between mothers and fathers in the strength of either between- or within-family level correlations were significant (see Supplementary Table S4).
-This includes the within-family correlation between warmth and emotional adjustment, whose effect size was nearly twice as large for fathers as compared to mothers.
 
-## Predictive effects between within-family changes in parenting and emotional adjustment
+## Do within-family fluctuations in parenting predict emotional adjustment or vice versa?
 
@@
 In support of hypothesis $H_{2a}$, the average child effect of adjustment on paternal warmth was positive and significant.
-None of the other intercepts ($H_{2b-d}$) for child effects on paternal parenting were significant.
+None of the other intercepts ($H_{2b-d}$) for child effects on fathers' parenting were significant, suggesting that fathers' parenting was relatively unresponsive to fluctuations in children's emotional adjustment.
 The variances of all child effects were significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
@@
 
-## Interactions between involvement quality and quantity
+## Does quantitative involvement moderate within-family predictive effects?
 
@@
 The novelty of this research area precluded us from formulating specific hypotheses.
-With regard to parenting effects, the only significant moderation we found was for maternal warmth.
-Specifically, in families where both parents were more involved, the cross-lagged effect of maternal warmth on emotional adjustment was *lower* than in families with less-involved parents.
-Specifically, $+1SD$ in joint involvement predicted a $-.14SD$ decrease in the within-families cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
+With regard to parenting effects, the only significant moderation we found was for maternal warmth: The positive effect of maternal warmth on adjustment was less pronounced in families in which both parents were highly involved.
+Specifically, a $+1SD$ increase in joint involvement predicted a $-.14SD$ decrease in the within-families cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
 This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
@@
 
-## Effects of child sex
+## Child sex differences
 
@@
 -->
-This study set out to obtain a more comprehensive and nuanced understanding of how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development.
-We used innovative methods to disentangle stable between-family differences from within-family fluctuations in the study variables, allowing us to investigate the direction of effects between parents and children.
+To recapitulate, we set out to obtain a more comprehensive and nuanced understanding of how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development.
+We used innovative methods to disentangle stable between-family differences from within-family fluctuations in the study variables, allowing us to investigate at which of these level associations occur, and establish the direction of effects between parents and children.
 Furthermore, we examined how joint and relative quantitative involvement predict parenting and emotional adjustment, and moderate the longitudinal interplay between qualitative parenting and emotional adjustment.
@@
 We found mixed evidence for our hypotheses that warmth, consistency, and inductive reasoning would correlate positively, and anger negatively, with emotional adjustment.
-Parental warmth was only correlated with emotional adjustment at the within-family level.
-Consistency was positively correlated with adjustment at the between-family level for both parents, and at the within-family level for mothers.
+Parental warmth correlated with emotional adjustment only at the within-family level.
+Consistency correlated positively with adjustment at the between-family level for both parents, and at the within-family level for mothers.
 Correlations with inductive reasoning were, paradoxically, opposite in sign at the between- and within-family levels.
@@
 Finally, for anger, we found the predicted negative associations at the between-family level only, which means that, in families where mothers or fathers display high anger, children tend to be less emotionally well-adjusted.
-Although most of these findings are in line with hypotheses based on prior research that did not disentangle between- and within-family variance, the few deviations from the expected pattern illustrate an important insight: Correlations do not necessarily occur at both of these levels, and can even be opposite in sign; a phenomenon known as the ecological fallacy [see @keijsersParentalMonitoringAdolescent2016].
+Most of these findings are in line with hypotheses based on prior research that did not disentangle between- and within-family variance.
+However, the few notable deviations from the expected pattern illustrate an important insight: Correlations do not necessarily occur at both of these levels, and can even be opposite in sign; a phenomenon known as the ecological fallacy [see @keijsersParentalMonitoringAdolescent2016].
 
@@
 
-In terms of the direction of within-family predictive effects, we found ample evidence for within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
+In terms of the direction of within-family predictive effects, we found within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
 This suggests that parental warmth and children's emotional adjustment reciprocally influence one another within families; potentially through causal mechanisms.
@@
 ## Differences between mothers and fathers
-<!-- All within-family predictive effects varied significantly across families, however, which highlights the importance of capturing and explaining these differences.
-This variability was notably greater for fathers than mothers, which suggests that fathers' influence is more context-dependent.
--->
 
-Overall, we found somewhat more evidence for associations of emotional adjustment with mothers' qualitative parenting, than with fathers' parenting, and we found some evidence for differences between mothers' and fathers' roles in emotional adjustment development at the within-family level.
-These findings are in line with existing theory [@faganShouldResearchersConceptualize2014; @paquetteTheorizingFatherchildRelationship2004].
-Stable between-family differences in mothers' and fathers' parenting behaviors were all similarly associated with children's adjustment, however.
-
-Within-family associations of emotional adjustment with consistency were stronger for mothers, and associations with warmth and reasoning were stronger for fathers.
+Correlations of parenting behaviors with children's adjustment were all similar for mothers and fathers at the between-family level.
+Within-family correlations, however, were stronger for mothers' consistency, and stronger for fathers' warmth and reasoning.
 As mentioned before, mothers tend to spend more time with their children than fathers; also in our sample.
@@
 
-With regard to parenting effects, the only difference between mothers and fathers was for the predictive effect of warmth on emotional adjustment, which was evident only for mothers.
-This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2018a].
+With regard to average parenting effects, the only difference between mothers and fathers was the predictive effect of warmth on emotional adjustment, which was evident only for mothers.
+This is in line with prior work, indicating that warmth is an important socialization mechanism for mothers [@vanlissaRoleFathersMothers2019].
+All parenting effects varied significantly across families, however,  and this variability was notably greater for fathers than mothers.
+This suggests that fathers' influence is more context-dependent than mothers' influence, and highlights the importance of exploring predictors of between-family differences in within-family processes in future research.
+
 With regard to child effects, we found substantial evidence that mothers' parenting was more responsive to children's emotional adjustment than fathers' parenting.
@@
 The present study mostly found evidence for correlations at the between-family level; evidence for within-family predictive effects was limited, and revealed mostly child effects.
-In a previous study with a comparable design, conducted in an adolescent sample, we similarly found mostly between-family level correlations and child effects [@vanlissaRoleFathersMothers2018a; see also @keijsersParentalMonitoringAdolescent2016].
+In a previous study with a comparable design, conducted in an adolescent sample, we similarly found mostly between-family level correlations and child effects [@vanlissaRoleFathersMothers2019; see also @keijsersParentalMonitoringAdolescent2016].
 We explained that finding in light of the relative equality of parent-adolescent relationships.
@@
 As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory.
+We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
 Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time.
-Our findings serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general.
-We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?
+Our findings further serve as a reminder that we should be critical of causal interpretations of research that does not account for stable between-family differences [e.g., @brandPerceivedParentalInconsistency1990].
 
-Future research might also devote greater attention to between-family level correlations as an outcome of interest.
-These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs.
+Future research should also devote greater attention to between-family level correlations as an outcome of interest.
+The intraclass correlations of parenting behaviors were high in our study, meaning that the brunt of variance in parenting behaviors is due to stable between-family differences.
+Between-family level correlations can reflect consolidated effects of parenting effects at an early age [@vanlissaRoleFathersMothers2019], or alternatively, they can reflect the influence of third variables that impact both constructs.
 We should seek to understand why between-family level correlations are so prominent by comparing between-family correlations across children of different ages, including very young children, and by examining the influence of between-family covariates, including sociological factors.
 For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
-Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.
+Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children [see also: @keizerRoleFathersIntergenerational2020].
 
@@
 A final potential limitation pertains to the generalizability of the findings.
-Our target sample consisted of native `r masked("Australian")`, non-minority, intact, biological families.
+From the total sample, we selected our target sample of intact, biological families of natural-born, non-minority, `r masked("Australian")` citizens.
 It is possible, but beyond the scope of this study, that the findings might have limited generalizability for Australian minorities, or for families in other industrialized nations.
@@
 
-Our findings suggest that between-family differences in qualitative parenting behaviors explain a substantial amount of between-family differences in children's emotional adjustment.
-We found limited evidence for within-family parent- or child effects that apply to most families; instead, we found that these effects vary between families.
-Some of this variance was explained by between-family differences in joint and relative quantitative involvement.
-In families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development.
+To recapitulate, we set out to investigate how fathers' and mothers' quantitative and qualitative involvement relate to children's emotional adjustment development.
+We used innovative methods to disentangle stable between-family differences from within-family fluctuations in qualitative parenting and emotional adjustment.
+Our results suggest that between-family differences in qualitative parenting behaviors explain a substantial amount of between-family differences in children's emotional adjustment.
+Moreover, hypotheses based on prior literature that did not disentangle between- and within-family variance were often supported at only one of these levels.
+This illustrates that analyses aggregating these two levels should not be over-interpreted as reflecting within-family processes.
+
+We found only limited evidence for within-family parenting effects that were significant on average, across families.
+Even at a relatively young age, child effects outweighed parenting effects in number and size.
+The direction of effects thus appears to flow primarily from children to parents.
+We did find that parenting effects varied substantially between families.
+Some of this variance was explained by between-family differences in joint and relative quantitative involvement: In families where mothers were more involved than fathers, there was greater differentiation between the roles mothers' versus fathers' qualitative parenting played in children's emotional adjustment development.
 In such families, maternal warmth was more strongly predictive of, and less dependent on, children's emotional adjustment, and fathers' inductive reasoning - a disciplinary strategy - was more contingent on children's emotional adjustment.
@@
 
-We found substantial evidence that, when joint involvement was higher, both parents reported engaging more in qualitative parenting behaviors that are positively associated with children's adjustment.
+We also found substantial evidence that, when joint involvement was higher, both parents engaged more in parenting behaviors positively associated with children's adjustment.
 Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth).
@@
 Evidence for a consistent unique role for fathers was limited, although within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role as disciplinarian.
-Effects of fathers' parenting varied significantly more than mothers', however.
+Fathers' parenting effects varied significantly more across families than mothers', however.
 Combined with the aforementioned interaction effects between qualitative and quantitative involvement, this suggests that - above all - the role of fathers is more context dependent than that of mothers.

### 2020-04-10 missingness_and_iccs.R

@@ -1,0 +1,14 @@
+library(MissMech)
+df <- read.table("Mplus/03-07-18_paper7.dat", stringsAsFactors = FALSE, na.strings =  ".")
+MissMech::TestMCARNormality(df[, -1])
+
+names(df) <- c("family", "sex", "time", "Man", "Fan", "Mco", "Fco", "Min", "Fin", "Mre", "Fre", "Mwa", "Fwa", "emo")
+
+tmp <- df[,c(1,3:7, 10:14)]
+tmp <- reshape(tmp, direction = "wide", v.names = names(tmp)[-c(1,2)], timevar = "time", idvar = "family")
+
+iccs <- sapply(gsub("\\.1", "", grep("\\.1", names(tmp), value = T)), function(i){
+  psych::ICC(tmp[, grep(i, names(tmp))])$results$ICC[2]
+})
+
+write.table(rep(iccs, each = 4), "clipboard", sep  ="\t")

### 2020-05-11 Revision_2.Rmd
version 1

--- 2020-04-10 Paper_shortened_pdf.Rmd
@@
   - id            : "1"
-    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
+    institution   : "Utrecht University, department of Methodology & Statistics"
   - id            : "2"
@@
   involvement, predict qualitative parenting behaviors and emotional
-  adjustment, and moderate the within-family interplay between them. Our
-  results primarily revealed between-family correlations of qualitative
-  involvement with emotional adjustment. Evidence for within-family effects was
+  adjustment, and moderate the within-family interplay between them. Results
+  revealed between-family correlations of qualitative
+  involvement with emotional adjustment. Evidence for significant average
+  within-family effects was
   limited to a parenting effect of maternal warmth, and several child effects,
@@
 lineno            : yes
-mask              : yes
+mask              : no
+urlcolor          : blue
 
@@
 \vspace{\baselineskip}
-Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
+Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
 It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
-Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child).
+Ample research has demonstrated associations between children's emotional adjustment and parental involvement; both quantitative involvement (the amount of time parents spend caring for their child), and qualitative involvement (specific parenting behaviors).
 Nevertheless, the existing literature has two major shortcomings.
-Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
-Yet it is plausible that quantitative involvement might moderate predictive effects between qualitative parenting and emotional adjustment.
-For example, children's emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
-Conversely, children's poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children's emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning.
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
 The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
 
-Secondly, although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014].
-Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development.
+Secondly, although fathers have been theorized to play a unique role in emotional adjustment development [@paquetteTheorizingFatherchildRelationship2004], they are not consistently included in research [@faganShouldResearchersConceptualize2014].
+Consequently, little is known about mothers' and fathers' joint and unique contributions to emotional adjustment development.
 The present study overcomes this limitation by incorporating both parents.
@@
 
-Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2019].
-Qualitative parenting behaviors are related to the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983].
-Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's emotional adjustment [@morrisRoleFamilyContext2007].
+Early developmental research focused on parental socialization; the notion that parents shape child development through qualitative parenting dimensions, such as responsiveness and demandingness [@maccobySocializationContextFamily1983].
+Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting behaviors, is positively associated with children's emotional adjustment [@morrisRoleFamilyContext2007].
 Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.<!--@khalequePerceivedParentalAcceptanceRejection2002; -->
@@
 The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007].
-Demandingness encompasses parenting behaviors related to setting and enforcing expectations for children.
+Demandingness encompasses parenting behaviors related to rule-setting and enforcement.
 Some of these behaviors help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013].
 We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments.
-Consistency correlates positively with emotional adjustment [@brandPerceivedParentalInconsistency1990].
-Therefore we also hypothesized that consistency would be positively associated with emotional adjustment, $H_{1b}$.
-There is less empirical evidence for links between inductive reasoning and emotional adjustment.
+As prior research found positive correlations between consistency and emotional adjustment [@brandPerceivedParentalInconsistency1990], we also hypothesized to find these positive associations, $H_{1b}$.
+There is less evidence for links between inductive reasoning and emotional adjustment.
 Given that inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.
 
-Negative parenting, which we operationalized as parental anger, is a qualitative parenting dimension not originally subsumed in the parenting styles typology.
-It is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+Finally, negative parenting, which we operationalized as parental anger, is a qualitative parenting behavior negatively assosicated with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
 Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
@@
 
-The family is commonly viewed as a dynamic system, whose members reciprocally influence one another [@kerrPerceivedParentingStyle2012].
-Nevertheless, bi-directional effects are not routinely investigated [@flouriFathersBehaviorsChildren2010].
-Most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could also support child effects.
+Contemporary theory views the family as a dynamic system, whose members reciprocally influence one another [@kerrPerceivedParentingStyle2012].
+Nevertheless, child effects on parents are not routinely investigated [@flouriFathersBehaviorsChildren2010].
+As most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, it could also support child effects.
 We thus hypothesize that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$.
-These predictions are in line with studies that investigated child effects specifically, and found that maladjustment elicits less warmth and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
+These predictions are in line with studies that specifically investigated child effects, which found that maladjustment elicits less warmth and more negative parenting [@gadeyneLongitudinalRelationsParenting2004].
 There is also evidence for bi-directional positive links between parental consistency and emotional adjustment [@lenguaBidirectionalAssociationsTemperament2005].
 The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011].
-Thus, in a way, parents respond positively to emotional adjustment, and negatively to maladjustment.
+This suggests that parents respond positively to emotional adjustment, and negatively to maladjustment.
 <!-- Neppl 2009: Adolescents' harsh behavior towards parents predicts harsh parenting, and positive behavior predicts positive parenting -->
@@
 Methods commonly used to examine direction of effects between parents and children have been criticized for failing to disentangle between-family differences from within-family predictive effects [@hamakerCritiqueCrosslaggedPanel2015].
-The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018], overcome this limitation by disentangling the two levels.
-
+This limitation is addressed by the random-intercept cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and the dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018].
 Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples.
@@
 The authors theorized that this finding was reflective of the relative equality of parent-child relationships in adolescence.
-If this is correct, we would expect to find more evidence of parenting effects in the present, younger, sample.
-These two studies also illustrate that applying these new methods changes the pattern of results in unexpected ways.
+If this is correct, we should expect more evidence of parenting effects in the present, younger, sample.
+However, these prior studies demonstrate that applying these new methods changes the pattern of results in unexpected ways.
 Consequently, it is impossible to make specific predictions about the pattern of between-family correlations and within-family predictive effects.
@@
 
-The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997].
+The notion that fathers should share child-rearing responsibilities with mothers has become commonplace in the Western world [@pleckFatherhoodIdealsUnited1997].
 Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
@@
 Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
-By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, which expands children's emotional repertoire and exercises their regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
 
-In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019].
+In line with this theory, research has found that mothers' respond more warmly and supportively to emotional expression, whereas fathers are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+In adolescence, maternal warmth was a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019].
 Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
 Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers.
-Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
+Based on this research, we hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
 
@@
 However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011].
-Indeed, the equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001].
+The equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001].
 Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006].
@@
 
-By the same reasoning, we might hypothesize that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004].
-However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
-Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
-One reason might be that parents are less able to monitor each other's child-rearing behavior when they are not equally involved [@mclanahanConsequencesFatherAbsence2014].
-Additionally, differential involvement might negatively impact the marital relationship, and thereby undermine both parents' involvement quality.
-Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001].
-An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental].
-When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious.
+By the same reasoning, we hypothesize that greater differential involvement will positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004].
+However, there are also reasons to expect that greater differential involvement will be negatively associated with both parents' parenting quality, and with children's emotional adjustment.
+Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and disengaged parenting in fathers [@sabattiniRelationMothersFathers2004].
+One reason might be that parents are less able to monitor each other when they are not equally involved [@mclanahanConsequencesFatherAbsence2014].
+Furthermore, an unfair distribution of the demanding task childcare strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental; @deutschPaternalParticipationChild2001].
+When care is distributed equally, mothers are more satisfied [@pharesFathersMothersInvolvement2009], and the family system is more harmonious.
 Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].
@@
 
-Bioecological theories argue that children require a certain amount of care; who provides it is somewhat flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
+Bioecological theories posit that children require a certain amount of care; who provides it is somewhat flexible [see @bronfenbrennerEcologyHumanDevelopment1979].
 We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$.
-Further, it has been argued that children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
+Further, it has been argued that children feel less supported when one parent is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001].
 We therefore predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$.
@@
 Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003].
-Consequently, it is not possible to substantiate specific hypotheses for such interactions.
 One notable exception is Feinberg's coparenting model [@feinbergInternalStructureEcological2003].
@@
 According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of relative involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs.
-Applying this insight from the coparenting literature, we might expect parents' quantitative involvement to moderate effects between qualitative parenting behaviors and emotional adjustment.
-Due to the limited literature, it is not possible to formulate specific hypotheses about the nature of these moderating effects.
-There is, however, preliminary evidence to indicate that interactions between involvement quality and quantity are relevant.
+Based on this insight, we might expect parents' quantitative involvement to moderate effects between qualitative parenting behaviors and emotional adjustment.
+As an example, parental anger might affect children's emotional adjustment differently when coming from a highly involved parent than from an absent parent.
+Conversely, children's poor emotional adjustment might elicit warm responses from hands-on parents, who are attuned to the reasons for these emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined.
+Preliminary evidence indicates that interactions between involvement quality and quantity are indeed relevant.
 For instance, fathers' playful and caring involvement positively predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012].
 Furthermore, the quantity of paternal school-related involvement interacted with paternal warmth to positively predict adolescents' academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014].
-We thus explored the moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.
+However, this literature is too limited to formulate specific hypotheses about moderating effects in the present study.
+We thus conduct an exploratory assessment of moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.
 
@@
 
-The present study examined whether specific qualitative parenting behaviors are associated with children's emotional adjustment, and whether these associations are moderated by quantitative involvement.
+The present study examined associations between qualitative parenting behaviors and children's emotional adjustment, and whether these associations are moderated by quantitative involvement.
 We aimed to make several contributions.
@@
 Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can establish at which level associations occur, and assess the within-family direction of effects between qualitative involvement and emotional adjustment.
-Finally, we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects.
+Finally, we examined child sex as a predictor of between-family differences and moderator of within-family effects.
 
@@
 `r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
-Ethics Committee registered with the `r masked("National Health and Medical Research Council")`.
-Documentation of the ethics approval is available online [@australianinstituteoffamilystudiesGrowingAustralia2020].
+Ethics Committee registered with the `r masked("National Health and Medical Research Council")` [@australianinstituteoffamilystudiesGrowingAustralia2020].
 `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population.
@@
 
-We calculated mean scores for all study variables with multiple indicators.
-Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), reliability, and intraclass correlations (ICCs) are reported in Supplementary Table 1; scale correlations are provided in Supplementary Table 2.
+Scale descriptive statistics, attrition (which ranged from 7-14% across respondents), reliability, and intraclass correlations (ICCs) are reported in Supplementary Table 1; scale correlations are provided in Supplementary Table 2.
 The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3).
-We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents).
+We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents), and calculated mean scores for all study variables.
 
@@
 
-Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001].
-Consistency and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
+Responsiveness and inductive reasoning were measured using the six-item warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and five-item inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001].
+Consistency and negative parenting were measured using the five-item consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and six-item anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000].
 These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".
@@
 We used the first wave reports as an indicator of between-family differences in quantitative involvement.
-Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].
+Both parents rated their involvement in four different child care activities on five item, five-point scales ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].
 
@@
 
-Emotional adjustment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].
+Emotional adjustment was measured using the reverse-coded five-item emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].
 
@@
 
-To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018].
-In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
-The ICCs (supplementary Table S1) indicate the proportion of variance at the between- versus within-family level.
-For warmth, consistency, and anger, the ICCs were high, ranging from [.60, .64].
+To address our research questions, we used dynamic structural equation modeling [DSEM, @asparouhovDynamicStructuralEquation2018].
+In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-family differences (the mean level for that family across all measurement occasions), and within-family fluctuations (deviations from this family-specific mean).
+Intraclass correlations (supplementary Table S1) for warmth, consistency, and anger were high, ranging from [.60, .64].
 This indicates that about 60% of the variance in these parenting behaviors is due to between-family differences.
@@
 
-Disentangling between- and within-family level variance allows us to determine at which of these levels associations occur, and to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015].
-Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between random intercepts.
-Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
-In this way, we included joint and relative quantitative involvement, and child sex as moderators.
-Figure \@ref(fig:figure1) shows a graphical representation of the resulting model.
+Using DSEM allows us to determine whether associations exist at the between- or at the within-family level, and to estimate the direction of cross-lagged effects more accurately [@hamakerCritiqueCrosslaggedPanel2015].
+Within-family level cross-lagged effects are estimated for every individual family, and the average and variance of these within-family effects is estimated across all families.
+Any stable factors that might bias these within-family effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the between-family level correlations.
+To examine whether between-family level variables moderate the within-family effects, we can include them as predictors of the family-specific within-family coefficients.
+In this way, we included joint and relative quantitative involvement, and child sex, as moderators of the within-family effects.
+See Figure \@ref(fig:figure1) for a graphical representation of the model.
 We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
@@
 
-Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation.
-This type of multilevel autoregressive model can only be estimated using Bayesian methods.
-We used two Gibbs-sampler chains of 50000 iterations each.
-Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parametric MCAR test resulted in $p > .05$).
-We used mildly informative inverse Wishart priors for covariance matrices, and normal priors centered on zero for regression coefficients.
+Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation; the tables and figure were created using tidySEM.
+All code and output are available online at <https://osf.io/678rq/>.
+DSEM can only be estimated using a Bayesian estimator.
+This is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parametric MCAR test resulted in $p > .05$).
+We used two Gibbs-sampler chains of 50000 iterations each, with normal priors centered on zero for regression coefficients, and mildly informative inverse Wishart priors for covariance matrices.
 We assessed convergence through visual inspection of the chains and posterior distributions.
@@
 Supplementary Table S4 reports all standardized and unstandardized model parameters.
-All code and output are available online at [masked].
 
@@
 At the between-family level, we examined correlations between families' average levels of qualitative parenting behaviors and emotional adjustment (see Table \@ref(tab:tabcors)).
-These correlations indicate whether between-family differences in the average levels of these parenting behaviors are associated with between-family differences in children's emotional adjustment.
+These correlations indicate whether between-family differences in the average levels of parenting behaviors are associated with between-family differences in children's emotional adjustment.
 Counter to hypotheses $H_{1,2a}$, we found no significant correlations between parental warmth and children's emotional adjustment.
@@
 table_within_cor <- table_within_cor[-3, -seq(2, 11, by = 3)]
-table_within_cor <- rbind(c("Within-families level correlations", rep(NA, ncol(table_within_cor)-1)), table_within_cor)
+table_within_cor <- rbind(c("Within-family level correlations", rep(NA, ncol(table_within_cor)-1)), table_within_cor)
 
@@
 
-We also examined within-family correlations, which are based on within-family fluctuations in the study variables (Table \@ref(tab:tabcors)).
-These correlations indicate whether fluctuations in parents' parenting behaviors are correlated with fluctuations in their children's emotional adjustment.
-In support of hypothesis $H_{1a}$, we found significant positive within-family level associations between fathers' and mothers' warmth, and emotional adjustment.
-This means that if, in any particular wave, parents report more warmth than usual, their children displayed better emotional adjustment than usual.
+We also examined within-family correlations (Table \@ref(tab:tabcors)), which indicate whether within-family fluctuations in parents' parenting behaviors are correlated with fluctuations in their children's emotional adjustment.
+In support of hypothesis $H_{1a}$, we found significant positive within-family correlations between fathers' and mothers' warmth, and emotional adjustment.
+This means that if, in any particular wave, parents reported more warmth than usual, their children displayed better emotional adjustment than usual.
 Furthermore, in partial support of hypotheses $H_{1b,c}$, we found significant positive correlations of emotional adjustment with  mothers', but not fathers', consistency, as well as with fathers', but not mothers', inductive reasoning.
@@
 
-In addition to the aforementioned within-family correlations, we also estimated over-time predictive effects between within-family fluctuations in qualitative parenting, and emotional adjustment.
+We estimated over-time predictive effects between within-family fluctuations in qualitative parenting, and emotional adjustment.
 Each family has family-specific regression coefficients, and across families, we estimate the intercept and variance of these cross-lagged effects.
@@
 To test these hypotheses, we examine the average standardized regression coefficients (see Table \@ref(tab:tabcrosspath)).
-With regard to parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave.
+With regard to average parenting effects, we found that, in line with hypothesis $H_{1a}$, maternal warmth positively predicted children's emotional adjustment in the next wave.
 The standardized regression coefficient of this effect is interpreted as follows: On average across families, a one standard deviation increase in maternal warmth in one wave predicts a $.99$ standard deviation increase in children's emotional adjustment in the next wave.
-We found no evidence for hypotheses $H_{1b-d}$ in terms of parenting effects.
-The variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
-Moreover, these variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
+Counter to hypotheses $H_{1b-d}$, we found no parenting effects for consistency, inductive reasoning, or anger.
+However, the variances of all parenting effects were significant, indicating that the effect of qualitative parenting behaviors on children's emotional adjustment varies across families.
+These variances were all significantly larger for fathers than for mothers (see Supplementary Table S4), indicating that fathers have a more variable effect on their children through specific parenting behaviors than mothers do.
 
@@
 Contrary to hypothesis $H_{2b}$, the average effect of emotional adjustment on maternal consistency was non-significant.
-As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive.
-Counter to hypothesis $H_{2d}$, the intercept for maternal anger was positive, not negative.
+As hypothesized in $H_{2c}$, the intercept for maternal inductive reasoning was positive and significant.
+Counter to hypothesis $H_{2d}$, the intercept for maternal anger was positive and significant, not negative.
 Thus, when children displayed better emotional adjustment than usual in one wave, their mothers displayed more anger than usual in the next wave.
@@
 None of the other intercepts ($H_{2b-d}$) for child effects on fathers' parenting were significant, suggesting that fathers' parenting was relatively unresponsive to fluctuations in children's emotional adjustment.
-The variances of all child effects were significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differs somewhat between families.
+The variances of all child effects were significant but small, indicating that the effect of children's emotional adjustment on their parents' child-rearing behaviors differ only somewhat between families.
 
@@
 
-To investigate our hypotheses $H_{4-6}$, we examined the predictive effects of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
-With regard to effects of joint quantitative involvement on parenting, we found partial support for our hypotheses $H_{4a-c}$: For both parents, greater joint involvement predicted increased warmth, consistency, and inductive reasoning.
+To investigate hypotheses $H_{4-6}$, we examined the predictive effects of joint and relative quantitative involvement on the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)).
+With regard to effects of joint quantitative involvement on parenting, we found partial support for hypotheses $H_{4a-c}$: For both parents, greater joint involvement predicted increased warmth, consistency, and inductive reasoning.
 Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers.
@@
 With regard to effects of relative quantitative involvement on parenting, we found limited support for hypotheses $H_{5a-d}$, that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent.
-For mothers, the only significant association we found was that mothers displayed more warmth when they were relatively more involved than fathers.
+For mothers, the only significant result was that mothers displayed more warmth when they were relatively more involved than fathers.
 We found that fathers displayed less warmth, consistency, and inductive reasoning, when mothers were relatively more involved than fathers.
@@
 The novelty of this research area precluded us from formulating specific hypotheses.
-With regard to parenting effects, the only significant moderation we found was for maternal warmth: The positive effect of maternal warmth on adjustment was less pronounced in families in which both parents were highly involved.
-Specifically, a $+1SD$ increase in joint involvement predicted a $-.14SD$ decrease in the within-families cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
-This same within-families effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
+With regard to parenting effects, the only significant interaction we found was for maternal warmth: The positive effect of maternal warmth on adjustment was less pronounced in families in which both parents were highly involved.
+Specifically, a $+1SD$ increase in joint involvement predicted a $-.14SD$ decrease in the within-family cross-lagged effect of maternal warmth on emotional adjustment (see Table \@ref(tab:tabcrosspath)).
+This same within-family effect increased by $.11$ of a standard deviation of the effect when relative involvement was one standard deviation higher, i.e., when mothers were more involved than fathers.
 We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.
@@
 We used innovative methods to disentangle stable between-family differences from within-family fluctuations in the study variables, allowing us to investigate at which of these level associations occur, and establish the direction of effects between parents and children.
-Furthermore, we examined how joint and relative quantitative involvement predict parenting and emotional adjustment, and moderate the longitudinal interplay between qualitative parenting and emotional adjustment.
+Furthermore, we examined whether joint and relative quantitative involvement predict parenting and emotional adjustment, and moderate the longitudinal interplay between qualitative parenting and emotional adjustment.
 
@@
 The negative between-family correlations between inductive reasoning and emotional adjustment suggest that poorly adjusted children are disciplined more often.
-Within-family fluctuations, by contrast, might tap into context-specific changes in *how* parents discipline their children to a greater extent [@critchleyParentDisciplinaryBehavior2006].
+Within-family fluctuations, by contrast, might tap into context-specific changes in *how* parents discipline their children [@critchleyParentDisciplinaryBehavior2006].
 When parents use more inductive reasoning than they usually do, talking over and reasoning about misbehavior, this might benefit emotional adjustment.
 
-## Direction of effects
+## Direction of within-family predictive effects
 
-In terms of the direction of within-family predictive effects, we found within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
-This suggests that parental warmth and children's emotional adjustment reciprocally influence one another within families; potentially through causal mechanisms.
+We found within-family predictive effects between parental warmth and emotional adjustment, which were bi-directional for mothers.
+This suggests that parents' warmth and children's emotional adjustment reciprocally influence one another, potentially through causal mechanisms.
 There was no other evidence for parenting effects however.
@@
 All parenting effects varied significantly across families, however,  and this variability was notably greater for fathers than mothers.
-This suggests that fathers' influence is more context-dependent than mothers' influence, and highlights the importance of exploring predictors of between-family differences in within-family processes in future research.
+This suggests that fathers' influence is more context-dependent than mothers' influence, and highlights the importance of exploring predictors of within-family processes in future research.
 
@@
 Turning to our findings from linkages between quantitative involvement and qualitative parenting behaviors, our results were in line with our hypothesis that greater involvement would be associated with more desirable parenting behaviors [@sabattiniRelationMothersFathers2004].
-Our results revealed that, when both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning.
+We found that, when both parents were more involved, they displayed greater warmth, consistency, and inductive reasoning.
 Fathers, in particular, reported substantially greater parental warmth when both parents were more involved.
@@
 
-Our operationalization of quantitative involvement, in terms of joint involvement and relative involvement, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001].
+Our operationalization of quantitative involvement in terms of joint- and relative involvement allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001].
 However, counter to our hypothesis, joint involvement was *negatively* associated with between-family differences in emotional adjustment.
-One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014].
-Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014].
-Another potential explanation is that the amount of quantitative care provided is driven, in part, by children's objective needs, such as atypical or delayed development.
+Self-determination theory suggests a potential explanation: Excessive involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014].
+This explanation is supported by research on "helicopter parenting", as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014].
+Another potential explanation is that quantitative involvement is driven, in part, by children's objective needs, such as atypical or delayed development.
 Such factors require increased quantitative care, and are also associated with children's emotional maladjustment, which could result in a negative correlation between the two.
@@
 This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved.
-When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment.
+When mothers were relatively more involved than fathers, fathers increased their inductive reasoning more strongly in response to children's emotional maladjustment.
 Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.
@@
 
-In line with prior research, we found indications that parents are stricter with sons than with daughters; specifically, parents displayed more anger and inductive reasoning, and less warmth, towards sons.
+In line with prior research, we found that parents are stricter with sons than with daughters; specifically, parents displayed more anger and inductive reasoning, and less warmth, towards sons.
 Sex also moderated child effects: Emotional adjustment predicted both parents' inductive reasoning more negatively when children were male.
@@
 
-The present study mostly found evidence for correlations at the between-family level; evidence for within-family predictive effects was limited, and revealed mostly child effects.
+The present study found substantial evidence for correlations at the between-family level; evidence for within-family predictive effects was limited, and revealed mostly child effects.
 In a previous study with a comparable design, conducted in an adolescent sample, we similarly found mostly between-family level correlations and child effects [@vanlissaRoleFathersMothers2019; see also @keijsersParentalMonitoringAdolescent2016].
@@
 
-Future research should also devote greater attention to between-family level correlations as an outcome of interest.
-The intraclass correlations of parenting behaviors were high in our study, meaning that the brunt of variance in parenting behaviors is due to stable between-family differences.
-Between-family level correlations can reflect consolidated effects of parenting effects at an early age [@vanlissaRoleFathersMothers2019], or alternatively, they can reflect the influence of third variables that impact both constructs.
-We should seek to understand why between-family level correlations are so prominent by comparing between-family correlations across children of different ages, including very young children, and by examining the influence of between-family covariates, including sociological factors.
-For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
-Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children [see also: @keizerRoleFathersIntergenerational2020].
+It has been argued that within-family predictive effects are of prime theoretical relevance in developmental psychology [@keijsersParentalMonitoringAdolescent2016].
+And yet, our study and others [e.g.
+@vanlissaRoleFathersMothers2019] have found high intraclass correlations, meaning that the brunt of variance in the variables studied is due to stable between-family differences.
+This suggests that between-family level correlations should be treated as an outcome of interest.
+
+Future research could seek to understand why evidence for between-family level correlations outweighs evidence for within-family parenting effects.
+We argued elsewhere that between-family level correlations can have different causes [@vanlissaRoleFathersMothers2019].
+They might, for example, reflect the consolidated effect of within-family parental socialization at an earlier age.
+This could be tested by comparing the size of between-family correlations across children of different ages, including very young children.
+Alternatively, these correlations might reflect the influence of stable third variables that impact both constructs: Economic hardship, for instance, undermines both parents’ capacity for consistent parenting, and children’s socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+Testing this explanation would require extensive investigation of confounding variables.
+This is common in sociology, but has rarely been done in conjunction with sophisticated models of development [see @keizerRoleFathersIntergenerational2020].
+Finally, future research might shed light on the extent to which between-family level correlations reflect genetic co-variation, or shared- versus non-shared environmental effects by using genetically informed designs with twin-families or adoptive versus biological families [see @knopikBehavioralGenetics2017].
 
@@
 Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level.
-Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately.
+Compared to methods that do not distinguish between these levels, our approach estimates within-family processes more accurately.
 Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood.
@@
 Consequently, our effect sizes might be attenuated by noisy measures of emotional adjustment.
-Teachers' reports do have clear benefits as well: Young children might lack the ability to report accurately on their own emotional adjustment, and using clinical assessments might be unfeasible for large-scale panel studies.
-Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of specific combinations of these behaviors [@maccobySocializationContextFamily1983].
+Teachers' reports have benefits as well, as young children might lack the ability to report accurately on their own emotional adjustment, and clinical assessments are typically unfeasible for large-scale panel studies.
+Another limitation to consider is the fact that we operationalized qualitative parenting in terms of specific parenting behaviors, whereas others have focused on parenting *styles* composed of combinations of these behaviors [@maccobySocializationContextFamily1983].
 We focused on specific parenting behaviors for two reasons.
-First, although aggregating specific parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction.
+First, although aggregating parenting behaviors into parenting styles is theoretically defensible, it also amounts to data reduction.
 This would prevent a fine-grained investigation of differences between mothers' and fathers' unique roles in child-rearing.
-Moreover, the longitudinal models we used are not suitable for categorical variables, such as parenting style.
-Still, we have to acknowledge the possibility that the constellation of parenting behaviors is more important than each dimension in isolation.
+Moreover, the longitudinal models we used are not suitable for categorical variables.
+Still, we have to acknowledge the possibility that a constellation of parenting behaviors has different effects than the sum of its constituent dimensions.
 Finally, one potential limitation is that we only included quantitative involvement in the first year of the study.
@@
 Consequently, we are unable to assess whether within-person fluctuations in qualitative and quantitative involvement interact.
-Future research intending to examine such interactions at the within-family level might instead choose to focus on a more restricted age range.
+Future research intending to examine such interactions at the within-family level might focus on a more restricted age range, or use an age-invariant assessment of quantitative involvement.
 A final potential limitation pertains to the generalizability of the findings.
@@
 Evidence for a consistent unique role for fathers was limited, although within-family correlations between paternal inductive reasoning and emotional adjustment suggest that fathers might play a role as disciplinarian.
-Fathers' parenting effects varied significantly more across families than mothers', however.
+Parenting effects also varied significantly more across families for fathers than mothers.
 Combined with the aforementioned interaction effects between qualitative and quantitative involvement, this suggests that - above all - the role of fathers is more context dependent than that of mothers.

### 2020-05-16 prepare_synthetic_data.R
Add synthetic data file

@@ -1,0 +1,27 @@
+library(worcs)
+library(digest)
+library(tidyr)
+library(MplusAutomation)
+library(tidySEM)
+tmp <- readModels("Mplus/dsem anger revision.out", what = "input")
+df <- read.table("Mplus/03-07-18_paper7.dat", na.strings = ".")
+names(df) <- mplus_expand_names(tmp$input$variable$names)
+
+lon <- pivot_longer(df, cols = grep("^([MF](an|co|re|wa)|emo)", names(df), value = TRUE), names_to = "var", values_to = "value")
+
+wid <- pivot_wider(lon, names_from = c("var", "time"), values_from = value)
+wid[c("family", "sex")] <- lapply(wid[c("family", "sex")], factor)
+set.seed(4618)
+syn <- synthetic(wid[,-1], missingness_expression = missRanger::missRanger(data = data))
+syn <- cbind(family = 1:nrow(syn), syn)
+
+synthetic <- pivot_longer(syn, 
+                          #names_to = "time",
+                          cols = grep("^([MF](an|co|re|wa)|emo)", names(syn), value = TRUE),
+                          names_pattern = "(.{3})_(.)",
+                          names_to = c(".value", "time"))
+                          
+synthetic$time <- as.integer(synthetic$time)
+synthetic <- synthetic[, names(df)]
+
+prepareMplusData(synthetic, "Mplus/synthetic_03-07-18_paper7.dat")

### 2020-05-16 Action-letter.Rmd
Final changes for print version

--- 2019-11-12 Action letter.Rmd
@@
 bibliography: ["DSEM_2017.bib"]  
-class       : "draft"
-
+class       : "final"
+header-includes:
+   - \usepackage{graphicx}
 output      : papaja::revision_letter_pdf
@@
 
-\subsection*{EDITOR'S NOTE: Please make sure your revision is concise.
-Our guidelines for length of manuscripts are 35 pages inclusive of everything except tables and figures.
-Because your manuscript reports on a longitudinal study, we could allow a few extra pages, but please keep close to this guideline.}
-
 Dear Dr. `r editor[3]`,
@@
 
-With regard to the Editorial Office's request that we upload a Word or RTF file: In line with the Submission Guidelines on the Developmental Psychology website, we have provided \latex as a `zip`-file, with accompanying `.pdf`.
+You requested that we make sure the revision is as concise as possible, and offered that you could allow a few extra pages given the complexity of the study.
+We have made extensive efforts to shorten the manuscript.
+We count 10.510 words including references, which is 10 words more than the length indicated for a  larger research report in the [Submission guidelines](https://www.apa.org/pubs/journals/dev/?tab=4).
+We therefore hope that you can allow us this extra space.
+
+With regard to the Editorial Office's request that we upload a Word or RTF file: In line with the [Submission guidelines](https://www.apa.org/pubs/journals/dev/?tab=4), we have provided \LaTeX\  as a `zip`-file, with accompanying `.pdf`.
 We do not have a Word version of this manuscript available.
-We could manually convert this to Word, but the typographical advantages of using Latex would then be lost.
-Please let us know how to proceed.
+We tried to convert this PDF to Word, but the result was completely unacceptable.
+Please let us know if further action is required.
+Relatedly, the `zip` contains an EPS file for Figure 1, to ensure that the published version will be rendered in high quality (we have previously published a PNG figure in *Developmental Psychology*, which was illegible in print).
 
@@
 
-\WorkInProgress
+We have done as requested; the first two paragraphs now introduce the two major goals of the paper, and follow the three-point argumentation structure you suggested.
+We have marked this argumentation structure in the paragraphs below:
+
+> [...] **(1)** Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child).
+**(2)** Nevertheless, the existing literature has two major shortcomings.
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
+For example, children’s emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
+Conversely, children’s poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning.
+**(3)** The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
+  Secondly, **(1)** although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014].
+**(2)** Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development.
+**(3)** The present study overcomes this limitation by incorporating both parents.
+We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
+We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.
+
+\Done
 
@@
 
-> Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
-The missing data handling is likelihood based and guarantees consistent estimation as long as the missing data is missing at random (MAR).
+> Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parameteric MCAR test resulted in $p > .05$).
 
@@
 
-We appreciate the suggestion, and have incorporated this Figure, see below:
-
-```{r figure1, out.width="100%", fig.cap="T-1: previous measurement occasion.
-ri: random intercept.
-s: random slope.
-M: Mother's qualitative involvement.
-E: Emotional adjustment.
-F: Father's qualitative involvement.
-inv: quantitative involvement.
-Thus, sME represents the random slope of Mothers' qualitative involvement on emotional adjustment."}
-knitr::include_graphics("Figure1.png")
-```
+We appreciate the suggestion, and have incorporated this Figure in the revised manuscript (Figure 1).
 
@@
 
-> In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+> In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2019].
 The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
@@
 
-\WorkInProgress
+First, to emphasize that we are studying emotional adjustment in children, we now begin the opening paragraph with *Children's emotional adjustment is defined as...*, and we conclude the second opening paragraph with *We [...] examine [...] mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.* The developmental age is also mentioned in the paper Title, and in the Abstract.
+
+Second, we have rewritten the first two paragraphs of the *Introduction*, following a structure suggested by the Editor.
+The Editor wrote: *"Readers want to know (1) what is already known about the association between the parent-child relationship and children's adjustment?
+(2) what is not known about the association between the parent-child relationship and children's adjustment?
+(3) how will your study fill that gap in knowledge?"* We have marked this argumentation structure in the paragraphs below:
+
+> Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
+It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
+**(1)** Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child).
+**(2)** Nevertheless, the existing literature has two major shortcomings.
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
+For example, children’s emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
+Conversely, children’s poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning.
+**(3)** The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
+  Secondly, **(1)** although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014].
+**(2)** Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development.
+**(3)** The present study overcomes this limitation by incorporating both parents.
+We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
+We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.
+
+\Done
 
@@
 
-\WorkInProgress
+We have done as requested in the re-written section on *Differences between mothers and fathers*:
+
+> The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997].
+Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010].
+Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014].
+Others have argued that mothers and fathers might contribute through somewhat different mechanisms.
+Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+  In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019].
+Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers.
+Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
+
+\Done
 
@@
 
-\WorkInProgress
+We have done as requested; the first two paragraphs now introduce the two major goals of the paper, in conceptual terms:
+
+> Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001].
+It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013].
+Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child).
+Nevertheless, the existing literature has two major shortcomings.
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+Therefore, we currently do not know how the interplay between quantitative and qualitative involvement relates to emotional development.
+For example, children’s emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent.
+Conversely, children’s poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children’s emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning.
+The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
+  Secondly, although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014].
+Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development.
+The present study overcomes this limitation by incorporating both parents.
+We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years.
+We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.
+
+\Done
 
@@
 
-We appreciate the suggestion; indeed, the idea that both quantitative and qualitative involvement are important (and might interact) is most important.
-The rough division along disciplinary lines is interesting, but not central to the argument.
-We have therefore edited our argumentation to focus on the importance of jointly investigating quantitative and qualitative involvement, and removed all references to disciplinary differences.
+We appreciate the suggestion; indeed, the idea that both quantitative and qualitative involvement are relevant for emotional adjustment development (and might interact) is paramount, and the rough division along disciplinary lines is secondary.
+To address this comment, we have rewritten the argument to focus on the importance of jointly investigating quantitative and qualitative involvement, and removed all references to disciplinary differences.
 
@@
 
-\WorkInProgress
+We have done as requested; the first two paragraphs introduce the study goals, and each subsequent section walks the reader through relevant existing literature, which are then synthesized intospecific labelled hypotheses.
+
+\Done
 
@@
 *[...]*
-Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.
+Consequently, we have no hypothesis regarding sex differences in average levels of emotional adjustment.
 
@@
 
-> Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing at random.
+> Bayes is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parameteric MCAR test resulted in $p > .05$).
 
@@
 > Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects.
-In this way, we included total joint childcare, the childcare distribution, __and child sex__ as moderators.
+In this way, we included joint and relative quantitative involvement, __and child sex__ as moderators.
 
@@
 
-> We examined differences between fathers and mothers in parent- and child effects the same way [by testing the significance of the difference between corresponding mother- and father effects (see Table 2 and Supplementary Table S5)].
+>We examined differences between fathers and mothers in parent- and child effects the same way.
+[by testing the significance of the difference between corresponding mother- and father effects (see Table 2 and Supplementary Table S5)].
 All effects that were significant for one parent but not significant for the other parent had significant difference tests, except for the child effect of emotional adjustment on parental anger, which did not differ significantly between parents.
@@
 
-> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
-Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
-Some have theorized that mothers and fathers might fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011].
-In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage and respond supportively to emotional expression to a greater extent than fathers, whereas fathers structure and sanction emotion expressions more than mothers [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
-Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
-Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
-We might therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning, $H_{3}$.
+> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010].
+Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014].
+Others have argued that mothers and fathers might contribute through somewhat different mechanisms.
+Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+  In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019].
+Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers.
+Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
 
@@
 
-> Negative parenting, which we operationalized as parental anger, is an additional qualitative parenting dimension that is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
-Parental anger expressions negatively impact children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because parental anger is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
+> Negative parenting, which we operationalized as parental anger, is a qualitative parenting dimension not originally subsumed in the parenting styles typology.
+It is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
+We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.
 
@@
 
-> we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+> we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2019].
 The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
@@
 We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016].
-In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a].
+In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2019].
 The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age.
@@
   Future research might also devote greater attention to between-family level correlations as an outcome of interest.
-These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs.
+These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2019], or alternatively, they can reflect the influence of third variables that impact both constructs.
 We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors.
 For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.
 
@@
 
-> We established direction of effects by testing the significance of the difference between corresponding parent- and child effects (see Table 2 and Supplementary Table S5).
+> We established direction of effects by testing the difference between corresponding parent- and child effects (see Table 2 and Supplementary Table S5).
 All effects that were significant in one direction, and non-significant in the opposite direction, had significant difference tests.
@@
 In this way, we included joint and relative quantitative involvement, and child sex as moderators.
-Figure \@ref(fig:figure1) shows a graphical representation of the resulting model.
+Figure 1 shows a graphical representation of the resulting model.
 We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.
@@
 
-> Counter to our hypothesis $H_{6a}$, we found that greater joint involvement was negatively associated with emotional adjustment.
-Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.
-However, hypothesis $H_{6b}$, that children's emotional adjustment would be lower in families where mothers were relatively more involved than fathers, was supported.
+>  Counter to hypothesis $H_{6a}$, we found that greater joint involvement negatively predicted emotional adjustment.
+Thus, children in families with highly involved parents generally displayed poorer emotional adjustment.
+Hypothesis $H_{6b}$ was supported: Children's emotional adjustment was lower in families where mothers were relatively more involved than fathers.
 
@@
 
+<!--
 \section*{References}
@@
 \setlength{\parindent}{-0.4in}
+-->

### 2020-05-25 Action letter.Rmd

@@
 
-\WorkInProgress
+We have done as requested in the re-written section on *Differences between mothers and fathers*:
+
+> The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997].
+Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004].
+Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010].
+Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014].
+Others have argued that mothers and fathers might contribute through somewhat different mechanisms.
+Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+  In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers.
+Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
+
+\Done
 
@@
 
-> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010].
-Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers.
-Some have theorized that mothers and fathers might fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011].
-In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage and respond supportively to emotional expression to a greater extent than fathers, whereas fathers structure and sanction emotion expressions more than mothers [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
-Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
-Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
-Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship.
-We might therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning, $H_{3}$.
+> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010].
+Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014].
+Others have argued that mothers and fathers might contribute through somewhat different mechanisms.
+Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998].
+Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013].
+By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].
+  In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999].
+In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a].
+Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003].
+Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers.
+Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.
 
@@
 For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.
 

### 2020-05-25 Paper_shortened_pdf.Rmd

@@
 To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018].
-In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean).
+In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-family differences (the mean level for each family across all measurement occasions), and within-family fluctuations (within-family deviations from the family-specific mean).
 The ICCs (supplementary Table S1) indicate the proportion of variance at the between- versus within-family level.
@@
 names(tab)[1] <- "Param."
-reg_table(tab, caption = "Standardized between-family level intercepts (average effect) and unstandardized variances of within-family predictive effects.", note = "Param.: Parameter.
+reg_table(tab, caption = "Average of the standardized within-family effects and variances of unstandardized within-family effects.", note = "Param.: Parameter.
 Est.: Estimate.
@@
 tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
-tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
-                              "Effects of relative quantitative involvement on random intercepts of:")
+tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on between-family differences in:",
+                              "Effects of relative quantitative involvement on between-family differences in:")
 tabinv$Parameter <- gsub("^(.+) on (?!random).*$", "\\1", tabinv$Parameter, perl= TRUE)
@@
 tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
-tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
-                              "Moderating effects of relative involvement on parenting effects of:",
-                              "Moderating effects of joint involvement on child effects on:",
-                              "Moderating effects of relative involvement on child effects on:")
+tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on within-family parenting effects",
+                              "Moderating effects of relative involvement on within-family parenting effects",
+                              "Moderating effects of joint involvement on within-family child effects",
+                              "Moderating effects of relative involvement on within-family child effects")
 tab <- rbind(tabinv, tabinvmod)
@@
 names(tab)[1] <- "Param."
-cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
+cat(as.character(reg_table(tab, caption = "Standardized effects of joint and relative quantitative involvement.", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.
 Param.: Parameter.
@@
 tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
-tabsex$Parameter[c(1,5,8)] <- c("Sex differences in random intercepts of parenting and adjustment",
-                              "Moderation of parenting effects by sex",
-                              "Moderation of child effects by sex")
+tabsex$Parameter[c(1,5,8)] <- c("Effect of sex on between-family differences in of parenting and adjustment",
+                              "Moderating effects of sex on within-family parenting effects",
+                              "Moderating effects of sex on within-family child effects")
 #tabsex[is.na(tabsex)] <- ""
@@
 names(tab)[1] <- "Param."
-reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Child sex is a binary predictor (-1 = female, 1 = male).
+reg_table(tabsex, caption = "Standardized effects of child sex.", note = "Child sex is a binary predictor (-1 = female, 1 = male).
 The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1.

### 2020-05-25 Revision_2.Rmd

@@
     institution   : "Erasmus University Rotterdam Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"
+
+authornote: |
+  All code and output are available from the OSF (https://osf.io/678rq/).
+  The first author is supported by a NWO Veni grant (NWO
+  grant number VI.Veni.191G.090). The second author is supported by a NWO Vidi
+  Grant (NWO grant number 452-17-006).
 
@@
 Nevertheless, the existing literature has two major shortcomings.
-Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014].
+Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together, let alone in interaction [@craigHowMothersFathers2011; @faganShouldResearchersConceptualize2014].
 The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.
@@
 The equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001].
-Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006].
+Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigHowMothersFathers2011].
 This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].
@@
 @vanlissaRoleFathersMothers2019] have found high intraclass correlations, meaning that the brunt of variance in the variables studied is due to stable between-family differences.
-This suggests that between-family level correlations should be treated as an outcome of interest.
-
-Future research could seek to understand why evidence for between-family level correlations outweighs evidence for within-family parenting effects.
+These data suggest that between-family level correlations should be treated as an outcome of interest.
 We argued elsewhere that between-family level correlations can have different causes [@vanlissaRoleFathersMothers2019].
-They might, for example, reflect the consolidated effect of within-family parental socialization at an earlier age.
+They might reflect the consolidated effect of within-family parental socialization at an earlier age.
 This could be tested by comparing the size of between-family correlations across children of different ages, including very young children.
-Alternatively, these correlations might reflect the influence of stable third variables that impact both constructs: Economic hardship, for instance, undermines both parents’ capacity for consistent parenting, and children’s socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
+Alternatively, they might reflect the influence of third variables that impact both constructs: Economic hardship, for instance, undermines both parents’ capacity for consistent parenting, and children’s socio-emotional adjustment [@mcloydImpactEconomicHardship1990].
 Testing this explanation would require extensive investigation of confounding variables.
 This is common in sociology, but has rarely been done in conjunction with sophisticated models of development [see @keizerRoleFathersIntergenerational2020].
-Finally, future research might shed light on the extent to which between-family level correlations reflect genetic co-variation, or shared- versus non-shared environmental effects by using genetically informed designs with twin-families or adoptive versus biological families [see @knopikBehavioralGenetics2017].
+Finally, future research might shed light on the extent to which between-family level correlations reflect genetic co-variation, or shared- versus non-shared environmental effects, by using genetically informed designs with twin- or adoptive families [see @knopikBehavioralGenetics2017].
 

### 2020-05-25 manuscript.Rmd

--- 2020-05-25 Revision_2.Rmd
@@
 authornote: |
-  All code and output are available from the OSF (https://osf.io/678rq/).
-  The first author is supported by a NWO Veni grant (NWO
-  grant number VI.Veni.191G.090). The second author is supported by a NWO Vidi
-  Grant (NWO grant number 452-17-006).
+  All code, output, and supplemental materials are available on the OSF
+  (https://osf.io/678rq/).
+  The present study was supported by grants from the Netherlands Organization
+  for Scientific Research (NWO MaGW VIDI; grant no. 452-17-005 to R.K.; NWO
+  SGW VENI; grant no. VI.VENI.191G.090 to C.J.), and by a grant from the 
+  European Research Council (ERC StG; grant no. 757210 to R.K.).
 
@@
 
-Finally, negative parenting, which we operationalized as parental anger, is a qualitative parenting behavior negatively assosicated with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
+Finally, negative parenting, which we operationalized as parental anger, is a qualitative parenting behavior negatively associated with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999].
 Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003].
@@
 This confound can be mitigated by estimating a latent mean and difference [@mcardleLatentVariableModeling2009].
-The latent mean reflects joint involvement; the total (average) amount of care provided to children, and tge latent difference reflects relative involvement; the extent to which the mother is more involved than the father (i.e., the childcare distribution).
+The latent mean reflects joint involvement; the total (average) amount of care provided to children, and the latent difference reflects relative involvement; the extent to which the mother is more involved than the father (i.e., the childcare distribution).
 
@@
 Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation; the tables and figure were created using tidySEM.
-All code and output are available online at <https://osf.io/678rq/>.
+All code, output, and supplemental materials are available on the OSF (https://osf.io/678rq/).
 DSEM can only be estimated using a Bayesian estimator.

### 2020-05-25 test2.Rmd

---
title: "test"
author: "Caspar J. van Lissa"
date: "3/27/2020"
output: html_document
runtime: shiny
---

### Select car(s)
```{r, echo=FALSE}

## import car data
data <- head(mtcars[1:6])
data$id <- seq(1:nrow(data))

## generate random car prices
set.seed(1)
data$price <- round(runif(nrow(data),5000,15000),0)

myZ <- reactive({
    addCheckboxButtons <- paste0('<input class = "mycheckbox" type="checkbox" name="row', data$id, '" value="op', data$id, '">',"")
  #Display table with checkbox buttons
  z <- cbind(Pick=addCheckboxButtons, data)
  z
})
```

```{r}
output$mytable = renderDataTable({
  myZ()
},
    callback = "function(table) {
    table.on('click.dt', 'tr', function() {
    Shiny.onInputChange('rows',$('.mycheckbox').map(function(){return this.checked*1;}).get().join(\",\"))
    });
}")

## display table
dataTableOutput({"mytable"})
```
